### 2018-11-22 figures.tex
All files and folders added

\documentclass[12pt]{article}
  \usepackage{geometry}
 \usepackage[round]{natbib}
 \usepackage{graphicx}
 \geometry{a4paper}
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{authblk}
  \usepackage[running]{lineno}
  \usepackage{setspace}
  \usepackage{booktabs}
  \usepackage{tabularx}
  \usepackage{chngpage}
  \usepackage{amsmath}
  \usepackage{amssymb}
  \usepackage[hidelinks]{hyperref}
 \usepackage[textsize=tiny]{todonotes}

\usepackage{natbib} %package for bibliography

%   \doublespacing
% \raggedright
\title{A Meta-analysis of Longevity Estimates of Mosquito Vectors of Disease: Figures}
\author{Ben Lambert, Ace North, Charles Godfray}

\begin{document}
\maketitle

\begin{figure}[h]
	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_individualEstimates_allSpecies_withoutBalabacensis.pdf}}
	\caption{\textbf{Individual time-series estimates of adult mosquito mean lifespan from the MRR analysis ordered by species.} The middle point shows the median estimates, and the left and right box whiskers show the 25\%, and 75\% posterior quantiles respectively. All estimates were obtained using the non-hierarchical exponential survival model as described in text.}
	\label{fig:mrr_lifespan_individualEstimates}
\end{figure}

\begin{figure}[h]
	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_lifetimes_final_female_without_sugar_not_blood.pdf}}
	\caption{\textbf{Posterior estimates of female mosquito mean lifespan across species, genus and overall groupings as determined from the MRR data.} The lifespans shown are for mosquitoes that were not fed with sugar or blood before release. The middle line in each box shows the median estimates and the solid dot indicates the mean. The left and right box edges show the 25\%, and 75\% posterior quantiles respectively. The whiskers show the range of the data, excluding points lying more than 1.5 times the interquartile range away from each edge of the box. The numbers before the start of the left whisker indicate the number of individual time-series within each species. All estimates were obtained using the hierarchical exponential survival model.}
	\label{fig:mrr_lifespans}
\end{figure}

\begin{figure}[h]
	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_sexDifferences_without_sugar_nor_blood.pdf}}
	\caption{\textbf{Posterior estimates of female and male mosquito mean lifespan across species, genus and overall groupings as determined from the MRR data.} The lifespans shown are for mosquitoes that were not fed with sugar or blood (for females) before release. The middle line in each box shows the median estimates and the solid dot indicates the mean. The left and right box edges show the 25\%, and 75\% posterior quantiles respectively. The whiskers show the range of the data, excluding points lying more than 1.5 times the interquartile range away from each edge of the box. The numbers before the start of the left whisker indicate the number of individual time-series within each species. All estimates were obtained using the hierarchical exponential survival model.}
	\label{fig:mrr_sexDifferences_without_sugar_nor_blood}
\end{figure}

\begin{figure}[h]
	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_lifeSpanVsTemperature.pdf}}
	\caption{\textbf{Posterior estimates of mean mosquito lifespan for each time-series versus the average monthly temperature for each study location and date.} The markers show the median posterior estimates, with the lower and upper bounds indicating the 25\% and 75\% quantiles, respectively. The black line shows a regression line with linear and quadratic terms estimated using the median posterior lifetimes, with the grey shading indicating 95\% confidence intervals. All estimates of lifespan were obtained using the non-hierarchical exponential survival model.}\label{fig:mrr_temperature}
\end{figure}

\begin{figure}[h]
	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/dissection_lifetimes_exponential.pdf}}
	\caption{\textbf{Posterior estimates of female mosquito mean physiological lifespan across species, genus and overall groupings as determined from the dissection data.} The lifespans shown are for mosquitoes that were not fed with sugar or blood (for females) before release. The middle line in each box shows the median estimates and the solid dot indicates the mean. The left and right box edges show the 25\%, and 75\% posterior quantiles respectively. The whiskers show the range of the data, excluding points lying more than 1.5 times the interquartile range away from each edge of the box. The numbers before the start of the left whisker indicate the number of individual time-series within each species. All estimates were obtained using the hierarchical exponential survival model.}
	\label{fig:dissection_lifetimes_exponential}
\end{figure}

\begin{figure}[h]
	\centerline{\includegraphics[width=1.3\textwidth]{./Figure_files/dissection_individualEstimates_allSpecies.pdf}}
	\caption{\textbf{Individual time-series estimates of adult mosquito mean lifespan ordered by species median as determined from the dissection data.} The middle line in each box shows the median estimates. The left and right box whiskers show the 25\%, and 75\% posterior quantiles respectively. All estimates were obtained using the non-hierarchical exponential survival model.}
	\label{fig:dissection_individualEstimates_allSpecies}
\end{figure}

\begin{figure}[h]
	\centerline{\includegraphics[width=1.3\textwidth]{./Figure_files/comparison.pdf}}
	\caption{\textbf{Comparison of estimates of female mosquito lifespan between MRR and dissection data.} For the MRR data, we show estimates for females that were neither fed with sugar or blood pre-release. The red points indicate the posterior mean estimates and the whiskers show the 25\% and 75\% quantiles. In both cases, the estimates were produced using the exponential survival model.}
	\label{fig:comparison}
\end{figure}

\begin{figure}[h]
	\centerline{\includegraphics[width=1.3\textwidth]{./Figure_files/mrr_elpd_vs_exponential_ordered.pdf}}
	\caption{\textbf{Predictive accuracy of age-dependent models of mortality versus the exponential model by species for the MRR analysis.} The predictive accuracy was determined using K-Fold cross-validation as described in SOM; the measure of accuracy presented here by the central dots is the difference in estimated expected log pointwise predictive density compared to the exponential model. The lower and upper whiskers represent the lower and upper bounds of the 95\% confidence interval in predictive accuracy. The species have been ordered so that those with the highest average predictive accuracy across all age-dependent models compared to the exponential model appear at the top. If all age-dependent models outperform the exponential, we deem this as evidence for age-dependent mortality (`+'); if a subsample of models perform better, we deem the evidence ambiguous (`?'); and if no age-dependent model exceeds the performance of the exponential, we conclude no evidence in favour of senescence (`-').}
	\label{fig:mrr_elpd}
\end{figure}

\begin{figure}[h]
	\centerline{\includegraphics[width=1.3\textwidth]{./Figure_files/dissection_elpd_vs_exponential_ordered.pdf}}
	\caption{\textbf{Predictive accuracy of age-dependent models of mortality versus the exponential model by species for the dissection analysis.} The predictive accuracy was determined using K-Fold cross-validation as described in SOM; the measure of accuracy presented here by the central dots is the difference in estimated expected log pointwise predictive density compared to the exponential model. The lower and upper whiskers represent the lower and upper bounds of the 95\% confidence interval in predictive accuracy. The species have been ordered so that those with the highest average predictive accuracy across all age-dependent models compared to the exponential model appear at the top. If all age-dependent models outperform the exponential, we deem this as evidence for age-dependent mortality (`+'); if a subsample of models perform better, we deem the evidence ambiguous (`?'); and if no age-dependent model exceeds the performance of the exponential, we conclude no evidence in favour of senescence (`-').}
	\label{fig:dissection_elpd}
\end{figure}

\begin{figure}[h]
	\centerline{\includegraphics[width=1.3\textwidth]{./Figure_files/eip_all_combined.pdf}}
	\caption{\textbf{The MRR and dissection analyses estimates of the proportion of mosquitoes surviving the EIP of malaria, chikungunya, dengue fever and Zika for the disease vectors in our sample.} We assumed that the EIPs were: malaria-10 days, chikungunya-2 days, dengue fever-6.5 days and Zika-12.5 days. The left and right whiskers indicate the 25\% and 75\% posterior quantiles, and the points indicate the mean. All estimates were obtained using the exponential survival model.}
	\label{fig:eip}
\end{figure}

 \bibliographystyle{authordate1}
 \bibliography{Malaria}
\newpage

\end{document}

### 2018-11-22 supplementary_materials.tex
All files and folders added

\documentclass[12pt]{article}
  \usepackage{geometry}
 \usepackage[round]{natbib}
 \usepackage{graphicx}
 \geometry{a4paper}
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{authblk}
  \usepackage[running]{lineno}
  \usepackage{setspace}
  \usepackage{booktabs}
  \usepackage{tabularx}
  \usepackage{chngpage}
  \usepackage{amsmath}
  \usepackage{amssymb}
  \usepackage[hidelinks]{hyperref}
 \usepackage[textsize=tiny]{todonotes}
\usepackage{pdfpages, caption}
\usepackage{listings}
\usepackage{minted}
\usepackage{framed}
\usepackage{mdframed}

\usepackage{natbib} %package for bibliography

\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thesection}{S\arabic{section}}

\newenvironment{problem}[2][Problem]{\begin{trivlist}
		\item[\hskip \labelsep {\bfseries #1}\hskip \labelsep {\bfseries #2.}]}{\end{trivlist}}

\DeclareMathOperator*{\argmax}{arg\,max}

\renewcommand\listingscaption{Stan code}
\definecolor{apricot}{rgb}{0.984,0.81,0.69}
\BeforeBeginEnvironment{minted}{\begin{mdframed}[backgroundcolor=apricot]}
	\AfterEndEnvironment{minted}{\end{mdframed}}
\newcommand{\Lim}[1]{\raisebox{0.5ex}{\scalebox{0.8}{$\displaystyle \lim_{#1}\;$}}}
\newcommand*\widefbox[1]{\fbox{\hspace{2em}#1\hspace{2em}}}
\interfootnotelinepenalty=10000

\usepackage{framed}
\makeatletter
\newenvironment{kframe}{%
	\def\at@end@of@kframe{}%
	\ifinner\ifhmode%
	\def\at@end@of@kframe{\end{minipage}}%
\begin{minipage}{\columnwidth}%
	\fi\fi%
	\def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
		\colorbox{shadecolor}{##1}\hskip-\fboxsep
		% There is no \\@totalrightmargin, so:
		\hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
	\MakeFramed {\advance\hsize-\width
		\@totalleftmargin\z@ \linewidth\hsize
		\@setminipage}}%
{\par\unskip\endMakeFramed%
	\at@end@of@kframe}
\makeatother

\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%

\definecolor{orangeBright}{RGB}{243,146,0}
\definecolor{blueBright}{RGB}{54,169,225}
\definecolor{pinkBright}{RGB}{214,11,82}
\definecolor{shadecolor}{rgb}{.97, .97, .97}
\definecolor{messagecolor}{rgb}{0, 0, 0}
\definecolor{warningcolor}{rgb}{1, 0, 1}
\definecolor{errorcolor}{rgb}{1, 0, 0}
\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX

%   \doublespacing
% \raggedright
\title{A Meta-analysis of Longevity Estimates of Mosquito Vectors of Disease: Supplementary Material}
\author{Ben Lambert, Ace North, Charles Godfray}

\begin{document}
\maketitle

\section{MRR experiments}
\subsection{Data}
The MRR database in \cite{guerra2014global} contains 393 individual time-series, along with meta-data for a range of factors for each experiment (for example, species, geography, and date of study). We  cleaned the data, making a number of amendments to it (see Section \ref{sec:amendments}), and transformed the data into a form amenable to our statistical model. We removed time-series with fewer than six separate recapture observations, resulting in 238 time series. For the species-level analyses, we required at least two time series per species, resulting in loss of six series, meaning that the data used in the Bayesian hierarchical analysis comprised 232 individual time series (see Section \ref{sec:appendix_mrrStudyList} for a list of the studies included). Tables \ref{tab:mrr_aggregateData} and \ref{tab:mrr_IndividualData} summarise the data used in the Bayesian hierarchical analysis. This data encompassed time-series from MRR experiments across 33 different species and three genera: \textit{Aedes} (91 separate time-series), \textit{Anopheles} (94), and \textit{Culex} (47), and spanning a wide geographical range (Figure \ref{fig:mrr_lifetimes_map}). The data comprise 177 female-only, 35 male-only and 18 mixed-sex release time-series. Where the age of the adult mosquitoes at release-time was available (102 time-series) we use this to increment our time variable (that represents mosquito age) in the statistical models. Where age was unavailable (130 time-series), we assume that the age at release is zero. We recognise this assumption will likely underestimate mosquito age for some studies and so we interpret our estimates as lower bounds on mosquito lifespan.

\subsection{Amendments by MRR ID}\label{sec:amendments}
This section describes the amendments made by MRR ID (the unique identifier that \cite{guerra2014global} use for each MRR record). Where there was ambiguity in the data, we attempted to consult the original publication. The amendments to the dataset are described below:

\begin{itemize}
	\item 69 and 70: Since the release was not disaggregated by sex, the recapture data were aggregated across both males and females, resulting in a single time series in each case.
	\item 482: There are actually two series (one with captured mosquitoes, the other with young reared) in the paper \cite[]{rawlings1981influence}, whereas the database contains only one recapture series. In this analysis, both time series are included.
	\item 483: There are three studies in the paper \cite[]{rawlings1981influence}, (by dye colour there is one for `M\&Y', another for white, and a third for green), whereas the database only contains one release series. In this analysis, we analyse all three time series.
	\item 578: A null observation was removed on day 12, since isn't present in the original paper \cite[]{eyles1943experiment}.
	\item 585: It is not clear from the original paper \cite[]{smith1941observations} whether the missing recapture observations on certain days are because no mosquitoes were found on those days, or because no efforts were made to recapture them on those days. We have assumed that the number recaptured were as stated in the database on those days (missing, because no recapture efforts were made on those days).
	\item 586: It was unclear from database whether the missing observations are due to lack of recapture, or no effort. Have assumed that zero mosquitoes were captured on these days as seems most likely (the paper could not be accessed).
	\item 342: Changed number of released from 6,000 to 12,000 following the original publication \cite[]{bryan1991anopheles}.
	\item 246: Moved the observations from captures to recaptures following the original publication \cite[]{arredondo1998gonotrophic}.
	\item 55: Corrected a mistake in release data; the actual number released is 739, not 749 \cite[]{midega2007estimating}.
	\item 478-481: All four of these series give releases not disaggregated by gender, whereas the recaptures are split this way. Have combined the recapture series to yield a single release-recapture series in each case.
	\item 154: Moved captured male series to recaptured following original paper \cite[]{tsuda2001movement}. 
	\item 334: Data starts on day 3, thus have added this number of days to our time variable in this case.
\end{itemize}

The R file \verb|s_clean_data.R| attached cleans the data, makes the above amendments and outputs the data in a form amenable for estimation using Stan \citep{carpenter2016stan}.

\begin{figure}[h]
	\centerline{\includegraphics[width=1.25\textwidth]{./Figure_files/mrr_mapGenusCropped.pdf}}
	\caption{\textbf{The location and number of separate time-series for the MRR studies included in this analysis.} The area of each circle indicates the number of time-series at each study site. The colour shows the genus of the mosquitoes in each study.}
	\label{fig:mrr_lifetimes_map}
\end{figure}

\begin{table}[htbp]
	\centering
	\begin{tabular}{lc}
		\toprule
		\textbf{Data characteristic} & \textbf{Total} \\
		\midrule
		Number of species & 33 \\
		Number of genera & 3 \\
		Number of \textit{Anopheles} & 94 \\
		Number of \textit{Aedes} & 91 \\
		Number of \textit{Culex} & 47 \\
		Number of female-only MRR & 179 \\
		Number of male-only MRR & 35 \\
		Number of mixed sex MRR & 18 \\
		Number of pre-release blood-feeding only MRR & 71 \\
		Number of pre-release sugar-feeding only MRR & 41 \\
		Number of pre-release both blood- and sugar-feeding MRR & 4 \\
		Number of pre-release neither blood- and sugar-feeding MRR & 116 \\
		Number of MRR time-series & 232 \\
		\bottomrule
	\end{tabular}%
	\caption{Summary of variables across all MRR time-series.}
	\label{tab:mrr_aggregateData}%
\end{table}%

\begin{table}[htbp]
	\centering
	\footnotesize
	\begin{adjustwidth}{-0.5in}{-0.5in}%
		\begin{tabularx}{1.25\textwidth}{l|ccccc}
			\toprule
			\textbf{Data characteristic} & \textbf{Min} & \textbf{Mean} & \textbf{Median} & \textbf{Max} & \textbf{Standard deviation} \\
			\midrule
			Study duration, days & 6.0   & 11.8  & 10.0  & 71.0    & 4.7 \\
			Number of days on which collections took place & 6.0   & 10.1  & 9.0   & 47    & 9.1 \\
			Number of separate release days & 1.0     & 1.9   & 1.0     & 23    & 3.0 \\
			Number released & 66    & 4,929  & 1,297  & 86,200 & 12,043 \\
			Number recaptured & 2     & 163   & 63    & 4,090  & 399 \\
			Recapture percentage & 0\% & 8.6\% & 5.2\% & 57.1\% & 10.04\% \\
			Age at release, days & 0   & 1.8   & 0   & 13    & 2.8 \\
			\bottomrule
		\end{tabularx}%
		\caption{Summary of data from individual MRR time-series.}
		\label{tab:mrr_IndividualData}%
	\end{adjustwidth}
\end{table}%

\subsection{Statistical model}\label{sec:mrr_statistical}
Data for a typical MRR experiment consists of a single release of $N_{R}$ marked mosquitoes followed by a series of marked mosquito recaptures on subsequent days (Figure \ref{fig:mrr_exampleMRRSeries} shows three such example series.) We model the number $y(t)$ of marked mosquitoes recaptured on day $t$ using a negative binomial sampling model,
%
\begin{equation}\label{eq:NB}
y(t) \sim \text{NB}\left((N_{R} - Y(t-1)) S(t) \psi, \kappa\right),
\end{equation}
%
where $Y(t-1)$ is the cumulative number of mosquitoes caught on all days before $t$, $S(t)$ is the probability that an individual mosquito survives and remains in the study area until time $t$, $\psi$ is the daily recapture probability for an individual mosquito which is assumed constant through time, and $\kappa$ is the time-independent shape parameter of the negative binomial distribution that controls the extent of variance in recapture rate likely due mostly to environmental heterogeneity. We chose a parameterisation of the negative binomial such that its mean is given by $\mu(t) = (N_{R} - Y(t-1)) S(t) \psi$, and its variance by $\sigma(t)^2 = \mu(t) + \frac{\mu(t)^2}{\kappa}$.

For some MRR experiments, there were a number of releases ($q\geq 2$) of marked mosquitoes throughout the duration of the study. In contrast to single releases, with multiple releases over time, it is not in general possible to determine the particular release to which a recaptured mosquito belongs. To avoid the complication of directly inferring this quantity, we choose to represent previous recaptures probabilistically. This results in a slightly different mean to that of the single release model,
%
\begin{equation}
\mu(t) = N_{Released}(1-\psi)^{t-1} S(t) \psi,
\end{equation}
%
where the factor of $(1-\psi)^{t-1}$ represents the probability that a mosquito is \textit{not} recaptured on a previous day. Where experiments consisted of two or more releases occurring at distinct points in time we assumed that recaptures of individual mosquitoes from either batch were independent of one another, although with the same sampling parameters ($\psi$ and $\kappa$). This results in an overall mean composed of the sum of those from all $q$ releases,
%
\begin{equation}
\mu(t) = \mu_1(t) + ... + \mu_q(t).
\end{equation}
%

\begin{figure}[h]
	\centerline{\includegraphics[width=1.25\textwidth]{./Figure_files/mrr_exampleMRRSeries.pdf}}
	\caption{\textbf{Example MRR time-series for three different studies from the \cite{guerra2014global} database.} In all the above experiments there was a single release of a batch of marked mosquitoes.}
	\label{fig:mrr_exampleMRRSeries}
\end{figure}

The simplest assumption is that death and dispersal rates are independent of mosquito age, giving an exponential `survival' function,
%
\begin{equation}
S(t) = e^{-\lambda t},
\end{equation}
%
where $\lambda$ is the sum of the rates of death and dispersal from the study area. To estimate the lifespan of mosquitoes we use this model because we find no evidence in support of age-dependence across all species (see below for the method used to determine this, and later the `Results').

We assume that the number of mosquitoes recaptured on one day is independent of the number recaptured on any other day, apart from their joint dependence on $\lambda$, $\psi$ and $\kappa$. This results in a likelihood of the data of the form,
%
\begin{equation}
\mathcal{L}(y(t_1),y(t_2),...,y(t_R)|\lambda,\psi,\kappa) = \prod\limits_{i=1}^{R} p(y(t_i)|\psi,\lambda,\kappa),
\end{equation}
%
where $R$ is the number of individual days where collections took place, and $p(y(t_i)|\psi,\lambda,\kappa)$ is the probability of recapturing $y(t_i)$ mosquitoes on day $t_i$ determined from a negative binomial distribution with parameters $\mu = e^{-\lambda t} \psi$, and $\kappa$.

\subsection{Individual time-series estimates}\label{sec:MRR_individual_analysis}
We first treat each time-series separately, and estimate individual $(\lambda,\psi,\kappa)$ parameters of the statistical model (described in the previous section) for each time-series. To use a Bayesian methodology we must specify priors on this set of parameters. Here we choose to specify independent priors on each parameter of the form: $\lambda\sim \mathcal{N}(-2.32,1)$, $\psi\sim \text{exp}(50)$ (with $\psi$ constrained to be less than 0.1) and $\kappa\sim \textnormal{log-normal}(2,1)$. This prior on the rate parameter of the exponential distribution ($\lambda$) corresponds to a wide range of possible lifespans (Figure \ref{fig:mrr_individualTimeSeries_priors}A), with a mean of 10 days. It was necessary to use an informative prior on $\psi$ to allow it to be estimated; the prior on $\kappa$ is fairly uninformative and allows a wide range of values (Figure \ref{fig:mrr_individualTimeSeries_priors}B,C).

\begin{figure}[h]
	\centerline{\includegraphics[width=1.25\textwidth]{./Figure_files/mrr_individualTimeSeries_priors.pdf}}
	\caption{\textbf{The prior probability distributions used for lifespan (A), recapture probability (B) and dispersal parameter (C) in the individual time-series analysis for the MRR studies.} Note: the prior for the recapture probability was truncated at $\psi=0.1$ since most MRR day one recapture fractions were significantly below this threshold.}
	\label{fig:mrr_individualTimeSeries_priors}
\end{figure}

\subsection{Estimating lifespan at the species, genus and overall groupings}\label{sec:MRR_hierarchical}
We synthesise information from across all MRR experiments to produce more robust estimates of mosquito lifespan than can be obtained from considering the individual time-series separately. However there exists considerable heterogeneity across the experiments. This heterogeneity has two sources. There is that arising from variability in experimental methodology. However there is also variability from actual differences in lifespan across the different mosquito cohorts; for example due to genetic differences between mosquito populations or due to climatic differences. Because of this heterogeneity we use a Bayesian hierarchical model which is akin to a `random effects' model in classical statistics. This type of model assumes that there is random variation in parameters at the individual time-series level, although each of the parameters is drawn from a common `population-level' distribution. In our case we separately estimate three different model groupings where the population-level distributions correspond to the species, genus and overall (across all studies) levels respectively (Figure \ref{fig:mrr_methodDiagramHierarchy}). This allows us to estimate average mosquito lifespan for a given species or genus, or alternatively across all time-series, by independent sampling from the posterior predictive distribution from which the individual $\lambda_i$ are drawn.

\begin{figure}[h]
	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_methodDiagramHierarchy.pdf}}
	\caption{\textbf{A representation of the hierarchical Bayesian model, for the constant mortality model case when considering the species-level grouping.} Here `pdf' indicates `probability density function', and $\lambda_i$, $\psi_i$, and $\kappa_i$ represent the force of mortality for the exponential distribution, daily recapture probability and over-dispersion parameter of the negative binomial distribution for time-series $i$ in the dataset; $\mu_i(t)= \psi_i e^{-\lambda_i t}$ is the mean number of mosquitoes recaptured on day $t$.}
	\label{fig:mrr_methodDiagramHierarchy}
\end{figure}

Whilst we assume that the variation in mosquito mortality across experiments is random in nature, we recognise that there may exist systematic sources of variation that we do not include in our model. To account for this potential source of bias we examined how the individual estimates of lifespan correlated with experimental covariates (number of traps, range of traps, trapping method), but found no such evidence of systematic variation in lifespan (data not shown).

A Bayesian framework requires that we incorporate our pre-analysis beliefs into our estimates through the use of priors for all parameters in a particular model. In our hierarchical model we are required to specify prior distributions at two levels of the analysis. The first of these links the individual time-series estimates with the overarching group-level distributions. The parameters of these group-level distributions are then set prior distributions (Figure \ref{fig:mrr_methodDiagramHierarchy}). As an example, we suppose that the rate parameter $\lambda_i$ of our exponential survival model in experiment $i$ is given by:
%
\begin{align}
\lambda_i = \text{exp}(&c_i + \delta^m_{\text{genus}[i]} \text{sex}^{\text{m}}_i + \delta^{\text{mix}}_{\text{genus}[i]} \text{sex}^{\text{mix}}_i +\\  & \delta^{\text{sugar}}_{\text{genus}[i]} \text{sugar}_i +  \delta^{\text{blood}}_{\text{genus}[i]} \text{blood}_i),
\end{align}
%
where $\delta^m_{\text{genus}[i]}$ measures a genus-specific effect of male-only releases ($\text{sex}^{\text{m}}_i=1$) versus female-only ($\text{sex}^{\text{m}}_i=0$); $\delta^{\text{mix}}_{\text{genus}[i]}$ is the analogous effect for mixed releases versus female-only; $\delta^{\text{sugar}}_{\text{genus}[i]}$ is the analogous effect for mosquitoes that were sugar-fed prior to release versus those that were not fed; and $\delta^{\text{blood}}_{\text{genus}[i]}$ is the analogous effect for mosquitoes that were blood-fed prior to release versus those that were not fed. The parameter $c_i$ is modelled hierarchically, being drawn from a group-level distribution assumed to be log-normal,
%
\begin{equation}
c_i \sim  \textnormal{normal}(\mu_\lambda,\sigma)
\end{equation}
%
We then specify priors on these group-level parameters, $\mu_\lambda \sim N(-2.32,1)$ and $\sigma \sim \textnormal{log-normal}(-3,1)$. On all $\delta$ parameters, except $\delta^{\text{mix}}_{\text{genus}[i]}$, we specify standard normal priors. We specify a uniform prior on $\delta^{\text{mix}}_{\text{genus}[i]}\sim U(0,\delta^m_{\text{genus}[i]})$ so that the effect of mixed releases lies between the female-only release effect (0) and the male-only release effect.

The relatively complex nature of priors in hierarchical models make it important to determine their influence on inferences. We chose the above -- somewhat uninformative -- priors to allow a range of mosquito lifespans in order to minimise their effect on the estimates we report (Figure \ref{fig:mrr_meanLifespanPrior}). We also chose to set hierarchical priors on the remaining parameters in our models -- $\psi$ the probability of daily recapture, and $\kappa$ the `over-dispersion' parameter of a negative binomial distribution. For $\psi$ we chose an informative prior that placed all probability mass $\psi\leq 10\%$, since in most experiments the fraction of mosquitoes captured was significantly below this value. For $\kappa$ we chose a fairly wide prior that had most of its probability mass below $\kappa=20$ (Figure \ref{fig:mrr_PsiKappaPriors}). In all cases the hierarchical priors were chosen to have similar implications on lifespan, recapture probability and over-dispersion at the individual time-series level as for the non-hierarchical analysis described previously.

\begin{figure}[h]
	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_lifespan_priors.pdf}}
	\caption{\textbf{Samples from the prior predictive distribution of mean lifespan across the six models of mosquito mortality introduced in Table \ref{tab:mrr_survivalDescription}.} In all cases 10,000 samples were generated using the priors indicated in Table \ref{tab:mrr_priors}, resulting in distributions with a mean close to 10 days.}
	\label{fig:mrr_meanLifespanPrior}
\end{figure}

\begin{figure}[h]
	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_priors_PsiKappa.pdf}}
	\caption{\textbf{Samples from the priors for daily recapture probability ($\psi$) and over-dispersion parameter ($\kappa$).} In both panels, 10,000 samples were generated using the following hierarchical priors: for the individual time series parameters, $\psi_i \sim \text{exp}(r)$ (with $\psi$ constrained to be less than 0.1 - encompassing the upper limit of day one recapture fractions observed in most data series) and $\kappa_i \sim \text{exp}(p)$; for the top-level parameters: $r\sim \text{gamma}(100, 2)$ and $p\sim \text{log-normal}(1.5, 1)$.}\label{fig:mrr_PsiKappaPriors}
\end{figure}

\subsection{Testing for age-dependent mortality in wild mosquitoes}\label{sec:mrr_age_dependence}
Previous work has found evidence of age-dependent mortality in lab populations \citep{styer2007mosquitoes,dawes2009anopheles}, and less-commonly in wild mosquitoes \citep{clements1981analysis}. Furthermore, recent modelling work has examined the implications of departures from a constant risk of mortality \citep{styer2007mosquitoes,hancock2009age,novoseltsev2012age}. To determine whether senescence occurs in wild mosquitoes we re-estimate our model using survival functions, $S(t)$, that allow for a rate of mortality that can vary with age. Specifically, we re-estimate our model using five other models that were previously used in the literature (see Table \ref{tab:mrr_survivalDescription} for a description of these). Each of these models makes different assumptions about how the rate of death and dispersal varies with age, but all can be represented by a general form,
%
\begin{equation}\label{eq:survivalInt}
S(t) = e^{-\int\limits_{0}^{t}\lambda(\tau) \mathrm{d}\tau}.
\end{equation}
%
where we constrain the parameters of our models to preclude the possibility of a hazard that decreases with age ($\frac{\mathrm{d}\lambda(t)}{\mathrm{d}t} \geq 0$). Whilst this could occur if older mosquitoes disperse less than younger mosquitoes, it is unlikely this would outweigh any declines in survival associated with old age.

The parameters of the survival function of each model were assigned hierarchical priors that allowed considerable variation in mosquito lifespan (see Figure \ref{fig:mrr_meanLifespanPrior}). Otherwise the statistical model we used remained the same as for the constant mortality case. In testing for age-dependent mortality, we did not account for sex or pre-release feeding effects due to the complexity of including these elements in more complex survival models.

\begin{table}
	\scriptsize
	\noindent\makebox[\textwidth]{%
		\begin{tabularx}{1.35\textwidth}{c|cXcX}
			\textbf{Survival function} & \textbf{Hazard rate} & \textbf{Interpretation} & \textbf{Mean time spent in study area, }\textbf{$\overline{T}$} & \textbf{Papers assuming this function} \\
			\midrule
			Exponential & $\lambda$ & Constant mortality risk & $\frac{1}{\lambda}$ & \cite{ross1910prevention}; \cite{anderson1992infectious}; \cite{smith2004statics} \\
			\midrule
			Gompertz &  $\alpha e^{\beta t}$     & Mortality increases with age at an ever-increasing rate &  $\frac{e^{\frac{\alpha}{\beta}} \Gamma(0,\frac{\alpha}{\beta})}{\beta}$    & \cite{clements1981analysis}; \cite{styer2007mosquitoes};\cite{novoseltsev2012age} \\
			\midrule
			Weibull &  $\alpha t^\beta$; $\;\beta\geq 1$     &  Mortality increases with age at an ever-increasing rate     &   $\alpha^{-\frac{1}{\beta}} \Gamma(1+\frac{1}{\beta})$    & \cite{hancock2009age}; and \cite{carey2001insect} considering general insect demography \\
			\midrule
			Gompertz-Makeham &  $\alpha e^{\beta t} + c$      &  Two additive mortality risks: one that increases with age, and another that is age-independent  &  No simple analytic form     & \cite{styer2007mosquitoes} \\
			\midrule
			Logistic & $\frac{\alpha e^{\beta t}}{1+\frac{\alpha s}{b}\left(e^{b t} - 1\right)}$   & Mortality risk increases with age, although at a declining rate  & No simple analytic form   & \cite{styer2007mosquitoes}; \cite{novoseltsev2012age} \\
			\midrule
			Logistic-Makeham &   $\frac{\alpha e^{\beta t}}{1+\frac{\alpha s}{b}\left(e^{b t} - 1\right)} + c$    & Two separate additive mortality risks: an age-dependent risk of the same form as the logistic model, and a constant hazard  &  No simple analytic form  & \cite{styer2007mosquitoes}; \cite{bellan2010importance} \\
			\bottomrule
	\end{tabularx}}
	\caption{\textbf{A description of the survival functions used in this study, arranged in rough order of model complexity (simple-complex from top-bottom.)} All parameters are defined to be non-negative. $\Gamma(\theta)$ and $\Gamma(\theta_1,\theta_2)$ refer to the Euler gamma function, and incomplete gamma function respectively. The `mean time spent in study area' is an estimate of the combined effects of mosquito mortality and dispersal from the area of the study where collections take place, since our data does not provide spatially-resolved data.}\label{tab:mrr_survivalDescription}
\end{table}

\begin{table}
	\footnotesize
	\noindent\makebox[\textwidth]{%
		\begin{tabularx}{1.2\textwidth}{l|c|X}
			\textbf{Survival function} & \textbf{time-series-level priors} & \textbf{Group-level priors}\\
			\midrule
			Exponential & $\lambda\sim \textnormal{log-normal}(\mu_\lambda,\sigma)$ &$\mu_\lambda \sim N(-2.32,1)\text{,}\; \sigma\sim \textnormal{log-normal}(-3,1)$\\
			\midrule
			Gompertz & $\alpha,\beta\sim$ log-normal$(\mu_{\alpha|\beta},0.2)$ &$\mu_\alpha\sim N(-3,1)\text{,}\; \mu_\beta\sim N(-3,1.1)$\\
			\midrule
			Weibull & $\alpha,(\beta-1)\sim$  log-normal$(\mu_{\alpha|\beta},0.2)$ &$\mu_\alpha\sim N(-4.8,1.75)\text{,}\; \mu_\beta\sim N(-4,2)$\\
			\midrule
			Gompertz-Makeham & $\alpha,\beta,c\sim$ log-normal$(\mu_{\alpha|\beta|c},0.2)$ &$\mu_\alpha\sim N(-3,1)\text{,}\; \mu_\beta\sim N(-3,0.5)\text{,}\; \mu_c\sim N(-4.5,0.5)$\\
			\midrule
			Logistic & $\alpha,\beta,s\sim$  log-normal$(\mu_{\alpha|\beta|s},0.2)$ &$\mu_\alpha\sim N(-2.8,1)\text{,}\; \mu_\beta\sim N(-3,1)\text{,}\; \mu_s\sim N(-3,1)$\\
			\midrule
			Logistic-Makeham & $\alpha,\beta,s,c\sim$ log-normal$(\mu_{\alpha|\beta|s|c},0.2)$ &$\mu_\alpha\sim N(-3.2,1)\text{,}\; \mu_\beta\sim N(-3.2,1)\text{,}\; \mu_s\sim N(-3,0.5)\text{,}\; \mu_c\sim N(-4,0.5)$\\
			\bottomrule
	\end{tabularx}}\caption{\textbf{The priors used on parameters of each different survival model.} For the exponential model the `group-level' priors were the same for the genus and `overall' models that were also estimated. The exponential model was the only model that was simple enough to allow the scale parameter of the log-normal ($\sigma$) to be estimated by the data. The notation $\alpha,\beta\sim \text{log-normal}(\mu_{\alpha|\beta},0.2)$ means that $\alpha$ and $\beta$ were assigned independent log-normal priors with location parameters $\mu_\alpha$ and $\mu_\beta$ respectively, and a scale parameter of 0.2 in both cases.}\label{tab:mrr_priors}
\end{table}

\subsection{Model estimation by MCMC}\label{sec:mrr_MCMC}
The likelihood and priors we use result in posterior distributions whose analytic form cannot be calculated with existent computational methods. Instead we use Markov Chain Monte Carlo (MCMC) methods to sample from each posterior distribution. In particular we used \textit{Stan} software \citep{carpenter2016stan} that implements an efficient MCMC algorithm known as NUTS \citep{hoffman2014no}.

To judge convergence of the sampling algorithm to the posterior density, we calculated $\hat{R}$ across all Markov chains -- a ratio that compares the between-chain variation to that within each chain that is commonly used to measure convergence in MCMC \citep{gelman1992inference}. For each class of model, we used the following MCMC parameters,
%
\begin{itemize}
	\item Individual lifespan models (Section \ref{sec:MRR_individual_analysis}): 12 independent chains with 1000 iterations per chain,
	\item Hierarchical lifespan estimates (Section \ref{sec:MRR_hierarchical}): 12 independent chains with 15,000 iterations per chain,
	\item Age-dependent analysis (Section \ref{sec:mrr_age_dependence}): 12 independent chains with 3000 iterations per chain.
\end{itemize}

In all cases, we discarded the first half of these iterations as warm-up \citep{gelman2014bayesian}. At the end of all runs for lifespan estimation, $\hat{R}<1.1$ for all model parameters. For the age-dependent cases, there remained a handful of parameters where $\hat{R}>1.1$ after running simulations for 3000 iterations but this did not affect our inferences, since repeated runs demonstrated the same pattern. We also ensured that across each MCMC run the number of divergent iterations (that can bias the MCMC away from the true posterior density) was minimal. In the majority of cases, the number of divergent iterations was far fewer than 1\% of the total number of samples.

The Stan scripts for the lifespan estimation (i.e. those using an exponential survival model) for the MRR analyses are provided below. The script for the individual time series analysis (Section \ref{sec:MRR_individual_analysis}):

\begin{minted}{stan}
data {
// Single release parameters
int<lower=0> N; // number of observations
int<lower=0> K; // number of groups
int Y[N]; // observations for the releases
int S[K]; // time series lengths
vector[N] t; // time observations (days) 
int Pos[K]; // starting position of each dataset

// Multiple release parameters
int RelFreq[K]; // frequency of releases in each study
// starting position of releases in each study
// in RelNumber
int PosRel[K];
int<lower=0> NReleases; // total number of releases
// numbers released in each study
vector[NReleases] RelNumber;
// time of each release in each study
vector[NReleases] RelTime;

real Age[K];
}

parameters {
real<lower=0.3, upper=20> kappa[K];
real<lower=0> cDecay[K];
real<lower=0, upper=1> psi[K];
}

model {
// Likelihood
for (i in 1:K) {
real lambda[S[i]];

if (RelFreq[i]< 2) // Single release
{
   real numberCaught;
   numberCaught = 0;	
   for (j in 1:S[i])
   {
     lambda[j]= (psi[i] *
                (RelNumber[PosRel[i]] - numberCaught) *
                exp(-cDecay[i] * (t[Pos[i] + j - 1] + Age[i])));
     //1e-5 to prevent loc=0
     Y[Pos[i] + j - 1] ~ neg_binomial_2(
        		lambda[j] + 1e-5, kappa[i]); 
     numberCaught = numberCaught + Y[Pos[i] + j - 1];
   }
} else // Multiple release
  {
    for (j in 1:S[i])
    {
      real lambdaTemp;			
      lambdaTemp = 0;
      for (kk in 1:RelFreq[i])
      {
        if (t[Pos[i] + j - 1]> RelTime[PosRel[i]+kk-1])
        {
           lambdaTemp = (
                   lambdaTemp +
                   RelNumber[PosRel[i] + kk - 1] * 
                   ((1 - psi[i])^(t[Pos[i] + j - 1] -
                   RelTime[PosRel[i]+kk-1]-1)) *
                   psi[i] * exp(-cDecay[i] * 
                   (t[Pos[i] + j - 1] -
                   RelTime[PosRel[i]+kk-1] + Age[i]))
           );
          }
   }
   lambda[j] = lambdaTemp;
   Y[Pos[i] + j - 1] ~ neg_binomial_2(lambda[j]+1e-5, kappa[i]);
   }
 }
}

// Priors
for (i in 1:K)
{
   cDecay[i] ~ lognormal(-2.32, 1);
   kappa[i] ~ lognormal(2, 1);  
   psi[i] ~ exponential(50);
}
}

generated quantities {
vector[K] lifespans;
  for(i in 1:K)
    lifespans[i] = 1 / cDecay[i];
}
\end{minted}

The script for the hierarchical lifespan estimation (Section \ref{sec:MRR_hierarchical}) is shown below:

\begin{minted}{stan}
data {
// Single release parameters
int<lower=0> N; // number of observations
int<lower=0> K; // number of groups
int Y[N]; // observations
int S[K]; // group sizes
vector[N] t; // time observations (days) 
int Pos[K]; // starting position of each dataset

// Multiple release parameters
int RelFreq[K]; // release frequency in each study
int PosRel[K]; // starting position of releases in each study
int<lower=0> NReleases; // release frequencies
vector[NReleases] RelNumber; // numbers released
vector[NReleases] RelTime; // release times

int<lower=0> nSpecies; // number of individual species
int SpeciesIndex[K]; // species index of each study

real Age[K];
int SexM[K];
int SexMix[K];
int Genus[K];
int Blood[K];
int Sugar[K];
}

transformed data{
int SpeciesToGenus[nSpecies];
for(i in 1:K){
  SpeciesToGenus[SpeciesIndex[i]] = Genus[i];
}
}

parameters {
real<lower=0.3,upper=100> kappa[K];
real a[nSpecies];
real<lower=0> p[nSpecies];
real<lower=0> r[nSpecies];
real cDecay[K];
real<lower=0,upper=0.1> psi[K];
real<lower=0> sigma_a[nSpecies];
real delta_M[3];
real delta_Blood[3];
real delta_Sugar[3];
real<lower=0, upper=1> u_mixed[3];
}

transformed parameters{
real delta_Mixed[3];
for(i in 1:3)
  delta_Mixed[i] = u_mixed[i] * delta_M[i];
}

model {
for (i in 1:K) {
  real lambda[S[i]];
  real decay_temp = exp(cDecay[i]+delta_M[Genus[i]] * SexM[i] +
    delta_Mixed[Genus[i]] * SexMix[i] + delta_Sugar[Genus[i]] *
    Sugar[i] + delta_Blood[Genus[i]] * Blood[i]);

  if (RelFreq[i]< 2) // Single release
  {
    real R_times_psi;
    real numberCaught;
    numberCaught = 0;	
    R_times_psi = psi[i];
    for (j in 1:S[i])
    {
    lambda[j]= R_times_psi *
            (RelNumber[PosRel[i]]-numberCaught) *
            exp(-decay_temp * (t[Pos[i] + j - 1] + Age[i]));
    Y[Pos[i] + j - 1] ~ neg_binomial_2(lambda[j] + 1e-5,
                                       kappa[i]);
    numberCaught = numberCaught + Y[Pos[i] + j - 1];
    }
  }
  else // Multiple release
  {
    for (j in 1:S[i])
    {
    real lambdaTemp;			
    lambdaTemp = 0;
    for (kk in 1:RelFreq[i])
    {
      if (t[Pos[i] + j - 1]> RelTime[PosRel[i] + kk - 1])
      {
        lambdaTemp = (lambdaTemp +
          RelNumber[PosRel[i] + kk - 1] * 
          ((1-psi[i])^(t[Pos[i] + j - 1] -
          RelTime[PosRel[i] + kk - 1] - 1)) *
          psi[i] * exp(-decay_temp *
          (t[Pos[i] + j - 1] -
          RelTime[PosRel[i] + kk - 1] + Age[i]));
         );
        }
      }
      lambda[j] = lambdaTemp;
      Y[Pos[i] + j - 1] ~ neg_binomial_2(lambda[j] + 1e-5,
                                         kappa[i]);
      }
    }
}

// Priors
for (i in 1:K)
{
  int speciesTemp;
  speciesTemp = SpeciesIndex[i];
  cDecay[i] ~ normal(a[speciesTemp], sigma_a[speciesTemp]);
  kappa[i] ~ exponential(p[speciesTemp]);  
  psi[i] ~ exponential(r[speciesTemp]);
}
a ~ normal(-2.32, 1);
sigma_a ~ lognormal(-3, 1);
p ~ lognormal(1.5, 1);
r ~ gamma(100, 2);
delta_M ~ normal(0, 1);
delta_Blood ~ normal(0, 1);
delta_Sugar ~ normal(0, 1);
}

generated quantities{
real overallcDecay[nSpecies];
real overallFLife[nSpecies];
real overallMLife[nSpecies];
real overallMixedLife[nSpecies];
real overallFBloodLife[nSpecies];
real overallFBothLife[nSpecies];
real overallFSugarLife[nSpecies];
real overallMSugarLife[nSpecies];

for(i in 1:nSpecies){
  overallcDecay[i] = normal_rng(a[i], sigma_a[i]);
  overallFLife[i] = 1 / exp(overallcDecay[i]);
  overallMLife[i] = 1 / exp(overallcDecay[i] +
                            delta_M[SpeciesToGenus[i]]);
  overallMixedLife[i] = 1 / exp(overallcDecay[i] +
                            delta_Mixed[SpeciesToGenus[i]]);
  overallFBloodLife[i] = 1 / exp(overallcDecay[i] +
                             delta_Blood[SpeciesToGenus[i]]);
  overallFSugarLife[i] = 1 / exp(overallcDecay[i] +
                             delta_Sugar[SpeciesToGenus[i]]);
  overallFBothLife[i] = 1 / exp(overallcDecay[i] +
                             delta_Sugar[SpeciesToGenus[i]] +
                             delta_Blood[SpeciesToGenus[i]]);
  overallMSugarLife[i] = 1 / exp(overallcDecay[i] +
                             delta_M[SpeciesToGenus[i]] +
                             delta_Sugar[SpeciesToGenus[i]]);
}
}
\end{minted}

\subsection{K-Fold cross validation}\label{sec:mrr_kFold}
One way to estimate a model's out-of-sample predictive capability is to use Akaike Information Criterion, Bayesian Information Criterion, Deviance Information Criterion or Watanabe Akaike Information Criterion that explicitly penalise a model in accordance to its complexity (typically indexed by its number of free parameters) to correct for model over-fit. The way in which this correction takes places however, is fairly heuristic, and less appropriate for hierarchical models. An alternative method, common in the machine learning literature, is known as `cross-validation' (see for example, \cite{kohavi1995study}). Here to estimate out of sample predictive capability, the original data set is partitioned into training and test sets. The model is then fitted to the training set, and its predictive performance measured on the independent test set.

We use cross-validation to compare the predictive power of the species, genus and overall models, and later to determine whether age-dependent mortality occurs. In particular, we use K-Fold cross validation where we repeatedly partition our dataset into a test set composed of a number of individual time-series, and a training set of the remainder (see for example, \cite{marsland2015machine}). We then use the fitted model to measure the predictive performance on the test set. The particular measure we use is the expected log point-wise predictive density (elpd)\citep{vehtari2015efficient} which sums the predictive performance for each data point averaged across all posterior samples.

Since the models are hierarchical we must specify how to draw parameters for a given test sample time-series. Here we chose to draw values of the individual time-series parameters as samples from the population distribution that corresponds to that particular sample's respective grouping. So if the particular test sample time-series pertains to \textit{Ae. aegypti} (and we are using the species-level model) then we draw independent samples for the statistical model's parameters from the estimated \textit{Ae. aegypti} population distribution,
%
\begin{align}
\theta_i \sim p(\theta^{\text{Ae. aegypti}}|data),
\end{align}
%
where $\theta_i$ is the value of the parameters used in test time-series $i$, and $p(\theta^{\text{Ae. aegypti}}|data)$ is the overall posterior distribution across all \textit{Ae. aegypti} experiments.

\subsection{Model checking}
To check the fit of the model to the data, we simulated from the posterior predictive distribution for each data point within each time series of recapture observations (see Figures \ref{fig:mrr_ppc1}, \ref{fig:mrr_ppc2} \&  \ref{fig:mrr_ppc3} and the attached file, \verb|mrr_ppcs_all.pdf| for the full graphs). In the majority of cases, the data lay within the 95\% predictive intervals indicating that the model was a good fit to the data.

{%
	\includegraphics[page=1,scale=.6]{./Figure_files/mrr_ppcs_all}
	\captionof{figure}{The numbers of mosquitoes recaptured (black lines) versus the 95\% central posterior interval of the posterior predictive distribution (blue shading) for a selection of the time series. For the rest of the posterior predictive checks for the MRR models, see the file referenced in the text.}
	\label{fig:mrr_ppc1}
	\par

{%
	\includegraphics[page=2,scale=.6]{./Figure_files/mrr_ppcs_all}
	\captionof{figure}{The numbers of mosquitoes recaptured (black lines) versus the 95\% central posterior interval of the posterior predictive distribution (blue shading) for a selection of the time series. For the rest of the posterior predictive checks for the MRR models, see the file referenced in the text.}
	\label{fig:mrr_ppc2}
	\par
}

{%
	\includegraphics[page=3,scale=.6]{./Figure_files/mrr_ppcs_all}
	\captionof{figure}{The numbers of mosquitoes recaptured (black lines) versus the 95\% central posterior interval of the posterior predictive distribution (blue shading) for a selection of the time series. For the rest of the posterior predictive checks for the MRR models, see the file referenced in the text.}
	\label{fig:mrr_ppc3}
	\par
}

\section{Dissection based estimates}

\subsection{Collection of dissection data}\label{sec:dissection_dissectionData}
A comprehensive search of the literature using Google Scholar (\url{scholar.google.co.uk}) was performed using various combinations of the following keywords: dissection, mosquito, parity, parous, age, and physiological age. No constraints were placed on publication date, location or type. This list was supplemented with a number of author-specific searches for those individuals most prevalent in the literature. In particular we searched for all articles authored by: Charlwood, Muller, Schlein, Samarawickrema, Reisen, Detinova, Polodova, Gillies, and Wilkes. An additional list of potential articles was provided by doing a forward article search on some of the most highly-cited articles in the database: \cite{polovodova1949determination,detinova1962age,gillies1965study,clements1981analysis}. The list of results was then filtered manually by examining the titles and abstracts to produce a candidate list of the articles most likely to contain data on the physiological age of dissected mosquitoes caught in the wild, as determined by the \cite{polovodova1949determination} method.

A relational database was used to store the raw data from the actual experiments, along with the meta-data associated with each of the experiments. In many of the published articles wild mosquitoes were caught and dissected over a period of time, and the raw data thus consists of snapshots of the age structure of the population at regular intervals in time. In the cases where the data was more sparse (fewer than ten individuals, on average, per date), we aggregated across dates and recorded this as a single entry; otherwise we recorded the snapshots of the population at each different date. We record separate series for each species that was captured, or for those that were recorded at separate capture locations (potentially with an alternative collection method), and do not aggregate over these datasets.

For each individual series we recorded the following meta-data: genus, species, collection method, whether or not insecticide was mentioned as being used in or around the time of mosquito collections, and the start and end dates of the experiment. At the article level we recorded the following meta-data: author, year, title, country, location (within country), start and end date, whether insecticide was used during any of the experimental replicates, whether a mrr experiment occurred alongside the dissection study, and the collection location (indoor or outdoor). At either the individual series or article levels we record additional meta-data describing the nature of data collection, for example explaining where the data was contained within the article, whether it was obtained by digitising graphs, and the number of separate dated series. For those few cases ($n=16$ series) where the data was obtained by digitising graphs, we used the WebPlotDigitizer online tool \citep{digitise}.

The data collection method resulted in 568 separate dissection series, across 72 published articles (see Section \ref{sec:appendix_dissectionStudyList} for a list of the studies included in our final dataset). The published datasets cover the period from 1960-2015 with comparable numbers of studies across each decade (Figure \ref{fig:dissection_timeSeries}). The statistical approach applied to the data relies on the assumption that there is a constant rate of recruitment into the adult population (see Section \ref{sec:dissection_dissectionStats}). If populations fluctuate strongly from month-to-month this assumption will likely be violated. To try to mitigate against such a risk, we aggregate the data across all dates to obtain a single series for each identifier, resulting in 201 such series. To obtain a reasonable level of accuracy on estimates we remove all those individual (aggregated) series where there are fewer than 100 mosquitoes in total in the series. Finally we remove data for any species where there was only one series resulting in 131 series across four genera (\textit{Anopheles}, \textit{Culex}, \textit{Aedes}, and \textit{Mansonia}) and 25 species (Table \ref{tab:dissection_speciesNumbers}). These studies are distributed across a wide range of geographies (Figure \ref{fig:dissection_map}). The raw data for the dissection analysis are shown in Figures \ref{fig:dissection_exampleData1}, \ref{fig:dissection_exampleData2} and \ref{fig:dissection_exampleData3}.

\begin{figure}[ht]
	\centerline{\includegraphics[width=0.8\textwidth]{./Figure_files/dissection_timeSeries.pdf}}
	\caption{\textbf{The numbers of published studies that estimate gonotrophic age by dissection across the study period.}}\label{fig:dissection_timeSeries}
\end{figure}

\begin{figure}[ht]
	\centerline{\includegraphics[width=1.25\textwidth]{./Figure_files/dissection_map.pdf}}
	\caption{\textbf{The geographic location of each of the dissection databases included in the meta-analysis.} The area of the bubbles indicates the number of unique data series available.}\label{fig:dissection_map}
\end{figure}

\begin{table}[htbp]
	\centering
	\begin{tabular}{lll}
		\toprule
		\textbf{Genus} & \textbf{Species} & \textbf{Frequency} \\
		\midrule
		\textit{Anopheles} & \textit{gambiae s.l.} & 19 \\
		\textit{Culex} & \textit{quinquefasciatus} & 12 \\
		\textit{Anopheles} & \textit{maculipennis} & 11 \\
		\textit{Anopheles} & \textit{farauti s.l.} & 10 \\
		\textit{Anopheles} & \textit{sergentii} & 8 \\
		\textit{Aedes} & \textit{polynesiensis} & 8 \\
		\textit{Culex} & \textit{pipiens} & 6 \\
		\textit{Anopheles} & \textit{culicifacies} & 5 \\
		\textit{Anopheles} & \textit{darlingi} & 5 \\
		\textit{Anopheles} & \textit{quadrimaculatus} & 5 \\
		\textit{Anopheles} & \textit{stephensi} & 4 \\
		\textit{Anopheles} & \textit{melas} & 4 \\
		\textit{Culex} & \textit{annulirostris} & 4 \\
		\textit{Aedes} & \textit{aegypti} & 3 \\
		\textit{Aedes} & \textit{samoanus} & 3 \\
		\textit{Anopheles} & \textit{minimus} & 3 \\
		\textit{Anopheles} & \textit{rivulorum} & 3 \\
		\textit{Culex} & \textit{thalassius} & 3 \\
		\textit{Mansonia} & \textit{uniformis} & 3 \\
		\textit{Anopheles} & \textit{subpictus} & 2 \\
		\textit{Aedes} & \textit{sollicitans} & 2 \\
		\textit{Aedes} & \textit{vexans} & 2 \\
		\textit{Anopheles} & \textit{bellator} & 2 \\
		\textit{Anopheles} & \textit{cruzii} & 2 \\
		\textit{Culex} & \textit{tritaeniorhynchus} & 2 \\
		\bottomrule
		\textit{\textbf{Anopheles}} &       & \textbf{83} \\
		\textit{\textbf{Culex}} &       & \textbf{27} \\
		\textit{\textbf{Aedes}} &       & \textbf{18} \\
		\textit{\textbf{Mansonia}} &       & \textbf{3} \\
		\bottomrule
		\textbf{Total} &  & \textbf{131} \\
	\end{tabular}%
	\caption{\textbf{The numbers of dissection series for each species or genus included in the overall dataset.}}\label{tab:dissection_speciesNumbers}%
\end{table}%

\begin{table}[htbp]
	\centering
	\footnotesize
	\begin{adjustwidth}{-0.5in}{-0.5in}%
		\begin{tabularx}{1.25\textwidth}{l|ccccc}
			\toprule
			\textbf{Variable} & \textbf{Min} & \textbf{Median} & \textbf{Mean} & \textbf{Max} & \textbf{Standard deviation} \\
			\midrule
			Min age of captures, gonotrophic cycles & 0  & 0  & 0.02  & 1.00  & 0.12 \\
			Median age of captures, gonotrophic cycles & 0  & 1.00  & 0.63  & 2.00  & 0.62 \\
			Mean age of captures, gonotrophic cycles & 0.17  & 0.90  & 1.00  & 3.03  & 0.58 \\
			Max age of captures, gonotrophic cycles & 1.00  & 5.00  & 5.77  & 13.00 & 2.94 \\
			Number captured & 100 & 565 & 1,317 & 14,012 & 1,885 \\
			\bottomrule
		\end{tabularx}%
	\end{adjustwidth}
	\caption{\textbf{Summaries of the characteristics of the 131 individual dissection data series that are included in this analysis.} Note when series are censored (see Section \ref{sec:dissection_censored}) we assume that all mosquitoes dissected over the threshold are of this age when calculating summary statistics.}\label{tab:dissection_summaryStats}
\end{table}

\begin{figure}[ht]
	\centerline{\includegraphics[width=1.3\textwidth]{./Figure_files/dissection_parity_data1.png}}
	\caption{\textbf{The normalised physiological age series for nine species in the database.} Each different coloured line represents an individual series. In each case the count for all ages has been normalised by the nulliparous count. In all cases we do not include any data for censored observations (see Section \ref{sec:dissection_censored}).}\label{fig:dissection_exampleData1}
\end{figure}

\begin{figure}[ht]
	\centerline{\includegraphics[width=1.3\textwidth]{./Figure_files/dissection_parity_data2.png}}
	\caption{\textbf{The normalised physiological age series for nine species in the database.} Each different coloured line represents an individual series. In each case the count for all ages has been normalised by the nulliparous count. In all cases we do not include any data for censored observations (see Section \ref{sec:dissection_censored}).}\label{fig:dissection_exampleData2}
\end{figure}

\begin{figure}[ht]
	\centerline{\includegraphics[width=1.3\textwidth]{./Figure_files/dissection_parity_data3.png}}
	\caption{\textbf{The normalised physiological age series for nine species in the database.} Each different coloured line represents an individual series. In each case the count for all ages has been normalised by the nulliparous count. In all cases we do not include any data for censored observations (see Section \ref{sec:dissection_censored}).}\label{fig:dissection_exampleData3}
\end{figure}

\subsection{Statistical analysis of dissection data}\label{sec:dissection_dissectionStats}
We suppose that the number of individuals recruited into the adult female population is constant over time, meaning that the age-structure of the population is stable. Whilst environmental heterogeneity will naturally lead to variation in adult recruitment over time, we hope that by aggregating data across studies undertaken over a range of dates we may reduce the impact of this effect on our results. Suppose that the probability an individual female mosquito survives until age $a$ is given by the survival function $S(a)$, then the number of individuals in the population surviving to this age is given by,
%
\begin{align}\label{eq:dissection_structured}
A(a) = A(0) S(a),
\end{align}
%
where $A(0)$ is the number of female mosquitoes recruited into the population per unit time. Consider one individual experiment where we randomly sample from a wild population that is structured as per eq. (\ref{eq:dissection_structured}). In this case the number of individuals sampled at each age is binomially-distributed,
%
\begin{align}
X(a) \sim \mathcal{B}(A(a),p(a)),
\end{align} 
%
where $p(a)$ is the probability of recapturing a single mosquito of age $a$. In what follows we typically assume that the probability of capturing a given mosquito is independent of their age, so that $p(a)=p=const$. Since in general $A(a)$ is large and $p(a)$ is small, we can approximate the above using a Poisson distribution,
%
\begin{align}
X(a) \sim \text{Poisson}(A(a)p).
\end{align}
%
However we believe that the assumption of \textit{independent} captures of individual mosquitoes, which underlies the binomial and Poisson models, is likely suspect for the same reasons as for the MRR analysis (see Section \ref{sec:mrr_statistical}). As before we choose to specify a negative binomial sampling distribution that allows for non-independent captures,
%
\begin{align}
X(a) \sim \text{NB}(A(a)p,\kappa),
\end{align} 
%
where we use the parameterisation such that the mean is given by $A(a)p$, and the over-dispersion parameter, $\kappa$, where as $\kappa\rightarrow \infty$ the above sampling distribution approaches a Poisson.

In the field, unfortunately, we do not know the number of mosquitoes recruited into the adult population, $A(a)$, nor the probability of capturing an individual at a given point in time, $p$. Instead we model their product $\Psi=A(0)p$ (the population of adult female mosquitoes of age zero that can be captured) probabilistically resulting in a model,
%
\begin{align}\label{eq:dissertation_negativeBinomial}
X(a) \sim \text{NB}(\Psi S(a),\kappa).
\end{align}
%
The resultant likelihood of a data series consisting of counts: $(y(a_1),y(a_2),...,y(a_R))$, is then calculated by assuming (conditional) independence of the observations,
%
\begin{equation}
\mathcal{L}(y(a_1),y(a_2),...,y(a_R)|S(.),\Psi,\kappa) = \prod\limits_{a=a_1}^{a_R} p(y(a)|S(a),\Psi,\kappa),
\end{equation}
%
where $p(y(a)|S(a),\Psi,\kappa)$ corresponds to the negative binomial probability mass function for a count of $y(a)$ mosquitoes aged $a$ as specified in eqn. (\ref{eq:dissertation_negativeBinomial}), and $R$ is the number of separate physiological age classes in which the count was non-zero.
%
Since we do not know $\Psi$ we must learn it from the data. One approach to estimate this parameter could be use the number of captures of nulliparous mosquitoes, $X(0)$ in place of $\Psi$. We prefer to allow for some uncertainty in this parameter and use the data to estimate its value. However, unfortunately, the negative binomial likelihood allows too much variation in this parameter (because the data are over-dispersed), and instead we specify a likelihood of the form,
%
\begin{align}
X(0) \sim \mathcal{N}(\Psi,\sqrt{\Psi}),
\end{align}
%
solely for the first data point $X(0)$. The above allows for some freedom in $\Psi$ whilst ensuring that the parameter's probability mass lies near enough to $X(0)$ to allow useful model estimates. $\Psi$ is set a uniform prior over the range $[0,\frac{3}{2}X(0)]$.

Since $S(a)$ is monotonically-decreasing we know that, if recruitment to the adult population is constant, then the numbers of nulliparous individuals should exceed the count in subsequent parity states. However in a number of data series there is a relative dearth of nulliparous mosquitoes versus uniparous individuals. This deficiency has been noted in a previously-published study where the authors hypothesise that it is due to the issue of sampling the nulliparous population, since they are more likely to rest outside, compared with parous individuals \citep{gillies1965study}. However there is evidence to suggest that, on average, the first gonotrophic cycle is longer than subsequent cycles (see Section \ref{sec:dissection_gonotrophicData}), meaning that a relative surplus of nulliparous mosquitoes may exist in captured samples \citep{clements1981analysis}.

In our analysis, we chose to remove the counts of nulliparous individuals from the series where the count was low relative to uniparous or higher parity individuals since this may indicate issues with sampling the nulliparous population. Specifically we stipulated that the nulliparous count should exceed 90\% of the count for the uniparous individuals. For those cases where this condition was not met we removed the nulliparous observation and analyse the series of counts for all subsequent pars (uniparous and subsequent parous states).

Here we estimate our model with one of six different survival functions (the same as for the MRR case; Table \ref{tab:mrr_survivalDescription}), each of which makes different assumptions regarding how the force of mortality is affected by mosquito age. To estimate lifespan at the species, genus and overall groupings we then use a hierarchical Bayesian model of the same mathematical form as for the analysis of MRR experiments (see Section \ref{sec:MRR_hierarchical}). However the priors for the analysis of dissection data were modified so that they represented a mean prior lifespan of three gonotrophic cycles, although allowed considerable variation in mean lifespan (Figure \ref{fig:dissection_lifespanPriorsParity}). The priors used for the over-dispersion parameter ($\kappa$) for the negative binomial likelihood were the same as for the MRR analysis.

To determine whether mosquitoes experience age-dependent mortality we compared the predictive power of each of the models that incorporate a hazard function that depends on age with that from the exponential model. As for the MRR analysis we also used K-Fold cross-validation to perform this comparison, where the data are randomly partitioned into training and test sets (see Section \ref{sec:mrr_kFold}). The model is then fitted to each training set and used to predict the data in the independent test set. Since there are fewer series than for the MRR dataset we used two partitions for each species, where each partition was of roughly the same number of individual series.

\begin{figure}[ht]
	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/dissection_lifespanPriorsParity.pdf}}
	\caption{\textbf{The priors over mean lifespan used to analyse the data for each of the hierarchical models.} The distributional form of the priors is given in Table \ref{tab:dissection_priors}. In all cases the plots show data for 10,000 samples from the relevant prior distribution.}\label{fig:dissection_lifespanPriorsParity}
\end{figure}

\begin{table}
	\footnotesize
	\noindent\makebox[\textwidth]{%
		\begin{tabularx}{1.2\textwidth}{l|c|X}
			\textbf{Survival function} & \textbf{Physiological age series-level priors} & \textbf{Group-level priors}\\
			\midrule
			Exponential & $\lambda\sim \textnormal{log-normal}(\mu_\lambda,\sigma)$ &$\mu_\lambda \sim N(-1.2,1)\text{,}\; \sigma\sim \textnormal{log-normal}(-2,1)$\\
			\midrule
			Gompertz & $\alpha,\beta\sim$ log-normal$(\mu_{\alpha|\beta},0.2)$ &$\mu_\alpha\sim N(-1.5,1)\text{,}\; \mu_\beta\sim N(-2.5,0.5)$\\
			\midrule
			Weibull & $\alpha,(\beta-1)\sim$  log-normal$(\mu_{\alpha|\beta},0.2)$ &$\mu_\alpha\sim N(-2.5,1)\text{,}\; \mu_\beta\sim N(-4,0.5)$\\
			\midrule
			Gompertz-Makeham & $\alpha,\beta,c\sim$ log-normal$(\mu_{\alpha|\beta|c},0.2)$ &$\mu_\alpha\sim N(-1.4,1)\text{,}\; \mu_\beta\sim N(-3,0.4)\text{,}\; \mu_c\sim N(-4.5,0.5)$\\
			\midrule
			Logistic & $\alpha,\beta,s\sim$  log-normal$(\mu_{\alpha|\beta|s},0.2)$ &$\mu_\alpha\sim N(-1.4,1)\text{,}\; \mu_\beta\sim N(-2,1)\text{,}\; \mu_s\sim N(-3,1)$\\
			\midrule
			Logistic-Makeham & $\alpha,\beta,s,c\sim$ log-normal$(\mu_{\alpha|\beta|s|c},0.2)$ &$\mu_\alpha\sim N(-3,1)\text{,}\; \mu_\beta\sim N(-1.3,1)\text{,}\; \mu_s\sim N(-2,0.5)\text{,}\; \mu_c\sim N(-4,0.5)$\\
			\bottomrule
	\end{tabularx}}\caption{\textbf{The priors used on parameters of each different survival model for the dissection data analysis.} For the exponential model the `group-level' priors were the same for the genus and `overall' models that were also estimated. The exponential model was the only model that was simple enough to allow the scale parameter of the log-normal ($\sigma$) to be estimated by the data. The notation $\alpha,\beta\sim \text{log-normal}(\mu_{\alpha|\beta},0.2)$ means that $\alpha$ and $\beta$ were assigned independent log-normal priors with location parameters $\mu_\alpha$ and $\mu_\beta$ respectively, and a scale parameter of 0.2 in both cases.}\label{tab:dissection_priors}
\end{table}

The estimates of mean mosquito lifespan that we present here (Figure \ref{fig:dissection_lifespan_gonotrophic}) resulted from the use of the exponential survival model (i.e. no age dependence). This choice was made because we found limited evidence in support of age-dependent mortality (see Section \ref{sec:dissection_ageDependence}). However, since the priors for each of the survival functions were specified to allow for a wide variety of lifespans (Figure \ref{fig:dissection_lifespanPriorsParity}) the specific choice of survival function made little difference to resultant estimates (data not shown).

To demonstrate that the results we obtain are not sensitive to the particular form of hierarchical prior structure chosen we also provide estimates of the lifespan for each series analysed on its own. To do so we assume priors on the rate parameter of the exponential distribution that provides support over a wide range of possible lifespans (Figure \ref{fig:dissection_lifespanPrior_individual}A). We also specified a prior on the over-dispersion parameter $\kappa$ that was comparable to the hierarchical case (Figure \ref{fig:dissection_lifespanPrior_individual}B).

\begin{figure}[ht]
	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/dissection_lifespanPrior_individual.pdf}}
	\caption{\textbf{The priors over A. mean lifespan and B. over-dispersion parameter used to analyse the individual dissection series.} The prior on the rate parameter of the exponential distribution was $\lambda\sim \text{log-normal}(-1.8,1)$. The prior on the dispersion parameter was $\kappa\sim \text{log-normal}(1,1)$.}\label{fig:dissection_lifespanPrior_individual}
\end{figure}

As for the MRR analysis we use \textit{Stan} that implements a form of MCMC algorithm known as NUTS \citep{hoffman2014no} to sample from the posterior. The protocol followed for the MCMC sampling was the same as described in Section \ref{sec:mrr_MCMC}.

\subsubsection{Analysis of censored data series}\label{sec:dissection_censored}
In some of the published papers the data are censored above a threshold age, $a_T$. In these cases we only know the number, $y(a_T)$, of individuals who were captured and dissected with an estimated physiological age that is greater than or equal to the threshold. Since we do not know the estimated physiological age of individual specimens $i$, we represent this as a parameter, $\aleph_i$, in our statistical model. This means that the joint distribution is a function of $y(a_T)$ different $\aleph_i$ parameters. Since these parameters are not directly of interest, and our chosen MCMC engine, Stan \citep{stan-software:2014}, does not directly allow discrete parameters in models, we marginalise these out of the joint distribution. To do this exactly would require an infinite sum over all the possible ages for all $y(a_T)$ mosquitoes that have been caught whose age exceeds the threshold. Rather than carry out this intractable number of summations we instead make the approximation that all mosquitoes in this group are of the same age $\aleph_i = \aleph,\; \forall i \in (1,...,y(a_T))$. We believe this assumption is justifiable, particularly since the numbers of mosquitoes in each subsequent age category is a strongly-decreasing function of age. This means that whilst we do not know with certainty individuals' ages, it is likely that most will be of the threshold age $a_T$.

This approximation means that to marginalise the parameter $\aleph$ out of the joint distribution, we are only required to do a single summation. Specifically if we have $y(a_T)$ captured individuals of age equal to, or exceeding, some threshold age, $a_T$, the probability of these observations is given by,
%
\begin{align}
q(y(a_T)|S(.),\Psi,\kappa) &= \sum_{\aleph=a_T}^{\infty} p(y(a_T),\aleph|S(\aleph),\Psi,\kappa)\\
&= \sum_{\aleph=a_T}^{\infty} p(y(a_T)|S(\aleph),\Psi,\kappa) \times p(\aleph)
\end{align}
%
where $q(y(a_T)|S(\aleph),\Psi,\kappa)$ is the probability of observing $y(a_T)$ counts for mosquitoes of an age $\aleph$. In practice, since we do not believe mosquitoes live for longer than 20 gonotrophic cycles (the maximum observed in the data was 13), we cut-off the summation at this point, and assume that the discrete prior probability distribution $p(\aleph)$ is uniform over this range. The overall likelihood for the cases where the series are censored therefore has the form,
%
\begin{equation}
\mathcal{L}(y(a_1),y(a_2),...,y(a_T-1),y(a_T)|S(.),\Psi,\kappa) = \left(\prod\limits_{a=a_1}^{a_T-1} p(y(a)|S(a),\Psi,\kappa)\right) q(y(a_T)|S(\aleph),\Psi,\kappa).
\end{equation}
%
In practice, the number of mosquitoes captured in those series that are censored represents a small percentage of total captures (the median is approximately 2\%), and so the effect of the $q(.)$ term above is likely minimal on resultant inferences.

\subsection{Model estimation by MCMC}\label{sec:dissection_MCMC}
As for the MRR analysis, we used \textit{Stan} software \citep{carpenter2016stan} to sample from the posterior distributions of model parameters.

To judge convergence of the sampling algorithm to the posterior density, we calculated $\hat{R}$ across all Markov chains -- a ratio that compares the between-chain variation to that within each chain that is commonly used to measure convergence in MCMC \citep{gelman1992inference}. For each model, we ran 16 independent Markov chains with 200 iterations per chain, discarding the first half of these iterations as warm-up \citep{gelman2014bayesian}. After running the algorithm for the given number of observations, $\hat{R}<1.1$ for all model parameters. We also ensured that across each MCMC run the number of divergent iterations (that can bias the MCMC away from the true posterior density) was minimal. In the majority of cases, the number of divergent iterations was far fewer than 1\% of the total number of samples.

The model used to estimate the lifespan in terms of gonotrophic cycles is provided below.

\begin{minted}{stan}
data{
int N; // number of obs
int K; // number of time series
int Y[N]; // number captured
int S[K]; // length of each series
int threshold[K]; // age of censoring; if -1 then no censoring
int Pos[K]; // position of start of each series in Y
int species[K]; // index variable of species
int nSpecies; // number of species
}

parameters{
real<lower=0> PopSize[K];
real<lower=0> alpha[K];
real alpha_mean[nSpecies];
real<lower=0> alpha_sigma[nSpecies];
real<lower=0> kappa[K];
real<lower=0> p[nSpecies];
}

model{
// Likelihood
for(i in 1:K){
  for(t in 2:S[i]){
    // uncensored
    if(threshold[i] < 0)
      Y[Pos[i]+t-1] ~ neg_binomial_2(PopSize[i] *
			exp(-alpha[i] * (t-1)), kappa[i]);
    // censoring
    else{
      if((t-1) < threshold[i])
        Y[Pos[i]+t-1] ~ neg_binomial_2(PopSize[i] *
			exp(-alpha[i] * (t-1)), kappa[i]);
      else{
        real lLogProb[20];
        for(j in 1:20){
          lLogProb[j] = neg_binomial_2_lpmf(Y[Pos[i]+t-1] |
			PopSize[i] * exp(-alpha[i] * (t-1)),
			kappa[i]);
        }
        target += log_sum_exp(lLogProb);
      }
    }
  }  
}

// Priors
for(i in 1:K){
  int aSpecies;
  aSpecies = species[i];
  Y[Pos[i]] ~ normal(PopSize[i], sqrt(PopSize[i]));
  alpha[i] ~ lognormal(alpha_mean[aSpecies],
  			alpha_sigma[aSpecies]);
  kappa[i] ~ exponential(p[aSpecies]);
}

alpha_mean ~ normal(-1.2, 1);
alpha_sigma ~ lognormal(-2, 1);
p ~ lognormal(1, 1);
}

generated quantities{
real alpha_average[nSpecies];
for(i in 1:nSpecies)
  alpha_average[i] = lognormal_rng(alpha_mean[i],alpha_sigma[i]);
}
\end{minted}

\subsection{Model checking}
As for the MRR experiments, we simulated from the posterior predictive distribution for each series of dissection data. In the majority of the cases, the capture data lay within the 95\% posterior predictive intervals, indicating that our model was a good fit to the data (see Figures \ref{fig:dissection_ppc_all1}, \ref{fig:dissection_ppc_all2} \&  \ref{fig:dissection_ppc_all3} and the attached file, \verb|dissection_ppcs_all.pdf| for the full graphs).

\newgeometry{scale=0.8}
\thispagestyle{empty}
{%
	\centering
	\includegraphics[page=1,scale=.7]{./Figure_files/dissection_ppcs_all}
	\captionof{figure}{The numbers of mosquitoes recaptured (black lines) versus the 95\% central posterior interval of the posterior predictive distribution (blue shading) for a selection of the dissection series. The titles correspond to the unique combination of "identifier - physiological-experiment-id" in the database. For the rest of the posterior predictive checks for the dissection data models, see the file referenced in the text.}
	\label{fig:dissection_ppc_all1}
	\par
}
\restoregeometry

\newgeometry{scale=0.8}
\thispagestyle{empty}
{%
	\centering
	\includegraphics[page=2,scale=.7]{./Figure_files/dissection_ppcs_all}
	\captionof{figure}{The numbers of mosquitoes recaptured (black lines) versus the 95\% central posterior interval of the posterior predictive distribution (blue shading) for a selection of the dissection series. The titles correspond to the unique combination of "identifier - physiological-experiment-id" in the database. For the rest of the posterior predictive checks for the dissection data models, see the file referenced in the text.}
	\label{fig:dissection_ppc_all2}
	\par
}
\restoregeometry

\newgeometry{scale=0.8}
\thispagestyle{empty}
{%
	\centering
	\includegraphics[page=3,scale=.7]{./Figure_files/dissection_ppcs_all}
	\captionof{figure}{The numbers of mosquitoes recaptured (black lines) versus the 95\% central posterior interval of the posterior predictive distribution (blue shading) for a selection of the dissection series. The titles correspond to the unique combination of "identifier - physiological-experiment-id" in the database. For the rest of the posterior predictive checks for the dissection data models, see the file referenced in the text.}
	\label{fig:dissection_ppc_all3}
	\par
}
\restoregeometry

\subsection{Data collection of gonotrophic cycle duration}\label{sec:dissection_gonotrophicData}
To convert the estimates of lifespan in physiological age into chronological age we require estimates of the duration of the gonotrophic cycle. To determine this characteristic we conducted a meta-analysis of previously-published studies that estimate the duration of the gonotrophic cycle. A search of the literature using Google Scholar (\url{scholar.google.co.uk}) was performed using the search term: `gonotrophic cycle duration'. The list of articles was then supplemented with a list of references discussed by \cite{silver2007mosquito}. Based on the abstracts of the resultant list of published studies we then decided whether to search each article for estimates of the duration of the gonotrophic cycle. Overall 79 separate estimates of this parameter were found across 42 published articles (see Section \ref{sec:appendix_gonotrophicDurationStudyList} for a list of the studies included in our dataset). Along with information about the estimates we also recorded study and series meta-data, including the location of the study, the method used for estimation, the species and genus, as well as the temperatures and/or seasons in which the experiments were carried out. The methods used to estimate gonotrophic cycle duration in the literature largely fell into two distinct categories: those based on MRR studies ($n=29$); and those based on observations of mosquitoes in a laboratory setting ($n=42$). In the MRR studies estimates are made of the duration of gonotrophic cycles by dissecting recaptured mosquitoes at each time point using the method of \cite{polovodova1949determination} to count ovariolar dilations. For the laboratory studies the duration of gonotrophic cycles is determined more directly by observing the time taken for mosquitoes to mate then blood-feed and finally oviposit.

Along with point estimates of the parameter we also collected information about the uncertainty in the estimates (if available). In many articles the duration of the gonotrophic cycle was estimated separately for the first versus subsequent cycles, and these estimates were recorded separately. Raw estimates of gonotrophic cycle duration were obtained for species across four different genera (\textit{Aedes}, \textit{Anopheles}, \textit{Culex} and \textit{Masonia}), although there was a bias towards \textit{Anopheles} with $n=47$ estimates out of the total of 79 (Figure \ref{fig:dissection_gonotrophicCycleRaw}).

\begin{figure}[ht]
	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/dissection_gonotrophicCycleRaw.pdf}}
	\caption{\textbf{Raw gonotrophic cycle duration estimates for A. the 1st cycle and B. subsequent cycles.} The confidence intervals represent a range of uncertainties since they were not available in a standardised form (see Section \ref{sec:dissection_gonotrophicMethod} for a discussion of how we handle this issue). If no statement of uncertainty was given in the estimates then we indicate this by a triangle marker. The studies are ordered according to the central estimate of gonotrophic duration.}\label{fig:dissection_gonotrophicCycleRaw}
\end{figure}

Since we collected data on the method used to estimate the duration of the gonotrophic cycle we show the raw estimates for the two most common approaches: MRR and laboratory (Figure \ref{fig:dissection_gonotrophicCycleRaw_MRRVsLab}). From these results it is evident that the estimates from the laboratory studies are, on average, higher than those using the MRR method, (although we wait until Section \ref{sec:dissection_gonotrophicDuration} for a quantitative comparison).

\begin{figure}[ht]
	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/dissection_gonotrophicCycleRaw_MRRVsLab.pdf}}
	\caption{\textbf{Raw gonotrophic cycle duration estimates from A. the MRR and B. from laboratory-based studies.} The confidence intervals represent heterogeneous measures since they were not available in a standardised form (see Section \ref{sec:dissection_gonotrophicMethod} for a discussion of we handle this issue). If no statement of uncertainty was given in the estimates then we indicate this by a triangle marker. The studies are ordered according to the central estimate of gonotrophic duration.}\label{fig:dissection_gonotrophicCycleRaw_MRRVsLab}
\end{figure}

\subsection{Statistical analysis of gonotrophic cycle data}\label{sec:dissection_gonotrophicMethod}
As discussed in Section \ref{sec:dissection_gonotrophicData}, there was considerable study-level heterogeneity in the information provided for the estimates of the gonotrophic cycle duration. A number of studies ($n=24$) provided no estimates of uncertainty in gonotrophic cycle duration, whereas the rest gave some indication of confidence or alternatively a range of possible estimates. However the types of uncertainty intervals that were specified varied considerably from study to study, from the vague (but common)  `4-6 days' to the more helpful `5$\pm 1$ day (95\% confidence interval)'.

The heterogeneous nature of the gonotrophic cycle estimates requires a method that explicitly accounts for this characteristic of the data. We decided to model the estimates as representing quantiles from an underlying normal distribution that represents uncertainty over possible durations of the gonotrophic cycle. In those circumstances where lower, central and upper bounds were given explicitly the data was fairly symmetric and so we believe assuming the observations come from an unskewed normal distribution is not unreasonable. Explicitly we treated each type of uncertainty interval as follows,

\begin{itemize}
	\item Simple range, for example, `4-6 days': treat the lower and upper bounds as the 2.5\% and 97.5\% quantiles of a normal distribution.
	\item Confidence intervals, `5$\pm 1$ day (95\% confidence interval)': treat the lower and upper bounds as the relevant quantiles of a normal distribution.
	\item Point estimates, for example `5 days': treat this as the median of a normal distribution.
\end{itemize}

The benefit of this approach is that we can convert the quantiles from our normal distribution parameterised by a mean ($\mu$) and standard deviation ($\sigma$) to equivalent quantiles from a standard normal,
%
\begin{align}
Z_{50} = \frac{G_{50} - \mu}{\sigma},
\end{align}
%
where $Z_{50}$ and $G_{50}$ indicate 50\% quantiles from a standard normal distribution and a normal distribution respectively. By manipulating the above expression we obtain the following,
%
\begin{align}
G_{50} = \sigma Z_{50} + \mu,
\end{align}
%
which forms a straight line in $(Z,G)$ space with slope $\sigma$ and y-intercept $\mu$. Therefore by plotting the raw observations of $(Z,G)$ quantiles then estimating a linear regression line we can characterise the underlying $\mathcal{N}(\mu,\sigma)$ distribution from which we assume they are drawn.

Due to the relative unavailability of estimates for gonotrophic cycle duration across the different genera we generated a single distribution to represent the uncertainty in this parameter by pooling all the data. This was justified because there was not found to be significant variation in the 1st gonotrophic cycle duration (ANOVA: $F_{3,75}=2.6,p>0.05$).

\subsection{Conversion of lifespan from physiological to calendar age}\label{sec:dissection_conversion}
The estimates of lifespan produced from analysing the dissection data are in terms of physiological age (the number of gonotrophic cycles undertaken). To allow comparison with the estimates from the MRR studies (Figure \ref{fig:mrr_lifespans}) as well as produce more useful estimates to inform disease transmission dynamical models we convert these estimates to calendar days. To do this we use the estimated parameters of the normal densities that we have assumed represent uncertainty in gonotrophic cycle duration (see Section \ref{sec:dissection_gonotrophicMethod}) and use them to convert our posterior samples of mean physiological lifespan $(L^p_1,...,L^p_S)$ to lifespan in calendar ages $(L^c_1,...,L^c_S)$. To do this we iterate the following for all $i\in (1,...,S)$ posterior samples,

\begin{enumerate}
	\item Sample $G_{1i} \sim \mathcal{N}(\mu_{1i},\sigma_{1i})$, to obtain a duration for the 1st gonotrophic cycle.
	\item Sample $G_{2i} \sim \mathcal{N}(\mu_{2i},\sigma_{2i})$, to obtain a duration for subsequent gonotrophic cycles.
	\item If $L^p_i > 1$, the mean lifespan is longer than one gonotrophic cycle:
	\subitem then $L^c_i = G_{1i} + G_{2i} \times (L^p_i - 1)$.
	\item Else:
	\subitem then $L^c_i = G_{1i}\times L^p_i$.
\end{enumerate}

In the using the above methodology to convert from physiological to calendar age we implicitly assume that for each gonotrophic cycle after the first that the cycles are of the same length (we choose one $G_{2i}$ per sample). However since we allow a different $G_{2i}$ for each sample we nonetheless believe that the above approach allows for sufficient uncertainty in the estimates of subsequent gonotrophic cycle durations. If instead we believed that significant variation in gonotrophic cycle occurred within a particular mosquito's life (as well as between mosquitoes) then we could draw a new value of $G_{2i}$ for each subsequent cycle. This produces results with a little more uncertainty than previously (data not shown).

\section{Results}
\subsection{MRR experments}
Figures \ref{fig:mrr_lifeSpanVsRange}, \ref{fig:mrr_lifeSpanVsTrapDensity} and \ref{fig:mrr_temperature} show the estimates of lifespan versus the trapping range, spatial density and temperature at study location. In these cases, the lifespans shown were obtained from the non-hierarchical model with the exponential survival model.

Figure \ref{fig:mrr_sex} shows the lifespans obtained by sex, for each genus and across all studies ("overall") using the hierarchical Bayesian model with exponential survival model. In the genus-level models, the sex-specific effects (male-only, female-only, mixed-releases) were estimated as independent random effects for each genus with standard normal priors. For the overall model, the sex-specific effects were estimated as independent random effects across all studies, with a standard normal prior.

Figure \ref{fig:mrr_sugarBlood} shows the lifespans obtained for each pre-release feeding category for each genus and across all studies using the hierarchical Bayesian model with exponential survival model. In the genus-level models, the feeding-specific effects (fed nothing, blood-fed-only, sugar-fed-only and fed both) were estimated as independent random effects for each genus with standard normal priors. For the overall model, the feeding-specific effects were estimated as independent random effects across all studies, with a standard normal prior.

\begin{figure}[h]
	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_lifeSpanVsRange.pdf}}
	\caption{\textbf{Posterior estimates of mean mosquito lifespan for each time-series versus the trapping range.} The markers show the median posterior estimates, with the lower and upper bounds indicating the 25\% and 75\% quantiles, respectively. The black line shows a linear regression line estimated using the median posterior lifetimes, with the grey shading indicating 95\% confidence intervals. All estimates were obtained using the non-hierarchical exponential survival model.}
	\label{fig:mrr_lifeSpanVsRange}
\end{figure}

\begin{figure}[h]
	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_lifespanVsTrapDensity.pdf}}
	\caption{\textbf{Posterior estimates of mean mosquito lifespan for each time-series versus the trap density.} The markers show the median posterior estimates, with the lower and upper bounds indicating the 25\% and 75\% quantiles, respectively. The trap density was estimated assuming traps were contained within a circular area with radius equal to the trapping range. The black line shows a linear regression line estimated using the median posterior lifetimes, with the grey shading indicating 95\% confidence intervals. All estimates were obtained using the non-hierarchical exponential survival model.}
	\label{fig:mrr_lifeSpanVsTrapDensity}
\end{figure}

\begin{figure}[h]
	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_ThreeSpeciesVersusTemperature.pdf}}
	\caption{\textbf{Posterior estimates of mean mosquito lifespan for each time-series versus the average monthly temperature for the four species with the most data.} The markers show the median posterior estimates, with the lower and upper bounds indicating the 25\% and 75\% quantiles, respectively. The black line shows a regression line with linear and quadratic terms estimated using the median posterior lifetimes, with the grey shading indicating 95\% confidence intervals. All estimates of lifespan were obtained using the non-hierarchical exponential survival model.}\label{fig:mrr_ThreeSpeciesVersusTemperature}
\end{figure}

\begin{figure}[h]
	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_female_blood_sugar.pdf}}
	\caption{\textbf{Posterior estimates of female mosquito mean lifespan according to pre-release feeding status across species, genus and overall groupings as determined from the MRR data.} The lifespans shown are for mosquitoes that were not fed with sugar or blood (for females) before release. The middle line in each box shows the median estimates and the solid dot indicates the mean. The left and right box edges show the 25\%, and 75\% posterior quantiles respectively. The whiskers show the range of the data, excluding points lying more than 1.5 times the interquartile range away from each edge of the box. The numbers before the start of the left whisker indicate the number of individual time-series within each species. All estimates were obtained using the hierarchical exponential survival model.}
	\label{fig:mrr_female_blood_sugar}
\end{figure}

\subsubsection{Table of estimated lifespans estimated from mark-release-recapture studies}

\begin{table}[htbp]
\begin{tabular}{l|l|l|l|l|l|l|l|l}
	\textbf{Genus} & \textbf{Species} & \textbf{5$\%$} & \textbf{25$\%$} & \textbf{50$\%$} & \textbf{75$\%$} & \textbf{95\%} &  \textbf{Mean} & \textbf{Std. dev.}\\
		\hline
 \textit{Aedes} & \textit{simpsoni s.l.} & 6.3 & 11.3 & 18.3 & 31.8 & 75.0 & 26.9 & 29.5 \\
\textit{Aedes} & \textit{communis} & 10.0 & 13.2 & 15.4 & 18.2 & 24.3 & 16.2 & 5.2 \\
\textit{Anopheles} & \textit{koliensis} & 6.7 & 10.5 & 15.2 & 24.0 & 56.5 & 21.9 & 25.2 \\
\textit{Anopheles} & \textit{punctulatus} & 7.2 & 10.7 & 15.1 & 22.8 & 48.9 & 20.3 & 28.6 \\
\textit{Aedes} & \textit{cantans} & 9.7 & 12.0 & 13.9 & 16.1 & 20.6 & 14.4 & 4.7 \\
\textit{Aedes} & \textit{albopictus} & 8.0 & 10.0 & 11.6 & 13.7 & 17.7 & 12.1 & 3.4 \\
\textit{Anopheles} & \textit{albimanus} & 5.9 & 8.6 & 11.6 & 16.9 & 35.7 & 15.2 & 18.7 \\
\textit{Anopheles} & \textit{darlingi} & 4.9 & 7.3 & 10.2 & 16.0 & 37.5 & 14.5 & 14.8 \\
\textit{Anopheles} & \textit{maculatus} & 5.7 & 7.9 & 10.2 & 13.7 & 24.3 & 12.2 & 8.7 \\
\textit{Anopheles} & \textit{culicifacies s.l.} & 4.6 & 7.3 & 9.1 & 11.4 & 18.2 & 10.1 & 6.9 \\
\textit{Anopheles} & \textit{sergenti} & 3.6 & 5.9 & 8.7 & 13.8 & 32.6 & 12.4 & 13.4 \\
\textit{Aedes} & \textit{triseriatus} & 5.6 & 6.6 & 7.5 & 8.5 & 10.7 & 7.7 & 1.8 \\
\textit{Anopheles} & \textit{farauti s.l.} & 4.2 & 5.4 & 6.4 & 7.7 & 10.5 & 6.8 & 2.1 \\
\textit{Aedes} & \textit{africanus} & 3.4 & 5.1 & 6.2 & 7.8 & 11.9 & 7.0 & 9.3 \\
\textit{Aedes} & \textit{aegypti} & 2.8 & 4.5 & 6.2 & 8.5 & 13.9 & 7.0 & 3.8 \\
\textit{Culex} & \textit{quinquefasciatus} & 4.3 & 5.2 & 5.9 & 7.0 & 9.4 & 6.3 & 1.8 \\
\textit{Culex} & \textit{nigripalpus} & 3.5 & 4.6 & 5.8 & 7.5 & 11.3 & 6.4 & 3.1 \\
\textit{Aedes} & \textit{notoscriptus} & 2.9 & 4.0 & 5.4 & 8.2 & 20.2 & 7.9 & 9.5 \\
\textit{Culex} & \textit{pipiens} & 3.4 & 4.4 & 5.2 & 6.4 & 9.0 & 5.6 & 2.0 \\
\textit{Anopheles} & \textit{saperoi} & 3.5 & 4.2 & 4.9 & 5.8 & 8.2 & 5.2 & 2.0 \\
\textit{Anopheles} & \textit{gambiae s.l.} & 3.0 & 3.8 & 4.4 & 5.1 & 6.4 & 4.5 & 1.1 \\
\textit{Aedes} & \textit{melanimon} & 3.3 & 3.9 & 4.3 & 4.8 & 5.7 & 4.4 & 1.0 \\
\textit{Anopheles} & \textit{funestus s.l.} & 2.9 & 3.6 & 4.2 & 4.8 & 6.0 & 4.3 & 1.0 \\
\textit{Anopheles} & \textit{lesteri} & 2.4 & 3.2 & 3.9 & 5.0 & 7.5 & 4.4 & 2.1 \\
\textit{Anopheles} & \textit{quadrimaculatus} & 3.3 & 3.6 & 3.9 & 4.2 & 4.8 & 4.0 & 0.5 \\
\textit{Anopheles} & \textit{pseudopunctipennis s.l.} & 2.3 & 2.9 & 3.5 & 4.3 & 6.7 & 3.8 & 1.9 \\
\textit{Aedes} & \textit{vexans} & 1.9 & 2.3 & 2.6 & 3.0 & 4.1 & 2.8 & 1.0 \\
\textit{Culex} & \textit{tarsalis} & 1.9 & 2.1 & 2.3 & 2.4 & 2.7 & 2.3 & 0.3 \\
\textit{Anopheles} & \textit{vestitipennis} & 1.3 & 1.8 & 2.2 & 2.9 & 4.4 & 2.5 & 1.2 \\
\textit{Anopheles} & \textit{albitarsis s.l.} & 1.3 & 1.6 & 1.9 & 2.2 & 2.9 & 2.0 & 1.0 \\
\textit{Culex} & \textit{tritaeniorhynchus} & 0.7 & 0.9 & 1.1 & 1.2 & 1.6 & 1.1 & 1.0 \\
\textit{Anopheles} & \textit{subpictus s.l.} & 0.4 & 0.6 & 0.8 & 1.1 & 2.4 & 1.1 & 2.6 \\
\hline
\textit{Aedes} & \text{} & 2.8 & 4.8 & 6.9 & 10.0 & 17.2 & 8.1 & 4.9 \\
\textit{Anopheles} & \text{} & 1.5 & 3.0 & 5.0 & 8.4 & 17.8 & 6.8 & 6.4 \\
\textit{Culex} & \text{} & 1.1 & 1.8 & 2.5 & 3.5 & 5.6 & 2.9 & 1.5 \\
\hline
\textbf{Overall} & \text{} & 1.4 & 2.8 & 4.6 & 7.5 & 15.4 & 6.0 & 5.2 \\
\end{tabular}
\caption{Posterior summaries of mean mosquito lifespan from the MRR analysis. The 5\%, 25\%, 50\%, 75\% and 95\% columns indicate the respective quantiles of the posterior distribution for mean lifespan, and the `Mean' and `Std. dev.' columns indicate the posterior mean and standard deviation of mean lifespan.}
\label{tab:mrr_estimated_lifespans}%
\end{table}%

\subsection{Dissection studies}

\subsubsection{Table of estimated lifespan in terms of gonotrophic cycles and chronological time using dissection data}

\begin{table}[htbp]
\begin{tabular}{l|l|l|l|l|l|l|l|l}
	\textbf{Genus} & \textbf{Species} & \textbf{5$\%$} & \textbf{25$\%$} & \textbf{50$\%$} & \textbf{75$\%$} & \textbf{95\%} &  \textbf{Mean} & \textbf{Std. dev.}\\
		\hline
		\textit{Anopheles} & \textit{sergentii} & 1.0 & 1.9 & 2.5 & 3.4 & 6.2 & 3.0 & 2.5 \\
		\textit{Anopheles} & \textit{gambiae s.l.} & 0.7 & 1.2 & 1.9 & 2.9 & 5.4 & 2.4 & 1.9 \\
		\textit{Culex} & \textit{thalassius} & 1.1 & 1.6 & 1.8 & 2.1 & 2.9 & 1.9 & 1.0 \\
		\textit{Anopheles} & \textit{farauti s.l.} & 1.5 & 1.6 & 1.7 & 1.7 & 1.9 & 1.7 & 0.1 \\
		\textit{Anopheles} & \textit{minimus} & 0.7 & 1.2 & 1.7 & 2.2 & 4.1 & 2.0 & 2.2 \\
		\textit{Anopheles} & \textit{maculipennis} & 0.8 & 1.1 & 1.4 & 1.8 & 2.9 & 1.6 & 0.7 \\
		\textit{Culex} & \textit{pipiens} & 1.1 & 1.3 & 1.4 & 1.5 & 1.7 & 1.4 & 0.2 \\
		\textit{Mansonia} & \textit{uniformis} & 0.9 & 1.1 & 1.1 & 1.2 & 1.4 & 1.1 & 0.2 \\
		\textit{Anopheles} & \textit{rivulorum} & 0.8 & 1.0 & 1.1 & 1.1 & 1.4 & 1.1 & 0.2 \\
		\textit{Anopheles} & \textit{melas} & 0.8 & 1.0 & 1.0 & 1.1 & 1.3 & 1.0 & 0.2 \\
		\textit{Anopheles} & \textit{culicifacies} & 0.9 & 1.0 & 1.0 & 1.1 & 1.2 & 1.0 & 0.1 \\
		\textit{Anopheles} & \textit{subpictus} & 0.6 & 0.9 & 1.0 & 1.2 & 2.1 & 1.2 & 2.0 \\
		\textit{Aedes} & \textit{aegypti} & 0.5 & 0.8 & 1.0 & 1.2 & 2.0 & 1.2 & 3.0 \\
		\textit{Anopheles} & \textit{quadrimaculatus} & 0.7 & 0.9 & 1.0 & 1.1 & 1.3 & 1.0 & 0.2 \\
		\textit{Anopheles} & \textit{stephensi} & 0.6 & 0.8 & 1.0 & 1.1 & 1.6 & 1.0 & 0.4 \\
		\textit{Anopheles} & \textit{darlingi} & 0.7 & 0.8 & 0.9 & 1.0 & 1.3 & 1.0 & 0.2 \\
		\textit{Culex} & \textit{quinquefasciatus} & 0.4 & 0.7 & 0.9 & 1.3 & 2.2 & 1.1 & 0.6 \\
		\textit{Aedes} & \textit{polynesiensis} & 0.8 & 0.9 & 0.9 & 0.9 & 1.0 & 0.9 & 0.1 \\
		\textit{Culex} & \textit{tritaeniorhynchus} & 0.3 & 0.6 & 0.9 & 1.3 & 4.2 & 2.5 & 30.8 \\
		\textit{Culex} & \textit{annulirostris} & 0.5 & 0.8 & 0.9 & 1.0 & 1.4 & 0.9 & 0.5 \\
		\textit{Aedes} & \textit{samoanus} & 0.6 & 0.7 & 0.8 & 0.9 & 1.0 & 0.8 & 0.2 \\
		\textit{Aedes} & \textit{sollicitans} & 0.4 & 0.5 & 0.6 & 0.8 & 1.3 & 0.7 & 0.5 \\
		\textit{Aedes} & \textit{vexans} & 0.4 & 0.5 & 0.6 & 0.7 & 0.9 & 0.6 & 0.7 \\
		\textit{Anopheles} & \textit{cruzii} & 0.3 & 0.5 & 0.5 & 0.7 & 1.1 & 0.6 & 0.4 \\
		\textit{Anopheles} & \textit{bellator} & 0.3 & 0.4 & 0.5 & 0.6 & 1.0 & 0.6 & 1.8 \\
		\hline
		\textit{Anopheles} & \textit{} & 0.6 & 1.0 & 1.4 & 1.9 & 3.2 & 1.6 & 0.9 \\
		\textit{Mansonia} & \textit{} & 0.9 & 1.1 & 1.1 & 1.2 & 1.4 & 1.1 & 0.2 \\
		\textit{Culex} & \textit{} & 0.5 & 0.8 & 1.0 & 1.4 & 2.3 & 1.2 & 0.6 \\
		\textit{Aedes} & \textit{} & 0.6 & 0.7 & 0.8 & 0.9 & 1.1 & 0.8 & 0.2 \\
		\hline
		\textbf{Overall} & \textit{} & 0.5 & 0.8 & 1.2 & 1.7 & 2.7 & 1.3 & 0.7 \\
	\end{tabular}
\caption{Posterior summaries of mean mosquito lifespan in terms of gonotrophic cycles from dissection studies. The 5\%, 25\%, 50\%, 75\% and 95\% columns indicate the respective quantiles of the posterior distribution for mean lifespan, and the `Mean' and `Std. dev.' columns indicate the posterior mean and standard deviation of mean lifespan.}
\label{tab:dissection_estimated_lifespans}%
\end{table}

\begin{table}[htbp]
	\begin{tabular}{l|l|l|l|l|l|l|l|l}
		\textbf{Genus} & \textbf{Species} & \textbf{5$\%$} & \textbf{25$\%$} & \textbf{50$\%$} & \textbf{75$\%$} & \textbf{95\%} &  \textbf{Mean} & \textbf{Std. dev.}\\
		\hline
		\textit{Anopheles} & \textit{sergentii} & 4.2 & 7.6 & 10.1 & 13.6 & 24.6 & 11.9 & 9.7 \\
		\textit{Anopheles} & \textit{gambiae s.l.} & 2.8 & 5.2 & 7.6 & 11.4 & 21.6 & 9.5 & 7.4 \\
		\textit{Culex} & \textit{thalassius} & 4.6 & 6.3 & 7.4 & 8.6 & 12.0 & 7.9 & 4.1 \\
		\textit{Anopheles} & \textit{farauti s.l.} & 5.9 & 6.5 & 6.9 & 7.3 & 8.0 & 6.9 & 0.7 \\
		\textit{Anopheles} & \textit{minimus} & 3.1 & 5.2 & 6.9 & 9.0 & 16.4 & 8.1 & 9.4 \\
		\textit{Anopheles} & \textit{maculipennis} & 3.3 & 4.8 & 5.9 & 7.5 & 11.4 & 6.5 & 2.7 \\
		\textit{Culex} & \textit{pipiens} & 4.6 & 5.3 & 5.7 & 6.2 & 6.9 & 5.8 & 0.8 \\
		\textit{Mansonia} & \textit{uniformis} & 3.8 & 4.4 & 4.8 & 5.2 & 6.0 & 4.8 & 0.8 \\
		\textit{Anopheles} & \textit{rivulorum} & 3.4 & 4.0 & 4.5 & 4.9 & 5.9 & 4.5 & 0.8 \\
		\textit{Anopheles} & \textit{melas} & 3.3 & 3.9 & 4.3 & 4.8 & 5.5 & 4.4 & 0.7 \\
		\textit{Anopheles} & \textit{culicifacies} & 3.4 & 4.0 & 4.3 & 4.7 & 5.4 & 4.4 & 0.6 \\
		\textit{Aedes} & \textit{aegypti} & 2.1 & 3.5 & 4.3 & 5.0 & 7.9 & 5.0 & 11.5 \\
		\textit{Anopheles} & \textit{subpictus} & 2.6 & 3.6 & 4.3 & 5.1 & 8.3 & 5.0 & 7.7 \\
		\textit{Anopheles} & \textit{quadrimaculatus} & 2.9 & 3.6 & 4.1 & 4.7 & 5.6 & 4.2 & 0.9 \\
		\textit{Anopheles} & \textit{stephensi} & 2.4 & 3.4 & 4.1 & 4.9 & 6.9 & 4.3 & 1.8 \\
		\textit{Culex} & \textit{quinquefasciatus} & 1.7 & 2.8 & 4.0 & 5.5 & 9.0 & 4.5 & 2.5 \\
		\textit{Anopheles} & \textit{darlingi} & 2.7 & 3.5 & 3.9 & 4.5 & 5.5 & 4.0 & 1.0 \\
		\textit{Aedes} & \textit{polynesiensis} & 3.1 & 3.5 & 3.8 & 4.1 & 4.5 & 3.8 & 0.5 \\
		\textit{Culex} & \textit{tritaeniorhynchus} & 1.3 & 2.6 & 3.8 & 5.7 & 16.8 & 9.8 & 123.0 \\
		\textit{Culex} & \textit{annulirostris} & 2.3 & 3.1 & 3.7 & 4.4 & 5.9 & 3.9 & 1.8 \\
		\textit{Aedes} & \textit{samoanus} & 2.6 & 3.0 & 3.3 & 3.7 & 4.4 & 3.4 & 0.9 \\
		\textit{Aedes} & \textit{sollicitans} & 1.5 & 2.2 & 2.7 & 3.3 & 5.4 & 3.0 & 2.0 \\
		\textit{Aedes} & \textit{vexans} & 1.7 & 2.1 & 2.5 & 2.9 & 3.9 & 2.7 & 2.7 \\
		\textit{Anopheles} & \textit{cruzii} & 1.4 & 1.9 & 2.3 & 2.9 & 4.6 & 2.6 & 1.5 \\
		\textit{Anopheles} & \textit{bellator} & 1.2 & 1.7 & 2.0 & 2.5 & 4.3 & 2.5 & 7.2 \\
		\hline
		\textit{Anopheles} & \textit{} & 2.5 & 4.1 & 5.7 & 7.9 & 12.7 & 6.4 & 3.4 \\
		\textit{Mansonia} & \textit{} & 3.7 & 4.4 & 4.8 & 5.1 & 5.9 & 4.8 & 0.8 \\
		\textit{Culex} & \textit{} & 2.1 & 3.2 & 4.4 & 5.9 & 9.5 & 4.9 & 2.4 \\
		\textit{Aedes} & \textit{} & 2.4 & 3.1 & 3.5 & 3.9 & 4.7 & 3.5 & 0.7 \\
		\hline
		\textbf{Overall} & \textit{} & 2.2 & 3.5 & 5.1 & 6.8 & 10.7 & 5.5 & 2.7 \\
	\end{tabular}
\caption{Posterior summaries of mean mosquito lifespan in terms of chronological time from dissection studies. The 5\%, 25\%, 50\%, 75\% and 95\% columns indicate the respective quantiles of the posterior distribution for mean lifespan, and the `Mean' and `Std. dev.' columns indicate the posterior mean and standard deviation of mean lifespan.}
\label{tab:dissection_estimated_lifespans_chron}%
\end{table}

\subsubsection{Chronological lifespan estimates}
\begin{figure}[h]
	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/dissection_lifetimes_exponential_chron.pdf}}
	\caption{\textbf{Posterior estimates of female mosquito mean chronological lifespan across species, genus and overall groupings as determined from the dissection data.} The lifespans shown are for mosquitoes that were not fed with sugar or blood (for females) before release. The middle line in each box shows the median estimates and the solid dot indicates the mean. The left and right box edges show the 25\%, and 75\% posterior quantiles respectively. The whiskers show the range of the data, excluding points lying more than 1.5 times the interquartile range away from each edge of the box. The numbers before the start of the left whisker indicate the number of individual time-series within each species. All estimates were obtained using the hierarchical exponential survival model.}
	\label{fig:dissection_lifetimes_exponential}
\end{figure}

\subsection{Comparison of LBLs between the MRR and dissection analyses}
To conduct a comparison between the LBLs from the MRR analysis and from the dissections, we converted the latter lifespans from gonotrophic cycles to chronological age as described in Section \ref{sec:dissection_conversion}. We conducted ANOVA using the individual time series estimates of LBL for each of the 10 species with estimates from both MRR- and dissection-based analysis. The results of this analysis are shown in Table \ref{tab:comparison}.

\begin{table}[htbp]
	\centering
	\begin{tabular}{l|l|l|l|l}
		species & \multicolumn{1}{l}{$\text{df}_1$} & \multicolumn{1}{l}{$\text{df}_2$} & \multicolumn{1}{l}{F} & \multicolumn{1}{l}{p} \\
		\midrule
		\textit{Ae. aegypti} & 1     & 55    & 2.2   & 0.14 \\
		\textit{A. darlingi} & 1     & 6     & 114.8   & 0.00 \\
		\textit{An. farauti s.l.} & 1     & 14    & 0.9  & 0.35 \\
		\textit{A. gambiae s.l.} & 1     & 37    & 3.9   & 0.06 \\
		\textit{Cx. pipiens} & 1     & 13    & 0.6   & 0.47 \\
		\textit{A. quadrimaculatus} & 1     & 8     & 2.4     & 0.16 \\
		\textit{Cx. quinquefasciatus} & 1     & 13    & 2.2    & 0.17 \\
		\textit{A. sergentii} & 1     & 8     & 0.4     & 0.55 \\
		\textit{Cx. tritaeniorhynchus} & 1     & 5     & 3.5   & 0.12 \\
		\textit{Ae. vexans} & 1     & 2     & 1.2   & 0.39 \\
	\end{tabular}%
	\caption{The results of comparisons between LBLs for those species with data for MRRs and dissection-based studies. The ANOVA tests were conducted using LBLs estimated at the time-series level.}
	\label{tab:comparison}%
\end{table}%

\subsection{EIPs of vector species of malaria, dengue fever, chikungunya and Zika}
The following anopheline species were identified as vector species of malaria,

\begin{itemize}
	\item \textbf{Africa}: \textit{A. gambiae s.l.}, \textit{A. funestus} and \textit{A. melas} \citep{sinka2012global}, and \textit{A. rivulorum} \citep{wilkes1996anopheles}.
	\item \textbf{Americas}: \textit{A. albimanus}, \textit{A. albitarsis}, \textit{A. darlingi}, \textit{A. pseudopunctipennis} and \textit{A. quadrimaculatus} \citep{sinka2012global}; \textit{A. bellator}  \citep{forattini1999role,lorenz2012morphometrical}; \textit{A. cruzii} \citep{lorenz2012morphometrical}; and \textit{A. vestitipennis} \citep{sinka2010dominant}.
	\item \textbf{Europe and the Middle-East}: \textit{A. sergentii} \citep{sinka2012global}; and \textit{A. maculipennis} \citep{hackett1935varieties}.
	\item \textbf{Asia}: \textit{A. farauti s.l.}, \textit{A. koliensis}, \textit{A. lesteri}, \textit{A. maculatus}, \textit{A. minimus}, \textit{A. punctulatus}, \textit{A. stephensi} and \textit{A. subpictus s.l.} \citep{sinka2012global}; and \textit{A. culicifacies} \citep{green1980chromosomal}.
\end{itemize}

For dengue fever, chikungunya and Zika, the main vector species are \textit{Ae. aegypti} and \textit{Ae. albopictus} \citep{kraemer2015global,grard2014zika,benelli2016declining}.

\section{Studies included in the MRR meta-analysis}\label{sec:appendix_mrrStudyList}
The following is the subset of studies from the original \cite{guerra2014global} database which were used in the MRR meta-analysis of Chapter \ref{chap:mrr}: \cite{marini2010study,baber2010population,lacroix2009dispersal,maciel2008calculating,midega2007estimating,maciel2007daily,elizondo2006gonotrophic,ba2005aspects,fabian2005mark,la2004anopheles,watson2000aedes,harrington2001analysis,tsuda2001movement,muir1998aedes,toure1998mark,quinones1997anopheles,costantini1996density,trpis1995estimates,jensen1994comparison,fernandez1994gonotrophic,jaal1992mark,rodriguez1992gonotrophic,chiang1991capture,jensen1991assessment,eldridge1990daily,macdonald1968mark,pumpuni1989population,charlwood1987mark,charlwood1989capture,birley1989effect,arredondo1998gonotrophic,hii1990estimation,renshaw1994host,milby1989estimation,maciel2007body,loong1990survival,nelson1980dispersal,maciel2006movement,mcdonald1977population,curtis1980preliminary,rawlings1982dispersal,conway1974population,reisen1979anopheles,nelson1978estimates,rawlings1981influence,sempala1981ecology,takagi1995movement,buei1980field,eyles1943experiment,ordonez2001use,pant1973field,charlwood1988evidence,reisen1982anopheles,nayar1980quantitative,carnevale1979etude,eyles1946long,reisen1984impact,charlwood1986capture,trpis1975demonstration,lutwama1994mark,wada1969dispersal,takken1998dispersal,abdel1966study,valerio2012dispersal,zetek1915behavior,takagi1995movement,yasuno1975migration,eyles1943measurement,germain1974evaluation}.

\section{Studies included in the dissection study meta-analysis}\label{sec:appendix_dissectionStudyList}
The following is a list of studies included in the dissection study meta-analysis of Chapter \ref{chap:dissection}: \cite{catangui1971studies,charlwood1979studies,chang1991comparative,charlwood2000dry,edalat2015vectorial,de2011survivorship,lines1991human,magesa1991trial,lines1991monitoring,forattini1996studies,samarawickrema1967study,samarawickrema1968biting,de2007parity,charlwood1985assessing,chandra1996age,hoc1995ovariole,vythilingam1997abundance,russell1986culex,chadee2010oviposition,ebsary1977biting,charlwood1980observations,shriram2011population,mahmood1981duration,samarawickrema1987seasonal,uttah2013physiological,ramaiah1992non,buei1982age,chadee1995seasonal,chandra2008age,foll1966conditions,kanda1975epidemiological,mala2014gonotrophic,shalaby1962studies,world1960preliminary,reisen1980anopheles,detinova1962age,smith1994age,jayanetti1987study,ch2013comparative,pant1962distribution,ghosh2010seasonal,mahanta1999temporal,schlein2012diurnal,penilla2002pteridine,schlein2015decrease,snow1978host,detinova1968age,nathan1981bancroftian,charlwood1985blood,schlein2008approach,beier2012attractive,muller2005plant,surendran2006anopheles,muller2010successful,mendis1998cordon,wilkes1996anopheles,chatterjee2000role,muller2010successful,tuchinda1969diurnal,chan1971life,aigbodion2011pelagia,qualls2015indoor,reisen1981anopheles,gillies1972range,gillies1965study,kay1979age,hitchcock1968age}.

\section{Studies included in the gonotrophic cycle duration meta-analysis}\label{sec:appendix_gonotrophicDurationStudyList}
The following is a list of studies included in the gonotrophic study meta-analysis of Chapter \ref{chap:dissection}: \cite{gillies1965study,sheppard1969dynamics,de1967time,pant1973field,germain1974evaluation,lowe1973reproductive,fernandez1994gonotrophic,buei1980field,rawlings1982tests,mori1977gonotrophic,sempala1981ecology,reisen1983population,birley1989effect,suzuki1978preliminary,charlwood1987mark,charlwood1986capture,charlwood1979studies,charlwood1995rise,chang1991comparative,edalat2015vectorial,de2011survivorship,samarawickrema1968biting,samarawickrema1967study,charlwood1985assessing,chandra1996age,russell1986culex,chadee2010oviposition,mahmood1981duration,ahumada2004modeling,kenawy1991development,rajagopalan1980population,chandra2008age,scholl1979aedes,mala2014gonotrophic,afrane2005effects,gillies1953duration,quinones1997anopheles,rua2005laboratory,delatte2009influence,arredondo1998gonotrophic,wong2014sampling,ijumba2002malaria}.

 \bibliographystyle{authordate1}
 \bibliography{Malaria}
\newpage

\end{document}

### 2018-11-23 MS.tex
Added tex version

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\PassOptionsToPackage{hyphens}{url} % url is loaded by hyperref
\usepackage[unicode=true]{hyperref}
\hypersetup{
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{longtable,booktabs}
% Fix footnotes in tables (requires footnote package)
\IfFileExists{footnote.sty}{\usepackage{footnote}\makesavenoteenv{long table}}{}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\date{}

\begin{document}

A Meta-analysis of Longevity Estimates of Mosquito Vectors of Disease

Ben Lambert\textsuperscript{1,2}, Ace North\textsuperscript{1} \& H.
Charles J. Godfray\textsuperscript{1}

\textsuperscript{1} Department of Zoology, University of Oxford, South
Parks Road, Oxford OX1 3PS, United Kingdom

Corresponding author:

Phone: 01865 271176

\textsuperscript{2} Present address: MRC Centre for Outbreak Analysis
and Modelling, Infectious Disease Epidemiology, Imperial College London,
London W2 1PG, UK.

\section{Abstract}\label{abstract}

Mosquitoes are responsible for more human deaths than any other
organism, yet we still know relatively little about their ecology.
Mosquito lifespan is a key determinant of transmission strength for the
diseases they vector, but the field experiments used to determine this
quantity -- mark-release-recapture (MRR) studies and wild-caught
dissection of female mosquitoes -- produce estimates with high
uncertainty. In this paper, we use Bayesian hierarchical models to
analyse a previously-published database of 232 MRR experiments and
compile then analyse a database of 131 dissection studies to produce the
first ever species- and genus-level estimates of mosquito lifespan. Due
to the assumptions required to analyse the field data, we term our
estimates lower bounds on lifespan (LBL). Notably, for the major African
malaria vector \emph{Anopheles gambiae s.l.}, we estimate LBLs of 4.4
days (median estimate; 25\%-75\% CI: 3.8-5.1 days for unfed female
mosquitoes from the MRR analysis) and 7.6 days (median estimate;
25\%-75\% CI: 5.2-11.0 days from the dissection analysis); and an LBL of
4.2 days (median estimate; 25\%-75\%
CI: 3.6-4.8 days, only present in the MRR database) for the
predominantly East-African vector \emph{A. funestus s.l}. We estimate
LBLs of 6.2 days (median estimate; 25\%-75\% CI: 4.5-8.5 days from the
MRR analysis) and 4.3 days (median estimate; 25\%-75\% CI: 3.5-5.1 days
from the dissection analysis) for \emph{Aedes aegypti}; and 11.6 days
(median estimate; 25\%-75\% CI: 10.0-13.7 days from the MRR analysis)
for \emph{Ae.} \emph{albopictus} -- the predominant vectors of dengue
fever, chikungunya and Zika. Our estimates indicate that there is
significant variation in lifespan across species, with most variation
explained by differences between genera. In correspondence with
laboratory studies, we estimate that female mosquitoes outlive males by
0.9 days on average (median estimate; 25\%-75\% CI: 0.3-1.6 days). We
fit models incorporating mosquito senescence to the data, which allows
us to assess evidence for age-dependent mortality in mosquitoes across
different species. We determine that 8 of 33 species included in the MRR
database indicated evidence for senescence, versus only 2 of 25 species
from the dissection database. Our analysis applies a common framework to
the analysis of databases of MRR and dissection-based experiments,
allowing us to produce robust estimates of lower bounds on lifespan. It
also enables us to critically appraise each field method, which
highlights a need for alternative field methods for measuring this
important mosquito characteristic.

\section{Keywords}\label{keywords}

mosquitoes, mortality, senescence, vector-borne disease, Bayesian,
hierarchical model.

\section{Introduction}\label{introduction}

Some of the most important infectious diseases afflicting humans are
transmitted by mosquitoes, including pathogens such as the causative
agent of malaria that have been associated with humans throughout our
evolutionary history as well recently emergent infections such as the
Zika virus. Most mosquito species have a ``gonotrophic cycle'' involving
successive episodes of vertebrate blood feeding, egg maturation and
oviposition. In order for a mosquito to transmit a pathogen it must feed
on an infectious person and live long enough to complete at least one
gonotrophic cycle and feed on an uninfected and susceptible individual.
Adult lifespan is thus a critical determinant of the ability of a
mosquito population to allow the persistence of an indirectly
transmitted infection. Lifespan can of course be straightforwardly
assessed in the laboratory, but it is generally accepted that
measurements under relatively benign laboratory conditions are likely to
have limited relevance in the field, and much effort has been directed
at estimating this parameter in the vector's natural environment. Most
work has focused on assessing average daily mortality rates, and the
simplest assumption is that these do not vary with mosquito age -- in
this case longevity is simply the reciprocal of mortality. Testing this
assumption and discovering whether mosquitoes senesce or show other
types of age-dependent mortality has also been studied in the field.

There are two main strategies to estimate mosquito mortality rates and
longevity. The first is through mark-release-recapture (MRR)
experiments, a technique that is widely applied to estimate these
parameters in many types of animal. As applied to mosquitoes, insects
are caught in the field or reared in the laboratory and then marked,
typically with fluorescent dust. The mosquitoes are then released into
the field and then efforts are made to recapture them, for example using
human baits or light traps, usually over an extended period of time.
Mortality rates can be statistically estimated from the numbers of
recaptures given certain assumptions. The main challenges with MRR is
ensuring the marking technique does not affect recapture probability,
and distinguishing mortality from mosquitoes dispersing out of range of
being recaptured. Also, releasing insects that can transmit disease
(especially if this increases ambient population levels) raises
important ethical issues.

The second technique is specific to mosquitoes and makes use of their
gonotrophic cycle. The mosquito ovary is made up of ovarioles, each of
which typically produces one egg every gonotrophic cycle. After the egg
passes into the oviduct the distended ovariole does not completely
recover its previous form but a discrete dilation remains which can be
detected by dissecting the female reproductive organs. Data on the
fraction of females that have oviposited provides some information about
mortality rates. However, a skilled dissector can distinguish the number
of dilations from multiple gonotrophic cycles so providing much richer
data on longevity. The challenges of this method include the amount of
time and expertise it takes to collect the data, establishing the
relationships between physiological and chronological time (though the
distribution of the number of gonotrophic cycles wild-caught mosquitoes
have gone through is of direct epidemiological relevance) and the fact
that it only applies to females.

An issue with both methods is that they require logistically difficult
and expensive field campaigns. There is thus value in conducting a
meta-analysis of existing data to explore consistency across studies,
identify correlates of lifespan and to learn lessons for further
studies. Here we analyse data from 232 MRR and 131 dissection studies
using a common statistical methodology. For MRR we make use of a very
valuable database of 394 mosquito studies assembled by Guerra et al.
(2014) while the dissection studies we extracted from the literature
ourselves. We concentrated on the three major genera of mosquito
vectors, \emph{Anopheles, Aedes} (in its traditional sense) and
\emph{Culex}, which constitute the majority of the data.

\section{Methods}\label{methods}

In recent years many important vectors of disease have been shown to be
complexes of very closely related species, biotypes or forms that cannot
be distinguished morphologically (for example the morphospecies
\emph{Anopheles gambiae sensu lato} is now separated into the widespread
\emph{gambiae, coluzzii, arabiensis} and a number of more local
species). As the majority of studies analysed here took place before
molecular techniques allowed these taxa to be separated we work here
chiefly with morphospecies.

\subsection{Mark-release-recapture}\label{mark-release-recapture}

Data from MRR experiments in the Guerra et al. (2014) database were
examined and those with fewer than six recaptures and species with only
a single MRR study were excluded for the hierarchical analysis. Of the
232 data sets, 177 involved only females, 35 males, and 18 both sex
releases. For 102 data sets the age of the released mosquitoes was known
(the average age of released mosquitoes was 4.0 days) while in the other
cases it was unknown or unrecorded; in these cases we assumed the
mosquitoes were newly emerged at the time of release and return to this
assumption later.

We analysed all MRR experiments within the same statistical framework
(for full details see the Supplementary Online Material). In the
simplest case \emph{N\textsubscript{R}} mosquitoes are released at day
zero and the probability that they remain in the recapture area until
day \emph{t} is \emph{S}(\emph{t}) when they are recaptured with
probability $\psi$. We model the number of mosquitoes recaptured on
day \emph{t} using a negative binomial sampling model with mean
\(\left( N_{R} - Y\left( t - 1 \right) \right)S\left( t \right)\psi\),
where \(Y\left( t - 1 \right)\) is cumulative captures before day
\emph{t}, and shape parameter $\kappa$. The negative binomial has been
used previously in MRR analyses (ref) because of its ability to
represent temporal over-dispersion in recaptures most likely caused by
variable weather. A slight modification was required for studies with
multiple releases (see SOM).

The simplest model for $S(t)$ assumes there is a constant
probability ($\lambda$) that a mosquito dies or leaves the recapture area
so that the numbers remaining after time \emph{t} are given by the
exponential distribution, \(exp\left\lbrack - \lambda t \right\rbrack\).
We utilised this form extensively but in testing for senescence used
five other models where $\lambda(t)$ varies with time so that

\begin{longtable}[]{@{}ll@{}}
\(S\left( t \right) = e^{- \int_{0}^{t}{\lambda\left( t \right) \mathrm{d}\tau}}.\)
\end{longtable}

Details of the five models (Gompertz, Weibull, Gompertz-Makeham,
Logistic and Logistic-Makeham), which vary in their ability to detect
different forms of age-dependent mortality, are given in the SOM. Using
multiple different types of models increased our chances of detecting
senescence though, as discussed below, also increases the likelihood of
false positives.

Parameters were estimated using Bayesian techniques with relative
uninformative priors for $\kappa$ and the parameters of
$\lambda(t)$, but assuming a prior for $\psi$ indicating a low
recapture probability (bounded in part by knowledge of the maximum daily
recapture rates). We used a Bayesian hierarchical model to estimate
distributions of lifespan at the species and the genus levels, and
across the complete data set. This procedure assumes that there is a
distribution of lifespan parameters for each species from which those
governing individual MRR time series are sampled, and similarly a
distribution at the genus level from which those for individual species
are derived (rather akin to random effects in classical statistics).
Within this framework we can also allow the parameters for individual
time series to be influenced by co-variates such as differences in
experimental methodology. As in the estimation of the parameters of the
individual experiments, relative uninformative priors were set for the
parameters of the hierarchical models except for $\psi$ where again a
distribution representing low recapture probabilities was assumed.
Posterior distributions were derived using Markov Chain Monte Carlo
(MCMC) methods with convergence assessed using the \(\hat{R}\) statistic
(Gelmin \& Rubin 1992). The predictive power of the model was assessed
using \emph{K}-fold cross validation which tests the ability of the
model fitted to part of the data to predict the rest using multiple
different partitions. Further details of the prior specification,
fitting and validation through posterior predictive checks are given in
the SOM.

Two studies of \emph{Anopheles balabacensis} reported capture rates
increasing with time, presumably reflecting a violation of our
assumption of constant recapture probabilities. We omitted this species
from the analysis.

The Guerra et al. database included the latitude and longitude of each
study, along with the date when the study began. We used this information to find
estimates of the air temperature for each study using the European Centre for Medium
Range Weather Forecasts' ERA Interim Daily historical database. For each
study we calculated the mean monthly temperature across a spatial area
of (latitude $\pm$ 1 degree, longitude $\pm$ 1 degree), for the month at which
each study was carried out. The records for this database begin in 1979,
which pre-dates the study date for 65 of our 232 MRR time-series. For
these time-series, we chose to estimate the air temperature by an
average of the corresponding monthly temperatures over the years
1979-89.

\subsection{Dissection}\label{dissection}

Studies using dissection to estimate mosquito longevity were located in
literature databases using relevant keyword, citation and author
searches, and by checking previous studies cited by the papers located
(see SOM). The list of studies located with associated metadata is
available as a Supplementary Online File.

Most dissection studies recorded the distribution of the number of
gonotrophic cycles in mosquito samples collected over a specific period
of time. Overall, we found 568 time series in 72 published articles.
Because seldom were sufficient ancillary data available data to analyse
trends over time, we aggregated the data from each time series. We
further omitted time series with fewer than 100 mosquitoes and for
species with only one data set leaving 131 studies of mosquitoes in the
genera \emph{Anopheles, Aedes}, \emph{Culex} and \emph{Mansonia}.

To compare lifespan estimates from dissection and MRR studies we need to
convert physiological age (the number of gonotrophic cycles) into
chronological age. Using a literature search and a review by Silver
(2007) we found 79 estimates in 42 published articles. Most estimates
were obtained by dissecting females recaptured in MRR studies or by
observations in the laboratory, the latter tending to give longer
durations. Studies differed greatly in how (if at all) they represented
uncertainty in their estimate of the duration of the gonotrophic cycle.
Where confidence limits were given we treated these as the relevant
quantiles of a normal distribution, where a range was stated (e.g. ``4-6
days'') we interpreted the bounds as the 2.5\% and 97.5\% quantiles of a
normal distribution, and where a single figure was quoted we assumed
this was the mean this distribution. Using the quantiles of the normal
distribution, we estimated its mean and standard deviation by regression
(see SOM). Initially we calculated distributions of gonotrophic cycle
lengths at the species and then genus levels, but because of the paucity
of data for many species and the lack of significant differences we
aggregated the data into a single distribution. We converted
physiological age to chronological age by sampling from this
distribution to obtain a particular gonotrophic cycle length for each
mosquito (we also explored sampling from this distribution to obtain the
duration of \emph{each} gonotrophic cycle which increased the
uncertainty in lifespan estimate but did not affect any of the
conclusions).

We modelled the number of mosquitoes found by dissection to be of age
\emph{a} using the negative binomial distribution with mean
$\Psi S(a)$ and shape parameter $\kappa$, where
$\Psi$ is the product of the recruitment rate of adult mosquitoes,
which we assume is constant over time, and the probability of being
captured for dissection, and \emph{S}(\emph{a}) is the probability of
surviving until age \emph{a}. We used the number of females that have
yet to lay eggs (nulliparous) to estimate the recruitment rate as
described further in the SOM. Initial examination revealed that in some
data sets the number of nulliparous females was anomalously low,
something that has been noticed before (Gillies \& Wilkes 1965). As some
studies have suggested that the first gonotrophic cycle tends to be
longer than the subsequent ones, this is probably due to differences in
capture probability. In data sets where the fraction of nulliparous
females was less than 90\% the uniparous (completed on gonotrophic
cycle) we excluded the nulliparous observation.

Data was analysed using a Bayesian framework as with the MRR data with
minor differences in the specification of the priors (see SOM). Some
published studies do not distinguish the number of gonotrophic studies
beyond a threshold (the more ovariole dilations there are the harder it
is to count them) which is akin to censoring the data. Because this
censoring involves only a relatively small number of mosquitoes in each
time series (median = 2\%), and because of the technical difficulties of
allowing for censoring in our Bayesian estimation procedures, we assume
the females died at the threshold age.

\section{Results}\label{results}

\subsection{Lifespan estimates from
MRR}\label{lifespan-estimates-from-mrr}

MRR estimates the length of time a mosquito remains alive and is still
in the area available for recapture. It thus provides a lower bound to
lifespan which we shall refer to as LBL. In 187 of the 230 MRR time
series the estimated LBL was less than 10 days (Fig. 1). The smallest
estimate was \textless{}1 day for the Asian malaria vector
\emph{Anopheles subpictus s.l.} which is unfeasibly short and almost
certainly reflects dispersal out of the recapture zone or a violation of
the assumptions of our analyses. The longest estimate was 18.3 days for
the temperate species \emph{Aedes simpsoni s.l.} which is a vector of
yellow fever in Africa. There are multiple data sets for the most
important vector species such as \emph{Anopheles gambiae, Aedes aegypti}
and \emph{albopictus} and \emph{Culex tarsalis} all of which show
considerable variation. For example, there are 54 estimates of LBL for
\emph{A. aegypti} which range from 2.2 days to 38.3 days with a mean of
8.3 days and coefficient of variation of 0.7. There are significant
differences in LBL amongst species (ANOVA on median LBL controlling for
sex and pre-release feeding: \emph{F}\textsubscript{37,194} = 2.5,
\emph{p} \textless{}0.01; the non-parametric Kruskal Wallace:
\({\chi^{2}}_{38}\), \emph{p}\textless{}0.01).

The estimated median LBL for female mosquitoes not pre-fed with blood or
sugar before release for \emph{Culex, Anopheles} and \emph{Aedes} were
2.5, 5.0 and 6.9 days respectively with an overall estimate of 4.6 days
(Fig. 2). Differences between genera were significant (ANOVA on median
LBL controlling for sex and pre-release feeding:
\emph{F}\textsubscript{2,229} = 12.4, \emph{p}\textless{}0.01; Kruskal
Wallace: \({\chi^{2}}_{2} = 30.8,p < 0.01\)). \emph{K}-fold cross
validation suggests that after the effect of genus is accounted for the
incorporation of a species term provides little predictive power (Fig.
S\#; in part explained by the latter model over-fitting the data where
there are few time series per species).

We reasoned that if dispersal out of the recapture area was reducing the
LBL below the true lifespan then there should be a positive correlation
between the spatial extent of the recapture zone and LBL. We found no
such pattern (Fig. S\#), although there was a positive correlation
between LBL and trap density (Fig. S\#).

The MRR experiments included a mixture of male-only and female-only
releases, and releases of both sexes. We estimated average male and
female LBL at the genus level (Fig. 3; there were too few studies to
make comparisons at the species level). There was a consistent trend for
females to live longer than males for each of the genera, with the
difference largest for \emph{Aedes} (2.5 days; fraction of pairwise
posterior samples of females versus males where difference was less than
zero, p\textless{}0.01), followed by \emph{Anopheles} (2.0 days; p=0.17)
and \emph{Culex} (0.3 days; p=0.34). Overall, female mosquitoes were
estimated to live 0.9 days longer than males (p=0.10).

The MRR experiments included information on whether mosquitoes were
pre-fed with sugar (41 time series), blood (71), both (4) or
alternatively unfed (116). We estimate that female mosquitoes that were
fed on sugar pre-release lived on average for 0.7 days longer than those
that were not fed (Fig. S\#; a pattern that was consistent across the
genera). There were insufficient males that were either fed or unfed
with sugar prior to release to make a meaningful comparison. Females
that were blood-fed prior to release on average lived 1.5 days longer
than those who were not fed for \emph{Aedes} but this trend was reversed
for \emph{Anopheles} meaning that there was little difference overall.

To access whether temperature is associated with LBL we used weather
records to calculate average temperatures at the MRR sites (see
Methods). Using both linear and quadratic temperature terms in
regressions, we found no significant relationship between study-site
temperature and LBL (overall or within genus) for the 238 datasets we
analysed (Fig. 4). This result held if, instead of pooling results from
all time series, we considered the four species with the most data
(\emph{Ae. aegypti}, \emph{Cx. tarsalis}, \emph{A. gambiae s.l.} and
\emph{A. culicifacies s.l.}) individually (Fig. S\#).

\subsection{Number of gonotrophic cycles estimates from
dissection}\label{number-of-gonotrophic-cycles-estimates-from-dissection}

Dissection allows the number of completed gonotrophic cycles to be
counted and from this the mean number of cycles before death was
estimated. Overall, the mean number of cycles completed in a lifetime
was 1.2 (Fig. 5; posterior median) and across the 131 studies, 95\% of
the individual time series estimates were less than 3 (Fig. 6). The
estimated greatest number of cycles was for \emph{Anopheles sergentii}
(2.5 cycles; posterior median) which is adapted to desert conditions (it
is known as the ``oasis vector'' of malaria) and may have evolved
greater longevity. The important African malaria vector \emph{An.
gambiae s.l.} was estimated to be the second longest living (1.9 cycles;
posterior median). The smallest estimated mean number of gonotrophic
cycles was for \emph{Anopheles} \emph{bellator} (0.5 cycles; posterior
median) which transmits malaria in Brazil's Atlantic Forest. There were
significant differences in estimated lifetime gonotrophic cycles amongst
species (ANOVA: \emph{F}\textsubscript{24,106} =2.2, \emph{p}
\textless{}0.01; the non-parametric Kruskal Wallace:
\({\chi^{2}}_{24}\), \emph{p}\textless{}0.01).

The estimated lifetime gonotrophic cycles for the different genera were
\emph{Anopheles,} 1.4; \emph{Mansonia}, 1.1; \emph{Culex,} 1.0; and
\emph{Aedes} 0.8 (Fig. 5) and the differences between the genera were
significant (ANOVA: \emph{F}\textsubscript{3,127} =3.4, \emph{p} =0.02;
the non-parametric Kruskal Wallace: \(\chi_{3}^{2} = 21.7\),
\emph{p}\textless{}0.01).

\subsection{Comparison of longevity estimates from two
methods}\label{comparison-of-longevity-estimates-from-two-methods}

To compare the two methods, we converted numbers of gonotrophic cycle
(physiological age) into lifespan (chronological age) as described in
the Methods (Fig. S\#). For ten species, we had enough data from both
species to make a comparison, and there was a positive correlation (not
statistically significant; Pearson correlation $\rho=0.42$, $n=10$, $p=0.23$)
between the two measures (Fig. 7), and in only one case -- for \emph{A.
darlingi} - (Table S\#) there was a significant difference in the
time-series level LBLs.

\subsection{Evidence for age-dependent
mortality}\label{evidence-for-age-dependent-mortality}

The survival model upon which the above analyses are based is the
single-parameter exponential model which assumes an age-invariant
mortality hazard. We also fitted five multi-parameter models that allow,
in different ways, mortality to vary with age. We did this to maximise
our chance of detecting age-varying mortality (though aware of the risks
of false positives with multiple estimations).

In Fig. 8 we compare the performance of the six models for describing
lifespan in MRR studies of 33 species using K-fold cross-validation. We
categorised the evidence for age-dependent mortality in each species
according to the performance of the five age-dependent models versus the
exponential: `+' indicated that all age-dependent models outperformed
the exponential; `?' indicated that the exponential outperformed one or
more age-dependent models; and `-' indicated that the exponential
performed at least as well as all other models. Overall, we estimated
that there were 8 `+' species, where age-dependent mortality fit the
data better; 11 `?' species where the evidence was mixed; and 14 species
where constant mortality models performed at least as well. The species
where age-dependent mortality best fit the data included the major
vector of dengue fever, Zika and chikungunya, \emph{Ae. Aegypti}. These
studies also tended to include multiple release MRR studies which, on
average, were conducted over a longer period of time than the others,
which may be why we failed to detect age-dependence in the latter (Fig
S\#).

In Fig. 9, we compare the performance of the six models for describing
lifespan in dissection studies of 25 species using K-fold
cross-validation, and categorise the evidence in the same way as for the
MRR analysis. By our metric, we determined that there were only two
species with evidence for age-dependent mortality (the major African
malaria vector \emph{A. gambiae s.l.} and \emph{A. minimus}, a malaria
vector in Asia).

Overall, we conclude that there is mixed evidence for age-dependent
mortality from studies of mosquitoes in the field. It is possible that
some of the sampled mosquito species did not live long enough in the
wild to experience physiological decline. A Spearman's rank correlation
test indicated that there was a correlation between the ranked estimated
LBLs of the species and the ranked mean predictive accuracy of
age-dependent models for the MRR analysis ($\rho$=0.19, p=0.01), however was
not significant for the dissection analysis ($\rho$=0.07, p=0.43). Similarly,
a recent study determined that the degree of senescence varies according
to season for semi-wild populations of \emph{Ae. aegypti} (Hugo et al.,
2014), and it is possible that by pooling data from different
geographies and seasons that we failed to detect age-dependent mortality
in some cases.

\subsection{Estimates of the fraction mosquitoes capable of transmitting
disease}\label{estimates-of-the-fraction-mosquitoes-capable-of-transmitting-disease}

We can use the posterior parameter estimates from our Bayesian analysis
to estimate the fraction of mosquitoes that live beyond a certain age.
In order to transmit a disease, a mosquito must live longer than the
length of the intrinsic incubation period (the time taken for a pathogen
ingested in one blood meal to be ready to be transmitted during a future
feeding event). This is a lower bound as it does not include the waiting
time to find a host after feeding or egg maturation. In Fig. 10 we plot
the fraction of the mosquito population that pass this threshold using
estimates from both MRR and dissection studies for vector species (see
SOM for references used to identify species as vectors) and their most
significant diseases.

For malaria, estimates of the minimum fraction of the population that
can transmit the disease vary from \textless{}0.1\% for \emph{A.
subpictus} (from the MRR analysis, as noted above likely to be due to
the LBL substantially underestimating lifespan) to 52\% for the
drought-adapted and long-lived \emph{A sergentii.} The proportions
surviving long enough to become infectious for the two major African
malaria vectors were: for \emph{A. gambiae s.l.}: 10\% (MRR) and 27\%
(dissection); and for \emph{A. funestus s.l.}: 9\% (MRR). Using the
individual time series estimates, however, there was inconclusive
evidence for a difference in EIP between the species (MRR ANOVA:
\emph{F}\textsubscript{14,58} = 1.8, \emph{p}=0.07; MRR Kruskal-Wallis:
\(\chi_{14}^{2} = 30.2\), \emph{p} \textless{}0.01; dissection ANOVA:
\emph{F}\textsubscript{11,64} = 1.8, \emph{p}\textless{}0.01; dissection
Kruskal-Wallis: \(\chi_{11}^{2} = 38.9\), p\textless{}0.01).

\emph{Ae. aegypti} and \emph{Ae. albopictus} are the main vectors of
dengue, chikungunya and Zika viruses. Because of their short intrinsic
incubation periods a greater fraction of mosquito potentially live long
enough to transmit diseases (Fig. 10), rising to a maximum of 84\% for
\emph{Ae. albopictus} transmitting chikungunya.

\section{Discussion}\label{discussion}

In this study, we applied a Bayesian hierarchical framework to the
analysis of a database of mark-release-recapture experiments and a
database of mosquito dissection studies to estimate mosquito lifespan.
By applying a single framework, this allows us to effectively synthesise
information from the disparate experiments which, individually, estimate
lifespan with considerable uncertainty. Due to the assumptions required
to analyse the field data, our estimates represent lower bounds on
lifespan (LBL). Across both meta-analyses, the estimated LBLs were
mostly less than 10 days, hinting that only a small proportion of
mosquitoes may live long enough to transmit disease. We determined that
LBL varies across species and genera, although most variance is
explained by genus. The MRR analysis includes experiments conducted on
each sex individually, and we estimate that, on average, males live
shorter lives than females. Pre-release feeding with sugar also
lengthens lifespan across all three genera, although this effect is less
marked than the sex differences. In contrast to a number of lab-based
experiments, temperature was not determined to significantly impact
lifespan. By fitting a range of survival models to the data in both
meta-analyses, we could assess evidence for age-dependent mortality.
Overall, we conclude that the evidence is mixed: in the MRR experiments,
in 8 of 33 species we found evidence for mosquito senescence, whereas in
only 2 of 25 species included in the dissection analysis were better fit
by a model incorporating an increasing risk of mortality with age.

MRR experiments are known to produce downwardly-biased estimates of
lifespan. Lab experiments have demonstrated that marking can negatively
impact survival (Verhurst et al., 2013; Dickens and Brant, 2014),
resulting in artificially depressed survival. MRR studies typically
cannot differentiate between a mosquito dying and dispersal from the
study area meaning that lifespan will be underestimated. In this study,
we found a positive correlation between lifespan estimates and the
density of traps, indicating that better trapping coverage likely raises
estimates towards their real value. We conducted an \emph{in silico}
Monte Carlo study to determine how accurately we could estimate mosquito
lifespan given study parameters in an ideal MRR experiment, where the
assumptions of no emigration and harmless marking are fully satisfied
(see SOM for full details). This work indicated that for many of the
experiments, the short study lengths or typical numbers of mosquitoes
released, results in considerable uncertainty in lifespan estimates (Fig
S\#). This indicates that statistical power can be substantially
increased by pooling data across experiments as we did using a Bayesian
hierarchical model.

The key assumptions of dissection based methods to determine
chronological age are: (i) physiological age can be accurately
determined by dissection of female specimens (unlike MRR, this method
can only be applied to one sex), (ii) the relationship between
physiological and chronological age is known, (iii) the population being
sampled is in equilibrium (recruitment matches mortality) and (iv)
individual mosquitoes can be randomly sampled from the population. The
reliability and accuracy of dissection has been questioned. The
objections include the impracticality of dissecting more than a small
proportion of ovarioles (Hoc and Wilkes, 1995), particularly in African
vector species (Gillies and Wilkes, 1965), the related issue of locating
ovarioles whose count of dilations represents true physiological age
(Fox and Brust, 1994), and the variation in numbers of ovariolar
dilations for mosquitoes of the same, known, physiological age (Kay,
1979; Russell, 1986; Hugo et al., 2008). Indeed there is considerable
uncertainty concerning the fundamental question of how dilations in
ovarioles form in the first place. Whilst the `Old School' of thought (a
term coined by Fox and Brust, 1994) headed by Polovodana (Polovodova,
1949) and Detinova (Detinova, 1962) considers dilations to result from
normal oogenesis, a `New School' headed by Lange and Hoc (Lange and Hoc,
1981) has challenged this assertion. The New School believe that only
abortive oogenesis results in follicular dilations because normal
oogenesis destroys the sack-like structures (Fox and Brust, 1994). This
means that Polovodana's method requires dissecting large numbers of
ovarioles to uncover those with the most dilations, where abortive
oogenesis has occurred in each gonotrophic cycle. They deem these
ovarioles `diagnostic' since only in these cases the number of dilations
equals the number of gonotrophic cycles that have occurred. As a
mosquito ages, the number of diagnostic ovarioles diminishes, since the
random occurrence of normal oogenesis in a particular ovariole means its
dilation count does not equal the number of gonotrophic cycles
undertaken. This increased difficulty of finding diagnostic ovarioles as
a mosquito ages would elevate the chance of age `hypodiagnosis' for
older specimens (Fox and Brust, 1994), and likely biases lifespan
estimates downwards. The difficulty of locating diagnostic ovarioles has
been investigated using lab populations of \emph{Culex} and \emph{Aedes}
mosquitoes by Hugo et al., 2008, who conclude that only a small
percentage of ovarioles are diagnostic. The exchange rate between
physiological age and chronological age is the duration of gonotrophic
cycles. Two methods are commonly used to estimate the duration of
gonotrophic cycles: MRR studies (see, for example, Gillies and Wilkes,
1965), where marked mosquitoes are recaptured and dissected to determine
the number of gonotrophic cycles occurring since release; and
laboratory-based observations of colonies of (typically) wild-caught
females, or their progeny (see, for example, Afrane et al., 2005).
Whilst it is unclear how each method could bias estimated gonotrophic
cycle duration, in our analysis, laboratory-based studies indicated a
longer gonotrophic cycle (Fig. S\#). The distributions we used to
convert physiological age into calendar age were calculated by pooling
data across both approaches, to incorporate uncertainty from both
experimental procedures. It is possible, however, that this aggregate
approach may induce biases in estimates and an approach more entrenched
in experimental knowledge would fare better. If a population of
mosquitoes is shrinking, this leads to a relative under-abundance of
young mosquitoes, and a flattening of the survival curve, resulting in
over-estimates of lifespan. For stable populations, periods when
shrinking occurs must result in equal changes in the population size
compared to those when it expands. If mosquito collections occur with
equal frequency in each of these two modes, then aggregating the data
across all sampling times and estimating a single model, as we do here,
should yield an approximately unbiased estimate of lifespan. The
additional uncertainty of a fluctuating population size, however, could
lead us to understate the uncertainty in estimates. Field entomologists
have challenged the assumption of random sampling the mosquito
population, although there are conflicting opinions as to whether this
results in a relative paucity (Gillies and Wilkes, 1965) or abundance
(Clements and Paterson, 1981) of nulliparous individuals. In our
database, there are cases where there was an obvious deficit of
nulliparous individuals, which has previously been ascribed to the
differing distribution of resting females between indoor and outdoor
traps (Detinova, 1962; Gillies and Wilkes, 1965). We chose to not
include those counts of nulliparous individuals in our analysis where
their number was less than 90\% of the uniparous. Whilst we see no
obvious differences in lifespan according to collection method (data not
shown) or location, it is possible that the assumption of random
sampling is violated, although the directionality of the bias induced by
this is unclear. Overall, the assumptions underpinning estimates from
dissection studies indicate that our estimates represent lower bounds on
lifespan. The alternative dissection-based approach of Detinova
(Detinova, 1962), based on dichotomous categorisation of female mosquito
specimens as `parous' or `unparous' relies on fewer assumptions, and is
widely used. Further work examining parity rates in field specimens may
be fruitful although, in principle, it offers less information on the
age structure of a population than Polovodova's approach.

By applying a common method to analysing all studies in our databases,
it is possible that we may have missed patterns of mortality that would
have been evident from using a more bespoke approach. As our \emph{in
silico} analysis of MRR experiments indicates, however, the
overdispersed data from single experiments results in high measurement
error (Fig. S\#). By applying different methods to each study, this
could lead us to falsely detect patterns when none are present, and we
prefer a pooled approach.

The different nature of the assumptions of each of the two methods means
they offer complimentary information on mosquito survival. We also note
that Polovodova's dissection-based studies require specialised expertise
which will often be unavailable, whereas MRR methods can more readily be
used. Furthermore, most if not all dissection methods that have been
used previously are only applicable to female mosquitoes, whereas MRR
can be applied to either sex and can additionally be used to determine
other ecological parameters (for example, population size and
dispersal). Although dissection data gives detailed of age-structure, we
thus foresee a continued reliance on MRR experiments in field
entomological experiments. Efforts to use both approaches concurrently
will be particularly useful and will allow quantification of the biases
induced by the assumptions of each. Similarly, MRR experiments releasing
large numbers of marked mosquitoes and recording
spatiotemporally-disaggregated captures of wild and re-caught marked
mosquitoes will be especially useful to estimate lifespan and dispersal.

To compare estimates of lifespan derived from MRR with those from
dissection-based methods, we display the estimates of lifespan from
those ten species occurring in both databases in a single plot (Fig. 7).
In is reassuring that there is correlation between estimates from both
approaches, although the small sample size likely hindered our ability
to determine statistical significance. In both cases, we estimate that
\emph{A. sergentii} was amongst the longest lived of the anopheline
species with a posterior median LBL of 8.7 days (median estimate;
25\%-75\% CI: 5.9-13.8 days) from the MRR analysis and 10.0 days (median
estimate; 25\%-50\% CI: 7.6-14.0 days) from the analysis of dissection
studies. This species is a major vector of malaria in the Sahara (Sinka
et al., 2010), where to act as a disease vector it must persevere
through these hard conditions. It is reasonable to hypothesise that this
species should live longer than those in environments where the
potential for blood-feeding and oviposition is greater. The species with
the greatest discrepancy in the estimates was the major African malaria
vector \emph{A. gambiae s.l.}, where we estimated LBLs of 4.4 days
(median estimate; 25\%-75\% CI: 3.8-5.1 days for unfed female) from the
MRR analysis and 7.6 days (median estimate; 25\%-75\% CI: 5.2-11.0) from
the dissection analysis. Across genera, the greatest discrepancy in
estimates was for \emph{Culex} (a posterior median of 2.5 days from the
MRR versus 4.4 days from the dissection analysis) with the smallest
discrepancy for \emph{Anopheles} (4.8 versus 5.0 days). By contrast for
\emph{Aedes} the estimates from the MRR studies (6.9 days) exceed those
of dissection-based studies (3.5 days). Across all studies we estimate
from the MRR analysis that mean mosquito lifespan is 4.6 days versus 5.1
days from the dissection-based studies. Some of the differences in these
group-level estimates between the two approaches is likely due to
environmental and genetic differences between mosquitoes in the
experiments that were analysed in each meta-analysis. However, we
believe that part of the discrepancy can be explained by the
methodological differences in approaches. We speculate that differences
in dispersal rate can explain some of the discrepancy. Both
\emph{Anopheles} and \emph{Culex} mosquitoes are generally thought to
fly farther during their lifetimes than \emph{Aedes} {[}Charles, do you
have a reference here?{]}, meaning that the estimates from MRR-based
approaches will be most downwardly-biased for these genera. This is
supported by our results since the dissection-based estimates
(themselves not reliant on assumptions about dispersal) exceed the MRR
estimates for \emph{Anopheles} and \emph{Culex} mosquitoes, but not for
\emph{Aedes}.

It is widely believed mosquitoes live artificially long under the benign
conditions of the laboratory. We find it informative to consider
estimates of lifespan derived from observations of such populations as
they constitute an upper bound on the lifespan of wild populations.
Also, since the numbers of mosquitoes involved in large cage experiments
often numbers in the thousands, these estimates have lower uncertainty
than those from field experiments although are typically conducted on
highly inbred mosquito strains. Styer et al. (2007), using colonies of
45,054 female and 55,997 male \emph{Ae. aegypti}, determined that
females lived nearly twice as long as males; the median lifespan was
estimated as 31.69 $\pm$ 0.06 days for females and 16.39 $\pm$ 0.03 days
for males. A similar study by Dawes et al., 2009 with a lab colony of
over 1000 female \emph{A. stephensi} found similar estimates for median
lifespan (31-42 days). These estimates are many multiples of the average
estimates that result from our analysis of field data which, as
discussed, represent lower bound estimates. Without an unbiased method
to measure mosquito lifespan, however, it is difficult to quantify and
explain the gap that exists between field and laboratory lifespans. The
development of additional methods to estimate mosquito age, such as
`Near-Infrared Spectroscopy' (Mayagaya et al., 2009; Sikulu et al.,
2011; Lambert et al., 2018), if they are proven to work in the field,
may be of considerable worth here.

We conducted a power analysis of MRR experiments (Fig. S\#) to determine
whether typical experimental characteristics could detect senescence.
Here we calculated the power of a maximum likelihood estimator of the
`senescence parameter' $\beta$ of the Gompertz survival function (see Table
S\#) for case study populations with three different levels of
senescence (Fig. S\#). This analysis indicated that power to detect
senescence strongly depends on study length (Fig. S\#) but is
insensitive to release size (Fig. S\#). Clements and Patterson (1981)
conducted a meta-analysis of MRR and dissection-based field experiments
and found evidence of an increasing risk of mortality hazard with age
that is similar in magnitude to that of the `mild' case considered
above. For this case, detecting senescence with a power of 80\% requires
a study length of at least 18 days. Since the median study length for
experiments included in our analysis was 10 days (Table S\#) this could
partly explain our failure to detect senescence at the species level. A
number of experiments have found evidence of age-dependence in
laboratory populations (Styer et al., 2007; Dawes et al., 2009).
However, the artificially benign environment of the laboratory means
mosquitoes live considerably longer than in the wild, where they may die
because of exogenous factors, before the effects of physiological
decline have had time to manifest. Field experiments have also found
evidence for age-dependent mortality. Harrington et al. (2008) conducted
a field experiment where mosquitoes reared under laboratory conditions
were marked and released at different ages. Analysis of the resultant
MRR time-series indicated that mosquito mortality increases with age at
release. It is possible, however, that this field experiment suffers
from the same biases as laboratory-based approaches, because the
released mosquitoes were often of ages considerably higher (up to 20
days) than typical estimates of wild mosquito lifespan.

As ethical concerns of contributing to disease burden are more often
considered, it is now less common for MRR experiments to release female
mosquitoes versus males than historically (Fig. S\#). Our analysis
indicates that females outlive male mosquitoes by approximately 0.9 days
(Fig. \#), meaning that differences between the sexes may exist for
other ecological parameters determinable by MRR. This suggests that
continued field entomological work on contained releases of mosquitoes
in semi-field sites or large microcosms may be a valuable source of
information on female mosquitoes.

Our estimates of LBL indicate that mosquitoes that were sugar-fed prior
to release lived on average 0.7 days longer than those that were unfed
(, suggesting the potential value of this underappreciated aspect of the
mosquito ecology to the insects. It may also partly explain the recent
successes in the use of Attractive Toxic Sugar Baits (ATSB) as a vector
control intervention (Schlein and Muller references). More research is
needed, however, to identify the sugar-feeding frequency and food
sources for wild populations.

There is considerable evidence from the laboratory that temperature
modulates mosquito ecology and behaviour (Matt Thomas references). The
locations and times of year over which the MRR studies were conducted
encompassed a large range of average air temperatures, from
approximately 10 \textsuperscript{o}C to 35 \textsuperscript{o}C and,
within this, we determined no relationship between lifespan and
temperature across all time series (Fig. 4) or, for any of the species
with the most data (Fig. S\#). It is possible that by considering a raw
average of air temperature across the month, this ignored other
temperature-dependent factors of importance for longevity, such as
temperature range. It is also possible that by ignoring the effects of
rainfall (the historical data on rainfall is less likely to be reliable
for a given location), that this masked a more complex interaction
between longevity and temperature. The observed laboratory relationship
between lifespan and temperature, however, may not be as robust in the
field if mosquitoes adjust their behaviours (such as, by seeking shade)
in reaction to changes in temperature. More work exploring the
relationship between mosquito ecology and temperature in semi-field
experiments may be useful in probing these interactions further.

In this work, we have used modern statistical methods to synthesise
precious field data conducted by entomologists past and present, to
produce lower bound estimates of mosquito lifespan. The importance of
vector mortality for disease transmission has long been recognised,
however, since even before 1957, when George Macdonald formulated the
now famous Ross-Macdonald equation of R\textsubscript{0} for malaria.
Indeed, the recent declines in malaria prevalence in Sub-Saharan Africa
were likely due to the upscaling of interventions (insecticide-treated
bednets and indoor residual spraying) that aim to reduce mosquito
lifespan (Bhatt et al., 2015). Worryingly, resistance to pyrethroids,
the only class of insecticide used in current insecticide-treated
bednets and likely the only product to come to market in the near
future, has been determined to be widespread and increasing in intensity
across Sub-Saharan Africa (WHO reference 2018). This alarming trend
highlights the need for continued MRR and dissection-based studies to
monitor the effectiveness of bednets and determine whether more
expensive alternatives, such as nets incorporating piperonyl butoxide be
deployed. It also emphasises the need for investment in new tools for
real time monitoring of mosquito populations. In recent years, there has
been considerable funding allocated to molecular and genomic research
into mosquitoes that strengthens existing interventions and suggest
novel control strategies. Without commensurate funding allocated to
applied vector ecology, our lack of knowledge in this area threatens our
opportunity to capitalise on molecular advances and potentially hinders
our ability to control of mosquito-borne disease.

\section{Acknowledgements}\label{acknowledgements}

The authors would like to thank the following for useful conversations
throughout the course of this work: Austin Burt, Thomas Churcher, Steve
Lindsay, Mike Bonsall and Ellie Sherrard-Smith.

\section{References}\label{references}

ADDIN EN.REFLISTX

\section{Figure Legends}\label{figure-legends}

Figure 1:

\end{document}

### 2018-11-23 MS.tex
Updates to make cross-referencing possible

@@
 \date{}
+
+\usepackage{xr}
+\externaldocument{figures}
 
@@
 lifespan which we shall refer to as LBL. In 187 of the 230 MRR time
-series the estimated LBL was less than 10 days (Fig. 1). The smallest
+series the estimated LBL was less than 10 days (Fig. \ref{fig:mrr_lifespan_individualEstimates}). The smallest
 estimate was \textless{}1 day for the Asian malaria vector
@@
 2.5, 5.0 and 6.9 days respectively with an overall estimate of 4.6 days
-(Fig. 2). Differences between genera were significant (ANOVA on median
+(Fig. \ref{fig:mrr_lifespans}). Differences between genera were significant (ANOVA on median
 LBL controlling for sex and pre-release feeding:
@@
 releases, and releases of both sexes. We estimated average male and
-female LBL at the genus level (Fig. 3; there were too few studies to
+female LBL at the genus level (Fig. \ref{fig:mrr_sexDifferences_without_sugar_nor_blood}; there were too few studies to
 make comparisons at the species level). There was a consistent trend for
@@
 temperature and LBL (overall or within genus) for the 238 datasets we
-analysed (Fig. 4). This result held if, instead of pooling results from
+analysed (Fig. \ref{fig:mrr_temperature}). This result held if, instead of pooling results from
 all time series, we considered the four species with the most data
@@
 estimated. Overall, the mean number of cycles completed in a lifetime
-was 1.2 (Fig. 5; posterior median) and across the 131 studies, 95\% of
-the individual time series estimates were less than 3 (Fig. 6). The
+was 1.2 (Fig. \ref{fig:dissection_lifetimes_exponential}; posterior median) and across the 131 studies, 95\% of
+the individual time series estimates were less than 3 (Fig. \ref{fig:dissection_individualEstimates_allSpecies}). The
 estimated greatest number of cycles was for \emph{Anopheles sergentii}
@@
 \emph{Anopheles,} 1.4; \emph{Mansonia}, 1.1; \emph{Culex,} 1.0; and
-\emph{Aedes} 0.8 (Fig. 5) and the differences between the genera were
+\emph{Aedes} 0.8 (Fig. \ref{fig:dissection_lifetimes_exponential}) and the differences between the genera were
 significant (ANOVA: \emph{F}\textsubscript{3,127} =3.4, \emph{p} =0.02;
@@
 statistically significant; Pearson correlation $\rho=0.42$, $n=10$, $p=0.23$)
-between the two measures (Fig. 7), and in only one case -- for \emph{A.
+between the two measures (Fig. \ref{fig:comparison}), and in only one case -- for \emph{A.
 darlingi} - (Table S\#) there was a significant difference in the
@@
 
-In Fig. 8 we compare the performance of the six models for describing
+In Fig. \ref{fig:mrr_elpd}, we compare the performance of the six models for describing
 lifespan in MRR studies of 33 species using K-fold cross-validation. We
@@
 
-In Fig. 9, we compare the performance of the six models for describing
+In Fig. \ref{fig:dissection_elpd}, we compare the performance of the six models for describing
 lifespan in dissection studies of 25 species using K-fold
@@
 feeding event). This is a lower bound as it does not include the waiting
-time to find a host after feeding or egg maturation. In Fig. 10 we plot
+time to find a host after feeding or egg maturation. In Fig. \ref{fig:eip}, we plot
 the fraction of the mosquito population that pass this threshold using
@@
 incubation periods a greater fraction of mosquito potentially live long
-enough to transmit diseases (Fig. 10), rising to a maximum of 84\% for
+enough to transmit diseases (Fig. \ref{fig:eip}), rising to a maximum of 84\% for
 \emph{Ae. albopictus} transmitting chikungunya.
@@
 dissection-based methods, we display the estimates of lifespan from
-those ten species occurring in both databases in a single plot (Fig. 7).
-In is reassuring that there is correlation between estimates from both
-approaches, although the small sample size likely hindered our ability
+those ten species occurring in both databases in a single plot (Fig.
+\ref{fig:comparison}).
+In is reassuring that there is correlation between estimates from both approaches, although the small sample size likely hindered our ability
 to determine statistical significance. In both cases, we estimate that
@@
 
-In this work, we have used modern statistical methods to synthesise
-precious field data conducted by entomologists past and present, to
-produce lower bound estimates of mosquito lifespan. The importance of
-vector mortality for disease transmission has long been recognised,
-however, since even before 1957, when George Macdonald formulated the
+The importance of vector mortality for disease transmission has long been recognised,
+since even before 1957, when George Macdonald formulated the
 now famous Ross-Macdonald equation of R\textsubscript{0} for malaria.
-Indeed, the recent declines in malaria prevalence in Sub-Saharan Africa
-were likely due to the upscaling of interventions (insecticide-treated
-bednets and indoor residual spraying) that aim to reduce mosquito
-lifespan (Bhatt et al., 2015). Worryingly, resistance to pyrethroids,
-the only class of insecticide used in current insecticide-treated
-bednets and likely the only product to come to market in the near
+Indeed, the recent declines in malaria prevalence in Sub-Saharan Africa were likely due to upscaling of interventions (insecticide-treated bednets and indoor residual spraying) that aim to reduce mosquito lifespan (Bhatt et al., 2015).
+Worryingly, resistance to pyrethroids, the only class of insecticide used in current insecticide-treated bednets and likely the only product to come to market in the near
 future, has been determined to be widespread and increasing in intensity
@@
 deployed. It also emphasises the need for investment in new tools for
-real time monitoring of mosquito populations. In recent years, there has
-been considerable funding allocated to molecular and genomic research
+real time monitoring of mosquito populations.
+In this work, we have used modern statistical methods to synthesise precious field data conducted by entomologists past and present, to produce lower bound estimates of mosquito lifespan.
+In recent years, there has been considerable funding allocated to molecular and genomic research
 into mosquitoes that strengthens existing interventions and suggest

### 2018-11-23 MS.tex
Cross references between files now should be working

@@
 \externaldocument{figures}
+\externaldocument[S-]{supplementary_materials}
 
@@
 between the spatial extent of the recapture zone and LBL. We found no
-such pattern (Fig. S\#), although there was a positive correlation
-between LBL and trap density (Fig. S\#).
+such pattern (Fig. \ref{S-fig:mrr_lifeSpanVsRange}), although there was a positive correlation
+between LBL and trap density (Fig. \ref{S-fig:mrr_lifeSpanVsTrapDensity}).
 
@@
 fed on sugar pre-release lived on average for 0.7 days longer than those
-that were not fed (Fig. S\#; a pattern that was consistent across the
+that were not fed (Fig. \ref{S-fig:mrr_female_blood_sugar}; a pattern that was consistent across the
 genera). There were insufficient males that were either fed or unfed
@@
 (\emph{Ae. aegypti}, \emph{Cx. tarsalis}, \emph{A. gambiae s.l.} and
-\emph{A. culicifacies s.l.}) individually (Fig. S\#).
+\emph{A. culicifacies s.l.}) individually (Fig. \ref{S-fig:mrr_ThreeSpeciesVersusTemperature}).
 
@@
 (physiological age) into lifespan (chronological age) as described in
-the Methods (Fig. S\#). For ten species, we had enough data from both
+the SOM using estimates of gonotrophic cycle duration (Fig. \ref{S-fig:dissection_gonotrophicCycleRaw}). For ten species, we had enough data from both
 species to make a comparison, and there was a positive correlation (not
@@
 which may be why we failed to detect age-dependence in the latter (Fig
-S\#).
+\ref{S-fig:mrr_mcPowerAnalysis_senescence}).
 
@@
 experiments, the short study lengths or typical numbers of mosquitoes
-released, results in considerable uncertainty in lifespan estimates (Fig
-S\#). This indicates that statistical power can be substantially
+released, results in considerable uncertainty in lifespan estimates (Fig.
+\ref{S-fig:mrr_mcPowerAnalysis}). This indicates that statistical power can be substantially
 increased by pooling data across experiments as we did using a Bayesian
@@
 cycle duration, in our analysis, laboratory-based studies indicated a
-longer gonotrophic cycle (Fig. S\#). The distributions we used to
+longer gonotrophic cycle (Fig. \ref{S-fig:dissection_gonotrophicCycleRaw_MRRVsLab}). The distributions we used to
 convert physiological age into calendar age were calculated by pooling
@@
 overdispersed data from single experiments results in high measurement
-error (Fig. S\#). By applying different methods to each study, this
+error (Fig. \ref{S-fig:mrr_mcPowerAnalysis}). By applying different methods to each study, this
 could lead us to falsely detect patterns when none are present, and we
@@
 
-We conducted a power analysis of MRR experiments (Fig. S\#) to determine
+We conducted a power analysis of MRR experiments to determine
 whether typical experimental characteristics could detect senescence.
@@
 S\#) for case study populations with three different levels of
-senescence (Fig. S\#). This analysis indicated that power to detect
-senescence strongly depends on study length (Fig. S\#) but is
-insensitive to release size (Fig. S\#). Clements and Patterson (1981)
+senescence (Fig. \ref{S-fig:mrr_mcPowerAnalysis_senescence}A). This analysis indicated that power to detect
+senescence strongly depends on study length (Fig. \ref{S-fig:mrr_mcPowerAnalysis_senescence}B) but is
+insensitive to release size (Fig. \ref{S-fig:mrr_mcPowerAnalysis_senescence}C). Clements and Patterson (1981)
 conducted a meta-analysis of MRR and dissection-based field experiments
@@
 considered, it is now less common for MRR experiments to release female
-mosquitoes versus males than historically (Fig. S\#). Our analysis
+mosquitoes versus males than historically (Fig. \ref{S-fig:mrr_sexReleasesOverTime}). Our analysis
 indicates that females outlive male mosquitoes by approximately 0.9 days
-(Fig. \#), meaning that differences between the sexes may exist for
+(Fig. \ref{fig:mrr_sexDifferences_without_sugar_nor_blood}), meaning that differences between the sexes may exist for
 other ecological parameters determinable by MRR. This suggests that
@@
 within this, we determined no relationship between lifespan and
-temperature across all time series (Fig. 4) or, for any of the species
-with the most data (Fig. S\#). It is possible that by considering a raw
-average of air temperature across the month, this ignored other
-temperature-dependent factors of importance for longevity, such as
-temperature range. It is also possible that by ignoring the effects of
+temperature across all time series (Fig. \ref{fig:mrr_temperature}) or, for any of the species
+with the most data (Fig. \ref{S-fig:mrr_ThreeSpeciesVersusTemperature}). It is possible that by considering a raw
+average of air temperature across the month, this ignored, more complex, interactions between temperature and lifespan. It is also possible that by ignoring the effects of
 rainfall (the historical data on rainfall is less likely to be reliable
@@
 
-The importance of vector mortality for disease transmission has long been recognised,
+In this work, we have used modern statistical methods to synthesise precious field data conducted by entomologists past and present, to produce lower bound estimates of mosquito lifespan.
+The importance of vector mortality for disease transmission has long been recognised, however,
 since even before 1957, when George Macdonald formulated the
@@
 deployed. It also emphasises the need for investment in new tools for
-real time monitoring of mosquito populations.
-In this work, we have used modern statistical methods to synthesise precious field data conducted by entomologists past and present, to produce lower bound estimates of mosquito lifespan.
-In recent years, there has been considerable funding allocated to molecular and genomic research
+real time monitoring of mosquito populations. In recent years, there has been considerable funding allocated to molecular and genomic research
 into mosquitoes that strengthens existing interventions and suggest
@@
 The authors would like to thank the following for useful conversations
-throughout the course of this work: Austin Burt, Thomas Churcher, Steve
-Lindsay, Mike Bonsall and Ellie Sherrard-Smith.
+throughout the course of this work: Austin Burt, Mike Bonsall, Thomas Churcher, Steve
+Lindsay and Ellie Sherrard-Smith.
 

### 2018-11-23 supplementary_materials.tex
Changed gir attributes
Cross references between files now should be working

@@
 \usepackage{mdframed}
+\usepackage[section]{placeins}
 
@@
 \newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX
+
+\usepackage{xr}
+\externaldocument{figures}
 
@@
 \end{table}%
+
+\begin{figure}[h]
+	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_sexReleasesOverTime.pdf}}
+	\caption{\textbf{The number of MRR time-series by date and sex of released mosquitoes.} The numbers represent by-decade totals of the 232 time-series we used in the main analysis.}
+	\label{fig:mrr_sexReleasesOverTime}
+\end{figure}
 
@@
 	\label{fig:mrr_female_blood_sugar}
+\end{figure}
+
+\begin{figure}[ht]
+	\centerline{\includegraphics[width=1.25\textwidth]{./Figure_files/mrr_mcPowerAnalysis.pdf}}
+	\caption{\textbf{The average error in predicting mean lifespan as a function of study length (left) and number of released mosquitoes (right) in a Monte Carlo analysis.} For each parameter set we generated 400 simulated data series using the negative binomial sampling model with an exponential survival function as described in the text, and estimated the mean lifespan using maximum likelihood.
+The error bars show the standard deviation in the prediction error.}
+	\label{fig:mrr_mcPowerAnalysis}
+\end{figure}
+
+\begin{figure}[ht]
+	\centerline{\includegraphics[width=1.25\textwidth]{./Figure_files/mrr_mcPowerAnalysis_senescence.pdf}}
+	\caption{\textbf{The statistical power to detect senescence for A.
+three case study populations as a function of B.
+study length and C.
+number released.} For each parameter set we generated 500 simulated data series using the negative binomial sampling model with a gompertz hazard function ($\lambda(t) = \alpha e^{\beta t}$), and estimated the $\beta$ parameter using maximum likelihood and tested the null hypothesis $\beta=0$ (constant mortality risk) against the alternative $\beta>0$ using an likelihood ratio test with a 5\% test size.
+The error bars show the standard deviation in the power for each parameter set.
+The recapture probability parameters ($\psi$ and $\kappa$) used were the averages of those estimated for the actual data.}
+	\label{fig:mrr_mcPowerAnalysis_senescence}
 \end{figure}
@@
 
+\subsection{The statistical power of MRR studies}\label{sec:mrr_discussion_MonteCarlo}
+Consider an ideal MRR experiment, where the assumptions of no emigration and harmless marking are fully satisfied.
+With these data, how accurately can we estimate mosquito lifespan?
+To address this question we generated artificial MRR data and attempted to estimate the (known) parameters by maximum likelihood (a `Monte Carlo' analysis).
+Specifically, we simulated releases of $N$ mosquitoes that are monitored for $m$ days in each case, and determined how statistical power to estimate lifespan depends on these two parameters (Figure \ref{fig:mrr_mcPowerAnalysis}).
+In order to focus on estimating lifespan, we assumed the mosquitoes experience constant mortality, and recaptures follow the negative binomial sampling model (eqn.
+\ref{eq:NB}) where the recapture probability parameters ($\psi$ and $\kappa$) were the averages of those estimated for the actual data.
+
+Unsurprisingly, the error in predicting lifespan declines as the duration of study is increased (Figure \ref{fig:mrr_mcPowerAnalysis}A).
+However once a study length is much longer than the lifespan of the majority of mosquitoes, the predictive power cannot be improved by extending the study duration.
+The effect of increasing release size (while holding study length constant) is similarly a case of diminishing returns (Figure \ref{fig:mrr_mcPowerAnalysis}B).
+For the parameters we use, there are significant gains in accuracy from releasing $1,000$ rather than $100$ mosquitoes, but modest gains from releasing $10,000$ rather than $1000$ mosquitoes.
+
+We also conducted a power analysis for the detection of senescence in MRR experiments (Figure \ref{fig:mrr_mcPowerAnalysis_senescence}).
+Here we calculated the power of a maximum likelihood estimator of the `senescence parameter' $\beta$ of the Gompertz survival function (see Table \ref{tab:mrr_survivalDescription}) for case study populations with three different levels of senescence (Figure \ref{fig:mrr_mcPowerAnalysis_senescence}A.).
+These indicate that the power to detect senescence is strongly dependent on study length (Figure \ref{fig:mrr_mcPowerAnalysis_senescence}B.), but is insensitive to release size (Figure \ref{fig:mrr_mcPowerAnalysis_senescence}C.).
+\cite{clements1981analysis} conducted a meta-analysis of MRR and dissection-based experiments and found evidence of senescence that is, at least, qualitatively similar in magnitude to that of the `mild' case considered above.
+For this case detecting senescence with a power of 80\% requires a study length of at least 18 days.
+Since the median study length for experiments included in our analysis was 10 days (Table \ref{tab:mrr_aggregateData}) this could partly explain our failure to detect senescence at the species level.
+
 \section{Studies included in the MRR meta-analysis}\label{sec:appendix_mrrStudyList}

### 2018-11-23 MS.tex
Changed bibliography to thesis version and started adding references

@@
 \usepackage{amssymb,amsmath}
+\usepackage[citestyle=authoryear,natbib=true,backend=bibtex,giveninits=true]{biblatex}
 \usepackage{ifxetex,ifluatex}
@@
 
+\bibliography{Malaria}
 \begin{document}
@@
 allowing us to produce robust estimates of lower bounds on lifespan. It
-also enables us to critically appraise each field method, which
-highlights a need for alternative field methods for measuring this
+also enables us to critically appraise each field method, highlighting a need for alternative field methods for measuring this
 important mosquito characteristic.
@@
 
-mosquitoes, mortality, senescence, vector-borne disease, Bayesian,
-hierarchical model.
+mosquitoes, mortality, senescence, mark-release-recapture, vector-borne disease, Bayesian, hierarchical model.
 
@@
 Some of the most important infectious diseases afflicting humans are
-transmitted by mosquitoes, including pathogens such as the causative
+transmitted by mosquitoes \citep{gates2014}, including pathogens such as the causative
 agent of malaria that have been associated with humans throughout our
-evolutionary history as well recently emergent infections such as the
-Zika virus. Most mosquito species have a ``gonotrophic cycle'' involving
+evolutionary history \citep{carter2002evolutionary} as well recently emergent infections such as the
+Zika virus \citep{world2016statement}. Most mosquito species have a ``gonotrophic cycle'' involving
 successive episodes of vertebrate blood feeding, egg maturation and
-oviposition. In order for a mosquito to transmit a pathogen it must feed
+oviposition \citep{silver2007mosquito}. In order for a mosquito to transmit a pathogen it must feed
 on an infectious person and live long enough to complete at least one
@@
 mosquito population to allow the persistence of an indirectly
-transmitted infection. Lifespan can of course be straightforwardly
+transmitted infection \citep{macdonald1957epidemiology}. Lifespan can of course be straightforwardly
 assessed in the laboratory, but it is generally accepted that
@@
 have limited relevance in the field, and much effort has been directed
-at estimating this parameter in the vector's natural environment. Most
+at estimating this parameter in the vector's natural environment \citep{clements1981analysis,guerra2014global}. Most
 work has focused on assessing average daily mortality rates, and the
@@
 assumption and discovering whether mosquitoes senesce or show other
-types of age-dependent mortality has also been studied in the field.
+types of age-dependent mortality has also been studied in the field \citep{clements1981analysis,harrington2008age,hugo2014adult}.
 
@@
 
-\section{References}\label{references}
-
-ADDIN EN.REFLISTX
-
-\section{Figure Legends}\label{figure-legends}
-
-Figure 1:
-
 \end{document}

### 2018-11-23 MS.tex
Some references added and read through to `Number of gonotrophic cycles estimates from dissection' subsection

@@
 estimates lower bounds on lifespan (LBL). Notably, for the major African
-malaria vector \emph{Anopheles gambiae s.l.}, we estimate LBLs of 4.4
-days (median estimate; 25\%-75\% CI: 3.8-5.1 days for unfed female
-mosquitoes from the MRR analysis) and 7.6 days (median estimate;
+malaria vector \emph{Anopheles gambiae s.l.}, we estimate LBLs of 4.5
+days (mean estimate; 25\%-75\% CI: 3.8-5.1 days for unfed female
+mosquitoes from the MRR analysis) and 9.5 days (mean estimate;
 25\%-75\% CI: 5.2-11.0 days from the dissection analysis); and an LBL of
-4.2 days (median estimate; 25\%-75\%
+4.3 days (mean estimate; 25\%-75\%
 CI: 3.6-4.8 days, only present in the MRR database) for the
 predominantly East-African vector \emph{A. funestus s.l}. We estimate
-LBLs of 6.2 days (median estimate; 25\%-75\% CI: 4.5-8.5 days from the
-MRR analysis) and 4.3 days (median estimate; 25\%-75\% CI: 3.5-5.1 days
-from the dissection analysis) for \emph{Aedes aegypti}; and 11.6 days
+LBLs of 7.0 days (mean estimate; 25\%-75\% CI: 4.5-8.5 days from the
+MRR analysis) and 5.0 days (mean estimate; 25\%-75\% CI: 3.5-5.1 days
+from the dissection analysis) for \emph{Aedes aegypti}; and 12.1 days
 (median estimate; 25\%-75\% CI: 10.0-13.7 days from the MRR analysis)
@@
 laboratory studies, we estimate that female mosquitoes outlive males by
-0.9 days on average (median estimate; 25\%-75\% CI: 0.3-1.6 days). We
+0.9 days on average (mean estimate; 25\%-75\% CI: 0.3-1.6 days). We
 fit models incorporating mosquito senescence to the data, which allows
@@
 Mortality rates can be statistically estimated from the numbers of
-recaptures given certain assumptions. The main challenges with MRR is
+recaptures given certain assumptions \citep{silver2007mosquito}. The main challenges with MRR is
 ensuring the marking technique does not affect recapture probability,
@@
 The second technique is specific to mosquitoes and makes use of their
-gonotrophic cycle. The mosquito ovary is made up of ovarioles, each of
+gonotrophic cycle \citep{polovodova1949determination,detinova1962age}. The mosquito ovary is made up of ovarioles, each of
 which typically produces one egg every gonotrophic cycle. After the egg
@@
 In recent years many important vectors of disease have been shown to be
-complexes of very closely related species, biotypes or forms that cannot
+complexes of closely related species, biotypes or forms that cannot
 be distinguished morphologically (for example the morphospecies
@@
 a single MRR study were excluded for the hierarchical analysis. Of the
-232 data sets, 177 involved only females, 35 males, and 18 both sex
+232 data sets, 179 involved only females, 35 males, and 18 both sex
 releases. For 102 data sets the age of the released mosquitoes was known
@@
 mosquitoes were newly emerged at the time of release and return to this
-assumption later.
+assumption later. See Table S\ref{S-tab:mrr_aggregateData} for a summary of other data characteristics.
 
 We analysed all MRR experiments within the same statistical framework
-(for full details see the Supplementary Online Material). In the
-simplest case \emph{N\textsubscript{R}} mosquitoes are released at day
+(for full details see the Supplementary Online Material (SOM)). In the
+simplest case \emph{N\textsubscript{R}} mosquitoes are released on day
 zero and the probability that they remain in the recapture area until
@@
 \emph{t}, and shape parameter $\kappa$. The negative binomial has been
-used previously in MRR analyses (ref) because of its ability to
+used previously in analyses of mosquito count data \citep{service1971studies,nedelman1983negative} because of its ability to
 represent temporal over-dispersion in recaptures most likely caused by
@@
 so that the numbers remaining after time \emph{t} are given by the
-exponential distribution, \(exp\left\lbrack - \lambda t \right\rbrack\).
+exponential distribution, $\text{exp}( -\lambda t )$.
 We utilised this form extensively but in testing for senescence used
-five other models where $\lambda(t)$ varies with time so that
+five other models where $\lambda(t)$ varies with time so that,
 
 \begin{longtable}[]{@{}ll@{}}
-\(S\left( t \right) = e^{- \int_{0}^{t}{\lambda\left( t \right) \mathrm{d}\tau}}.\)
+\(S\left( t \right) = e^{- \int_{0}^{t}{\lambda\left( \tau \right) \mathrm{d}\tau}}.\)
 \end{longtable}
@@
 recapture probability (bounded in part by knowledge of the maximum daily
-recapture rates). We used a Bayesian hierarchical model to estimate
+recapture rates; see SOM). We used a Bayesian hierarchical model to estimate
 distributions of lifespan at the species and the genus levels, and
@@
 (MCMC) methods with convergence assessed using the \(\hat{R}\) statistic
-(Gelmin \& Rubin 1992). The predictive power of the model was assessed
+\citep{gelman1992inference}. The predictive power of the model was assessed
 using \emph{K}-fold cross validation which tests the ability of the
@@
 
-The Guerra et al. database included the latitude and longitude of each
-study, along with the date when the study began. We used this information to find
+The Guerra et al., (2014) database included the latitude and longitude of each
+study along with the date when the study began. We used this information to find
 estimates of the air temperature for each study using the European Centre for Medium
@@
 of time. Overall, we found 568 time series in 72 published articles.
-Because seldom were sufficient ancillary data available data to analyse
+Because seldom were sufficient ancillary data available to analyse
 trends over time, we aggregated the data from each time series. We
@@
 data sets the number of nulliparous females was anomalously low,
-something that has been noticed before (Gillies \& Wilkes 1965). As some
+something that has been noticed before \citep{gillies1965study}. As some
 studies have suggested that the first gonotrophic cycle tends to be
@@
 MRR estimates the length of time a mosquito remains alive and is still
-in the area available for recapture. It thus provides a lower bound to
+in the area available for recapture. It thus provides a lower bound to mean
 lifespan which we shall refer to as LBL. In 187 of the 230 MRR time
 series the estimated LBL was less than 10 days (Fig. \ref{fig:mrr_lifespan_individualEstimates}). The smallest
-estimate was \textless{}1 day for the Asian malaria vector
+estimate was 1.1 days for the Asian malaria vector
 \emph{Anopheles subpictus s.l.} which is unfeasibly short and almost
 certainly reflects dispersal out of the recapture zone or a violation of
-the assumptions of our analyses. The longest estimate was 18.3 days for
+the assumptions of our analyses. The longest estimate was 26.9 days for
 the temperate species \emph{Aedes simpsoni s.l.} which is a vector of
@@
 considerable variation. For example, there are 54 estimates of LBL for
-\emph{A. aegypti} which range from 2.2 days to 38.3 days with a mean of
-8.3 days and coefficient of variation of 0.7. There are significant
+\emph{Ae. aegypti} which range from 2.2 days to 38.3 days with a mean of
+8.3 days and coefficient of variation of 0.7. There were significant
 differences in LBL amongst species (ANOVA on median LBL controlling for
@@
 
-The estimated median LBL for female mosquitoes not pre-fed with blood or
+The estimated mean LBL for female mosquitoes not pre-fed with blood or
 sugar before release for \emph{Culex, Anopheles} and \emph{Aedes} were
-2.5, 5.0 and 6.9 days respectively with an overall estimate of 4.6 days
-(Fig. \ref{fig:mrr_lifespans}). Differences between genera were significant (ANOVA on median
+2.9, 6.8 and 8.1 days respectively with an overall estimate of 6.0 days
+(Fig. \ref{fig:mrr_lifespans}; Table S\ref{S-tab:mrr_estimated_lifespans}). Differences between genera were significant (ANOVA on median
 LBL controlling for sex and pre-release feeding:
@@
 all time series, we considered the four species with the most data
-(\emph{Ae. aegypti}, \emph{Cx. tarsalis}, \emph{A. gambiae s.l.} and
-\emph{A. culicifacies s.l.}) individually (Fig. \ref{S-fig:mrr_ThreeSpeciesVersusTemperature}).
+individually (\emph{Ae. aegypti}, \emph{Cx. tarsalis}, \emph{A. gambiae s.l.} and
+\emph{A. culicifacies s.l.}; Fig. \ref{S-fig:mrr_ThreeSpeciesVersusTemperature}).
 

### 2018-11-23 MS.tex
Read up until `Comparison of longevity estimates from two methods'

@@
 estimated. Overall, the mean number of cycles completed in a lifetime
-was 1.2 (Fig. \ref{fig:dissection_lifetimes_exponential}; posterior median) and across the 131 studies, 95\% of
+was 1.3 (posterior mean; Fig. \ref{fig:dissection_lifetimes_exponential}; Table S\ref{S-tab:dissection_estimated_lifespans}) and across the 131 studies, 95\% of
 the individual time series estimates were less than 3 (Fig. \ref{fig:dissection_individualEstimates_allSpecies}). The
 estimated greatest number of cycles was for \emph{Anopheles sergentii}
-(2.5 cycles; posterior median) which is adapted to desert conditions (it
+(3.0 cycles; posterior mean) which is adapted to desert conditions (it
 is known as the ``oasis vector'' of malaria) and may have evolved
-greater longevity. The important African malaria vector \emph{An.
-gambiae s.l.} was estimated to be the second longest living (1.9 cycles;
-posterior median). The smallest estimated mean number of gonotrophic
-cycles was for \emph{Anopheles} \emph{bellator} (0.5 cycles; posterior
-median) which transmits malaria in Brazil's Atlantic Forest. There were
+greater longevity. The important African malaria vector \emph{A.
+gambiae s.l.} was estimated to be the second longest living (2.4 cycles;
+posterior mean). The smallest estimated mean number of gonotrophic
+cycles was for \emph{Anopheles} \emph{bellator} (0.6 cycles; posterior
+mean) which transmits malaria in Brazil's Atlantic Forest. There were
 significant differences in estimated lifetime gonotrophic cycles amongst
@@
 The estimated lifetime gonotrophic cycles for the different genera were
-\emph{Anopheles,} 1.4; \emph{Mansonia}, 1.1; \emph{Culex,} 1.0; and
-\emph{Aedes} 0.8 (Fig. \ref{fig:dissection_lifetimes_exponential}) and the differences between the genera were
+\emph{Anopheles,} 1.6; \emph{Culex,} 1.2; \emph{Mansonia}, 1.1; and
+\emph{Aedes} 0.8 (Fig. \ref{fig:dissection_lifetimes_exponential}; Table S\ref{S-tab:dissection_estimated_lifespans}) and the differences between the genera were
 significant (ANOVA: \emph{F}\textsubscript{3,127} =3.4, \emph{p} =0.02;

### 2018-11-23 MS.tex
Read up until discussion

@@
 (physiological age) into lifespan (chronological age) as described in
-the SOM using estimates of gonotrophic cycle duration (Fig. \ref{S-fig:dissection_gonotrophicCycleRaw}). For ten species, we had enough data from both
-species to make a comparison, and there was a positive correlation (not
+the SOM using estimates of gonotrophic cycle duration (Fig.
+\ref{S-fig:dissection_gonotrophicCycleRaw}).
+Table S\ref{S-tab:dissection_estimated_lifespans_chron} provides posterior summaries of chronological for the species and genera in the dissection dataset.
+For ten species, we had enough data from both species to make a comparison, and there was a positive correlation (not
 statistically significant; Pearson correlation $\rho=0.42$, $n=10$, $p=0.23$)
 between the two measures (Fig. \ref{fig:comparison}), and in only one case -- for \emph{A.
-darlingi} - (Table S\#) there was a significant difference in the
-time-series level LBLs.
+darlingi} - there was a significant difference in the
+time-series level LBLs (Table S\ref{S-tab:comparison}).
 
@@
 can transmit the disease vary from \textless{}0.1\% for \emph{A.
-subpictus} (from the MRR analysis, as noted above likely to be due to
-the LBL substantially underestimating lifespan) to 52\% for the
+subpictus} (posterior median; from the MRR analysis, as noted above likely to be due to
+the LBL substantially underestimating lifespan) to 52\% (posterior median) for the
 drought-adapted and long-lived \emph{A sergentii.} The proportions
@@
 (dissection); and for \emph{A. funestus s.l.}: 9\% (MRR). Using the
-individual time series estimates, however, there was inconclusive
-evidence for a difference in EIP between the species (MRR ANOVA:
-\emph{F}\textsubscript{14,58} = 1.8, \emph{p}=0.07; MRR Kruskal-Wallis:
-\(\chi_{14}^{2} = 30.2\), \emph{p} \textless{}0.01; dissection ANOVA:
-\emph{F}\textsubscript{11,64} = 1.8, \emph{p}\textless{}0.01; dissection
-Kruskal-Wallis: \(\chi_{11}^{2} = 38.9\), p\textless{}0.01).
+individual time series estimates, there
+evidence for a difference in EIP between the species (Kruskal-Wallis used due to non-normality of data; MRR:
+\(\chi_{14}^{2} = 30.2\), \emph{p} \textless{}0.01; dissection: \(\chi_{11}^{2} = 38.9\), p\textless{}0.01).
 

### 2018-11-23 MS.tex
First draft of article

@@
 from the dissection analysis) for \emph{Aedes aegypti}; and 12.1 days
-(median estimate; 25\%-75\% CI: 10.0-13.7 days from the MRR analysis)
+(mean estimate; 25\%-75\% CI: 10.0-13.7 days from the MRR analysis)
 for \emph{Ae.} \emph{albopictus} -- the predominant vectors of dengue
@@
 different partitions. Further details of the prior specification,
-fitting and validation through posterior predictive checks are given in
+fitting and validation through posterior predictive checks \citep{lambert2018student} are given in
 the SOM.
@@
 In this study, we applied a Bayesian hierarchical framework to the
-analysis of a database of mark-release-recapture experiments and a
-database of mosquito dissection studies to estimate mosquito lifespan.
+analysis of a database of mark-release-recapture experiments and another for mosquito dissection studies to estimate mosquito lifespan.
 By applying a single framework, this allows us to effectively synthesise
@@
 marked than the sex differences. In contrast to a number of lab-based
-experiments, temperature was not determined to significantly impact
+experiments \citep{yang2009assessing,brady2013modelling}, temperature was not determined to significantly impact
 lifespan. By fitting a range of survival models to the data in both
@@
 lifespan. Lab experiments have demonstrated that marking can negatively
-impact survival (Verhurst et al., 2013; Dickens and Brant, 2014),
+impact survival \citep{verhulst2013advances,dickens2014effects}
 resulting in artificially depressed survival. MRR studies typically
@@
 objections include the impracticality of dissecting more than a small
-proportion of ovarioles (Hoc and Wilkes, 1995), particularly in African
-vector species (Gillies and Wilkes, 1965), the related issue of locating
+proportion of ovarioles \citep{hoc1995ovariole}, particularly in African
+vector species \citep{gillies1965study}, the related issue of locating
 ovarioles whose count of dilations represents true physiological age
-(Fox and Brust, 1994), and the variation in numbers of ovariolar
-dilations for mosquitoes of the same, known, physiological age (Kay,
-1979; Russell, 1986; Hugo et al., 2008). Indeed there is considerable
+\citep{fox1994dilatations}, and the variation in numbers of ovariolar
+dilations for mosquitoes of the same, known, physiological age \citep{kay1979age,russell1986population,hugo2008evaluations}. Indeed there is considerable
 uncertainty concerning the fundamental question of how dilations in
 ovarioles form in the first place. Whilst the `Old School' of thought (a
-term coined by Fox and Brust, 1994) headed by Polovodana (Polovodova,
-1949) and Detinova (Detinova, 1962) considers dilations to result from
-normal oogenesis, a `New School' headed by Lange and Hoc (Lange and Hoc,
-1981) has challenged this assertion. The New School believe that only
+term coined by Fox and Brust, 1994) headed by Polovodana \citep{polovodova1949determination} and Detinova \citep{detinova1962age} considers dilations to result from
+normal oogenesis, a `New School' headed by Lange and Hoc \citep{lange1981abortive} has challenged this assertion. The New School believe that only
 abortive oogenesis results in follicular dilations because normal
-oogenesis destroys the sack-like structures (Fox and Brust, 1994). This
+oogenesis destroys the sack-like structures \citep{fox1994dilatations}. This
 means that Polovodana's method requires dissecting large numbers of
@@
 a mosquito ages would elevate the chance of age `hypodiagnosis' for
-older specimens (Fox and Brust, 1994), and likely biases lifespan
+older specimens \citep{fox1994dilatations}, and likely biases lifespan
 estimates downwards. The difficulty of locating diagnostic ovarioles has
 been investigated using lab populations of \emph{Culex} and \emph{Aedes}
-mosquitoes by Hugo et al., 2008, who conclude that only a small
+mosquitoes by Hugo et al. (2008), who conclude that only a small
 percentage of ovarioles are diagnostic. The exchange rate between
@@
 cycles. Two methods are commonly used to estimate the duration of
-gonotrophic cycles: MRR studies (see, for example, Gillies and Wilkes,
-1965), where marked mosquitoes are recaptured and dissected to determine
+gonotrophic cycles: MRR studies (see, for example, \cite{gillies1965study}), where marked mosquitoes are recaptured and dissected to determine
 the number of gonotrophic cycles occurring since release; and
 laboratory-based observations of colonies of (typically) wild-caught
-females, or their progeny (see, for example, Afrane et al., 2005).
+females, or their progeny (see, for example, \cite{afrane2005effects}).
 Whilst it is unclear how each method could bias estimated gonotrophic
@@
 population, although there are conflicting opinions as to whether this
-results in a relative paucity (Gillies and Wilkes, 1965) or abundance
-(Clements and Paterson, 1981) of nulliparous individuals. In our
+results in a relative paucity \citep{gillies1965study} or abundance
+\citep{clements1981analysis} of nulliparous individuals. In our
 database, there are cases where there was an obvious deficit of
@@
 differing distribution of resting females between indoor and outdoor
-traps (Detinova, 1962; Gillies and Wilkes, 1965). We chose to not
+traps \citep{detinova1962age,clements1981analysis}. We chose to not
 include those counts of nulliparous individuals in our analysis where
@@
 lifespan. The alternative dissection-based approach of Detinova
-(Detinova, 1962), based on dichotomous categorisation of female mosquito
+\cite{detinova1962age}, based on dichotomous categorisation of female mosquito
 specimens as `parous' or `unparous' relies on fewer assumptions, and is
@@
 spatiotemporally-disaggregated captures of wild and re-caught marked
-mosquitoes will be especially useful to estimate lifespan and dispersal.
+mosquitoes will continue be useful in estimating lifespan and dispersal.
 
@@
 \emph{A. sergentii} was amongst the longest lived of the anopheline
-species with a posterior median LBL of 8.7 days (median estimate;
-25\%-75\% CI: 5.9-13.8 days) from the MRR analysis and 10.0 days (median
+species with a posterior median LBL of 12.4 days (mean estimate;
+25\%-75\% CI: 5.9-13.8 days) from the MRR analysis and 11.9 days (mean
 estimate; 25\%-50\% CI: 7.6-14.0 days) from the analysis of dissection
-studies. This species is a major vector of malaria in the Sahara (Sinka
-et al., 2010), where to act as a disease vector it must persevere
+studies. This species is a major vector of malaria in the Sahara \citep{sinka2010dominant}, where to act as a disease vector it must persevere
 through these hard conditions. It is reasonable to hypothesise that this
@@
 the greatest discrepancy in the estimates was the major African malaria
-vector \emph{A. gambiae s.l.}, where we estimated LBLs of 4.4 days
-(median estimate; 25\%-75\% CI: 3.8-5.1 days for unfed female) from the
-MRR analysis and 7.6 days (median estimate; 25\%-75\% CI: 5.2-11.0) from
-the dissection analysis. Across genera, the greatest discrepancy in
-estimates was for \emph{Culex} (a posterior median of 2.5 days from the
-MRR versus 4.4 days from the dissection analysis) with the smallest
-discrepancy for \emph{Anopheles} (4.8 versus 5.0 days). By contrast for
-\emph{Aedes} the estimates from the MRR studies (6.9 days) exceed those
-of dissection-based studies (3.5 days). Across all studies we estimate
-from the MRR analysis that mean mosquito lifespan is 4.6 days versus 5.1
+vector \emph{A. gambiae s.l.}, where we estimated LBLs of 4.5 days
+(mean estimate; 25\%-75\% CI: 3.8-5.1 days for unfed female) from the
+MRR analysis and 9.5 days (mean estimate; 25\%-75\% CI: 5.2-11.0) from
+the dissection analysis. Across genera, the greatest discrepancy in estimates was for
+\emph{Aedes}, where the estimates from the MRR studies (8.1 days) are considerably longer than those of dissection-based studies (3.5 days).
+This was followed by \emph{Culex} (a posterior mean of 2.9 days from the
+MRR versus 4.9 days from the dissection analysis) with the smallest
+discrepancy for \emph{Anopheles} (6.8 versus 6.4 days). Across all studies we estimate
+from the MRR analysis that mean mosquito lifespan is 6.0 days versus 5.5
 days from the dissection-based studies. Some of the differences in these
@@
 supported by our results since the dissection-based estimates
-(themselves not reliant on assumptions about dispersal) exceed the MRR
+(themselves not reliant on assumptions about dispersal) are similar or exceed the MRR
 estimates for \emph{Anopheles} and \emph{Culex} mosquitoes, but not for
@@
 than those from field experiments although are typically conducted on
-highly inbred mosquito strains. Styer et al. (2007), using colonies of
+highly inbred mosquito strains. \cite{styer2007mosquitoes}, using colonies of
 45,054 female and 55,997 male \emph{Ae. aegypti}, determined that
@@
 estimated as 31.69 $\pm$ 0.06 days for females and 16.39 $\pm$ 0.03 days
-for males. A similar study by Dawes et al., 2009 with a lab colony of
+for males. A similar study by \cite{dawes2009anopheles} with a lab colony of
 over 1000 female \emph{A. stephensi} found similar estimates for median
@@
 development of additional methods to estimate mosquito age, such as
-`Near-Infrared Spectroscopy' (Mayagaya et al., 2009; Sikulu et al.,
-2011; Lambert et al., 2018), if they are proven to work in the field,
+`Near-Infrared Spectroscopy' \citep{mayagaya2009non,sikulu2011evaluating,lambert2018monitoring} if they are proven to work in the field,
 may be of considerable worth here.
@@
 `senescence parameter' $\beta$ of the Gompertz survival function (see Table
-S\#) for case study populations with three different levels of
+S\ref{S-tab:mrr_survivalDescription}) for case study populations with three different levels of
 senescence (Fig. \ref{S-fig:mrr_mcPowerAnalysis_senescence}A). This analysis indicated that power to detect
@@
 above. For this case, detecting senescence with a power of 80\% requires
-a study length of at least 18 days. Since the median study length for
-experiments included in our analysis was 10 days (Table S\#) this could
+a study length of at least 18 days. Since the median study duration for
+experiments included in our analysis was 10 days (Table S\ref{S-tab:mrr_IndividualData}) this could
 partly explain our failure to detect senescence at the species level. A
 number of experiments have found evidence of age-dependence in
-laboratory populations (Styer et al., 2007; Dawes et al., 2009).
+laboratory populations \citep{styer2007mosquitoes,dawes2009anopheles}.
 However, the artificially benign environment of the laboratory means
@@
 in semi-field sites or large microcosms may be a valuable source of
-information on female mosquitoes.
+information on female mosquito ecology.
 
 Our estimates of LBL indicate that mosquitoes that were sugar-fed prior
-to release lived on average 0.7 days longer than those that were unfed
-(, suggesting the potential value of this underappreciated aspect of the
+to release lived on average 0.7 days longer than those that were unfed (Fig. \ref{S-fig:mrr_female_blood_sugar})
+suggesting the potential value of this underappreciated aspect of the
 mosquito ecology to the insects. It may also partly explain the recent
-successes in the use of Attractive Toxic Sugar Baits (ATSB) as a vector
-control intervention (Schlein and Muller references). More research is
+successes in the use of Attractive Toxic Sugar Baits as a vector
+control intervention \citep{muller2008decline,muller2010effective,muller2010field,muller2010successful,beier2012attractive}. More research is
 needed, however, to identify the sugar-feeding frequency and food
@@
 
-There is considerable evidence from the laboratory that temperature
-modulates mosquito ecology and behaviour (Matt Thomas references). The
+There is evidence mainly from laboratory studies that temperature
+modulates mosquito ecology and behaviour \citep{yang2009assessing,brady2013modelling,murdock2012complex,beck2013effect}. The
 locations and times of year over which the MRR studies were conducted
@@
 now famous Ross-Macdonald equation of R\textsubscript{0} for malaria.
-Indeed, the recent declines in malaria prevalence in Sub-Saharan Africa were likely due to upscaling of interventions (insecticide-treated bednets and indoor residual spraying) that aim to reduce mosquito lifespan (Bhatt et al., 2015).
+Indeed, the recent declines in malaria prevalence in Sub-Saharan Africa were likely due to upscaling of interventions (insecticide-treated bednets and indoor residual spraying) that aim to reduce mosquito lifespan \citep{bhatt2015effect}.
 Worryingly, resistance to pyrethroids, the only class of insecticide used in current insecticide-treated bednets and likely the only product to come to market in the near
 future, has been determined to be widespread and increasing in intensity
-across Sub-Saharan Africa (WHO reference 2018). This alarming trend
+across Sub-Saharan Africa \citep{world2018global}. This alarming trend
 highlights the need for continued MRR and dissection-based studies to
@@
 deployed. It also emphasises the need for investment in new tools for
-real time monitoring of mosquito populations. In recent years, there has been considerable funding allocated to molecular and genomic research
+real time monitoring of mosquito populations. In recent years, considerable funding has been allocated to molecular and genomic research
 into mosquitoes that strengthens existing interventions and suggest
@@
 
+\printbibliography
+
 \end{document}

### 2018-11-23 supplementary_materials.tex
Some references added and read through to `Number of gonotrophic cycles estimates from dissection' subsection
Read up until discussion
Going through supplementary up Collection of dissection data section

@@
 The MRR database in \cite{guerra2014global} contains 393 individual time-series, along with meta-data for a range of factors for each experiment (for example, species, geography, and date of study).
-We  cleaned the data, making a number of amendments to it (see Section \ref{sec:amendments}), and transformed the data into a form amenable to our statistical model.
+We  cleaned the data, making a number of amendments to it (see Section \ref{sec:amendments}), and transformed the data into a form amenable for our statistical model.
 We removed time-series with fewer than six separate recapture observations, resulting in 238 time series.
@@
 Tables \ref{tab:mrr_aggregateData} and \ref{tab:mrr_IndividualData} summarise the data used in the Bayesian hierarchical analysis.
-This data encompassed time-series from MRR experiments across 33 different species and three genera: \textit{Aedes} (91 separate time-series), \textit{Anopheles} (94), and \textit{Culex} (47), and spanning a wide geographical range (Figure \ref{fig:mrr_lifetimes_map}).
-The data comprise 177 female-only, 35 male-only and 18 mixed-sex release time-series.
-Where the age of the adult mosquitoes at release-time was available (102 time-series) we use this to increment our time variable (that represents mosquito age) in the statistical models.
-Where age was unavailable (130 time-series), we assume that the age at release is zero.
-We recognise this assumption will likely underestimate mosquito age for some studies and so we interpret our estimates as lower bounds on mosquito lifespan.
+This data encompassed time-series from MRR experiments across 33 different species and three genera: \textit{Aedes} (91 separate time-series), \textit{Anopheles} (94), and \textit{Culex} (47), and spanning a wide geographical range (Fig.
+\ref{fig:mrr_lifetimes_map}).
 
@@
 		\toprule
-		\textbf{Data characteristic} & \textbf{Total} \\
-		\midrule
-		Number of species & 33 \\
-		Number of genera & 3 \\
-		Number of \textit{Anopheles} & 94 \\
-		Number of \textit{Aedes} & 91 \\
-		Number of \textit{Culex} & 47 \\
-		Number of female-only MRR & 179 \\
-		Number of male-only MRR & 35 \\
-		Number of mixed sex MRR & 18 \\
-		Number of pre-release blood-feeding only MRR & 71 \\
-		Number of pre-release sugar-feeding only MRR & 41 \\
-		Number of pre-release both blood- and sugar-feeding MRR & 4 \\
-		Number of pre-release neither blood- and sugar-feeding MRR & 116 \\
-		Number of MRR time-series & 232 \\
+		\textbf{Data characteristic} & \textbf{Count} \\
+		\midrule
+		Species & 33 \\
+		Genera & 3 \\
+		\textit{Anopheles} & 94 \\
+		\textit{Aedes} & 91 \\
+		\textit{Culex} & 47 \\
+		Female-only MRR & 179 \\
+		Male-only MRR & 35 \\
+		Mixed sex MRR & 18 \\
+		Pre-release blood-feeding only MRR & 71 \\
+		Pre-release sugar-feeding only MRR & 41 \\
+		Pre-release both blood- and sugar-feeding MRR & 4 \\
+		Pre-release neither blood- and sugar-feeding MRR & 116 \\
+		MRR time-series & 232 \\
 		\bottomrule
 	\end{tabular}%
-	\caption{Summary of variables across all MRR time-series.}
+	\caption{\textbf{Summary of variables across all MRR time-series.} For the `Species' and `Genera' rows, the counts represent the number of distinct entries in the database; otherwise the counts represent the number of time series.}
 	\label{tab:mrr_aggregateData}%
@@
 		\end{tabularx}%
-		\caption{Summary of data from individual MRR time-series.}
+		\caption{\textbf{Summary of data from individual MRR time-series.}}
 		\label{tab:mrr_IndividualData}%
@@
 \subsection{Statistical model}\label{sec:mrr_statistical}
-Data for a typical MRR experiment consists of a single release of $N_{R}$ marked mosquitoes followed by a series of marked mosquito recaptures on subsequent days (Figure \ref{fig:mrr_exampleMRRSeries} shows three such example series.) We model the number $y(t)$ of marked mosquitoes recaptured on day $t$ using a negative binomial sampling model,
+Data for a typical MRR experiment consists of a single release of $N_{R}$ marked mosquitoes followed by a series of marked mosquito recaptures on subsequent days (Fig.
+\ref{fig:mrr_exampleMRRSeries} shows three such example series).
+We model the number $y(t)$ of marked mosquitoes recaptured on day $t$ using a negative binomial sampling model,
 %
@@
 where $\lambda$ is the sum of the rates of death and dispersal from the study area.
-To estimate the lifespan of mosquitoes we use this model because we find no evidence in support of age-dependence across all species (see below for the method used to determine this, and later the `Results').
+To estimate the lifespan of mosquitoes we use this model, partly because of the limited evidence we determine in favour of senescence (Figs.
+\ref{fig:mrr_elpd} and \ref{fig:dissection_elpd}).
 
@@
 To use a Bayesian methodology we must specify priors on this set of parameters.
-Here we choose to specify independent priors on each parameter of the form: $\lambda\sim \mathcal{N}(-2.32,1)$, $\psi\sim \text{exp}(50)$ (with $\psi$ constrained to be less than 0.1) and $\kappa\sim \textnormal{log-normal}(2,1)$.
-This prior on the rate parameter of the exponential distribution ($\lambda$) corresponds to a wide range of possible lifespans (Figure \ref{fig:mrr_individualTimeSeries_priors}A), with a mean of 10 days.
-It was necessary to use an informative prior on $\psi$ to allow it to be estimated; the prior on $\kappa$ is fairly uninformative and allows a wide range of values (Figure \ref{fig:mrr_individualTimeSeries_priors}B,C).
+Here we choose to specify independent priors on each parameter of the form: $\lambda\sim \mathcal{N}(-2.32,1)$, $\psi\sim \text{exp}(50)$ (with $\psi$ constrained to be less than 0.1, a threshold below which the majority of daily recapture probabilities in the data are observed) and $\kappa\sim \textnormal{log-normal}(2,1)$.
+This prior on the rate parameter of the exponential distribution ($\lambda$) corresponds to a wide range of possible lifespans (Fig.
+\ref{fig:mrr_individualTimeSeries_priors}A), with a mean of 10 days.
+It was necessary to use an informative prior on $\psi$ to bound it to sensible values and to prevent the correlation of estimates with $\lambda$ leading to unreasonably short estimates of $\lambda$; the prior on $\kappa$ is fairly uninformative and allows a wide range of values (Fig.
+\ref{fig:mrr_individualTimeSeries_priors}B,C).
 
@@
 There is that arising from variability in experimental methodology.
-However there is also variability from actual differences in lifespan across the different mosquito cohorts; for example due to genetic differences between mosquito populations or due to climatic differences.
+There is also variability from actual differences in lifespan across the different mosquito cohorts; for example due to genetic differences between mosquito populations or due to climatic differences.
 Because of this heterogeneity we use a Bayesian hierarchical model which is akin to a `random effects' model in classical statistics.
 This type of model assumes that there is random variation in parameters at the individual time-series level, although each of the parameters is drawn from a common `population-level' distribution.
-In our case we separately estimate three different model groupings where the population-level distributions correspond to the species, genus and overall (across all studies) levels respectively (Figure \ref{fig:mrr_methodDiagramHierarchy}).
+In our case we separately estimate three different model groupings where the population-level distributions correspond to the species, genus and overall (across all studies) levels respectively (Fig.
+\ref{fig:mrr_methodDiagramHierarchy}).
 This allows us to estimate average mosquito lifespan for a given species or genus, or alternatively across all time-series, by independent sampling from the posterior predictive distribution from which the individual $\lambda_i$ are drawn.
@@
 Whilst we assume that the variation in mosquito mortality across experiments is random in nature, we recognise that there may exist systematic sources of variation that we do not include in our model.
-To account for this potential source of bias we examined how the individual estimates of lifespan correlated with experimental covariates (number of traps, range of traps, trapping method), but found no such evidence of systematic variation in lifespan (data not shown).
+To account for this potential source of bias we examined how the individual estimates of lifespan correlated with experimental covariates (trapping method and number or density of traps).
+The data on trapping method in the Guerra et al., (2014) database is less reliable than other data entries because published MRR studies may use multiple trapping methods and typically do not disaggregate captures by method.
+When we examined the correlation of trapping method with lifespan estimates, however, we did not find systematic variation by method (data not shown).
+We determined a positive correlation between trapping density and lifespan estimates (Fig.
+\ref{fig:mrr_lifeSpanVsTrapDensity}), which indicates the impact of mosquito dispersal on estimated lifespan.
 
@@
 The first of these links the individual time-series estimates with the overarching group-level distributions.
-The parameters of these group-level distributions are then set prior distributions (Figure \ref{fig:mrr_methodDiagramHierarchy}).
+The parameters of these group-level distributions are then set prior distributions (Fig.
+\ref{fig:mrr_methodDiagramHierarchy}).
 As an example, we suppose that the rate parameter $\lambda_i$ of our exponential survival model in experiment $i$ is given by:
@@
 where $\delta^m_{\text{genus}[i]}$ measures a genus-specific effect of male-only releases ($\text{sex}^{\text{m}}_i=1$) versus female-only ($\text{sex}^{\text{m}}_i=0$); $\delta^{\text{mix}}_{\text{genus}[i]}$ is the analogous effect for mixed releases versus female-only; $\delta^{\text{sugar}}_{\text{genus}[i]}$ is the analogous effect for mosquitoes that were sugar-fed prior to release versus those that were not fed; and $\delta^{\text{blood}}_{\text{genus}[i]}$ is the analogous effect for mosquitoes that were blood-fed prior to release versus those that were not fed.
-The parameter $c_i$ is modelled hierarchically, being drawn from a group-level distribution assumed to be log-normal,
+The parameter $c_i$ is modelled hierarchically, being drawn from a group-level distribution assumed to be normal,
 %
 \begin{equation}
-c_i \sim  \textnormal{normal}(\mu_\lambda,\sigma)
+c_i \sim  \mathcal{N}(\mu_\lambda,\sigma)
 \end{equation}
 %
-We then specify priors on these group-level parameters, $\mu_\lambda \sim N(-2.32,1)$ and $\sigma \sim \textnormal{log-normal}(-3,1)$.
+We then specify priors on these group-level parameters, $\mu_\lambda \sim \mathcal{N}(-2.32,1)$ and $\sigma \sim \textnormal{log-normal}(-3,1)$.
 On all $\delta$ parameters, except $\delta^{\text{mix}}_{\text{genus}[i]}$, we specify standard normal priors.
-We specify a uniform prior on $\delta^{\text{mix}}_{\text{genus}[i]}\sim U(0,\delta^m_{\text{genus}[i]})$ so that the effect of mixed releases lies between the female-only release effect (0) and the male-only release effect.
+We specify a uniform prior on $\delta^{\text{mix}}_{\text{genus}[i]}\sim U(0,\delta^m_{\text{genus}[i]})$ so that the effect of mixed releases lies between the female-only release effect (0) and the male-only release effect ($\delta^m_{\text{genus}[i]}$).
 
 The relatively complex nature of priors in hierarchical models make it important to determine their influence on inferences.
-We chose the above -- somewhat uninformative -- priors to allow a range of mosquito lifespans in order to minimise their effect on the estimates we report (Figure \ref{fig:mrr_meanLifespanPrior}).
+We chose the above -- somewhat uninformative -- priors to allow a range of mosquito lifespans in order to minimise their effect on the estimates we report (Fig.
+\ref{fig:mrr_meanLifespanPrior}).
 We also chose to set hierarchical priors on the remaining parameters in our models -- $\psi$ the probability of daily recapture, and $\kappa$ the `over-dispersion' parameter of a negative binomial distribution.
 For $\psi$ we chose an informative prior that placed all probability mass $\psi\leq 10\%$, since in most experiments the fraction of mosquitoes captured was significantly below this value.
-For $\kappa$ we chose a fairly wide prior that had most of its probability mass below $\kappa=20$ (Figure \ref{fig:mrr_PsiKappaPriors}).
+For $\kappa$ we chose a fairly wide prior that had most of its probability mass below $\kappa=20$ (Fig.
+\ref{fig:mrr_PsiKappaPriors}).
 In all cases the hierarchical priors were chosen to have similar implications on lifespan, recapture probability and over-dispersion at the individual time-series level as for the non-hierarchical analysis described previously.
@@
 
-The parameters of the survival function of each model were assigned hierarchical priors that allowed considerable variation in mosquito lifespan (see Figure \ref{fig:mrr_meanLifespanPrior}).
+The parameters of the survival function of each model were assigned hierarchical priors that allowed considerable variation in mosquito lifespan (see Fig.
+\ref{fig:mrr_meanLifespanPrior}).
 Otherwise the statistical model we used remained the same as for the constant mortality case.
@@
 			\midrule
-			Exponential & $\lambda\sim \textnormal{log-normal}(\mu_\lambda,\sigma)$ &$\mu_\lambda \sim N(-2.32,1)\text{,}\; \sigma\sim \textnormal{log-normal}(-3,1)$\\
+			Exponential & $c\sim \mathcal{N}(\mu_\lambda,\sigma)$ &$\mu_\lambda \sim N(-2.32,1)\text{,}\; \sigma\sim \textnormal{log-normal}(-3,1)$\\
 			\midrule
@@
 Instead we use Markov Chain Monte Carlo (MCMC) methods to sample from each posterior distribution.
-In particular we used \textit{Stan} software \citep{carpenter2016stan} that implements an efficient MCMC algorithm known as NUTS \citep{hoffman2014no}.
+We used \textit{Stan} software \citep{carpenter2016stan} that implements an efficient MCMC algorithm known as NUTS \citep{hoffman2014no}.
 
@@
 \subsection{K-Fold cross validation}\label{sec:mrr_kFold}
-One way to estimate a model's out-of-sample predictive capability is to use Akaike Information Criterion, Bayesian Information Criterion, Deviance Information Criterion or Watanabe Akaike Information Criterion that explicitly penalise a model in accordance to its complexity (typically indexed by its number of free parameters) to correct for model over-fit.
+One way to estimate a model's out-of-sample predictive capability is to use the Akaike Information Criterion that explicitly penalises a model in accordance to its complexity (typically indexed by its number of free parameters) to correct for model over-fit.
 The way in which this correction takes places however, is fairly heuristic, and less appropriate for hierarchical models.
@@
 
-For each individual series we recorded the following meta-data: genus, species, collection method, whether or not insecticide was mentioned as being used in or around the time of mosquito collections, and the start and end dates of the experiment.
+For each individual series we recorded the following meta-data: genus, species, collection method, whether or not insecticide was mentioned as being used during or shortly before the time of mosquito collections, and the start and end dates of the experiment.
 At the article level we recorded the following meta-data: author, year, title, country, location (within country), start and end date, whether insecticide was used during any of the experimental replicates, whether a mrr experiment occurred alongside the dissection study, and the collection location (indoor or outdoor).
@@
 The data collection method resulted in 568 separate dissection series, across 72 published articles (see Section \ref{sec:appendix_dissectionStudyList} for a list of the studies included in our final dataset).
-The published datasets cover the period from 1960-2015 with comparable numbers of studies across each decade (Figure \ref{fig:dissection_timeSeries}).
+The published datasets cover the period from 1960-2015 with comparable numbers of studies across each decade (Fig.
+\ref{fig:dissection_timeSeries}).
 The statistical approach applied to the data relies on the assumption that there is a constant rate of recruitment into the adult population (see Section \ref{sec:dissection_dissectionStats}).
@@
 Finally we remove data for any species where there was only one series resulting in 131 series across four genera (\textit{Anopheles}, \textit{Culex}, \textit{Aedes}, and \textit{Mansonia}) and 25 species (Table \ref{tab:dissection_speciesNumbers}).
-These studies are distributed across a wide range of geographies (Figure \ref{fig:dissection_map}).
+These studies are distributed across a wide range of geographies (Fig.
+\ref{fig:dissection_map}).
 The raw data for the dissection analysis are shown in Figures \ref{fig:dissection_exampleData1}, \ref{fig:dissection_exampleData2} and \ref{fig:dissection_exampleData3}.
@@
 To estimate lifespan at the species, genus and overall groupings we then use a hierarchical Bayesian model of the same mathematical form as for the analysis of MRR experiments (see Section \ref{sec:MRR_hierarchical}).
-However the priors for the analysis of dissection data were modified so that they represented a mean prior lifespan of three gonotrophic cycles, although allowed considerable variation in mean lifespan (Figure \ref{fig:dissection_lifespanPriorsParity}).
+However the priors for the analysis of dissection data were modified so that they represented a mean prior lifespan of three gonotrophic cycles, although allowed considerable variation in mean lifespan (Fig.
+\ref{fig:dissection_lifespanPriorsParity}).
 The priors used for the over-dispersion parameter ($\kappa$) for the negative binomial likelihood were the same as for the MRR analysis.
@@
 
-The estimates of mean mosquito lifespan that we present here (Figure \ref{fig:dissection_lifespan_gonotrophic}) resulted from the use of the exponential survival model (i.e.
+The estimates of mean mosquito lifespan that we present here (Fig.
+\ref{fig:dissection_lifespan_gonotrophic}) resulted from the use of the exponential survival model (i.e.
 no age dependence).
 This choice was made because we found limited evidence in support of age-dependent mortality (see Section \ref{sec:dissection_ageDependence}).
-However, since the priors for each of the survival functions were specified to allow for a wide variety of lifespans (Figure \ref{fig:dissection_lifespanPriorsParity}) the specific choice of survival function made little difference to resultant estimates (data not shown).
+However, since the priors for each of the survival functions were specified to allow for a wide variety of lifespans (Fig.
+\ref{fig:dissection_lifespanPriorsParity}) the specific choice of survival function made little difference to resultant estimates (data not shown).
 
 To demonstrate that the results we obtain are not sensitive to the particular form of hierarchical prior structure chosen we also provide estimates of the lifespan for each series analysed on its own.
-To do so we assume priors on the rate parameter of the exponential distribution that provides support over a wide range of possible lifespans (Figure \ref{fig:dissection_lifespanPrior_individual}A).
-We also specified a prior on the over-dispersion parameter $\kappa$ that was comparable to the hierarchical case (Figure \ref{fig:dissection_lifespanPrior_individual}B).
+To do so we assume priors on the rate parameter of the exponential distribution that provides support over a wide range of possible lifespans (Fig.
+\ref{fig:dissection_lifespanPrior_individual}A).
+We also specified a prior on the over-dispersion parameter $\kappa$ that was comparable to the hierarchical case (Fig.
+\ref{fig:dissection_lifespanPrior_individual}B).
 
@@
 In many articles the duration of the gonotrophic cycle was estimated separately for the first versus subsequent cycles, and these estimates were recorded separately.
-Raw estimates of gonotrophic cycle duration were obtained for species across four different genera (\textit{Aedes}, \textit{Anopheles}, \textit{Culex} and \textit{Masonia}), although there was a bias towards \textit{Anopheles} with $n=47$ estimates out of the total of 79 (Figure \ref{fig:dissection_gonotrophicCycleRaw}).
+Raw estimates of gonotrophic cycle duration were obtained for species across four different genera (\textit{Aedes}, \textit{Anopheles}, \textit{Culex} and \textit{Masonia}), although there was a bias towards \textit{Anopheles} with $n=47$ estimates out of the total of 79 (Fig.
+\ref{fig:dissection_gonotrophicCycleRaw}).
 
@@
 
-Since we collected data on the method used to estimate the duration of the gonotrophic cycle we show the raw estimates for the two most common approaches: MRR and laboratory (Figure \ref{fig:dissection_gonotrophicCycleRaw_MRRVsLab}).
+Since we collected data on the method used to estimate the duration of the gonotrophic cycle we show the raw estimates for the two most common approaches: MRR and laboratory (Fig.
+\ref{fig:dissection_gonotrophicCycleRaw_MRRVsLab}).
 From these results it is evident that the estimates from the laboratory studies are, on average, higher than those using the MRR method, (although we wait until Section \ref{sec:dissection_gonotrophicDuration} for a quantitative comparison).
@@
 The estimates of lifespan produced from analysing the dissection data are in terms of physiological age (the number of gonotrophic cycles undertaken).
-To allow comparison with the estimates from the MRR studies (Figure \ref{fig:mrr_lifespans}) as well as produce more useful estimates to inform disease transmission dynamical models we convert these estimates to calendar days.
+To allow comparison with the estimates from the MRR studies (Fig.
+\ref{fig:mrr_lifespans}) as well as produce more useful estimates to inform disease transmission dynamical models we convert these estimates to calendar days.
 To do this we use the estimated parameters of the normal densities that we have assumed represent uncertainty in gonotrophic cycle duration (see Section \ref{sec:dissection_gonotrophicMethod}) and use them to convert our posterior samples of mean physiological lifespan $(L^p_1,...,L^p_S)$ to lifespan in calendar ages $(L^c_1,...,L^c_S)$.
@@
 
-Figure \ref{fig:mrr_sex} shows the lifespans obtained by sex, for each genus and across all studies ("overall") using the hierarchical Bayesian model with exponential survival model.
+Fig.
+\ref{fig:mrr_sex} shows the lifespans obtained by sex, for each genus and across all studies ("overall") using the hierarchical Bayesian model with exponential survival model.
 In the genus-level models, the sex-specific effects (male-only, female-only, mixed-releases) were estimated as independent random effects for each genus with standard normal priors.
@@
 
-Figure \ref{fig:mrr_sugarBlood} shows the lifespans obtained for each pre-release feeding category for each genus and across all studies using the hierarchical Bayesian model with exponential survival model.
+Fig.
+\ref{fig:mrr_sugarBlood} shows the lifespans obtained for each pre-release feeding category for each genus and across all studies using the hierarchical Bayesian model with exponential survival model.
 In the genus-level models, the feeding-specific effects (fed nothing, blood-fed-only, sugar-fed-only and fed both) were estimated as independent random effects for each genus with standard normal priors.
@@
 \end{tabular}
-\caption{Posterior summaries of mean mosquito lifespan from the MRR analysis.
+\caption{Posterior summaries of mean unfed female mosquito lifespan from the MRR analysis.
 The 5\%, 25\%, 50\%, 75\% and 95\% columns indicate the respective quantiles of the posterior distribution for mean lifespan, and the `Mean' and `Std.
@@
 
+\subsubsection{Chronological lifespan estimates}
+\begin{figure}[h]
+	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/dissection_lifetimes_exponential_chron.pdf}}
+	\caption{\textbf{Posterior estimates of female mosquito mean chronological lifespan across species, genus and overall groupings as determined from the dissection data.} The lifespans shown are for mosquitoes that were not fed with sugar or blood (for females) before release.
+The middle line in each box shows the median estimates and the solid dot indicates the mean.
+The left and right box edges show the 25\%, and 75\% posterior quantiles respectively.
+The whiskers show the range of the data, excluding points lying more than 1.5 times the interquartile range away from each edge of the box.
+The numbers before the start of the left whisker indicate the number of individual time-series within each species.
+All estimates were obtained using the hierarchical exponential survival model.}
+	\label{fig:dissection_lifetimes_exponential}
+\end{figure}
+
 \begin{table}[htbp]
@@
 
-\subsubsection{Chronological lifespan estimates}
-\begin{figure}[h]
-	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/dissection_lifetimes_exponential_chron.pdf}}
-	\caption{\textbf{Posterior estimates of female mosquito mean chronological lifespan across species, genus and overall groupings as determined from the dissection data.} The lifespans shown are for mosquitoes that were not fed with sugar or blood (for females) before release.
-The middle line in each box shows the median estimates and the solid dot indicates the mean.
-The left and right box edges show the 25\%, and 75\% posterior quantiles respectively.
-The whiskers show the range of the data, excluding points lying more than 1.5 times the interquartile range away from each edge of the box.
-The numbers before the start of the left whisker indicate the number of individual time-series within each species.
-All estimates were obtained using the hierarchical exponential survival model.}
-	\label{fig:dissection_lifetimes_exponential}
-\end{figure}
-
 \subsection{Comparison of LBLs between the MRR and dissection analyses}
@@
 To address this question we generated artificial MRR data and attempted to estimate the (known) parameters by maximum likelihood (a `Monte Carlo' analysis).
-Specifically, we simulated releases of $N$ mosquitoes that are monitored for $m$ days in each case, and determined how statistical power to estimate lifespan depends on these two parameters (Figure \ref{fig:mrr_mcPowerAnalysis}).
+Specifically, we simulated releases of $N$ mosquitoes that are monitored for $m$ days in each case, and determined how statistical power to estimate lifespan depends on these two parameters (Fig.
+\ref{fig:mrr_mcPowerAnalysis}).
 In order to focus on estimating lifespan, we assumed the mosquitoes experience constant mortality, and recaptures follow the negative binomial sampling model (eqn.
@@
 
-Unsurprisingly, the error in predicting lifespan declines as the duration of study is increased (Figure \ref{fig:mrr_mcPowerAnalysis}A).
+Unsurprisingly, the error in predicting lifespan declines as the duration of study is increased (Fig.
+\ref{fig:mrr_mcPowerAnalysis}A).
 However once a study length is much longer than the lifespan of the majority of mosquitoes, the predictive power cannot be improved by extending the study duration.
-The effect of increasing release size (while holding study length constant) is similarly a case of diminishing returns (Figure \ref{fig:mrr_mcPowerAnalysis}B).
+The effect of increasing release size (while holding study length constant) is similarly a case of diminishing returns (Fig.
+\ref{fig:mrr_mcPowerAnalysis}B).
 For the parameters we use, there are significant gains in accuracy from releasing $1,000$ rather than $100$ mosquitoes, but modest gains from releasing $10,000$ rather than $1000$ mosquitoes.
 
-We also conducted a power analysis for the detection of senescence in MRR experiments (Figure \ref{fig:mrr_mcPowerAnalysis_senescence}).
-Here we calculated the power of a maximum likelihood estimator of the `senescence parameter' $\beta$ of the Gompertz survival function (see Table \ref{tab:mrr_survivalDescription}) for case study populations with three different levels of senescence (Figure \ref{fig:mrr_mcPowerAnalysis_senescence}A.).
-These indicate that the power to detect senescence is strongly dependent on study length (Figure \ref{fig:mrr_mcPowerAnalysis_senescence}B.), but is insensitive to release size (Figure \ref{fig:mrr_mcPowerAnalysis_senescence}C.).
+We also conducted a power analysis for the detection of senescence in MRR experiments (Fig.
+\ref{fig:mrr_mcPowerAnalysis_senescence}).
+Here we calculated the power of a maximum likelihood estimator of the `senescence parameter' $\beta$ of the Gompertz survival function (see Table \ref{tab:mrr_survivalDescription}) for case study populations with three different levels of senescence (Fig.
+\ref{fig:mrr_mcPowerAnalysis_senescence}A.).
+These indicate that the power to detect senescence is strongly dependent on study length (Fig.
+\ref{fig:mrr_mcPowerAnalysis_senescence}B.), but is insensitive to release size (Fig.
+\ref{fig:mrr_mcPowerAnalysis_senescence}C.).
 \cite{clements1981analysis} conducted a meta-analysis of MRR and dissection-based experiments and found evidence of senescence that is, at least, qualitatively similar in magnitude to that of the `mild' case considered above.

### 2018-11-23 supplementary_materials.tex
Read up to `Here we estimate our model with one of six different survival functions '

@@
 To obtain a reasonable level of accuracy on estimates we remove all those individual (aggregated) series where there are fewer than 100 mosquitoes in total in the series.
-Finally we remove data for any species where there was only one series resulting in 131 series across four genera (\textit{Anopheles}, \textit{Culex}, \textit{Aedes}, and \textit{Mansonia}) and 25 species (Table \ref{tab:dissection_speciesNumbers}).
+Finally we remove data for any species where there was only one series resulting in 131 series across four genera (\textit{Anopheles}, \textit{Culex}, \textit{Aedes}, and \textit{Mansonia}) and 25 species (Table S\ref{tab:dissection_speciesNumbers}).
 These studies are distributed across a wide range of geographies (Fig.
@@
 where $p(a)$ is the probability of recapturing a single mosquito of age $a$.
-In what follows we typically assume that the probability of capturing a given mosquito is independent of their age, so that $p(a)=p=const$.
+In what follows we assume that the probability of capturing a given mosquito is independent of their age, so that $p(a)=p=const$.
 Since in general $A(a)$ is large and $p(a)$ is small, we can approximate the above using a Poisson distribution,
@@
 
-Here we estimate our model with one of six different survival functions (the same as for the MRR case; Table \ref{tab:mrr_survivalDescription}), each of which makes different assumptions regarding how the force of mortality is affected by mosquito age.
+Here we estimate our model with one of six different survival functions (the same as for the MRR case; Table S\ref{tab:mrr_survivalDescription}), each of which makes different assumptions regarding how the force of mortality is affected by mosquito age.
 To estimate lifespan at the species, genus and overall groupings we then use a hierarchical Bayesian model of the same mathematical form as for the analysis of MRR experiments (see Section \ref{sec:MRR_hierarchical}).

### 2018-11-23 MS.tex
Sorted table references

@@
 mosquitoes were newly emerged at the time of release and return to this
-assumption later. See Table S\ref{S-tab:mrr_aggregateData} for a summary of other data characteristics.
+assumption later. See Table \ref{S-tab:mrr_aggregateData} for a summary of other data characteristics.
 
@@
 2.9, 6.8 and 8.1 days respectively with an overall estimate of 6.0 days
-(Fig. \ref{fig:mrr_lifespans}; Table S\ref{S-tab:mrr_estimated_lifespans}). Differences between genera were significant (ANOVA on median
+(Fig. \ref{fig:mrr_lifespans}; Table \ref{S-tab:mrr_estimated_lifespans}). Differences between genera were significant (ANOVA on median
 LBL controlling for sex and pre-release feeding:
@@
 estimated. Overall, the mean number of cycles completed in a lifetime
-was 1.3 (posterior mean; Fig. \ref{fig:dissection_lifetimes_exponential}; Table S\ref{S-tab:dissection_estimated_lifespans}) and across the 131 studies, 95\% of
+was 1.3 (posterior mean; Fig. \ref{fig:dissection_lifetimes_exponential}; Table \ref{S-tab:dissection_estimated_lifespans}) and across the 131 studies, 95\% of
 the individual time series estimates were less than 3 (Fig. \ref{fig:dissection_individualEstimates_allSpecies}). The
@@
 \emph{Anopheles,} 1.6; \emph{Culex,} 1.2; \emph{Mansonia}, 1.1; and
-\emph{Aedes} 0.8 (Fig. \ref{fig:dissection_lifetimes_exponential}; Table S\ref{S-tab:dissection_estimated_lifespans}) and the differences between the genera were
+\emph{Aedes} 0.8 (Fig. \ref{fig:dissection_lifetimes_exponential}; Table \ref{S-tab:dissection_estimated_lifespans}) and the differences between the genera were
 significant (ANOVA: \emph{F}\textsubscript{3,127} =3.4, \emph{p} =0.02;
@@
 \ref{S-fig:dissection_gonotrophicCycleRaw}).
-Table S\ref{S-tab:dissection_estimated_lifespans_chron} provides posterior summaries of chronological for the species and genera in the dissection dataset.
+Table \ref{S-tab:dissection_estimated_lifespans_chron} provides posterior summaries of chronological for the species and genera in the dissection dataset.
 For ten species, we had enough data from both species to make a comparison, and there was a positive correlation (not
@@
 darlingi} - there was a significant difference in the
-time-series level LBLs (Table S\ref{S-tab:comparison}).
+time-series level LBLs (Table \ref{S-tab:comparison}).
 
@@
 `senescence parameter' $\beta$ of the Gompertz survival function (see Table
-S\ref{S-tab:mrr_survivalDescription}) for case study populations with three different levels of
+\ref{S-tab:mrr_survivalDescription}) for case study populations with three different levels of
 senescence (Fig. \ref{S-fig:mrr_mcPowerAnalysis_senescence}A). This analysis indicated that power to detect
@@
 a study length of at least 18 days. Since the median study duration for
-experiments included in our analysis was 10 days (Table S\ref{S-tab:mrr_IndividualData}) this could
+experiments included in our analysis was 10 days (Table \ref{S-tab:mrr_IndividualData}) this could
 partly explain our failure to detect senescence at the species level. A

### 2018-11-23 MS.tex
Read up till `Analysis of censored data series' section

@@
 \ref{S-fig:dissection_gonotrophicCycleRaw}).
-Table \ref{S-tab:dissection_estimated_lifespans_chron} provides posterior summaries of chronological for the species and genera in the dissection dataset.
+Table \ref{S-tab:dissection_estimated_lifespans_chron} provides posterior summaries of chronological for the species and genera in the dissection dataset (see also Fig.
+).
 For ten species, we had enough data from both species to make a comparison, and there was a positive correlation (not

### 2018-11-23 supplementary_materials.tex
Sorted table references
Read up till `Analysis of censored data series' section

--- 2018-11-23 supplementary_materials.tex
@@
 \renewcommand{\thefigure}{S\arabic{figure}}
+\renewcommand{\thetable}{S\arabic{table}}
 \renewcommand{\thesection}{S\arabic{section}}
@@
 The estimates of mean mosquito lifespan that we present here (Fig.
-\ref{fig:dissection_lifespan_gonotrophic}) resulted from the use of the exponential survival model (i.e.
+\ref{fig:dissection_lifetimes_exponential}) resulted from the use of the exponential survival model (i.e.
 no age dependence).
-This choice was made because we found limited evidence in support of age-dependent mortality (see Section \ref{sec:dissection_ageDependence}).
+This choice was made because we found limited evidence in support of age-dependent mortality (see `Results').
 However, since the priors for each of the survival functions were specified to allow for a wide variety of lifespans (Fig.
@@
 
-To demonstrate that the results we obtain are not sensitive to the particular form of hierarchical prior structure chosen we also provide estimates of the lifespan for each series analysed on its own.
+To demonstrate that the results we obtain are not sensitive to the particular form of hierarchical prior structure chosen we also provide estimates of the lifespan for each series analysed separately (i.e.
+non-hierarchically).
 To do so we assume priors on the rate parameter of the exponential distribution that provides support over a wide range of possible lifespans (Fig.
@@
 All estimates were obtained using the hierarchical exponential survival model.}
-	\label{fig:dissection_lifetimes_exponential}
+	\label{fig:dissection_lifetimes_exponential_chron}
 \end{figure}

### 2018-11-23 supplementary_materials.tex
Read up to `Statistical analysis of dissection data'

@@
 	\centerline{\includegraphics[width=1.3\textwidth]{./Figure_files/dissection_parity_data3.png}}
-	\caption{\textbf{The normalised physiological age series for nine species in the database.} Each different coloured line represents an individual series.
+	\caption{\textbf{The normalised physiological age series for seven species in the database.} Each different coloured line represents an individual series.
 In each case the count for all ages has been normalised by the nulliparous count.
@@
 \end{figure}
-
-As for the MRR analysis we use \textit{Stan} that implements a form of MCMC algorithm known as NUTS \citep{hoffman2014no} to sample from the posterior.
-The protocol followed for the MCMC sampling was the same as described in Section \ref{sec:mrr_MCMC}.
 

### 2018-11-23 supplementary_materials.tex
Finished run through of SOM

@@
 	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/dissection_lifespanPriorsParity.pdf}}
-	\caption{\textbf{The priors over mean lifespan used to analyse the data for each of the hierarchical models.} The distributional form of the priors is given in Table \ref{tab:dissection_priors}.
+	\caption{\textbf{The priors for mean lifespan used to analyse the data for each of the hierarchical models.} The distributional form of the priors is given in Table \ref{tab:dissection_priors}.
 In all cases the plots show data for 10,000 samples from the relevant prior distribution.}\label{fig:dissection_lifespanPriorsParity}
@@
 	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/dissection_lifespanPrior_individual.pdf}}
-	\caption{\textbf{The priors over A.
+	\caption{\textbf{The priors for A.
 mean lifespan and B.
-over-dispersion parameter used to analyse the individual dissection series.} The prior on the rate parameter of the exponential distribution was $\lambda\sim \text{log-normal}(-1.8,1)$.
+the over-dispersion parameter used to analyse the individual dissection series.} The prior on the rate parameter of the exponential distribution was $\lambda\sim \text{log-normal}(-1.8,1)$.
 The prior on the dispersion parameter was $\kappa\sim \text{log-normal}(1,1)$.}\label{fig:dissection_lifespanPrior_individual}
@@
 	\includegraphics[page=1,scale=.7]{./Figure_files/dissection_ppcs_all}
-	\captionof{figure}{The numbers of mosquitoes recaptured (black lines) versus the 95\% central posterior interval of the posterior predictive distribution (blue shading) for a selection of the dissection series.
-The titles correspond to the unique combination of "identifier - physiological-experiment-id" in the database.
+	\captionof{figure}{\textbf{The numbers of mosquitoes recaptured (black lines) versus the 95\% central posterior interval of the posterior predictive distribution (blue shading) for a selection of the dissection series.} The titles correspond to the unique combination of "identifier - physiological-experiment-id" in the database.
 For the rest of the posterior predictive checks for the dissection data models, see the file referenced in the text.}
@@
 	\includegraphics[page=2,scale=.7]{./Figure_files/dissection_ppcs_all}
-	\captionof{figure}{The numbers of mosquitoes recaptured (black lines) versus the 95\% central posterior interval of the posterior predictive distribution (blue shading) for a selection of the dissection series.
-The titles correspond to the unique combination of "identifier - physiological-experiment-id" in the database.
+	\captionof{figure}{\textbf{The numbers of mosquitoes recaptured (black lines) versus the 95\% central posterior interval of the posterior predictive distribution (blue shading) for a selection of the dissection series.} The titles correspond to the unique combination of "identifier - physiological-experiment-id" in the database.
 For the rest of the posterior predictive checks for the dissection data models, see the file referenced in the text.}
@@
 	\includegraphics[page=3,scale=.7]{./Figure_files/dissection_ppcs_all}
-	\captionof{figure}{The numbers of mosquitoes recaptured (black lines) versus the 95\% central posterior interval of the posterior predictive distribution (blue shading) for a selection of the dissection series.
-The titles correspond to the unique combination of "identifier - physiological-experiment-id" in the database.
+	\captionof{figure}{\textbf{The numbers of mosquitoes recaptured (black lines) versus the 95\% central posterior interval of the posterior predictive distribution (blue shading) for a selection of the dissection series.} The titles correspond to the unique combination of "identifier - physiological-experiment-id" in the database.
 For the rest of the posterior predictive checks for the dissection data models, see the file referenced in the text.}
@@
 \ref{fig:dissection_gonotrophicCycleRaw_MRRVsLab}).
-From these results it is evident that the estimates from the laboratory studies are, on average, higher than those using the MRR method, (although we wait until Section \ref{sec:dissection_gonotrophicDuration} for a quantitative comparison).
+From these results it is evident that the estimates from the laboratory studies are, on average, higher than those using the MRR method.
 
@@
 We decided to model the estimates as representing quantiles from an underlying normal distribution that represents uncertainty over possible durations of the gonotrophic cycle.
-In those circumstances where lower, central and upper bounds were given explicitly the data was fairly symmetric and so we believe assuming the observations come from an unskewed normal distribution is not unreasonable.
+In those circumstances where lower, central and upper bounds were given explicitly the data was fairly symmetric and so we believe assuming the observations come from an unskewed normal distribution is reasonable.
 Explicitly we treated each type of uncertainty interval as follows,
@@
 This was justified because there was not found to be significant variation in the 1st gonotrophic cycle duration (ANOVA: $F_{3,75}=2.6,p>0.05$).
+
+Using the collected data, we estimated that uncertainty in the first gonotrophic cycle duration is estimated as $\mathcal{N}(4.26, 0.38)$ and, for subsequent cycles, $\mathcal{N}(3.85, 0.35)$.
 
@@
 \subsection{MRR experments}
-Figures \ref{fig:mrr_lifeSpanVsRange}, \ref{fig:mrr_lifeSpanVsTrapDensity} and \ref{fig:mrr_temperature} show the estimates of lifespan versus the trapping range, spatial density and temperature at study location.
+Figures \ref{fig:mrr_lifeSpanVsRange} and \ref{fig:mrr_lifeSpanVsTrapDensity} show the estimates of lifespan versus the trapping range and spatial density of traps at study location.
 In these cases, the lifespans shown were obtained from the non-hierarchical model with the exponential survival model.
+The spatial density of traps was obtained assuming that the study area was a circular area of radius equal to the maximum trapping distance.
 
 Fig.
-\ref{fig:mrr_sex} shows the lifespans obtained by sex, for each genus and across all studies ("overall") using the hierarchical Bayesian model with exponential survival model.
-In the genus-level models, the sex-specific effects (male-only, female-only, mixed-releases) were estimated as independent random effects for each genus with standard normal priors.
-For the overall model, the sex-specific effects were estimated as independent random effects across all studies, with a standard normal prior.
-
-Fig.
-\ref{fig:mrr_sugarBlood} shows the lifespans obtained for each pre-release feeding category for each genus and across all studies using the hierarchical Bayesian model with exponential survival model.
-In the genus-level models, the feeding-specific effects (fed nothing, blood-fed-only, sugar-fed-only and fed both) were estimated as independent random effects for each genus with standard normal priors.
+\ref{fig:mrr_female_blood_sugar} shows the lifespans obtained for each pre-release feeding category for each genus and across all studies using the hierarchical Bayesian model with exponential survival model.
+In the genus-level models, the feeding-specific effects (not pre-fed, blood-fed-only and sugar-fed-only) were estimated as independent random effects for each genus with standard normal priors.
 For the overall model, the feeding-specific effects were estimated as independent random effects across all studies, with a standard normal prior.
@@
 \end{figure}
-
-\subsubsection{Table of estimated lifespans estimated from mark-release-recapture studies}
 
@@
 \end{tabular}
-\caption{Posterior summaries of mean unfed female mosquito lifespan from the MRR analysis.
-The 5\%, 25\%, 50\%, 75\% and 95\% columns indicate the respective quantiles of the posterior distribution for mean lifespan, and the `Mean' and `Std.
+\caption{\textbf{Posterior summaries of mean unfed female mosquito lifespan from the MRR analysis.} The 5\%, 25\%, 50\%, 75\% and 95\% columns indicate the respective quantiles of the posterior distribution for mean lifespan, and the `Mean' and `Std.
 dev.' columns indicate the posterior mean and standard deviation of mean lifespan.}
@@
 \subsection{Dissection studies}
-
-\subsubsection{Table of estimated lifespan in terms of gonotrophic cycles and chronological time using dissection data}
+Tables \ref{tab:dissection_estimated_lifespans} and \ref{tab:dissection_estimated_lifespans_chron} show posterior summaries of lifespan estimated from the dissection analysis, in terms of gonotrophic cycles and chronological time, respectively.
 
@@
 	\end{tabular}
-\caption{Posterior summaries of mean mosquito lifespan in terms of gonotrophic cycles from dissection studies.
-The 5\%, 25\%, 50\%, 75\% and 95\% columns indicate the respective quantiles of the posterior distribution for mean lifespan, and the `Mean' and `Std.
+\caption{\textbf{Posterior summaries of mean female mosquito lifespan in terms of gonotrophic cycles from dissection studies.} The 5\%, 25\%, 50\%, 75\% and 95\% columns indicate the respective quantiles of the posterior distribution for mean lifespan, and the `Mean' and `Std.
 dev.' columns indicate the posterior mean and standard deviation of mean lifespan.}
@@
 
-\subsubsection{Chronological lifespan estimates}
 \begin{figure}[h]
@@
 
-\begin{table}[htbp]
+\begin{table}[htbp!]
 	\begin{tabular}{l|l|l|l|l|l|l|l|l}
@@
 	\end{tabular}
-	\caption{Posterior summaries of mean mosquito lifespan in terms of chronological time from dissection studies.
-The 5\%, 25\%, 50\%, 75\% and 95\% columns indicate the respective quantiles of the posterior distribution for mean lifespan, and the `Mean' and `Std.
+	\caption{\textbf{Posterior summaries of mean female mosquito lifespan in terms of chronological time from dissection studies.} The 5\%, 25\%, 50\%, 75\% and 95\% columns indicate the respective quantiles of the posterior distribution for mean lifespan, and the `Mean' and `Std.
 dev.' columns indicate the posterior mean and standard deviation of mean lifespan.}
@@
 	\end{tabular}%
-	\caption{The results of comparisons between LBLs for those species with data for MRRs and dissection-based studies. The ANOVA tests were conducted using LBLs estimated at the time-series level.}
+	\caption{\textbf{The results of comparisons between LBLs for those species with data for MRRs and dissection-based studies.} The ANOVA tests were conducted using median posterior LBLs estimated at the time-series level.}
 	\label{tab:comparison}%

### 2018-11-24 MS.tex
Added top vs genus vs species elpd graph and changed a few estimates from medians to means

@@
 laboratory studies, we estimate that female mosquitoes outlive males by
-0.9 days on average (mean estimate; 25\%-75\% CI: 0.3-1.6 days). We
+1.2 days on average (mean estimate; 25\%-75\% CI: 0.3-1.6 days). We
 fit models incorporating mosquito senescence to the data, which allows
@@
 considerable variation. For example, there are 54 estimates of LBL for
-\emph{Ae. aegypti} which range from 2.2 days to 38.3 days with a mean of
-8.3 days and coefficient of variation of 0.7. There were significant
+\emph{Ae. aegypti} which range from 2.5 days to 42.1 days with a mean of
+11.4 days and coefficient of variation of 0.6 (all estimates are posterior mean). There were significant
 differences in LBL amongst species (ANOVA on median LBL controlling for
@@
 incorporation of a species term provides little predictive power (Fig.
-S\#; in part explained by the latter model over-fitting the data where
+\ref{S-fig:mrr_genusTopLevel}; in part explained by the latter model over-fitting the data where
 there are few time series per species).
@@
 females to live longer than males for each of the genera, with the
-difference largest for \emph{Aedes} (2.5 days; fraction of pairwise
+difference largest for \emph{Aedes} (2.9 days; fraction of pairwise
 posterior samples of females versus males where difference was less than
-zero, p\textless{}0.01), followed by \emph{Anopheles} (2.0 days; p=0.17)
-and \emph{Culex} (0.3 days; p=0.34). Overall, female mosquitoes were
-estimated to live 0.9 days longer than males (p=0.10).
+zero, p\textless{}0.01), followed by \emph{Anopheles} (2.2 days; p=0.17)
+and \emph{Culex} (0.2 days; p=0.34). Overall, female mosquitoes were
+estimated to live 1.2 days longer than males (p=0.10).
 
@@
 alternatively unfed (116). We estimate that female mosquitoes that were
-fed on sugar pre-release lived on average for 0.7 days longer than those
-that were not fed (Fig. \ref{S-fig:mrr_female_blood_sugar}; a pattern that was consistent across the
+fed on sugar pre-release lived on average for 1.0 days (posterior mean) longer than those
+that were not fed ($p>0.05$; Fig. \ref{S-fig:mrr_female_blood_sugar}; a pattern that was consistent across the
 genera). There were insufficient males that were either fed or unfed
 with sugar prior to release to make a meaningful comparison. Females
-that were blood-fed prior to release on average lived 1.5 days longer
+that were blood-fed prior to release on average lived 1.7 days (posterior mean) longer
 than those who were not fed for \emph{Aedes} but this trend was reversed
-for \emph{Anopheles} meaning that there was little difference overall.
+for \emph{Anopheles} meaning that there was little difference overall (0.15 days; posterior mean; $p=0.44$).
 
@@
 Table \ref{S-tab:dissection_estimated_lifespans_chron} provides posterior summaries of chronological for the species and genera in the dissection dataset (see also Fig.
-).
+\ref{S-fig:dissection_lifetimes_exponential_chron}).
 For ten species, we had enough data from both species to make a comparison, and there was a positive correlation (not
@@
 \emph{A. sergentii} was amongst the longest lived of the anopheline
-species with a posterior median LBL of 12.4 days (mean estimate;
+species with an LBL of 12.4 days (mean estimate;
 25\%-75\% CI: 5.9-13.8 days) from the MRR analysis and 11.9 days (mean
@@
 mosquitoes versus males than historically (Fig. \ref{S-fig:mrr_sexReleasesOverTime}). Our analysis
-indicates that females outlive male mosquitoes by approximately 0.9 days
+indicates that females outlive male mosquitoes by approximately 1.2 days
 (Fig. \ref{fig:mrr_sexDifferences_without_sugar_nor_blood}), meaning that differences between the sexes may exist for

### 2018-11-24 supplementary_materials.tex
Added top vs genus vs species elpd graph and changed a few estimates from medians to means

@@
 	\label{fig:mrr_female_blood_sugar}
+\end{figure}
+
+\begin{figure}[h]
+	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_genus_vs_species_vs_top.pdf}}
+	\caption{\textbf{The estimated predictive accuracy of the species- and overall-level models relative to the genus-level model.} The predictive accuracy was determined using K-Fold cross-validation as described in SOM; the measure of accuracy presented here by the central dots is the difference in estimated expected log pointwise predictive density compared to the exponential model.
+The lower and upper whiskers represent the lower and upper bounds of the 95\% confidence interval in predictive accuracy.}
+	\label{fig:mrr_genusTopLevel}
 \end{figure}

### 2018-11-28 MS.tex
Major reformatting following meeting with Ace

@@
 \externaldocument[S-]{supplementary_materials}
+\externaldocument[SM-]{supplementary_materials_methods}
 
@@
 important mosquito characteristic.
+
+\section{Author summary}
+
+\section{Author contributions}
+HCJG, AN and BL were involved in conceptualising this study.
+BL was responsible for data curation and the formal analysis of the data.
+BL and AN developed the statistical methodology and conducted the investigation.
+All authors were involved in drafting the original manuscript and revising it.
 
@@
 
-\section{Methods}\label{methods}
-
-In recent years many important vectors of disease have been shown to be
-complexes of closely related species, biotypes or forms that cannot
-be distinguished morphologically (for example the morphospecies
-\emph{Anopheles gambiae sensu lato} is now separated into the widespread
-\emph{gambiae, coluzzii, arabiensis} and a number of more local
-species). As the majority of studies analysed here took place before
-molecular techniques allowed these taxa to be separated we work here
-chiefly with morphospecies.
-
-\subsection{Mark-release-recapture}\label{mark-release-recapture}
-
-Data from MRR experiments in the Guerra et al. (2014) database were
-examined and those with fewer than six recaptures and species with only
-a single MRR study were excluded for the hierarchical analysis. Of the
-232 data sets, 179 involved only females, 35 males, and 18 both sex
-releases. For 102 data sets the age of the released mosquitoes was known
-(the average age of released mosquitoes was 4.0 days) while in the other
-cases it was unknown or unrecorded; in these cases we assumed the
-mosquitoes were newly emerged at the time of release and return to this
-assumption later. See Table \ref{S-tab:mrr_aggregateData} for a summary of other data characteristics.
-
-We analysed all MRR experiments within the same statistical framework
-(for full details see the Supplementary Online Material (SOM)). In the
-simplest case \emph{N\textsubscript{R}} mosquitoes are released on day
-zero and the probability that they remain in the recapture area until
-day \emph{t} is \emph{S}(\emph{t}) when they are recaptured with
-probability $\psi$. We model the number of mosquitoes recaptured on
-day \emph{t} using a negative binomial sampling model with mean
-\(\left( N_{R} - Y\left( t - 1 \right) \right)S\left( t \right)\psi\),
-where \(Y\left( t - 1 \right)\) is cumulative captures before day
-\emph{t}, and shape parameter $\kappa$. The negative binomial has been
-used previously in analyses of mosquito count data \citep{service1971studies,nedelman1983negative} because of its ability to
-represent temporal over-dispersion in recaptures most likely caused by
-variable weather. A slight modification was required for studies with
-multiple releases (see SOM).
-
-The simplest model for $S(t)$ assumes there is a constant
-probability ($\lambda$) that a mosquito dies or leaves the recapture area
-so that the numbers remaining after time \emph{t} are given by the
-exponential distribution, $\text{exp}( -\lambda t )$.
-We utilised this form extensively but in testing for senescence used
-five other models where $\lambda(t)$ varies with time so that,
-
-\begin{longtable}[]{@{}ll@{}}
-\(S\left( t \right) = e^{- \int_{0}^{t}{\lambda\left( \tau \right) \mathrm{d}\tau}}.\)
-\end{longtable}
-
-Details of the five models (Gompertz, Weibull, Gompertz-Makeham,
-Logistic and Logistic-Makeham), which vary in their ability to detect
-different forms of age-dependent mortality, are given in the SOM. Using
-multiple different types of models increased our chances of detecting
-senescence though, as discussed below, also increases the likelihood of
-false positives.
-
-Parameters were estimated using Bayesian techniques with relative
-uninformative priors for $\kappa$ and the parameters of
-$\lambda(t)$, but assuming a prior for $\psi$ indicating a low
-recapture probability (bounded in part by knowledge of the maximum daily
-recapture rates; see SOM). We used a Bayesian hierarchical model to estimate
-distributions of lifespan at the species and the genus levels, and
-across the complete data set. This procedure assumes that there is a
-distribution of lifespan parameters for each species from which those
-governing individual MRR time series are sampled, and similarly a
-distribution at the genus level from which those for individual species
-are derived (rather akin to random effects in classical statistics).
-Within this framework we can also allow the parameters for individual
-time series to be influenced by co-variates such as differences in
-experimental methodology. As in the estimation of the parameters of the
-individual experiments, relative uninformative priors were set for the
-parameters of the hierarchical models except for $\psi$ where again a
-distribution representing low recapture probabilities was assumed.
-Posterior distributions were derived using Markov Chain Monte Carlo
-(MCMC) methods with convergence assessed using the \(\hat{R}\) statistic
-\citep{gelman1992inference}. The predictive power of the model was assessed
-using \emph{K}-fold cross validation which tests the ability of the
-model fitted to part of the data to predict the rest using multiple
-different partitions. Further details of the prior specification,
-fitting and validation through posterior predictive checks \citep{lambert2018student} are given in
-the SOM.
-
-Two studies of \emph{Anopheles balabacensis} reported capture rates
-increasing with time, presumably reflecting a violation of our
-assumption of constant recapture probabilities. We omitted this species
-from the analysis.
-
-The Guerra et al., (2014) database included the latitude and longitude of each
-study along with the date when the study began. We used this information to find
-estimates of the air temperature for each study using the European Centre for Medium
-Range Weather Forecasts' ERA Interim Daily historical database. For each
-study we calculated the mean monthly temperature across a spatial area
-of (latitude $\pm$ 1 degree, longitude $\pm$ 1 degree), for the month at which
-each study was carried out. The records for this database begin in 1979,
-which pre-dates the study date for 65 of our 232 MRR time-series. For
-these time-series, we chose to estimate the air temperature by an
-average of the corresponding monthly temperatures over the years
-1979-89.
-
-\subsection{Dissection}\label{dissection}
-
-Studies using dissection to estimate mosquito longevity were located in
-literature databases using relevant keyword, citation and author
-searches, and by checking previous studies cited by the papers located
-(see SOM). The list of studies located with associated metadata is
-available as a Supplementary Online File.
-
-Most dissection studies recorded the distribution of the number of
-gonotrophic cycles in mosquito samples collected over a specific period
-of time. Overall, we found 568 time series in 72 published articles.
-Because seldom were sufficient ancillary data available to analyse
-trends over time, we aggregated the data from each time series. We
-further omitted time series with fewer than 100 mosquitoes and for
-species with only one data set leaving 131 studies of mosquitoes in the
-genera \emph{Anopheles, Aedes}, \emph{Culex} and \emph{Mansonia}.
-
-To compare lifespan estimates from dissection and MRR studies we need to
-convert physiological age (the number of gonotrophic cycles) into
-chronological age. Using a literature search and a review by Silver
-(2007) we found 79 estimates in 42 published articles. Most estimates
-were obtained by dissecting females recaptured in MRR studies or by
-observations in the laboratory, the latter tending to give longer
-durations. Studies differed greatly in how (if at all) they represented
-uncertainty in their estimate of the duration of the gonotrophic cycle.
-Where confidence limits were given we treated these as the relevant
-quantiles of a normal distribution, where a range was stated (e.g. ``4-6
-days'') we interpreted the bounds as the 2.5\% and 97.5\% quantiles of a
-normal distribution, and where a single figure was quoted we assumed
-this was the mean this distribution. Using the quantiles of the normal
-distribution, we estimated its mean and standard deviation by regression
-(see SOM). Initially we calculated distributions of gonotrophic cycle
-lengths at the species and then genus levels, but because of the paucity
-of data for many species and the lack of significant differences we
-aggregated the data into a single distribution. We converted
-physiological age to chronological age by sampling from this
-distribution to obtain a particular gonotrophic cycle length for each
-mosquito (we also explored sampling from this distribution to obtain the
-duration of \emph{each} gonotrophic cycle which increased the
-uncertainty in lifespan estimate but did not affect any of the
-conclusions).
-
-We modelled the number of mosquitoes found by dissection to be of age
-\emph{a} using the negative binomial distribution with mean
-$\Psi S(a)$ and shape parameter $\kappa$, where
-$\Psi$ is the product of the recruitment rate of adult mosquitoes,
-which we assume is constant over time, and the probability of being
-captured for dissection, and \emph{S}(\emph{a}) is the probability of
-surviving until age \emph{a}. We used the number of females that have
-yet to lay eggs (nulliparous) to estimate the recruitment rate as
-described further in the SOM. Initial examination revealed that in some
-data sets the number of nulliparous females was anomalously low,
-something that has been noticed before \citep{gillies1965study}. As some
-studies have suggested that the first gonotrophic cycle tends to be
-longer than the subsequent ones, this is probably due to differences in
-capture probability. In data sets where the fraction of nulliparous
-females was less than 90\% the uniparous (completed on gonotrophic
-cycle) we excluded the nulliparous observation.
-
-Data was analysed using a Bayesian framework as with the MRR data with
-minor differences in the specification of the priors (see SOM). Some
-published studies do not distinguish the number of gonotrophic studies
-beyond a threshold (the more ovariole dilations there are the harder it
-is to count them) which is akin to censoring the data. Because this
-censoring involves only a relatively small number of mosquitoes in each
-time series (median = 2\%), and because of the technical difficulties of
-allowing for censoring in our Bayesian estimation procedures, we assume
-the females died at the threshold age.
-
 \section{Results}\label{results}
@@
 temperature and LBL (overall or within genus) for the 238 datasets we
-analysed (Fig. \ref{fig:mrr_temperature}). This result held if, instead of pooling results from
+analysed (Fig. \ref{S-fig:mrr_temperature}). This result held if, instead of pooling results from
 all time series, we considered the four species with the most data
@@
 
-\subsection{Number of gonotrophic cycles estimates from
-dissection}\label{number-of-gonotrophic-cycles-estimates-from-dissection}
+\subsection{Number of gonotrophic cycles estimates from dissection}\label{number-of-gonotrophic-cycles-estimates-from-dissection}
 
@@
 was 1.3 (posterior mean; Fig. \ref{fig:dissection_lifetimes_exponential}; Table \ref{S-tab:dissection_estimated_lifespans}) and across the 131 studies, 95\% of
-the individual time series estimates were less than 3 (Fig. \ref{fig:dissection_individualEstimates_allSpecies}). The
+the individual time series estimates were less than 3 (Fig. \ref{S-fig:dissection_individualEstimates_allSpecies}). The
 estimated greatest number of cycles was for \emph{Anopheles sergentii}
@@
 
-To compare the two methods, we converted numbers of gonotrophic cycle
+To compare the two methods, we converted numbers of gonotrophic cycles
 (physiological age) into lifespan (chronological age) as described in
@@
 `senescence parameter' $\beta$ of the Gompertz survival function (see Table
-\ref{S-tab:mrr_survivalDescription}) for case study populations with three different levels of
+\ref{SM-tab:mrr_survivalDescription}) for case study populations with three different levels of
 senescence (Fig. \ref{S-fig:mrr_mcPowerAnalysis_senescence}A). This analysis indicated that power to detect
@@
 a study length of at least 18 days. Since the median study duration for
-experiments included in our analysis was 10 days (Table \ref{S-tab:mrr_IndividualData}) this could
+experiments included in our analysis was 10 days (Table \ref{SM-tab:mrr_IndividualData}) this could
 partly explain our failure to detect senescence at the species level. A
@@
 considered, it is now less common for MRR experiments to release female
-mosquitoes versus males than historically (Fig. \ref{S-fig:mrr_sexReleasesOverTime}). Our analysis
+mosquitoes versus males than historically (Fig. \ref{SM-fig:mrr_sexReleasesOverTime}). Our analysis
 indicates that females outlive male mosquitoes by approximately 1.2 days
@@
 within this, we determined no relationship between lifespan and
-temperature across all time series (Fig. \ref{fig:mrr_temperature}) or, for any of the species
+temperature across all time series (Fig. \ref{S-fig:mrr_temperature}) or, for any of the species
 with the most data (Fig. \ref{S-fig:mrr_ThreeSpeciesVersusTemperature}). It is possible that by considering a raw
@@
 
+\section{Methods}\label{methods}
+
+In recent years many important vectors of disease have been shown to be
+complexes of closely related species, biotypes or forms that cannot
+be distinguished morphologically (for example the morphospecies
+\emph{Anopheles gambiae sensu lato} is now separated into the widespread
+\emph{gambiae, coluzzii, arabiensis} and a number of more local
+species). As the majority of studies analysed here took place before
+molecular techniques allowed these taxa to be separated we work here
+chiefly with morphospecies.
+
+\subsection{Mark-release-recapture}\label{mark-release-recapture}
+
+Data from MRR experiments in the Guerra et al. (2014) database were
+examined and those with fewer than six recaptures and species with only
+a single MRR study were excluded for the hierarchical analysis. Of the
+232 data sets, 179 involved only females, 35 males, and 18 both sex
+releases. For 102 data sets the age of the released mosquitoes was known
+(the average age of released mosquitoes was 4.0 days) while in the other
+cases it was unknown or unrecorded; in these cases we assumed the
+mosquitoes were newly emerged at the time of release and return to this
+assumption later. See Table \ref{SM-tab:mrr_aggregateData} for a summary of other data characteristics.
+
+We analysed all MRR experiments within the same statistical framework
+(for full details see the Supplementary Online Material (SOM)). In the
+simplest case \emph{N\textsubscript{R}} mosquitoes are released on day
+zero and the probability that they remain in the recapture area until
+day \emph{t} is \emph{S}(\emph{t}) when they are recaptured with
+probability $\psi$. We model the number of mosquitoes recaptured on
+day \emph{t} using a negative binomial sampling model with mean
+\(\left( N_{R} - Y\left( t - 1 \right) \right)S\left( t \right)\psi\),
+where \(Y\left( t - 1 \right)\) is cumulative captures before day
+\emph{t}, and shape parameter $\kappa$. The negative binomial has been
+used previously in analyses of mosquito count data \citep{service1971studies,nedelman1983negative} because of its ability to
+represent temporal over-dispersion in recaptures most likely caused by
+variable weather. A slight modification was required for studies with
+multiple releases (see SOM).
+
+The simplest model for $S(t)$ assumes there is a constant
+probability ($\lambda$) that a mosquito dies or leaves the recapture area
+so that the numbers remaining after time \emph{t} are given by the
+exponential distribution, $\text{exp}( -\lambda t )$.
+We utilised this form extensively but in testing for senescence used
+five other models where $\lambda(t)$ varies with time so that,
+
+\begin{longtable}[]{@{}ll@{}}
+	\(S\left( t \right) = e^{- \int_{0}^{t}{\lambda\left( \tau \right) \mathrm{d}\tau}}.\)
+\end{longtable}
+
+Details of the five models (Gompertz, Weibull, Gompertz-Makeham,
+Logistic and Logistic-Makeham), which vary in their ability to detect
+different forms of age-dependent mortality, are given in the SOM. Using
+multiple different types of models increased our chances of detecting
+senescence though, as discussed below, also increases the likelihood of
+false positives.
+
+Parameters were estimated using Bayesian techniques with relative
+uninformative priors for $\kappa$ and the parameters of
+$\lambda(t)$, but assuming a prior for $\psi$ indicating a low
+recapture probability (bounded in part by knowledge of the maximum daily
+recapture rates; see SOM). We used a Bayesian hierarchical model to estimate
+distributions of lifespan at the species and the genus levels, and
+across the complete data set. This procedure assumes that there is a
+distribution of lifespan parameters for each species from which those
+governing individual MRR time series are sampled, and similarly a
+distribution at the genus level from which those for individual species
+are derived (rather akin to random effects in classical statistics).
+Within this framework we can also allow the parameters for individual
+time series to be influenced by co-variates such as differences in
+experimental methodology. As in the estimation of the parameters of the
+individual experiments, relative uninformative priors were set for the
+parameters of the hierarchical models except for $\psi$ where again a
+distribution representing low recapture probabilities was assumed.
+Posterior distributions were derived using Markov Chain Monte Carlo
+(MCMC) methods with convergence assessed using the \(\hat{R}\) statistic
+\citep{gelman1992inference}. The predictive power of the model was assessed
+using \emph{K}-fold cross validation which tests the ability of the
+model fitted to part of the data to predict the rest using multiple
+different partitions. Further details of the prior specification,
+fitting and validation through posterior predictive checks \citep{lambert2018student} are given in
+the SOM.
+
+Two studies of \emph{Anopheles balabacensis} reported capture rates
+increasing with time, presumably reflecting a violation of our
+assumption of constant recapture probabilities. We omitted this species
+from the analysis.
+
+The Guerra et al., (2014) database included the latitude and longitude of each
+study along with the date when the study began. We used this information to find
+estimates of the air temperature for each study using the European Centre for Medium
+Range Weather Forecasts' ERA Interim Daily historical database. For each
+study we calculated the mean monthly temperature across a spatial area
+of (latitude $\pm$ 1 degree, longitude $\pm$ 1 degree), for the month at which
+each study was carried out. The records for this database begin in 1979,
+which pre-dates the study date for 65 of our 232 MRR time-series. For
+these time-series, we chose to estimate the air temperature by an
+average of the corresponding monthly temperatures over the years
+1979-89.
+
+\subsection{Dissection}\label{dissection}
+
+Studies using dissection to estimate mosquito longevity were located in
+literature databases using relevant keyword, citation and author
+searches, and by checking previous studies cited by the papers located
+(see SOM). The list of studies located with associated metadata is
+available as a Supplementary Online File.
+
+Most dissection studies recorded the distribution of the number of
+gonotrophic cycles in mosquito samples collected over a specific period
+of time. Overall, we found 568 time series in 72 published articles.
+Because seldom were sufficient ancillary data available to analyse
+trends over time, we aggregated the data from each time series. We
+further omitted time series with fewer than 100 mosquitoes and for
+species with only one data set leaving 131 studies of mosquitoes in the
+genera \emph{Anopheles, Aedes}, \emph{Culex} and \emph{Mansonia}.
+
+To compare lifespan estimates from dissection and MRR studies we need to
+convert physiological age (the number of gonotrophic cycles) into
+chronological age. Using a literature search and a review by Silver
+(2007) we found 79 estimates in 42 published articles. Most estimates
+were obtained by dissecting females recaptured in MRR studies or by
+observations in the laboratory, the latter tending to give longer
+durations. Studies differed greatly in how (if at all) they represented
+uncertainty in their estimate of the duration of the gonotrophic cycle.
+Where confidence limits were given we treated these as the relevant
+quantiles of a normal distribution, where a range was stated (e.g. ``4-6
+days'') we interpreted the bounds as the 2.5\% and 97.5\% quantiles of a
+normal distribution, and where a single figure was quoted we assumed
+this was the mean this distribution. Using the quantiles of the normal
+distribution, we estimated its mean and standard deviation by regression
+(see SOM). Initially we calculated distributions of gonotrophic cycle
+lengths at the species and then genus levels, but because of the paucity
+of data for many species and the lack of significant differences we
+aggregated the data into a single distribution. We converted
+physiological age to chronological age by sampling from this
+distribution to obtain a particular gonotrophic cycle length for each
+mosquito (we also explored sampling from this distribution to obtain the
+duration of \emph{each} gonotrophic cycle which increased the
+uncertainty in lifespan estimate but did not affect any of the
+conclusions).
+
+We modelled the number of mosquitoes found by dissection to be of age
+\emph{a} using the negative binomial distribution with mean
+$\Psi S(a)$ and shape parameter $\kappa$, where
+$\Psi$ is the product of the recruitment rate of adult mosquitoes,
+which we assume is constant over time, and the probability of being
+captured for dissection, and \emph{S}(\emph{a}) is the probability of
+surviving until age \emph{a}. We used the number of females that have
+yet to lay eggs (nulliparous) to estimate the recruitment rate as
+described further in the SOM. Initial examination revealed that in some
+data sets the number of nulliparous females was anomalously low,
+something that has been noticed before \citep{gillies1965study}. As some
+studies have suggested that the first gonotrophic cycle tends to be
+longer than the subsequent ones, this is probably due to differences in
+capture probability. In data sets where the fraction of nulliparous
+females was less than 90\% the uniparous (completed on gonotrophic
+cycle) we excluded the nulliparous observation.
+
+Data was analysed using a Bayesian framework as with the MRR data with
+minor differences in the specification of the priors (see SOM). Some
+published studies do not distinguish the number of gonotrophic studies
+beyond a threshold (the more ovariole dilations there are the harder it
+is to count them) which is akin to censoring the data. Because this
+censoring involves only a relatively small number of mosquitoes in each
+time series (median = 2\%), and because of the technical difficulties of
+allowing for censoring in our Bayesian estimation procedures, we assume
+the females died at the threshold age.
+
 \section{Acknowledgements}\label{acknowledgements}

### 2018-11-28 figures.tex
Major reformatting following meeting with Ace

@@
 \begin{figure}[h]
-	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_lifeSpanVsTemperature.pdf}}
-	\caption{\textbf{Posterior estimates of mean mosquito lifespan for each time-series versus the average monthly temperature for each study location and date.} The markers show the median posterior estimates, with the lower and upper bounds indicating the 25\% and 75\% quantiles, respectively.
-The black line shows a regression line with linear and quadratic terms estimated using the median posterior lifetimes, with the grey shading indicating 95\% confidence intervals.
-All estimates of lifespan were obtained using the non-hierarchical exponential survival model.}\label{fig:mrr_temperature}
-\end{figure}
-
-\begin{figure}[h]
 	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/dissection_lifetimes_exponential.pdf}}
@@
 	\label{fig:dissection_lifetimes_exponential}
-\end{figure}
-
-\begin{figure}[h]
-	\centerline{\includegraphics[width=1.3\textwidth]{./Figure_files/dissection_individualEstimates_allSpecies.pdf}}
-	\caption{\textbf{Individual time-series estimates of adult mosquito mean lifespan ordered by species median as determined from the dissection data.} The middle line in each box shows the median estimates.
-The left and right box whiskers show the 25\%, and 75\% posterior quantiles respectively.
-All estimates were obtained using the non-hierarchical exponential survival model.}
-	\label{fig:dissection_individualEstimates_allSpecies}
 \end{figure}
@@
 
- \bibliographystyle{authordate1}
- \bibliography{Malaria}
-\newpage
-
 \end{document}

### 2018-11-28 supplementary_materials.tex
Major reformatting following meeting with Ace

@@
 \externaldocument{figures}
+\externaldocument{supplementary_materials_methods}
 
@@
 % \raggedright
-\title{A Meta-analysis of Longevity Estimates of Mosquito Vectors of Disease: Supplementary Material}
+\title{A Meta-analysis of Longevity Estimates of Mosquito Vectors of Disease: Supplementary Figures}
 \author{Ben Lambert, Ace North, Charles Godfray}
@@
 
-\section{MRR experiments}
-\subsection{Data}
-The MRR database in \cite{guerra2014global} contains 393 individual time-series, along with meta-data for a range of factors for each experiment (for example, species, geography, and date of study).
-We  cleaned the data, making a number of amendments to it (see Section \ref{sec:amendments}), and transformed the data into a form amenable for our statistical model.
-We removed time-series with fewer than six separate recapture observations, resulting in 238 time series.
-For the species-level analyses, we required at least two time series per species, resulting in loss of six series, meaning that the data used in the Bayesian hierarchical analysis comprised 232 individual time series (see Section \ref{sec:appendix_mrrStudyList} for a list of the studies included).
-Tables \ref{tab:mrr_aggregateData} and \ref{tab:mrr_IndividualData} summarise the data used in the Bayesian hierarchical analysis.
-This data encompassed time-series from MRR experiments across 33 different species and three genera: \textit{Aedes} (91 separate time-series), \textit{Anopheles} (94), and \textit{Culex} (47), and spanning a wide geographical range (Fig.
-\ref{fig:mrr_lifetimes_map}).
-
-\subsection{Amendments by MRR ID}\label{sec:amendments}
-This section describes the amendments made by MRR ID (the unique identifier that \cite{guerra2014global} use for each MRR record).
-Where there was ambiguity in the data, we attempted to consult the original publication.
-The amendments to the dataset are described below:
-
-\begin{itemize}
-	\item 69 and 70: Since the release was not disaggregated by sex, the recapture data were aggregated across both males and females, resulting in a single time series in each case.
-	\item 482: There are actually two series (one with captured mosquitoes, the other with young reared) in the paper \cite[]{rawlings1981influence}, whereas the database contains only one recapture series.
-In this analysis, both time series are included.
-	\item 483: There are three studies in the paper \cite[]{rawlings1981influence}, (by dye colour there is one for `M\&Y', another for white, and a third for green), whereas the database only contains one release series.
-In this analysis, we analyse all three time series.
-	\item 578: A null observation was removed on day 12, since isn't present in the original paper \cite[]{eyles1943experiment}.
-	\item 585: It is not clear from the original paper \cite[]{smith1941observations} whether the missing recapture observations on certain days are because no mosquitoes were found on those days, or because no efforts were made to recapture them on those days.
-We have assumed that the number recaptured were as stated in the database on those days (missing, because no recapture efforts were made on those days).
-	\item 586: It was unclear from database whether the missing observations are due to lack of recapture, or no effort.
-Have assumed that zero mosquitoes were captured on these days as seems most likely (the paper could not be accessed).
-	\item 342: Changed number of released from 6,000 to 12,000 following the original publication \cite[]{bryan1991anopheles}.
-	\item 246: Moved the observations from captures to recaptures following the original publication \cite[]{arredondo1998gonotrophic}.
-	\item 55: Corrected a mistake in release data; the actual number released is 739, not 749 \cite[]{midega2007estimating}.
-	\item 478-481: All four of these series give releases not disaggregated by gender, whereas the recaptures are split this way.
-Have combined the recapture series to yield a single release-recapture series in each case.
-	\item 154: Moved captured male series to recaptured following original paper \cite[]{tsuda2001movement}. 
-	\item 334: Data starts on day 3, thus have added this number of days to our time variable in this case.
-\end{itemize}
-
-The R file \verb|s_clean_data.R| attached cleans the data, makes the above amendments and outputs the data in a form amenable for estimation using Stan \citep{carpenter2016stan}.
-
-\begin{figure}[h]
-	\centerline{\includegraphics[width=1.25\textwidth]{./Figure_files/mrr_mapGenusCropped.pdf}}
-	\caption{\textbf{The location and number of separate time-series for the MRR studies included in this analysis.} The area of each circle indicates the number of time-series at each study site.
-The colour shows the genus of the mosquitoes in each study.}
-	\label{fig:mrr_lifetimes_map}
-\end{figure}
-
-\begin{table}[htbp]
-	\centering
-	\begin{tabular}{lc}
-		\toprule
-		\textbf{Data characteristic} & \textbf{Count} \\
-		\midrule
-		Species & 33 \\
-		Genera & 3 \\
-		\textit{Anopheles} & 94 \\
-		\textit{Aedes} & 91 \\
-		\textit{Culex} & 47 \\
-		Female-only MRR & 179 \\
-		Male-only MRR & 35 \\
-		Mixed sex MRR & 18 \\
-		Pre-release blood-feeding only MRR & 71 \\
-		Pre-release sugar-feeding only MRR & 41 \\
-		Pre-release both blood- and sugar-feeding MRR & 4 \\
-		Pre-release neither blood- and sugar-feeding MRR & 116 \\
-		MRR time-series & 232 \\
-		\bottomrule
-	\end{tabular}%
-	\caption{\textbf{Summary of variables across all MRR time-series.} For the `Species' and `Genera' rows, the counts represent the number of distinct entries in the database; otherwise the counts represent the number of time series.}
-	\label{tab:mrr_aggregateData}%
-\end{table}%
-
-\begin{table}[htbp]
-	\centering
-	\footnotesize
-	\begin{adjustwidth}{-0.5in}{-0.5in}%
-		\begin{tabularx}{1.25\textwidth}{l|ccccc}
-			\toprule
-			\textbf{Data characteristic} & \textbf{Min} & \textbf{Mean} & \textbf{Median} & \textbf{Max} & \textbf{Standard deviation} \\
-			\midrule
-			Study duration, days & 6.0   & 11.8  & 10.0  & 71.0    & 4.7 \\
-			Number of days on which collections took place & 6.0   & 10.1  & 9.0   & 47    & 9.1 \\
-			Number of separate release days & 1.0     & 1.9   & 1.0     & 23    & 3.0 \\
-			Number released & 66    & 4,929  & 1,297  & 86,200 & 12,043 \\
-			Number recaptured & 2     & 163   & 63    & 4,090  & 399 \\
-			Recapture percentage & 0\% & 8.6\% & 5.2\% & 57.1\% & 10.04\% \\
-			Age at release, days & 0   & 1.8   & 0   & 13    & 2.8 \\
-			\bottomrule
-		\end{tabularx}%
-		\caption{\textbf{Summary of data from individual MRR time-series.}}
-		\label{tab:mrr_IndividualData}%
-	\end{adjustwidth}
-\end{table}%
-
-\begin{figure}[h]
-	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_sexReleasesOverTime.pdf}}
-	\caption{\textbf{The number of MRR time-series by date and sex of released mosquitoes.} The numbers represent by-decade totals of the 232 time-series we used in the main analysis.}
-	\label{fig:mrr_sexReleasesOverTime}
-\end{figure}
-
-\subsection{Statistical model}\label{sec:mrr_statistical}
-Data for a typical MRR experiment consists of a single release of $N_{R}$ marked mosquitoes followed by a series of marked mosquito recaptures on subsequent days (Fig.
-\ref{fig:mrr_exampleMRRSeries} shows three such example series).
-We model the number $y(t)$ of marked mosquitoes recaptured on day $t$ using a negative binomial sampling model,
-%
-\begin{equation}\label{eq:NB}
-y(t) \sim \text{NB}\left((N_{R} - Y(t-1)) S(t) \psi, \kappa\right),
-\end{equation}
-%
-where $Y(t-1)$ is the cumulative number of mosquitoes caught on all days before $t$, $S(t)$ is the probability that an individual mosquito survives and remains in the study area until time $t$, $\psi$ is the daily recapture probability for an individual mosquito which is assumed constant through time, and $\kappa$ is the time-independent shape parameter of the negative binomial distribution that controls the extent of variance in recapture rate likely due mostly to environmental heterogeneity.
-We chose a parameterisation of the negative binomial such that its mean is given by $\mu(t) = (N_{R} - Y(t-1)) S(t) \psi$, and its variance by $\sigma(t)^2 = \mu(t) + \frac{\mu(t)^2}{\kappa}$.
-
-For some MRR experiments, there were a number of releases ($q\geq 2$) of marked mosquitoes throughout the duration of the study.
-In contrast to single releases, with multiple releases over time, it is not in general possible to determine the particular release to which a recaptured mosquito belongs.
-To avoid the complication of directly inferring this quantity, we choose to represent previous recaptures probabilistically.
-This results in a slightly different mean to that of the single release model,
-%
-\begin{equation}
-\mu(t) = N_{Released}(1-\psi)^{t-1} S(t) \psi,
-\end{equation}
-%
-where the factor of $(1-\psi)^{t-1}$ represents the probability that a mosquito is \textit{not} recaptured on a previous day.
-Where experiments consisted of two or more releases occurring at distinct points in time we assumed that recaptures of individual mosquitoes from either batch were independent of one another, although with the same sampling parameters ($\psi$ and $\kappa$).
-This results in an overall mean composed of the sum of those from all $q$ releases,
-%
-\begin{equation}
-\mu(t) = \mu_1(t) + ... + \mu_q(t).
-\end{equation}
-%
-
-\begin{figure}[h]
-	\centerline{\includegraphics[width=1.25\textwidth]{./Figure_files/mrr_exampleMRRSeries.pdf}}
-	\caption{\textbf{Example MRR time-series for three different studies from the \cite{guerra2014global} database.} In all the above experiments there was a single release of a batch of marked mosquitoes.}
-	\label{fig:mrr_exampleMRRSeries}
-\end{figure}
-
-The simplest assumption is that death and dispersal rates are independent of mosquito age, giving an exponential `survival' function,
-%
-\begin{equation}
-S(t) = e^{-\lambda t},
-\end{equation}
-%
-where $\lambda$ is the sum of the rates of death and dispersal from the study area.
-To estimate the lifespan of mosquitoes we use this model, partly because of the limited evidence we determine in favour of senescence (Figs.
-\ref{fig:mrr_elpd} and \ref{fig:dissection_elpd}).
-
-We assume that the number of mosquitoes recaptured on one day is independent of the number recaptured on any other day, apart from their joint dependence on $\lambda$, $\psi$ and $\kappa$.
-This results in a likelihood of the data of the form,
-%
-\begin{equation}
-\mathcal{L}(y(t_1),y(t_2),...,y(t_R)|\lambda,\psi,\kappa) = \prod\limits_{i=1}^{R} p(y(t_i)|\psi,\lambda,\kappa),
-\end{equation}
-%
-where $R$ is the number of individual days where collections took place, and $p(y(t_i)|\psi,\lambda,\kappa)$ is the probability of recapturing $y(t_i)$ mosquitoes on day $t_i$ determined from a negative binomial distribution with parameters $\mu = e^{-\lambda t} \psi$, and $\kappa$.
-
-\subsection{Individual time-series estimates}\label{sec:MRR_individual_analysis}
-We first treat each time-series separately, and estimate individual $(\lambda,\psi,\kappa)$ parameters of the statistical model (described in the previous section) for each time-series.
-To use a Bayesian methodology we must specify priors on this set of parameters.
-Here we choose to specify independent priors on each parameter of the form: $\lambda\sim \mathcal{N}(-2.32,1)$, $\psi\sim \text{exp}(50)$ (with $\psi$ constrained to be less than 0.1, a threshold below which the majority of daily recapture probabilities in the data are observed) and $\kappa\sim \textnormal{log-normal}(2,1)$.
-This prior on the rate parameter of the exponential distribution ($\lambda$) corresponds to a wide range of possible lifespans (Fig.
-\ref{fig:mrr_individualTimeSeries_priors}A), with a mean of 10 days.
-It was necessary to use an informative prior on $\psi$ to bound it to sensible values and to prevent the correlation of estimates with $\lambda$ leading to unreasonably short estimates of $\lambda$; the prior on $\kappa$ is fairly uninformative and allows a wide range of values (Fig.
-\ref{fig:mrr_individualTimeSeries_priors}B,C).
-
-\begin{figure}[h]
-	\centerline{\includegraphics[width=1.25\textwidth]{./Figure_files/mrr_individualTimeSeries_priors.pdf}}
-	\caption{\textbf{The prior probability distributions used for lifespan (A), recapture probability (B) and dispersal parameter (C) in the individual time-series analysis for the MRR studies.} Note: the prior for the recapture probability was truncated at $\psi=0.1$ since most MRR day one recapture fractions were significantly below this threshold.}
-	\label{fig:mrr_individualTimeSeries_priors}
-\end{figure}
-
-\subsection{Estimating lifespan at the species, genus and overall groupings}\label{sec:MRR_hierarchical}
-We synthesise information from across all MRR experiments to produce more robust estimates of mosquito lifespan than can be obtained from considering the individual time-series separately.
-However there exists considerable heterogeneity across the experiments.
-This heterogeneity has two sources.
-There is that arising from variability in experimental methodology.
-There is also variability from actual differences in lifespan across the different mosquito cohorts; for example due to genetic differences between mosquito populations or due to climatic differences.
-Because of this heterogeneity we use a Bayesian hierarchical model which is akin to a `random effects' model in classical statistics.
-This type of model assumes that there is random variation in parameters at the individual time-series level, although each of the parameters is drawn from a common `population-level' distribution.
-In our case we separately estimate three different model groupings where the population-level distributions correspond to the species, genus and overall (across all studies) levels respectively (Fig.
-\ref{fig:mrr_methodDiagramHierarchy}).
-This allows us to estimate average mosquito lifespan for a given species or genus, or alternatively across all time-series, by independent sampling from the posterior predictive distribution from which the individual $\lambda_i$ are drawn.
-
-\begin{figure}[h]
-	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_methodDiagramHierarchy.pdf}}
-	\caption{\textbf{A representation of the hierarchical Bayesian model, for the constant mortality model case when considering the species-level grouping.} Here `pdf' indicates `probability density function', and $\lambda_i$, $\psi_i$, and $\kappa_i$ represent the force of mortality for the exponential distribution, daily recapture probability and over-dispersion parameter of the negative binomial distribution for time-series $i$ in the dataset; $\mu_i(t)= \psi_i e^{-\lambda_i t}$ is the mean number of mosquitoes recaptured on day $t$.}
-	\label{fig:mrr_methodDiagramHierarchy}
-\end{figure}
-
-Whilst we assume that the variation in mosquito mortality across experiments is random in nature, we recognise that there may exist systematic sources of variation that we do not include in our model.
-To account for this potential source of bias we examined how the individual estimates of lifespan correlated with experimental covariates (trapping method and number or density of traps).
-The data on trapping method in the Guerra et al., (2014) database is less reliable than other data entries because published MRR studies may use multiple trapping methods and typically do not disaggregate captures by method.
-When we examined the correlation of trapping method with lifespan estimates, however, we did not find systematic variation by method (data not shown).
-We determined a positive correlation between trapping density and lifespan estimates (Fig.
-\ref{fig:mrr_lifeSpanVsTrapDensity}), which indicates the impact of mosquito dispersal on estimated lifespan.
-
-A Bayesian framework requires that we incorporate our pre-analysis beliefs into our estimates through the use of priors for all parameters in a particular model.
-In our hierarchical model we are required to specify prior distributions at two levels of the analysis.
-The first of these links the individual time-series estimates with the overarching group-level distributions.
-The parameters of these group-level distributions are then set prior distributions (Fig.
-\ref{fig:mrr_methodDiagramHierarchy}).
-As an example, we suppose that the rate parameter $\lambda_i$ of our exponential survival model in experiment $i$ is given by:
-%
-\begin{align}
-\lambda_i = \text{exp}(&c_i + \delta^m_{\text{genus}[i]} \text{sex}^{\text{m}}_i + \delta^{\text{mix}}_{\text{genus}[i]} \text{sex}^{\text{mix}}_i +\\  & \delta^{\text{sugar}}_{\text{genus}[i]} \text{sugar}_i +  \delta^{\text{blood}}_{\text{genus}[i]} \text{blood}_i),
-\end{align}
-%
-where $\delta^m_{\text{genus}[i]}$ measures a genus-specific effect of male-only releases ($\text{sex}^{\text{m}}_i=1$) versus female-only ($\text{sex}^{\text{m}}_i=0$); $\delta^{\text{mix}}_{\text{genus}[i]}$ is the analogous effect for mixed releases versus female-only; $\delta^{\text{sugar}}_{\text{genus}[i]}$ is the analogous effect for mosquitoes that were sugar-fed prior to release versus those that were not fed; and $\delta^{\text{blood}}_{\text{genus}[i]}$ is the analogous effect for mosquitoes that were blood-fed prior to release versus those that were not fed.
-The parameter $c_i$ is modelled hierarchically, being drawn from a group-level distribution assumed to be normal,
-%
-\begin{equation}
-c_i \sim  \mathcal{N}(\mu_\lambda,\sigma)
-\end{equation}
-%
-We then specify priors on these group-level parameters, $\mu_\lambda \sim \mathcal{N}(-2.32,1)$ and $\sigma \sim \textnormal{log-normal}(-3,1)$.
-On all $\delta$ parameters, except $\delta^{\text{mix}}_{\text{genus}[i]}$, we specify standard normal priors.
-We specify a uniform prior on $\delta^{\text{mix}}_{\text{genus}[i]}\sim U(0,\delta^m_{\text{genus}[i]})$ so that the effect of mixed releases lies between the female-only release effect (0) and the male-only release effect ($\delta^m_{\text{genus}[i]}$).
-
-The relatively complex nature of priors in hierarchical models make it important to determine their influence on inferences.
-We chose the above -- somewhat uninformative -- priors to allow a range of mosquito lifespans in order to minimise their effect on the estimates we report (Fig.
-\ref{fig:mrr_meanLifespanPrior}).
-We also chose to set hierarchical priors on the remaining parameters in our models -- $\psi$ the probability of daily recapture, and $\kappa$ the `over-dispersion' parameter of a negative binomial distribution.
-For $\psi$ we chose an informative prior that placed all probability mass $\psi\leq 10\%$, since in most experiments the fraction of mosquitoes captured was significantly below this value.
-For $\kappa$ we chose a fairly wide prior that had most of its probability mass below $\kappa=20$ (Fig.
-\ref{fig:mrr_PsiKappaPriors}).
-In all cases the hierarchical priors were chosen to have similar implications on lifespan, recapture probability and over-dispersion at the individual time-series level as for the non-hierarchical analysis described previously.
-
-\begin{figure}[h]
-	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_lifespan_priors.pdf}}
-	\caption{\textbf{Samples from the prior predictive distribution of mean lifespan across the six models of mosquito mortality introduced in Table \ref{tab:mrr_survivalDescription}.} In all cases 10,000 samples were generated using the priors indicated in Table \ref{tab:mrr_priors}, resulting in distributions with a mean close to 10 days.}
-	\label{fig:mrr_meanLifespanPrior}
-\end{figure}
-
-\begin{figure}[h]
-	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_priors_PsiKappa.pdf}}
-	\caption{\textbf{Samples from the priors for daily recapture probability ($\psi$) and over-dispersion parameter ($\kappa$).} In both panels, 10,000 samples were generated using the following hierarchical priors: for the individual time series parameters, $\psi_i \sim \text{exp}(r)$ (with $\psi$ constrained to be less than 0.1 - encompassing the upper limit of day one recapture fractions observed in most data series) and $\kappa_i \sim \text{exp}(p)$; for the top-level parameters: $r\sim \text{gamma}(100, 2)$ and $p\sim \text{log-normal}(1.5, 1)$.}\label{fig:mrr_PsiKappaPriors}
-\end{figure}
-
-\subsection{Testing for age-dependent mortality in wild mosquitoes}\label{sec:mrr_age_dependence}
-Previous work has found evidence of age-dependent mortality in lab populations \citep{styer2007mosquitoes,dawes2009anopheles}, and less-commonly in wild mosquitoes \citep{clements1981analysis}.
-Furthermore, recent modelling work has examined the implications of departures from a constant risk of mortality \citep{styer2007mosquitoes,hancock2009age,novoseltsev2012age}.
-To determine whether senescence occurs in wild mosquitoes we re-estimate our model using survival functions, $S(t)$, that allow for a rate of mortality that can vary with age.
-Specifically, we re-estimate our model using five other models that were previously used in the literature (see Table \ref{tab:mrr_survivalDescription} for a description of these).
-Each of these models makes different assumptions about how the rate of death and dispersal varies with age, but all can be represented by a general form,
-%
-\begin{equation}\label{eq:survivalInt}
-S(t) = e^{-\int\limits_{0}^{t}\lambda(\tau) \mathrm{d}\tau}.
-\end{equation}
-%
-where we constrain the parameters of our models to preclude the possibility of a hazard that decreases with age ($\frac{\mathrm{d}\lambda(t)}{\mathrm{d}t} \geq 0$).
-Whilst this could occur if older mosquitoes disperse less than younger mosquitoes, it is unlikely this would outweigh any declines in survival associated with old age.
-
-The parameters of the survival function of each model were assigned hierarchical priors that allowed considerable variation in mosquito lifespan (see Fig.
-\ref{fig:mrr_meanLifespanPrior}).
-Otherwise the statistical model we used remained the same as for the constant mortality case.
-In testing for age-dependent mortality, we did not account for sex or pre-release feeding effects due to the complexity of including these elements in more complex survival models.
-
-\begin{table}
-	\scriptsize
-	\noindent\makebox[\textwidth]{%
-		\begin{tabularx}{1.35\textwidth}{c|cXcX}
-			\textbf{Survival function} & \textbf{Hazard rate} & \textbf{Interpretation} & \textbf{Mean time spent in study area, }\textbf{$\overline{T}$} & \textbf{Papers assuming this function} \\
-			\midrule
-			Exponential & $\lambda$ & Constant mortality risk & $\frac{1}{\lambda}$ & \cite{ross1910prevention}; \cite{anderson1992infectious}; \cite{smith2004statics} \\
-			\midrule
-			Gompertz &  $\alpha e^{\beta t}$     & Mortality increases with age at an ever-increasing rate &  $\frac{e^{\frac{\alpha}{\beta}} \Gamma(0,\frac{\alpha}{\beta})}{\beta}$    & \cite{clements1981analysis}; \cite{styer2007mosquitoes};\cite{novoseltsev2012age} \\
-			\midrule
-			Weibull &  $\alpha t^\beta$; $\;\beta\geq 1$     &  Mortality increases with age at an ever-increasing rate     &   $\alpha^{-\frac{1}{\beta}} \Gamma(1+\frac{1}{\beta})$    & \cite{hancock2009age}; and \cite{carey2001insect} considering general insect demography \\
-			\midrule
-			Gompertz-Makeham &  $\alpha e^{\beta t} + c$      &  Two additive mortality risks: one that increases with age, and another that is age-independent  &  No simple analytic form     & \cite{styer2007mosquitoes} \\
-			\midrule
-			Logistic & $\frac{\alpha e^{\beta t}}{1+\frac{\alpha s}{b}\left(e^{b t} - 1\right)}$   & Mortality risk increases with age, although at a declining rate  & No simple analytic form   & \cite{styer2007mosquitoes}; \cite{novoseltsev2012age} \\
-			\midrule
-			Logistic-Makeham &   $\frac{\alpha e^{\beta t}}{1+\frac{\alpha s}{b}\left(e^{b t} - 1\right)} + c$    & Two separate additive mortality risks: an age-dependent risk of the same form as the logistic model, and a constant hazard  &  No simple analytic form  & \cite{styer2007mosquitoes}; \cite{bellan2010importance} \\
-			\bottomrule
-	\end{tabularx}}
-	\caption{\textbf{A description of the survival functions used in this study, arranged in rough order of model complexity (simple-complex from top-bottom.)} All parameters are defined to be non-negative.
-$\Gamma(\theta)$ and $\Gamma(\theta_1,\theta_2)$ refer to the Euler gamma function, and incomplete gamma function respectively.
-The `mean time spent in study area' is an estimate of the combined effects of mosquito mortality and dispersal from the area of the study where collections take place, since our data does not provide spatially-resolved data.}\label{tab:mrr_survivalDescription}
-\end{table}
-
-\begin{table}
-	\footnotesize
-	\noindent\makebox[\textwidth]{%
-		\begin{tabularx}{1.2\textwidth}{l|c|X}
-			\textbf{Survival function} & \textbf{time-series-level priors} & \textbf{Group-level priors}\\
-			\midrule
-			Exponential & $c\sim \mathcal{N}(\mu_\lambda,\sigma)$ &$\mu_\lambda \sim N(-2.32,1)\text{,}\; \sigma\sim \textnormal{log-normal}(-3,1)$\\
-			\midrule
-			Gompertz & $\alpha,\beta\sim$ log-normal$(\mu_{\alpha|\beta},0.2)$ &$\mu_\alpha\sim N(-3,1)\text{,}\; \mu_\beta\sim N(-3,1.1)$\\
-			\midrule
-			Weibull & $\alpha,(\beta-1)\sim$  log-normal$(\mu_{\alpha|\beta},0.2)$ &$\mu_\alpha\sim N(-4.8,1.75)\text{,}\; \mu_\beta\sim N(-4,2)$\\
-			\midrule
-			Gompertz-Makeham & $\alpha,\beta,c\sim$ log-normal$(\mu_{\alpha|\beta|c},0.2)$ &$\mu_\alpha\sim N(-3,1)\text{,}\; \mu_\beta\sim N(-3,0.5)\text{,}\; \mu_c\sim N(-4.5,0.5)$\\
-			\midrule
-			Logistic & $\alpha,\beta,s\sim$  log-normal$(\mu_{\alpha|\beta|s},0.2)$ &$\mu_\alpha\sim N(-2.8,1)\text{,}\; \mu_\beta\sim N(-3,1)\text{,}\; \mu_s\sim N(-3,1)$\\
-			\midrule
-			Logistic-Makeham & $\alpha,\beta,s,c\sim$ log-normal$(\mu_{\alpha|\beta|s|c},0.2)$ &$\mu_\alpha\sim N(-3.2,1)\text{,}\; \mu_\beta\sim N(-3.2,1)\text{,}\; \mu_s\sim N(-3,0.5)\text{,}\; \mu_c\sim N(-4,0.5)$\\
-			\bottomrule
-	\end{tabularx}}\caption{\textbf{The priors used on parameters of each different survival model.} For the exponential model the `group-level' priors were the same for the genus and `overall' models that were also estimated.
-The exponential model was the only model that was simple enough to allow the scale parameter of the log-normal ($\sigma$) to be estimated by the data.
-The notation $\alpha,\beta\sim \text{log-normal}(\mu_{\alpha|\beta},0.2)$ means that $\alpha$ and $\beta$ were assigned independent log-normal priors with location parameters $\mu_\alpha$ and $\mu_\beta$ respectively, and a scale parameter of 0.2 in both cases.}\label{tab:mrr_priors}
-\end{table}
-
-\subsection{Model estimation by MCMC}\label{sec:mrr_MCMC}
-The likelihood and priors we use result in posterior distributions whose analytic form cannot be calculated with existent computational methods.
-Instead we use Markov Chain Monte Carlo (MCMC) methods to sample from each posterior distribution.
-We used \textit{Stan} software \citep{carpenter2016stan} that implements an efficient MCMC algorithm known as NUTS \citep{hoffman2014no}.
-
-To judge convergence of the sampling algorithm to the posterior density, we calculated $\hat{R}$ across all Markov chains -- a ratio that compares the between-chain variation to that within each chain that is commonly used to measure convergence in MCMC \citep{gelman1992inference}.
-For each class of model, we used the following MCMC parameters,
-%
-\begin{itemize}
-	\item Individual lifespan models (Section \ref{sec:MRR_individual_analysis}): 12 independent chains with 1000 iterations per chain,
-	\item Hierarchical lifespan estimates (Section \ref{sec:MRR_hierarchical}): 12 independent chains with 15,000 iterations per chain,
-	\item Age-dependent analysis (Section \ref{sec:mrr_age_dependence}): 12 independent chains with 3000 iterations per chain.
-\end{itemize}
-
-In all cases, we discarded the first half of these iterations as warm-up \citep{gelman2014bayesian}.
-At the end of all runs for lifespan estimation, $\hat{R}<1.1$ for all model parameters.
-For the age-dependent cases, there remained a handful of parameters where $\hat{R}>1.1$ after running simulations for 3000 iterations but this did not affect our inferences, since repeated runs demonstrated the same pattern.
-We also ensured that across each MCMC run the number of divergent iterations (that can bias the MCMC away from the true posterior density) was minimal.
-In the majority of cases, the number of divergent iterations was far fewer than 1\% of the total number of samples.
-
-The Stan scripts for the lifespan estimation (i.e.
-those using an exponential survival model) for the MRR analyses are provided below.
-The script for the individual time series analysis (Section \ref{sec:MRR_individual_analysis}):
-
-\begin{minted}{stan}
-data {
-// Single release parameters
-int<lower=0> N; // number of observations
-int<lower=0> K; // number of groups
-int Y[N]; // observations for the releases
-int S[K]; // time series lengths
-vector[N] t; // time observations (days) 
-int Pos[K]; // starting position of each dataset
-
-// Multiple release parameters
-int RelFreq[K]; // frequency of releases in each study
-// starting position of releases in each study
-// in RelNumber
-int PosRel[K];
-int<lower=0> NReleases; // total number of releases
-// numbers released in each study
-vector[NReleases] RelNumber;
-// time of each release in each study
-vector[NReleases] RelTime;
-
-real Age[K];
-}
-
-parameters {
-real<lower=0.3, upper=20> kappa[K];
-real<lower=0> cDecay[K];
-real<lower=0, upper=1> psi[K];
-}
-
-model {
-// Likelihood
-for (i in 1:K) {
-real lambda[S[i]];
-
-if (RelFreq[i]< 2) // Single release
-{
-   real numberCaught;
-   numberCaught = 0;	
-   for (j in 1:S[i])
-   {
-     lambda[j]= (psi[i] *
-                (RelNumber[PosRel[i]] - numberCaught) *
-                exp(-cDecay[i] * (t[Pos[i] + j - 1] + Age[i])));
-     //1e-5 to prevent loc=0
-     Y[Pos[i] + j - 1] ~ neg_binomial_2(
-        		lambda[j] + 1e-5, kappa[i]); 
-     numberCaught = numberCaught + Y[Pos[i] + j - 1];
-   }
-} else // Multiple release
-  {
-    for (j in 1:S[i])
-    {
-      real lambdaTemp;			
-      lambdaTemp = 0;
-      for (kk in 1:RelFreq[i])
-      {
-        if (t[Pos[i] + j - 1]> RelTime[PosRel[i]+kk-1])
-        {
-           lambdaTemp = (
-                   lambdaTemp +
-                   RelNumber[PosRel[i] + kk - 1] * 
-                   ((1 - psi[i])^(t[Pos[i] + j - 1] -
-                   RelTime[PosRel[i]+kk-1]-1)) *
-                   psi[i] * exp(-cDecay[i] * 
-                   (t[Pos[i] + j - 1] -
-                   RelTime[PosRel[i]+kk-1] + Age[i]))
-           );
-          }
-   }
-   lambda[j] = lambdaTemp;
-   Y[Pos[i] + j - 1] ~ neg_binomial_2(lambda[j]+1e-5, kappa[i]);
-   }
- }
-}
-
-// Priors
-for (i in 1:K)
-{
-   cDecay[i] ~ lognormal(-2.32, 1);
-   kappa[i] ~ lognormal(2, 1);  
-   psi[i] ~ exponential(50);
-}
-}
-
-generated quantities {
-vector[K] lifespans;
-  for(i in 1:K)
-    lifespans[i] = 1 / cDecay[i];
-}
-\end{minted}
-
-The script for the hierarchical lifespan estimation (Section \ref{sec:MRR_hierarchical}) is shown below:
-
-\begin{minted}{stan}
-data {
-// Single release parameters
-int<lower=0> N; // number of observations
-int<lower=0> K; // number of groups
-int Y[N]; // observations
-int S[K]; // group sizes
-vector[N] t; // time observations (days) 
-int Pos[K]; // starting position of each dataset
-
-// Multiple release parameters
-int RelFreq[K]; // release frequency in each study
-int PosRel[K]; // starting position of releases in each study
-int<lower=0> NReleases; // release frequencies
-vector[NReleases] RelNumber; // numbers released
-vector[NReleases] RelTime; // release times
-
-int<lower=0> nSpecies; // number of individual species
-int SpeciesIndex[K]; // species index of each study
-
-real Age[K];
-int SexM[K];
-int SexMix[K];
-int Genus[K];
-int Blood[K];
-int Sugar[K];
-}
-
-transformed data{
-int SpeciesToGenus[nSpecies];
-for(i in 1:K){
-  SpeciesToGenus[SpeciesIndex[i]] = Genus[i];
-}
-}
-
-parameters {
-real<lower=0.3,upper=100> kappa[K];
-real a[nSpecies];
-real<lower=0> p[nSpecies];
-real<lower=0> r[nSpecies];
-real cDecay[K];
-real<lower=0,upper=0.1> psi[K];
-real<lower=0> sigma_a[nSpecies];
-real delta_M[3];
-real delta_Blood[3];
-real delta_Sugar[3];
-real<lower=0, upper=1> u_mixed[3];
-}
-
-transformed parameters{
-real delta_Mixed[3];
-for(i in 1:3)
-  delta_Mixed[i] = u_mixed[i] * delta_M[i];
-}
-
-model {
-for (i in 1:K) {
-  real lambda[S[i]];
-  real decay_temp = exp(cDecay[i]+delta_M[Genus[i]] * SexM[i] +
-    delta_Mixed[Genus[i]] * SexMix[i] + delta_Sugar[Genus[i]] *
-    Sugar[i] + delta_Blood[Genus[i]] * Blood[i]);
-
-  if (RelFreq[i]< 2) // Single release
-  {
-    real R_times_psi;
-    real numberCaught;
-    numberCaught = 0;	
-    R_times_psi = psi[i];
-    for (j in 1:S[i])
-    {
-    lambda[j]= R_times_psi *
-            (RelNumber[PosRel[i]]-numberCaught) *
-            exp(-decay_temp * (t[Pos[i] + j - 1] + Age[i]));
-    Y[Pos[i] + j - 1] ~ neg_binomial_2(lambda[j] + 1e-5,
-                                       kappa[i]);
-    numberCaught = numberCaught + Y[Pos[i] + j - 1];
-    }
-  }
-  else // Multiple release
-  {
-    for (j in 1:S[i])
-    {
-    real lambdaTemp;			
-    lambdaTemp = 0;
-    for (kk in 1:RelFreq[i])
-    {
-      if (t[Pos[i] + j - 1]> RelTime[PosRel[i] + kk - 1])
-      {
-        lambdaTemp = (lambdaTemp +
-          RelNumber[PosRel[i] + kk - 1] * 
-          ((1-psi[i])^(t[Pos[i] + j - 1] -
-          RelTime[PosRel[i] + kk - 1] - 1)) *
-          psi[i] * exp(-decay_temp *
-          (t[Pos[i] + j - 1] -
-          RelTime[PosRel[i] + kk - 1] + Age[i]));
-         );
-        }
-      }
-      lambda[j] = lambdaTemp;
-      Y[Pos[i] + j - 1] ~ neg_binomial_2(lambda[j] + 1e-5,
-                                         kappa[i]);
-      }
-    }
-}
-
-// Priors
-for (i in 1:K)
-{
-  int speciesTemp;
-  speciesTemp = SpeciesIndex[i];
-  cDecay[i] ~ normal(a[speciesTemp], sigma_a[speciesTemp]);
-  kappa[i] ~ exponential(p[speciesTemp]);  
-  psi[i] ~ exponential(r[speciesTemp]);
-}
-a ~ normal(-2.32, 1);
-sigma_a ~ lognormal(-3, 1);
-p ~ lognormal(1.5, 1);
-r ~ gamma(100, 2);
-delta_M ~ normal(0, 1);
-delta_Blood ~ normal(0, 1);
-delta_Sugar ~ normal(0, 1);
-}
-
-generated quantities{
-real overallcDecay[nSpecies];
-real overallFLife[nSpecies];
-real overallMLife[nSpecies];
-real overallMixedLife[nSpecies];
-real overallFBloodLife[nSpecies];
-real overallFBothLife[nSpecies];
-real overallFSugarLife[nSpecies];
-real overallMSugarLife[nSpecies];
-
-for(i in 1:nSpecies){
-  overallcDecay[i] = normal_rng(a[i], sigma_a[i]);
-  overallFLife[i] = 1 / exp(overallcDecay[i]);
-  overallMLife[i] = 1 / exp(overallcDecay[i] +
-                            delta_M[SpeciesToGenus[i]]);
-  overallMixedLife[i] = 1 / exp(overallcDecay[i] +
-                            delta_Mixed[SpeciesToGenus[i]]);
-  overallFBloodLife[i] = 1 / exp(overallcDecay[i] +
-                             delta_Blood[SpeciesToGenus[i]]);
-  overallFSugarLife[i] = 1 / exp(overallcDecay[i] +
-                             delta_Sugar[SpeciesToGenus[i]]);
-  overallFBothLife[i] = 1 / exp(overallcDecay[i] +
-                             delta_Sugar[SpeciesToGenus[i]] +
-                             delta_Blood[SpeciesToGenus[i]]);
-  overallMSugarLife[i] = 1 / exp(overallcDecay[i] +
-                             delta_M[SpeciesToGenus[i]] +
-                             delta_Sugar[SpeciesToGenus[i]]);
-}
-}
-\end{minted}
-
-\subsection{K-Fold cross validation}\label{sec:mrr_kFold}
-One way to estimate a model's out-of-sample predictive capability is to use the Akaike Information Criterion that explicitly penalises a model in accordance to its complexity (typically indexed by its number of free parameters) to correct for model over-fit.
-The way in which this correction takes places however, is fairly heuristic, and less appropriate for hierarchical models.
-An alternative method, common in the machine learning literature, is known as `cross-validation' (see for example, \cite{kohavi1995study}).
-Here to estimate out of sample predictive capability, the original data set is partitioned into training and test sets.
-The model is then fitted to the training set, and its predictive performance measured on the independent test set.
-
-We use cross-validation to compare the predictive power of the species, genus and overall models, and later to determine whether age-dependent mortality occurs.
-In particular, we use K-Fold cross validation where we repeatedly partition our dataset into a test set composed of a number of individual time-series, and a training set of the remainder (see for example, \cite{marsland2015machine}).
-We then use the fitted model to measure the predictive performance on the test set.
-The particular measure we use is the expected log point-wise predictive density (elpd)\citep{vehtari2015efficient} which sums the predictive performance for each data point averaged across all posterior samples.
-
-Since the models are hierarchical we must specify how to draw parameters for a given test sample time-series.
-Here we chose to draw values of the individual time-series parameters as samples from the population distribution that corresponds to that particular sample's respective grouping.
-So if the particular test sample time-series pertains to \textit{Ae.
-aegypti} (and we are using the species-level model) then we draw independent samples for the statistical model's parameters from the estimated \textit{Ae.
-aegypti} population distribution,
-%
-\begin{align}
-\theta_i \sim p(\theta^{\text{Ae. aegypti}}|data),
-\end{align}
-%
-where $\theta_i$ is the value of the parameters used in test time-series $i$, and $p(\theta^{\text{Ae.
-aegypti}}|data)$ is the overall posterior distribution across all \textit{Ae.
-aegypti} experiments.
-
-\subsection{Model checking}
-To check the fit of the model to the data, we simulated from the posterior predictive distribution for each data point within each time series of recapture observations (see Figures \ref{fig:mrr_ppc1}, \ref{fig:mrr_ppc2} \&  \ref{fig:mrr_ppc3} and the attached file, \verb|mrr_ppcs_all.pdf| for the full graphs).
-In the majority of cases, the data lay within the 95\% predictive intervals indicating that the model was a good fit to the data.
-
-{%
-	\includegraphics[page=1,scale=.6]{./Figure_files/mrr_ppcs_all}
-	\captionof{figure}{The numbers of mosquitoes recaptured (black lines) versus the 95\% central posterior interval of the posterior predictive distribution (blue shading) for a selection of the time series.
-For the rest of the posterior predictive checks for the MRR models, see the file referenced in the text.}
-	\label{fig:mrr_ppc1}
-	\par
-
-{%
-	\includegraphics[page=2,scale=.6]{./Figure_files/mrr_ppcs_all}
-	\captionof{figure}{The numbers of mosquitoes recaptured (black lines) versus the 95\% central posterior interval of the posterior predictive distribution (blue shading) for a selection of the time series.
-For the rest of the posterior predictive checks for the MRR models, see the file referenced in the text.}
-	\label{fig:mrr_ppc2}
-	\par
-}
-
-{%
-	\includegraphics[page=3,scale=.6]{./Figure_files/mrr_ppcs_all}
-	\captionof{figure}{The numbers of mosquitoes recaptured (black lines) versus the 95\% central posterior interval of the posterior predictive distribution (blue shading) for a selection of the time series.
-For the rest of the posterior predictive checks for the MRR models, see the file referenced in the text.}
-	\label{fig:mrr_ppc3}
-	\par
-}
-
-\section{Dissection based estimates}
-
-\subsection{Collection of dissection data}\label{sec:dissection_dissectionData}
-A comprehensive search of the literature using Google Scholar (\url{scholar.google.co.uk}) was performed using various combinations of the following keywords: dissection, mosquito, parity, parous, age, and physiological age.
-No constraints were placed on publication date, location or type.
-This list was supplemented with a number of author-specific searches for those individuals most prevalent in the literature.
-In particular we searched for all articles authored by: Charlwood, Muller, Schlein, Samarawickrema, Reisen, Detinova, Polodova, Gillies, and Wilkes.
-An additional list of potential articles was provided by doing a forward article search on some of the most highly-cited articles in the database: \cite{polovodova1949determination,detinova1962age,gillies1965study,clements1981analysis}.
-The list of results was then filtered manually by examining the titles and abstracts to produce a candidate list of the articles most likely to contain data on the physiological age of dissected mosquitoes caught in the wild, as determined by the \cite{polovodova1949determination} method.
-
-A relational database was used to store the raw data from the actual experiments, along with the meta-data associated with each of the experiments.
-In many of the published articles wild mosquitoes were caught and dissected over a period of time, and the raw data thus consists of snapshots of the age structure of the population at regular intervals in time.
-In the cases where the data was more sparse (fewer than ten individuals, on average, per date), we aggregated across dates and recorded this as a single entry; otherwise we recorded the snapshots of the population at each different date.
-We record separate series for each species that was captured, or for those that were recorded at separate capture locations (potentially with an alternative collection method), and do not aggregate over these datasets.
-
-For each individual series we recorded the following meta-data: genus, species, collection method, whether or not insecticide was mentioned as being used during or shortly before the time of mosquito collections, and the start and end dates of the experiment.
-At the article level we recorded the following meta-data: author, year, title, country, location (within country), start and end date, whether insecticide was used during any of the experimental replicates, whether a mrr experiment occurred alongside the dissection study, and the collection location (indoor or outdoor).
-At either the individual series or article levels we record additional meta-data describing the nature of data collection, for example explaining where the data was contained within the article, whether it was obtained by digitising graphs, and the number of separate dated series.
-For those few cases ($n=16$ series) where the data was obtained by digitising graphs, we used the WebPlotDigitizer online tool \citep{digitise}.
-
-The data collection method resulted in 568 separate dissection series, across 72 published articles (see Section \ref{sec:appendix_dissectionStudyList} for a list of the studies included in our final dataset).
-The published datasets cover the period from 1960-2015 with comparable numbers of studies across each decade (Fig.
-\ref{fig:dissection_timeSeries}).
-The statistical approach applied to the data relies on the assumption that there is a constant rate of recruitment into the adult population (see Section \ref{sec:dissection_dissectionStats}).
-If populations fluctuate strongly from month-to-month this assumption will likely be violated.
-To try to mitigate against such a risk, we aggregate the data across all dates to obtain a single series for each identifier, resulting in 201 such series.
-To obtain a reasonable level of accuracy on estimates we remove all those individual (aggregated) series where there are fewer than 100 mosquitoes in total in the series.
-Finally we remove data for any species where there was only one series resulting in 131 series across four genera (\textit{Anopheles}, \textit{Culex}, \textit{Aedes}, and \textit{Mansonia}) and 25 species (Table \ref{tab:dissection_speciesNumbers}).
-These studies are distributed across a wide range of geographies (Fig.
-\ref{fig:dissection_map}).
-The raw data for the dissection analysis are shown in Figures \ref{fig:dissection_exampleData1}, \ref{fig:dissection_exampleData2} and \ref{fig:dissection_exampleData3}.
-
-\begin{figure}[ht]
-	\centerline{\includegraphics[width=0.8\textwidth]{./Figure_files/dissection_timeSeries.pdf}}
-	\caption{\textbf{The numbers of published studies that estimate gonotrophic age by dissection across the study period.}}\label{fig:dissection_timeSeries}
-\end{figure}
-
-\begin{figure}[ht]
-	\centerline{\includegraphics[width=1.25\textwidth]{./Figure_files/dissection_map.pdf}}
-	\caption{\textbf{The geographic location of each of the dissection databases included in the meta-analysis.} The area of the bubbles indicates the number of unique data series available.}\label{fig:dissection_map}
-\end{figure}
-
-\begin{table}[htbp]
-	\centering
-	\begin{tabular}{lll}
-		\toprule
-		\textbf{Genus} & \textbf{Species} & \textbf{Frequency} \\
-		\midrule
-		\textit{Anopheles} & \textit{gambiae s.l.} & 19 \\
-		\textit{Culex} & \textit{quinquefasciatus} & 12 \\
-		\textit{Anopheles} & \textit{maculipennis} & 11 \\
-		\textit{Anopheles} & \textit{farauti s.l.} & 10 \\
-		\textit{Anopheles} & \textit{sergentii} & 8 \\
-		\textit{Aedes} & \textit{polynesiensis} & 8 \\
-		\textit{Culex} & \textit{pipiens} & 6 \\
-		\textit{Anopheles} & \textit{culicifacies} & 5 \\
-		\textit{Anopheles} & \textit{darlingi} & 5 \\
-		\textit{Anopheles} & \textit{quadrimaculatus} & 5 \\
-		\textit{Anopheles} & \textit{stephensi} & 4 \\
-		\textit{Anopheles} & \textit{melas} & 4 \\
-		\textit{Culex} & \textit{annulirostris} & 4 \\
-		\textit{Aedes} & \textit{aegypti} & 3 \\
-		\textit{Aedes} & \textit{samoanus} & 3 \\
-		\textit{Anopheles} & \textit{minimus} & 3 \\
-		\textit{Anopheles} & \textit{rivulorum} & 3 \\
-		\textit{Culex} & \textit{thalassius} & 3 \\
-		\textit{Mansonia} & \textit{uniformis} & 3 \\
-		\textit{Anopheles} & \textit{subpictus} & 2 \\
-		\textit{Aedes} & \textit{sollicitans} & 2 \\
-		\textit{Aedes} & \textit{vexans} & 2 \\
-		\textit{Anopheles} & \textit{bellator} & 2 \\
-		\textit{Anopheles} & \textit{cruzii} & 2 \\
-		\textit{Culex} & \textit{tritaeniorhynchus} & 2 \\
-		\bottomrule
-		\textit{\textbf{Anopheles}} &       & \textbf{83} \\
-		\textit{\textbf{Culex}} &       & \textbf{27} \\
-		\textit{\textbf{Aedes}} &       & \textbf{18} \\
-		\textit{\textbf{Mansonia}} &       & \textbf{3} \\
-		\bottomrule
-		\textbf{Total} &  & \textbf{131} \\
-	\end{tabular}%
-	\caption{\textbf{The numbers of dissection series for each species or genus included in the overall dataset.}}\label{tab:dissection_speciesNumbers}%
-\end{table}%
-
-\begin{table}[htbp]
-	\centering
-	\footnotesize
-	\begin{adjustwidth}{-0.5in}{-0.5in}%
-		\begin{tabularx}{1.25\textwidth}{l|ccccc}
-			\toprule
-			\textbf{Variable} & \textbf{Min} & \textbf{Median} & \textbf{Mean} & \textbf{Max} & \textbf{Standard deviation} \\
-			\midrule
-			Min age of captures, gonotrophic cycles & 0  & 0  & 0.02  & 1.00  & 0.12 \\
-			Median age of captures, gonotrophic cycles & 0  & 1.00  & 0.63  & 2.00  & 0.62 \\
-			Mean age of captures, gonotrophic cycles & 0.17  & 0.90  & 1.00  & 3.03  & 0.58 \\
-			Max age of captures, gonotrophic cycles & 1.00  & 5.00  & 5.77  & 13.00 & 2.94 \\
-			Number captured & 100 & 565 & 1,317 & 14,012 & 1,885 \\
-			\bottomrule
-		\end{tabularx}%
-	\end{adjustwidth}
-	\caption{\textbf{Summaries of the characteristics of the 131 individual dissection data series that are included in this analysis.} Note when series are censored (see Section \ref{sec:dissection_censored}) we assume that all mosquitoes dissected over the threshold are of this age when calculating summary statistics.}\label{tab:dissection_summaryStats}
-\end{table}
-
-\begin{figure}[ht]
-	\centerline{\includegraphics[width=1.3\textwidth]{./Figure_files/dissection_parity_data1.png}}
-	\caption{\textbf{The normalised physiological age series for nine species in the database.} Each different coloured line represents an individual series.
-In each case the count for all ages has been normalised by the nulliparous count.
-In all cases we do not include any data for censored observations (see Section \ref{sec:dissection_censored}).}\label{fig:dissection_exampleData1}
-\end{figure}
-
-\begin{figure}[ht]
-	\centerline{\includegraphics[width=1.3\textwidth]{./Figure_files/dissection_parity_data2.png}}
-	\caption{\textbf{The normalised physiological age series for nine species in the database.} Each different coloured line represents an individual series.
-In each case the count for all ages has been normalised by the nulliparous count.
-In all cases we do not include any data for censored observations (see Section \ref{sec:dissection_censored}).}\label{fig:dissection_exampleData2}
-\end{figure}
-
-\begin{figure}[ht]
-	\centerline{\includegraphics[width=1.3\textwidth]{./Figure_files/dissection_parity_data3.png}}
-	\caption{\textbf{The normalised physiological age series for seven species in the database.} Each different coloured line represents an individual series.
-In each case the count for all ages has been normalised by the nulliparous count.
-In all cases we do not include any data for censored observations (see Section \ref{sec:dissection_censored}).}\label{fig:dissection_exampleData3}
-\end{figure}
-
-\subsection{Statistical analysis of dissection data}\label{sec:dissection_dissectionStats}
-We suppose that the number of individuals recruited into the adult female population is constant over time, meaning that the age-structure of the population is stable.
-Whilst environmental heterogeneity will naturally lead to variation in adult recruitment over time, we hope that by aggregating data across studies undertaken over a range of dates we may reduce the impact of this effect on our results.
-Suppose that the probability an individual female mosquito survives until age $a$ is given by the survival function $S(a)$, then the number of individuals in the population surviving to this age is given by,
-%
-\begin{align}\label{eq:dissection_structured}
-A(a) = A(0) S(a),
-\end{align}
-%
-where $A(0)$ is the number of female mosquitoes recruited into the population per unit time.
-Consider one individual experiment where we randomly sample from a wild population that is structured as per eq.
-(\ref{eq:dissection_structured}).
-In this case the number of individuals sampled at each age is binomially-distributed,
-%
-\begin{align}
-X(a) \sim \mathcal{B}(A(a),p(a)),
-\end{align} 
-%
-where $p(a)$ is the probability of recapturing a single mosquito of age $a$.
-In what follows we assume that the probability of capturing a given mosquito is independent of their age, so that $p(a)=p=const$.
-Since in general $A(a)$ is large and $p(a)$ is small, we can approximate the above using a Poisson distribution,
-%
-\begin{align}
-X(a) \sim \text{Poisson}(A(a)p).
-\end{align}
-%
-However we believe that the assumption of \textit{independent} captures of individual mosquitoes, which underlies the binomial and Poisson models, is likely suspect for the same reasons as for the MRR analysis (see Section \ref{sec:mrr_statistical}).
-As before we choose to specify a negative binomial sampling distribution that allows for non-independent captures,
-%
-\begin{align}
-X(a) \sim \text{NB}(A(a)p,\kappa),
-\end{align} 
-%
-where we use the parameterisation such that the mean is given by $A(a)p$, and the over-dispersion parameter, $\kappa$, where as $\kappa\rightarrow \infty$ the above sampling distribution approaches a Poisson.
-
-In the field, unfortunately, we do not know the number of mosquitoes recruited into the adult population, $A(a)$, nor the probability of capturing an individual at a given point in time, $p$.
-Instead we model their product $\Psi=A(0)p$ (the population of adult female mosquitoes of age zero that can be captured) probabilistically resulting in a model,
-%
-\begin{align}\label{eq:dissertation_negativeBinomial}
-X(a) \sim \text{NB}(\Psi S(a),\kappa).
-\end{align}
-%
-The resultant likelihood of a data series consisting of counts: $(y(a_1),y(a_2),...,y(a_R))$, is then calculated by assuming (conditional) independence of the observations,
-%
-\begin{equation}
-\mathcal{L}(y(a_1),y(a_2),...,y(a_R)|S(.),\Psi,\kappa) = \prod\limits_{a=a_1}^{a_R} p(y(a)|S(a),\Psi,\kappa),
-\end{equation}
-%
-where $p(y(a)|S(a),\Psi,\kappa)$ corresponds to the negative binomial probability mass function for a count of $y(a)$ mosquitoes aged $a$ as specified in eqn.
-(\ref{eq:dissertation_negativeBinomial}), and $R$ is the number of separate physiological age classes in which the count was non-zero.
-%
-Since we do not know $\Psi$ we must learn it from the data.
-One approach to estimate this parameter could be use the number of captures of nulliparous mosquitoes, $X(0)$ in place of $\Psi$.
-We prefer to allow for some uncertainty in this parameter and use the data to estimate its value.
-However, unfortunately, the negative binomial likelihood allows too much variation in this parameter (because the data are over-dispersed), and instead we specify a likelihood of the form,
-%
-\begin{align}
-X(0) \sim \mathcal{N}(\Psi,\sqrt{\Psi}),
-\end{align}
-%
-solely for the first data point $X(0)$.
-The above allows for some freedom in $\Psi$ whilst ensuring that the parameter's probability mass lies near enough to $X(0)$ to allow useful model estimates.
-$\Psi$ is set a uniform prior over the range $[0,\frac{3}{2}X(0)]$.
-
-Since $S(a)$ is monotonically-decreasing we know that, if recruitment to the adult population is constant, then the numbers of nulliparous individuals should exceed the count in subsequent parity states.
-However in a number of data series there is a relative dearth of nulliparous mosquitoes versus uniparous individuals.
-This deficiency has been noted in a previously-published study where the authors hypothesise that it is due to the issue of sampling the nulliparous population, since they are more likely to rest outside, compared with parous individuals \citep{gillies1965study}.
-However there is evidence to suggest that, on average, the first gonotrophic cycle is longer than subsequent cycles (see Section \ref{sec:dissection_gonotrophicData}), meaning that a relative surplus of nulliparous mosquitoes may exist in captured samples \citep{clements1981analysis}.
-
-In our analysis, we chose to remove the counts of nulliparous individuals from the series where the count was low relative to uniparous or higher parity individuals since this may indicate issues with sampling the nulliparous population.
-Specifically we stipulated that the nulliparous count should exceed 90\% of the count for the uniparous individuals.
-For those cases where this condition was not met we removed the nulliparous observation and analyse the series of counts for all subsequent pars (uniparous and subsequent parous states).
-
-Here we estimate our model with one of six different survival functions (the same as for the MRR case; Table \ref{tab:mrr_survivalDescription}), each of which makes different assumptions regarding how the force of mortality is affected by mosquito age.
-To estimate lifespan at the species, genus and overall groupings we then use a hierarchical Bayesian model of the same mathematical form as for the analysis of MRR experiments (see Section \ref{sec:MRR_hierarchical}).
-However the priors for the analysis of dissection data were modified so that they represented a mean prior lifespan of three gonotrophic cycles, although allowed considerable variation in mean lifespan (Fig.
-\ref{fig:dissection_lifespanPriorsParity}).
-The priors used for the over-dispersion parameter ($\kappa$) for the negative binomial likelihood were the same as for the MRR analysis.
-
-To determine whether mosquitoes experience age-dependent mortality we compared the predictive power of each of the models that incorporate a hazard function that depends on age with that from the exponential model.
-As for the MRR analysis we also used K-Fold cross-validation to perform this comparison, where the data are randomly partitioned into training and test sets (see Section \ref{sec:mrr_kFold}).
-The model is then fitted to each training set and used to predict the data in the independent test set.
-Since there are fewer series than for the MRR dataset we used two partitions for each species, where each partition was of roughly the same number of individual series.
-
-\begin{figure}[ht]
-	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/dissection_lifespanPriorsParity.pdf}}
-	\caption{\textbf{The priors for mean lifespan used to analyse the data for each of the hierarchical models.} The distributional form of the priors is given in Table \ref{tab:dissection_priors}.
-In all cases the plots show data for 10,000 samples from the relevant prior distribution.}\label{fig:dissection_lifespanPriorsParity}
-\end{figure}
-
-\begin{table}
-	\footnotesize
-	\noindent\makebox[\textwidth]{%
-		\begin{tabularx}{1.2\textwidth}{l|c|X}
-			\textbf{Survival function} & \textbf{Physiological age series-level priors} & \textbf{Group-level priors}\\
-			\midrule
-			Exponential & $\lambda\sim \textnormal{log-normal}(\mu_\lambda,\sigma)$ &$\mu_\lambda \sim N(-1.2,1)\text{,}\; \sigma\sim \textnormal{log-normal}(-2,1)$\\
-			\midrule
-			Gompertz & $\alpha,\beta\sim$ log-normal$(\mu_{\alpha|\beta},0.2)$ &$\mu_\alpha\sim N(-1.5,1)\text{,}\; \mu_\beta\sim N(-2.5,0.5)$\\
-			\midrule
-			Weibull & $\alpha,(\beta-1)\sim$  log-normal$(\mu_{\alpha|\beta},0.2)$ &$\mu_\alpha\sim N(-2.5,1)\text{,}\; \mu_\beta\sim N(-4,0.5)$\\
-			\midrule
-			Gompertz-Makeham & $\alpha,\beta,c\sim$ log-normal$(\mu_{\alpha|\beta|c},0.2)$ &$\mu_\alpha\sim N(-1.4,1)\text{,}\; \mu_\beta\sim N(-3,0.4)\text{,}\; \mu_c\sim N(-4.5,0.5)$\\
-			\midrule
-			Logistic & $\alpha,\beta,s\sim$  log-normal$(\mu_{\alpha|\beta|s},0.2)$ &$\mu_\alpha\sim N(-1.4,1)\text{,}\; \mu_\beta\sim N(-2,1)\text{,}\; \mu_s\sim N(-3,1)$\\
-			\midrule
-			Logistic-Makeham & $\alpha,\beta,s,c\sim$ log-normal$(\mu_{\alpha|\beta|s|c},0.2)$ &$\mu_\alpha\sim N(-3,1)\text{,}\; \mu_\beta\sim N(-1.3,1)\text{,}\; \mu_s\sim N(-2,0.5)\text{,}\; \mu_c\sim N(-4,0.5)$\\
-			\bottomrule
-	\end{tabularx}}\caption{\textbf{The priors used on parameters of each different survival model for the dissection data analysis.} For the exponential model the `group-level' priors were the same for the genus and `overall' models that were also estimated.
-The exponential model was the only model that was simple enough to allow the scale parameter of the log-normal ($\sigma$) to be estimated by the data.
-The notation $\alpha,\beta\sim \text{log-normal}(\mu_{\alpha|\beta},0.2)$ means that $\alpha$ and $\beta$ were assigned independent log-normal priors with location parameters $\mu_\alpha$ and $\mu_\beta$ respectively, and a scale parameter of 0.2 in both cases.}\label{tab:dissection_priors}
-\end{table}
-
-The estimates of mean mosquito lifespan that we present here (Fig.
-\ref{fig:dissection_lifetimes_exponential}) resulted from the use of the exponential survival model (i.e.
-no age dependence).
-This choice was made because we found limited evidence in support of age-dependent mortality (see `Results').
-However, since the priors for each of the survival functions were specified to allow for a wide variety of lifespans (Fig.
-\ref{fig:dissection_lifespanPriorsParity}) the specific choice of survival function made little difference to resultant estimates (data not shown).
-
-To demonstrate that the results we obtain are not sensitive to the particular form of hierarchical prior structure chosen we also provide estimates of the lifespan for each series analysed separately (i.e.
-non-hierarchically).
-To do so we assume priors on the rate parameter of the exponential distribution that provides support over a wide range of possible lifespans (Fig.
-\ref{fig:dissection_lifespanPrior_individual}A).
-We also specified a prior on the over-dispersion parameter $\kappa$ that was comparable to the hierarchical case (Fig.
-\ref{fig:dissection_lifespanPrior_individual}B).
-
-\begin{figure}[ht]
-	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/dissection_lifespanPrior_individual.pdf}}
-	\caption{\textbf{The priors for A.
-mean lifespan and B.
-the over-dispersion parameter used to analyse the individual dissection series.} The prior on the rate parameter of the exponential distribution was $\lambda\sim \text{log-normal}(-1.8,1)$.
-The prior on the dispersion parameter was $\kappa\sim \text{log-normal}(1,1)$.}\label{fig:dissection_lifespanPrior_individual}
-\end{figure}
-
-\subsubsection{Analysis of censored data series}\label{sec:dissection_censored}
-In some of the published papers the data are censored above a threshold age, $a_T$.
-In these cases we only know the number, $y(a_T)$, of individuals who were captured and dissected with an estimated physiological age that is greater than or equal to the threshold.
-Since we do not know the estimated physiological age of individual specimens $i$, we represent this as a parameter, $\aleph_i$, in our statistical model.
-This means that the joint distribution is a function of $y(a_T)$ different $\aleph_i$ parameters.
-Since these parameters are not directly of interest, and our chosen MCMC engine, Stan \citep{stan-software:2014}, does not directly allow discrete parameters in models, we marginalise these out of the joint distribution.
-To do this exactly would require an infinite sum over all the possible ages for all $y(a_T)$ mosquitoes that have been caught whose age exceeds the threshold.
-Rather than carry out this intractable number of summations we instead make the approximation that all mosquitoes in this group are of the same age $\aleph_i = \aleph,\; \forall i \in (1,...,y(a_T))$.
-We believe this assumption is justifiable, particularly since the numbers of mosquitoes in each subsequent age category is a strongly-decreasing function of age.
-This means that whilst we do not know with certainty individuals' ages, it is likely that most will be of the threshold age $a_T$.
-
-This approximation means that to marginalise the parameter $\aleph$ out of the joint distribution, we are only required to do a single summation.
-Specifically if we have $y(a_T)$ captured individuals of age equal to, or exceeding, some threshold age, $a_T$, the probability of these observations is given by,
-%
-\begin{align}
-q(y(a_T)|S(.),\Psi,\kappa) &= \sum_{\aleph=a_T}^{\infty} p(y(a_T),\aleph|S(\aleph),\Psi,\kappa)\\
-&= \sum_{\aleph=a_T}^{\infty} p(y(a_T)|S(\aleph),\Psi,\kappa) \times p(\aleph)
-\end{align}
-%
-where $q(y(a_T)|S(\aleph),\Psi,\kappa)$ is the probability of observing $y(a_T)$ counts for mosquitoes of an age $\aleph$.
-In practice, since we do not believe mosquitoes live for longer than 20 gonotrophic cycles (the maximum observed in the data was 13), we cut-off the summation at this point, and assume that the discrete prior probability distribution $p(\aleph)$ is uniform over this range.
-The overall likelihood for the cases where the series are censored therefore has the form,
-%
-\begin{equation}
-\mathcal{L}(y(a_1),y(a_2),...,y(a_T-1),y(a_T)|S(.),\Psi,\kappa) = \left(\prod\limits_{a=a_1}^{a_T-1} p(y(a)|S(a),\Psi,\kappa)\right) q(y(a_T)|S(\aleph),\Psi,\kappa).
-\end{equation}
-%
-In practice, the number of mosquitoes captured in those series that are censored represents a small percentage of total captures (the median is approximately 2\%), and so the effect of the $q(.)$ term above is likely minimal on resultant inferences.
-
-\subsection{Model estimation by MCMC}\label{sec:dissection_MCMC}
-As for the MRR analysis, we used \textit{Stan} software \citep{carpenter2016stan} to sample from the posterior distributions of model parameters.
-
-To judge convergence of the sampling algorithm to the posterior density, we calculated $\hat{R}$ across all Markov chains -- a ratio that compares the between-chain variation to that within each chain that is commonly used to measure convergence in MCMC \citep{gelman1992inference}.
-For each model, we ran 16 independent Markov chains with 200 iterations per chain, discarding the first half of these iterations as warm-up \citep{gelman2014bayesian}.
-After running the algorithm for the given number of observations, $\hat{R}<1.1$ for all model parameters.
-We also ensured that across each MCMC run the number of divergent iterations (that can bias the MCMC away from the true posterior density) was minimal.
-In the majority of cases, the number of divergent iterations was far fewer than 1\% of the total number of samples.
-
-The model used to estimate the lifespan in terms of gonotrophic cycles is provided below.
-
-\begin{minted}{stan}
-data{
-int N; // number of obs
-int K; // number of time series
-int Y[N]; // number captured
-int S[K]; // length of each series
-int threshold[K]; // age of censoring; if -1 then no censoring
-int Pos[K]; // position of start of each series in Y
-int species[K]; // index variable of species
-int nSpecies; // number of species
-}
-
-parameters{
-real<lower=0> PopSize[K];
-real<lower=0> alpha[K];
-real alpha_mean[nSpecies];
-real<lower=0> alpha_sigma[nSpecies];
-real<lower=0> kappa[K];
-real<lower=0> p[nSpecies];
-}
-
-model{
-// Likelihood
-for(i in 1:K){
-  for(t in 2:S[i]){
-    // uncensored
-    if(threshold[i] < 0)
-      Y[Pos[i]+t-1] ~ neg_binomial_2(PopSize[i] *
-			exp(-alpha[i] * (t-1)), kappa[i]);
-    // censoring
-    else{
-      if((t-1) < threshold[i])
-        Y[Pos[i]+t-1] ~ neg_binomial_2(PopSize[i] *
-			exp(-alpha[i] * (t-1)), kappa[i]);
-      else{
-        real lLogProb[20];
-        for(j in 1:20){
-          lLogProb[j] = neg_binomial_2_lpmf(Y[Pos[i]+t-1] |
-			PopSize[i] * exp(-alpha[i] * (t-1)),
-			kappa[i]);
-        }
-        target += log_sum_exp(lLogProb);
-      }
-    }
-  }  
-}
-
-// Priors
-for(i in 1:K){
-  int aSpecies;
-  aSpecies = species[i];
-  Y[Pos[i]] ~ normal(PopSize[i], sqrt(PopSize[i]));
-  alpha[i] ~ lognormal(alpha_mean[aSpecies],
-  			alpha_sigma[aSpecies]);
-  kappa[i] ~ exponential(p[aSpecies]);
-}
-
-alpha_mean ~ normal(-1.2, 1);
-alpha_sigma ~ lognormal(-2, 1);
-p ~ lognormal(1, 1);
-}
-
-generated quantities{
-real alpha_average[nSpecies];
-for(i in 1:nSpecies)
-  alpha_average[i] = lognormal_rng(alpha_mean[i],alpha_sigma[i]);
-}
-\end{minted}
-
-\subsection{Model checking}
-As for the MRR experiments, we simulated from the posterior predictive distribution for each series of dissection data.
-In the majority of the cases, the capture data lay within the 95\% posterior predictive intervals, indicating that our model was a good fit to the data (see Figures \ref{fig:dissection_ppc_all1}, \ref{fig:dissection_ppc_all2} \&  \ref{fig:dissection_ppc_all3} and the attached file, \verb|dissection_ppcs_all.pdf| for the full graphs).
-
-\newgeometry{scale=0.8}
-\thispagestyle{empty}
-{%
-	\centering
-	\includegraphics[page=1,scale=.7]{./Figure_files/dissection_ppcs_all}
-	\captionof{figure}{\textbf{The numbers of mosquitoes recaptured (black lines) versus the 95\% central posterior interval of the posterior predictive distribution (blue shading) for a selection of the dissection series.} The titles correspond to the unique combination of "identifier - physiological-experiment-id" in the database.
-For the rest of the posterior predictive checks for the dissection data models, see the file referenced in the text.}
-	\label{fig:dissection_ppc_all1}
-	\par
-}
-\restoregeometry
-
-\newgeometry{scale=0.8}
-\thispagestyle{empty}
-{%
-	\centering
-	\includegraphics[page=2,scale=.7]{./Figure_files/dissection_ppcs_all}
-	\captionof{figure}{\textbf{The numbers of mosquitoes recaptured (black lines) versus the 95\% central posterior interval of the posterior predictive distribution (blue shading) for a selection of the dissection series.} The titles correspond to the unique combination of "identifier - physiological-experiment-id" in the database.
-For the rest of the posterior predictive checks for the dissection data models, see the file referenced in the text.}
-	\label{fig:dissection_ppc_all2}
-	\par
-}
-\restoregeometry
-
-\newgeometry{scale=0.8}
-\thispagestyle{empty}
-{%
-	\centering
-	\includegraphics[page=3,scale=.7]{./Figure_files/dissection_ppcs_all}
-	\captionof{figure}{\textbf{The numbers of mosquitoes recaptured (black lines) versus the 95\% central posterior interval of the posterior predictive distribution (blue shading) for a selection of the dissection series.} The titles correspond to the unique combination of "identifier - physiological-experiment-id" in the database.
-For the rest of the posterior predictive checks for the dissection data models, see the file referenced in the text.}
-	\label{fig:dissection_ppc_all3}
-	\par
-}
-\restoregeometry
-
-\subsection{Data collection of gonotrophic cycle duration}\label{sec:dissection_gonotrophicData}
-To convert the estimates of lifespan in physiological age into chronological age we require estimates of the duration of the gonotrophic cycle.
-To determine this characteristic we conducted a meta-analysis of previously-published studies that estimate the duration of the gonotrophic cycle.
-A search of the literature using Google Scholar (\url{scholar.google.co.uk}) was performed using the search term: `gonotrophic cycle duration'.
-The list of articles was then supplemented with a list of references discussed by \cite{silver2007mosquito}.
-Based on the abstracts of the resultant list of published studies we then decided whether to search each article for estimates of the duration of the gonotrophic cycle.
-Overall 79 separate estimates of this parameter were found across 42 published articles (see Section \ref{sec:appendix_gonotrophicDurationStudyList} for a list of the studies included in our dataset).
-Along with information about the estimates we also recorded study and series meta-data, including the location of the study, the method used for estimation, the species and genus, as well as the temperatures and/or seasons in which the experiments were carried out.
-The methods used to estimate gonotrophic cycle duration in the literature largely fell into two distinct categories: those based on MRR studies ($n=29$); and those based on observations of mosquitoes in a laboratory setting ($n=42$).
-In the MRR studies estimates are made of the duration of gonotrophic cycles by dissecting recaptured mosquitoes at each time point using the method of \cite{polovodova1949determination} to count ovariolar dilations.
-For the laboratory studies the duration of gonotrophic cycles is determined more directly by observing the time taken for mosquitoes to mate then blood-feed and finally oviposit.
-
-Along with point estimates of the parameter we also collected information about the uncertainty in the estimates (if available).
-In many articles the duration of the gonotrophic cycle was estimated separately for the first versus subsequent cycles, and these estimates were recorded separately.
-Raw estimates of gonotrophic cycle duration were obtained for species across four different genera (\textit{Aedes}, \textit{Anopheles}, \textit{Culex} and \textit{Masonia}), although there was a bias towards \textit{Anopheles} with $n=47$ estimates out of the total of 79 (Fig.
-\ref{fig:dissection_gonotrophicCycleRaw}).
-
-\begin{figure}[ht]
-	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/dissection_gonotrophicCycleRaw.pdf}}
-	\caption{\textbf{Raw gonotrophic cycle duration estimates for A.
-the 1st cycle and B.
-subsequent cycles.} The confidence intervals represent a range of uncertainties since they were not available in a standardised form (see Section \ref{sec:dissection_gonotrophicMethod} for a discussion of how we handle this issue).
-If no statement of uncertainty was given in the estimates then we indicate this by a triangle marker.
-The studies are ordered according to the central estimate of gonotrophic duration.}\label{fig:dissection_gonotrophicCycleRaw}
-\end{figure}
-
-Since we collected data on the method used to estimate the duration of the gonotrophic cycle we show the raw estimates for the two most common approaches: MRR and laboratory (Fig.
-\ref{fig:dissection_gonotrophicCycleRaw_MRRVsLab}).
-From these results it is evident that the estimates from the laboratory studies are, on average, higher than those using the MRR method.
-
-\begin{figure}[ht]
-	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/dissection_gonotrophicCycleRaw_MRRVsLab.pdf}}
-	\caption{\textbf{Raw gonotrophic cycle duration estimates from A.
-the MRR and B.
-from laboratory-based studies.} The confidence intervals represent heterogeneous measures since they were not available in a standardised form (see Section \ref{sec:dissection_gonotrophicMethod} for a discussion of we handle this issue).
-If no statement of uncertainty was given in the estimates then we indicate this by a triangle marker.
-The studies are ordered according to the central estimate of gonotrophic duration.}\label{fig:dissection_gonotrophicCycleRaw_MRRVsLab}
-\end{figure}
-
-\subsection{Statistical analysis of gonotrophic cycle data}\label{sec:dissection_gonotrophicMethod}
-As discussed in Section \ref{sec:dissection_gonotrophicData}, there was considerable study-level heterogeneity in the information provided for the estimates of the gonotrophic cycle duration.
-A number of studies ($n=24$) provided no estimates of uncertainty in gonotrophic cycle duration, whereas the rest gave some indication of confidence or alternatively a range of possible estimates.
-However the types of uncertainty intervals that were specified varied considerably from study to study, from the vague (but common)  `4-6 days' to the more helpful `5$\pm 1$ day (95\% confidence interval)'.
-
-The heterogeneous nature of the gonotrophic cycle estimates requires a method that explicitly accounts for this characteristic of the data.
-We decided to model the estimates as representing quantiles from an underlying normal distribution that represents uncertainty over possible durations of the gonotrophic cycle.
-In those circumstances where lower, central and upper bounds were given explicitly the data was fairly symmetric and so we believe assuming the observations come from an unskewed normal distribution is reasonable.
-Explicitly we treated each type of uncertainty interval as follows,
-
-\begin{itemize}
-	\item Simple range, for example, `4-6 days': treat the lower and upper bounds as the 2.5\% and 97.5\% quantiles of a normal distribution.
-	\item Confidence intervals, `5$\pm 1$ day (95\% confidence interval)': treat the lower and upper bounds as the relevant quantiles of a normal distribution.
-	\item Point estimates, for example `5 days': treat this as the median of a normal distribution.
-\end{itemize}
-
-The benefit of this approach is that we can convert the quantiles from our normal distribution parameterised by a mean ($\mu$) and standard deviation ($\sigma$) to equivalent quantiles from a standard normal,
-%
-\begin{align}
-Z_{50} = \frac{G_{50} - \mu}{\sigma},
-\end{align}
-%
-where $Z_{50}$ and $G_{50}$ indicate 50\% quantiles from a standard normal distribution and a normal distribution respectively. By manipulating the above expression we obtain the following,
-%
-\begin{align}
-G_{50} = \sigma Z_{50} + \mu,
-\end{align}
-%
-which forms a straight line in $(Z,G)$ space with slope $\sigma$ and y-intercept $\mu$.
-Therefore by plotting the raw observations of $(Z,G)$ quantiles then estimating a linear regression line we can characterise the underlying $\mathcal{N}(\mu,\sigma)$ distribution from which we assume they are drawn.
-
-Due to the relative unavailability of estimates for gonotrophic cycle duration across the different genera we generated a single distribution to represent the uncertainty in this parameter by pooling all the data.
-This was justified because there was not found to be significant variation in the 1st gonotrophic cycle duration (ANOVA: $F_{3,75}=2.6,p>0.05$).
-
-Using the collected data, we estimated that uncertainty in the first gonotrophic cycle duration is estimated as $\mathcal{N}(4.26, 0.38)$ and, for subsequent cycles, $\mathcal{N}(3.85, 0.35)$.
-
-\subsection{Conversion of lifespan from physiological to calendar age}\label{sec:dissection_conversion}
-The estimates of lifespan produced from analysing the dissection data are in terms of physiological age (the number of gonotrophic cycles undertaken).
-To allow comparison with the estimates from the MRR studies (Fig.
-\ref{fig:mrr_lifespans}) as well as produce more useful estimates to inform disease transmission dynamical models we convert these estimates to calendar days.
-To do this we use the estimated parameters of the normal densities that we have assumed represent uncertainty in gonotrophic cycle duration (see Section \ref{sec:dissection_gonotrophicMethod}) and use them to convert our posterior samples of mean physiological lifespan $(L^p_1,...,L^p_S)$ to lifespan in calendar ages $(L^c_1,...,L^c_S)$.
-To do this we iterate the following for all $i\in (1,...,S)$ posterior samples,
-
-\begin{enumerate}
-	\item Sample $G_{1i} \sim \mathcal{N}(\mu_{1i},\sigma_{1i})$, to obtain a duration for the 1st gonotrophic cycle.
-	\item Sample $G_{2i} \sim \mathcal{N}(\mu_{2i},\sigma_{2i})$, to obtain a duration for subsequent gonotrophic cycles.
-	\item If $L^p_i > 1$, the mean lifespan is longer than one gonotrophic cycle:
-	\subitem then $L^c_i = G_{1i} + G_{2i} \times (L^p_i - 1)$.
-	\item Else:
-	\subitem then $L^c_i = G_{1i}\times L^p_i$.
-\end{enumerate}
-
-In the using the above methodology to convert from physiological to calendar age we implicitly assume that for each gonotrophic cycle after the first that the cycles are of the same length (we choose one $G_{2i}$ per sample).
-However since we allow a different $G_{2i}$ for each sample we nonetheless believe that the above approach allows for sufficient uncertainty in the estimates of subsequent gonotrophic cycle durations.
-If instead we believed that significant variation in gonotrophic cycle occurred within a particular mosquito's life (as well as between mosquitoes) then we could draw a new value of $G_{2i}$ for each subsequent cycle.
-This produces results with a little more uncertainty than previously (data not shown).
-
-\section{Results}
-\subsection{MRR experments}
+\section{MRR experments}
 Figures \ref{fig:mrr_lifeSpanVsRange} and \ref{fig:mrr_lifeSpanVsTrapDensity} show the estimates of lifespan versus the trapping range and spatial density of traps at study location.
@@
 For the overall model, the feeding-specific effects were estimated as independent random effects across all studies, with a standard normal prior.
-
-\begin{figure}[h]
-	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_lifeSpanVsRange.pdf}}
-	\caption{\textbf{Posterior estimates of mean mosquito lifespan for each time-series versus the trapping range.} The markers show the median posterior estimates, with the lower and upper bounds indicating the 25\% and 75\% quantiles, respectively.
-The black line shows a linear regression line estimated using the median posterior lifetimes, with the grey shading indicating 95\% confidence intervals.
-All estimates were obtained using the non-hierarchical exponential survival model.}
-	\label{fig:mrr_lifeSpanVsRange}
-\end{figure}
-
-\begin{figure}[h]
-	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_lifespanVsTrapDensity.pdf}}
-	\caption{\textbf{Posterior estimates of mean mosquito lifespan for each time-series versus the trap density.} The markers show the median posterior estimates, with the lower and upper bounds indicating the 25\% and 75\% quantiles, respectively.
-The trap density was estimated assuming traps were contained within a circular area with radius equal to the trapping range.
-The black line shows a linear regression line estimated using the median posterior lifetimes, with the grey shading indicating 95\% confidence intervals.
-All estimates were obtained using the non-hierarchical exponential survival model.}
-	\label{fig:mrr_lifeSpanVsTrapDensity}
-\end{figure}
-
-\begin{figure}[h]
-	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_ThreeSpeciesVersusTemperature.pdf}}
-	\caption{\textbf{Posterior estimates of mean mosquito lifespan for each time-series versus the average monthly temperature for the four species with the most data.} The markers show the median posterior estimates, with the lower and upper bounds indicating the 25\% and 75\% quantiles, respectively.
-The black line shows a regression line with linear and quadratic terms estimated using the median posterior lifetimes, with the grey shading indicating 95\% confidence intervals.
-All estimates of lifespan were obtained using the non-hierarchical exponential survival model.}\label{fig:mrr_ThreeSpeciesVersusTemperature}
-\end{figure}
-
-\begin{figure}[h]
-	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_female_blood_sugar.pdf}}
-	\caption{\textbf{Posterior estimates of female mosquito mean lifespan according to pre-release feeding status across species, genus and overall groupings as determined from the MRR data.} The lifespans shown are for mosquitoes that were not fed with sugar or blood (for females) before release.
-The middle line in each box shows the median estimates and the solid dot indicates the mean.
-The left and right box edges show the 25\%, and 75\% posterior quantiles respectively.
-The whiskers show the range of the data, excluding points lying more than 1.5 times the interquartile range away from each edge of the box.
-The numbers before the start of the left whisker indicate the number of individual time-series within each species.
-All estimates were obtained using the hierarchical exponential survival model.}
-	\label{fig:mrr_female_blood_sugar}
-\end{figure}
-
-\begin{figure}[h]
-	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_genus_vs_species_vs_top.pdf}}
-	\caption{\textbf{The estimated predictive accuracy of the species- and overall-level models relative to the genus-level model.} The predictive accuracy was determined using K-Fold cross-validation as described in SOM; the measure of accuracy presented here by the central dots is the difference in estimated expected log pointwise predictive density compared to the exponential model.
-The lower and upper whiskers represent the lower and upper bounds of the 95\% confidence interval in predictive accuracy.}
-	\label{fig:mrr_genusTopLevel}
-\end{figure}
-
-\begin{figure}[ht]
-	\centerline{\includegraphics[width=1.25\textwidth]{./Figure_files/mrr_mcPowerAnalysis.pdf}}
-	\caption{\textbf{The average error in predicting mean lifespan as a function of study length (left) and number of released mosquitoes (right) in a Monte Carlo analysis.} For each parameter set we generated 400 simulated data series using the negative binomial sampling model with an exponential survival function as described in the text, and estimated the mean lifespan using maximum likelihood.
-The error bars show the standard deviation in the prediction error.}
-	\label{fig:mrr_mcPowerAnalysis}
-\end{figure}
-
-\begin{figure}[ht]
-	\centerline{\includegraphics[width=1.25\textwidth]{./Figure_files/mrr_mcPowerAnalysis_senescence.pdf}}
-	\caption{\textbf{The statistical power to detect senescence for A.
-three case study populations as a function of B.
-study length and C.
-number released.} For each parameter set we generated 500 simulated data series using the negative binomial sampling model with a gompertz hazard function ($\lambda(t) = \alpha e^{\beta t}$), and estimated the $\beta$ parameter using maximum likelihood and tested the null hypothesis $\beta=0$ (constant mortality risk) against the alternative $\beta>0$ using an likelihood ratio test with a 5\% test size.
-The error bars show the standard deviation in the power for each parameter set.
-The recapture probability parameters ($\psi$ and $\kappa$) used were the averages of those estimated for the actual data.}
-	\label{fig:mrr_mcPowerAnalysis_senescence}
-\end{figure}
 
@@
 
-\subsection{Dissection studies}
-Tables \ref{tab:dissection_estimated_lifespans} and \ref{tab:dissection_estimated_lifespans_chron} show posterior summaries of lifespan estimated from the dissection analysis, in terms of gonotrophic cycles and chronological time, respectively.
+\begin{figure}[h]
+	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_genus_vs_species_vs_top.pdf}}
+	\caption{\textbf{The estimated predictive accuracy of the species- and overall-level models relative to the genus-level model.} The predictive accuracy was determined using K-Fold cross-validation as described in SOM; the measure of accuracy presented here by the central dots is the difference in estimated expected log pointwise predictive density compared to the exponential model.
+The lower and upper whiskers represent the lower and upper bounds of the 95\% confidence interval in predictive accuracy.}
+	\label{fig:mrr_genusTopLevel}
+\end{figure}
+
+\begin{figure}[h]
+	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_lifeSpanVsRange.pdf}}
+	\caption{\textbf{Posterior estimates of mean mosquito lifespan for each time-series versus the trapping range.} The markers show the median posterior estimates, with the lower and upper bounds indicating the 25\% and 75\% quantiles, respectively.
+The black line shows a linear regression line estimated using the median posterior lifetimes, with the grey shading indicating 95\% confidence intervals.
+All estimates were obtained using the non-hierarchical exponential survival model.}
+	\label{fig:mrr_lifeSpanVsRange}
+\end{figure}
+
+\begin{figure}[h]
+	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_lifespanVsTrapDensity.pdf}}
+	\caption{\textbf{Posterior estimates of mean mosquito lifespan for each time-series versus the trap density.} The markers show the median posterior estimates, with the lower and upper bounds indicating the 25\% and 75\% quantiles, respectively.
+The trap density was estimated assuming traps were contained within a circular area with radius equal to the trapping range.
+The black line shows a linear regression line estimated using the median posterior lifetimes, with the grey shading indicating 95\% confidence intervals.
+All estimates were obtained using the non-hierarchical exponential survival model.}
+	\label{fig:mrr_lifeSpanVsTrapDensity}
+\end{figure}
+
+\begin{figure}[h]
+	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_female_blood_sugar.pdf}}
+	\caption{\textbf{Posterior estimates of female mosquito mean lifespan according to pre-release feeding status across species, genus and overall groupings as determined from the MRR data.} The lifespans shown are for mosquitoes that were not fed with sugar or blood (for females) before release.
+The middle line in each box shows the median estimates and the solid dot indicates the mean.
+The left and right box edges show the 25\%, and 75\% posterior quantiles respectively.
+The whiskers show the range of the data, excluding points lying more than 1.5 times the interquartile range away from each edge of the box.
+The numbers before the start of the left whisker indicate the number of individual time-series within each species.
+All estimates were obtained using the hierarchical exponential survival model.}
+	\label{fig:mrr_female_blood_sugar}
+\end{figure}
+
+\begin{figure}[h]
+	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_lifeSpanVsTemperature.pdf}}
+	\caption{\textbf{Posterior estimates of mean mosquito lifespan for each time-series versus the average monthly temperature for each study location and date.} The markers show the median posterior estimates, with the lower and upper bounds indicating the 25\% and 75\% quantiles, respectively.
+The black line shows a regression line with linear and quadratic terms estimated using the median posterior lifetimes, with the grey shading indicating 95\% confidence intervals.
+All estimates of lifespan were obtained using the non-hierarchical exponential survival model.}\label{fig:mrr_temperature}
+\end{figure}
+
+\begin{figure}[h]
+	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_ThreeSpeciesVersusTemperature.pdf}}
+	\caption{\textbf{Posterior estimates of mean mosquito lifespan for each time-series versus the average monthly temperature for the four species with the most data.} The markers show the median posterior estimates, with the lower and upper bounds indicating the 25\% and 75\% quantiles, respectively.
+The black line shows a regression line with linear and quadratic terms estimated using the median posterior lifetimes, with the grey shading indicating 95\% confidence intervals.
+All estimates of lifespan were obtained using the non-hierarchical exponential survival model.}\label{fig:mrr_ThreeSpeciesVersusTemperature}
+\end{figure}
+
+\section{Dissection analysis}
+
+\begin{figure}[h]
+	\centerline{\includegraphics[width=1.3\textwidth]{./Figure_files/dissection_individualEstimates_allSpecies.pdf}}
+	\caption{\textbf{Individual time-series estimates of adult mosquito mean lifespan ordered by species median as determined from the dissection data.} The middle line in each box shows the median estimates.
+The left and right box whiskers show the 25\%, and 75\% posterior quantiles respectively.
+All estimates were obtained using the non-hierarchical exponential survival model.}
+	\label{fig:dissection_individualEstimates_allSpecies}
+\end{figure}
 
@@
 \end{table}
+
+\begin{figure}[ht]
+	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/dissection_gonotrophicCycleRaw.pdf}}
+	\caption{\textbf{Raw gonotrophic cycle duration estimates for A.
+the 1st cycle and B.
+subsequent cycles.} The confidence intervals represent a range of uncertainties since they were not available in a standardised form (see Section \ref{sec:dissection_gonotrophicMethod} for a discussion of how we handle this issue).
+If no statement of uncertainty was given in the estimates then we indicate this by a triangle marker.
+The studies are ordered according to the central estimate of gonotrophic duration.}\label{fig:dissection_gonotrophicCycleRaw}
+\end{figure}
+
+\begin{figure}[ht]
+	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/dissection_gonotrophicCycleRaw_MRRVsLab.pdf}}
+	\caption{\textbf{Raw gonotrophic cycle duration estimates from A.
+the MRR and B.
+from laboratory-based studies.} The confidence intervals represent heterogeneous measures since they were not available in a standardised form (see Section \ref{sec:dissection_gonotrophicMethod} for a discussion of we handle this issue).
+If no statement of uncertainty was given in the estimates then we indicate this by a triangle marker.
+The studies are ordered according to the central estimate of gonotrophic duration.}\label{fig:dissection_gonotrophicCycleRaw_MRRVsLab}
+\end{figure}
 
@@
 
-\subsection{EIPs of vector species of malaria, dengue fever, chikungunya and Zika}
-The following anopheline species were identified as vector species of malaria,
+\section{Evidence for senescence}
 
-\begin{itemize}
-	\item \textbf{Africa}: \textit{A. gambiae s.l.}, \textit{A. funestus} and \textit{A. melas} \citep{sinka2012global}, and \textit{A. rivulorum} \citep{wilkes1996anopheles}.
-	\item \textbf{Americas}: \textit{A.
-albimanus}, \textit{A.
-albitarsis}, \textit{A.
-darlingi}, \textit{A.
-pseudopunctipennis} and \textit{A.
-quadrimaculatus} \citep{sinka2012global}; \textit{A.
-bellator}  \citep{forattini1999role,lorenz2012morphometrical}; \textit{A.
-cruzii} \citep{lorenz2012morphometrical}; and \textit{A.
-vestitipennis} \citep{sinka2010dominant}.
-	\item \textbf{Europe and the Middle-East}: \textit{A. sergentii} \citep{sinka2012global}; and \textit{A. maculipennis} \citep{hackett1935varieties}.
-	\item \textbf{Asia}: \textit{A.
-farauti s.l.}, \textit{A.
-koliensis}, \textit{A.
-lesteri}, \textit{A.
-maculatus}, \textit{A.
-minimus}, \textit{A.
-punctulatus}, \textit{A.
-stephensi} and \textit{A.
-subpictus s.l.} \citep{sinka2012global}; and \textit{A.
-culicifacies} \citep{green1980chromosomal}.
-\end{itemize}
+\begin{figure}[ht]
+	\centerline{\includegraphics[width=1.25\textwidth]{./Figure_files/mrr_mcPowerAnalysis_senescence.pdf}}
+	\caption{\textbf{The statistical power to detect senescence for A.
+three case study populations as a function of B.
+study length and C.
+number released.} For each parameter set we generated 500 simulated data series using the negative binomial sampling model with a gompertz hazard function ($\lambda(t) = \alpha e^{\beta t}$), and estimated the $\beta$ parameter using maximum likelihood and tested the null hypothesis $\beta=0$ (constant mortality risk) against the alternative $\beta>0$ using an likelihood ratio test with a 5\% test size.
+The error bars show the standard deviation in the power for each parameter set.
+The recapture probability parameters ($\psi$ and $\kappa$) used were the averages of those estimated for the actual data.}
+	\label{fig:mrr_mcPowerAnalysis_senescence}
+\end{figure}
 
-For dengue fever, chikungunya and Zika, the main vector species are \textit{Ae. aegypti} and \textit{Ae. albopictus} \citep{kraemer2015global,grard2014zika,benelli2016declining}.
-
-\subsection{The statistical power of MRR studies}\label{sec:mrr_discussion_MonteCarlo}
+\section{Power analysis of MRR studies}
 Consider an ideal MRR experiment, where the assumptions of no emigration and harmless marking are fully satisfied.
@@
 
-\section{Studies included in the MRR meta-analysis}\label{sec:appendix_mrrStudyList}
-The following is the subset of studies from the original \cite{guerra2014global} database which were used in the MRR meta-analysis of Chapter \ref{chap:mrr}: \cite{marini2010study,baber2010population,lacroix2009dispersal,maciel2008calculating,midega2007estimating,maciel2007daily,elizondo2006gonotrophic,ba2005aspects,fabian2005mark,la2004anopheles,watson2000aedes,harrington2001analysis,tsuda2001movement,muir1998aedes,toure1998mark,quinones1997anopheles,costantini1996density,trpis1995estimates,jensen1994comparison,fernandez1994gonotrophic,jaal1992mark,rodriguez1992gonotrophic,chiang1991capture,jensen1991assessment,eldridge1990daily,macdonald1968mark,pumpuni1989population,charlwood1987mark,charlwood1989capture,birley1989effect,arredondo1998gonotrophic,hii1990estimation,renshaw1994host,milby1989estimation,maciel2007body,loong1990survival,nelson1980dispersal,maciel2006movement,mcdonald1977population,curtis1980preliminary,rawlings1982dispersal,conway1974population,reisen1979anopheles,nelson1978estimates,rawlings1981influence,sempala1981ecology,takagi1995movement,buei1980field,eyles1943experiment,ordonez2001use,pant1973field,charlwood1988evidence,reisen1982anopheles,nayar1980quantitative,carnevale1979etude,eyles1946long,reisen1984impact,charlwood1986capture,trpis1975demonstration,lutwama1994mark,wada1969dispersal,takken1998dispersal,abdel1966study,valerio2012dispersal,zetek1915behavior,takagi1995movement,yasuno1975migration,eyles1943measurement,germain1974evaluation}.
-
-\section{Studies included in the dissection study meta-analysis}\label{sec:appendix_dissectionStudyList}
-The following is a list of studies included in the dissection study meta-analysis of Chapter \ref{chap:dissection}: \cite{catangui1971studies,charlwood1979studies,chang1991comparative,charlwood2000dry,edalat2015vectorial,de2011survivorship,lines1991human,magesa1991trial,lines1991monitoring,forattini1996studies,samarawickrema1967study,samarawickrema1968biting,de2007parity,charlwood1985assessing,chandra1996age,hoc1995ovariole,vythilingam1997abundance,russell1986culex,chadee2010oviposition,ebsary1977biting,charlwood1980observations,shriram2011population,mahmood1981duration,samarawickrema1987seasonal,uttah2013physiological,ramaiah1992non,buei1982age,chadee1995seasonal,chandra2008age,foll1966conditions,kanda1975epidemiological,mala2014gonotrophic,shalaby1962studies,world1960preliminary,reisen1980anopheles,detinova1962age,smith1994age,jayanetti1987study,ch2013comparative,pant1962distribution,ghosh2010seasonal,mahanta1999temporal,schlein2012diurnal,penilla2002pteridine,schlein2015decrease,snow1978host,detinova1968age,nathan1981bancroftian,charlwood1985blood,schlein2008approach,beier2012attractive,muller2005plant,surendran2006anopheles,muller2010successful,mendis1998cordon,wilkes1996anopheles,chatterjee2000role,muller2010successful,tuchinda1969diurnal,chan1971life,aigbodion2011pelagia,qualls2015indoor,reisen1981anopheles,gillies1972range,gillies1965study,kay1979age,hitchcock1968age}.
-
-\section{Studies included in the gonotrophic cycle duration meta-analysis}\label{sec:appendix_gonotrophicDurationStudyList}
-The following is a list of studies included in the gonotrophic study meta-analysis of Chapter \ref{chap:dissection}: \cite{gillies1965study,sheppard1969dynamics,de1967time,pant1973field,germain1974evaluation,lowe1973reproductive,fernandez1994gonotrophic,buei1980field,rawlings1982tests,mori1977gonotrophic,sempala1981ecology,reisen1983population,birley1989effect,suzuki1978preliminary,charlwood1987mark,charlwood1986capture,charlwood1979studies,charlwood1995rise,chang1991comparative,edalat2015vectorial,de2011survivorship,samarawickrema1968biting,samarawickrema1967study,charlwood1985assessing,chandra1996age,russell1986culex,chadee2010oviposition,mahmood1981duration,ahumada2004modeling,kenawy1991development,rajagopalan1980population,chandra2008age,scholl1979aedes,mala2014gonotrophic,afrane2005effects,gillies1953duration,quinones1997anopheles,rua2005laboratory,delatte2009influence,arredondo1998gonotrophic,wong2014sampling,ijumba2002malaria}.
-
- \bibliographystyle{authordate1}
- \bibliography{Malaria}
-\newpage
+\begin{figure}[ht]
+	\centerline{\includegraphics[width=1.25\textwidth]{./Figure_files/mrr_mcPowerAnalysis.pdf}}
+	\caption{\textbf{The average error in predicting mean lifespan as a function of study length (left) and number of released mosquitoes (right) in a Monte Carlo analysis.} For each parameter set we generated 400 simulated data series using the negative binomial sampling model with an exponential survival function as described in the text, and estimated the mean lifespan using maximum likelihood.
+The error bars show the standard deviation in the prediction error.}
+	\label{fig:mrr_mcPowerAnalysis}
+\end{figure}
 

### 2018-11-28 supplementary_materials_methods.tex
Major reformatting following meeting with Ace

\documentclass[12pt]{article}
  \usepackage{geometry}
 \usepackage[round]{natbib}
 \usepackage{graphicx}
 \geometry{a4paper}
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{authblk}
  \usepackage[running]{lineno}
  \usepackage{setspace}
  \usepackage{booktabs}
  \usepackage{tabularx}
  \usepackage{chngpage}
  \usepackage{amsmath}
  \usepackage{amssymb}
  \usepackage[hidelinks]{hyperref}
 \usepackage[textsize=tiny]{todonotes}
\usepackage{pdfpages, caption}
\usepackage{listings}
\usepackage{minted}
\usepackage{framed}
\usepackage{mdframed}
\usepackage[section]{placeins}

\usepackage{natbib} %package for bibliography

\renewcommand{\thefigure}{SM\arabic{figure}}
\renewcommand{\thetable}{SM\arabic{table}}
\renewcommand{\thesection}{SM\arabic{section}}

\newenvironment{problem}[2][Problem]{\begin{trivlist}
		\item[\hskip \labelsep {\bfseries #1}\hskip \labelsep {\bfseries #2.}]}{\end{trivlist}}

\DeclareMathOperator*{\argmax}{arg\,max}

\renewcommand\listingscaption{Stan code}
\definecolor{apricot}{rgb}{0.984,0.81,0.69}
\BeforeBeginEnvironment{minted}{\begin{mdframed}[backgroundcolor=apricot]}
	\AfterEndEnvironment{minted}{\end{mdframed}}
\newcommand{\Lim}[1]{\raisebox{0.5ex}{\scalebox{0.8}{$\displaystyle \lim_{#1}\;$}}}
\newcommand*\widefbox[1]{\fbox{\hspace{2em}#1\hspace{2em}}}
\interfootnotelinepenalty=10000

\usepackage{framed}
\makeatletter
\newenvironment{kframe}{%
	\def\at@end@of@kframe{}%
	\ifinner\ifhmode%
	\def\at@end@of@kframe{\end{minipage}}%
\begin{minipage}{\columnwidth}%
	\fi\fi%
	\def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
		\colorbox{shadecolor}{##1}\hskip-\fboxsep
		% There is no \\@totalrightmargin, so:
		\hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
	\MakeFramed {\advance\hsize-\width
		\@totalleftmargin\z@ \linewidth\hsize
		\@setminipage}}%
{\par\unskip\endMakeFramed%
	\at@end@of@kframe}
\makeatother

\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%

\definecolor{orangeBright}{RGB}{243,146,0}
\definecolor{blueBright}{RGB}{54,169,225}
\definecolor{pinkBright}{RGB}{214,11,82}
\definecolor{shadecolor}{rgb}{.97, .97, .97}
\definecolor{messagecolor}{rgb}{0, 0, 0}
\definecolor{warningcolor}{rgb}{1, 0, 1}
\definecolor{errorcolor}{rgb}{1, 0, 0}
\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX

\usepackage{xr}
\externaldocument{figures}
\externaldocument[S-]{supplementary_materials}

%   \doublespacing
% \raggedright
\title{A Meta-analysis of Longevity Estimates of Mosquito Vectors of Disease: Additional Methods}
\author{Ben Lambert, Ace North, Charles Godfray}

\begin{document}
\maketitle

\section{MRR experiments}
\subsection{Data}
The MRR database in \cite{guerra2014global} contains 393 individual time-series, along with meta-data for a range of factors for each experiment (for example, species, geography, and date of study). We  cleaned the data, making a number of amendments to it (see Section \ref{sec:amendments}), and transformed the data into a form amenable for our statistical model. We removed time-series with fewer than six separate recapture observations, resulting in 238 time series. For the species-level analyses, we required at least two time series per species, resulting in loss of six series, meaning that the data used in the Bayesian hierarchical analysis comprised 232 individual time series (see Section \ref{sec:appendix_mrrStudyList} for a list of the studies included). Tables \ref{tab:mrr_aggregateData} and \ref{tab:mrr_IndividualData} summarise the data used in the Bayesian hierarchical analysis. This data encompassed time-series from MRR experiments across 33 different species and three genera: \textit{Aedes} (91 separate time-series), \textit{Anopheles} (94), and \textit{Culex} (47), and spanning a wide geographical range (Fig. \ref{fig:mrr_lifetimes_map}).

\subsection{Amendments by MRR ID}\label{sec:amendments}
This section describes the amendments made by MRR ID (the unique identifier that \cite{guerra2014global} use for each MRR record). Where there was ambiguity in the data, we attempted to consult the original publication. The amendments to the dataset are described below:

\begin{itemize}
	\item 69 and 70: Since the release was not disaggregated by sex, the recapture data were aggregated across both males and females, resulting in a single time series in each case.
	\item 482: There are actually two series (one with captured mosquitoes, the other with young reared) in the paper \cite[]{rawlings1981influence}, whereas the database contains only one recapture series. In this analysis, both time series are included.
	\item 483: There are three studies in the paper \cite[]{rawlings1981influence}, (by dye colour there is one for `M\&Y', another for white, and a third for green), whereas the database only contains one release series. In this analysis, we analyse all three time series.
	\item 578: A null observation was removed on day 12, since isn't present in the original paper \cite[]{eyles1943experiment}.
	\item 585: It is not clear from the original paper \cite[]{smith1941observations} whether the missing recapture observations on certain days are because no mosquitoes were found on those days, or because no efforts were made to recapture them on those days. We have assumed that the number recaptured were as stated in the database on those days (missing, because no recapture efforts were made on those days).
	\item 586: It was unclear from database whether the missing observations are due to lack of recapture, or no effort. Have assumed that zero mosquitoes were captured on these days as seems most likely (the paper could not be accessed).
	\item 342: Changed number of released from 6,000 to 12,000 following the original publication \cite[]{bryan1991anopheles}.
	\item 246: Moved the observations from captures to recaptures following the original publication \cite[]{arredondo1998gonotrophic}.
	\item 55: Corrected a mistake in release data; the actual number released is 739, not 749 \cite[]{midega2007estimating}.
	\item 478-481: All four of these series give releases not disaggregated by gender, whereas the recaptures are split this way. Have combined the recapture series to yield a single release-recapture series in each case.
	\item 154: Moved captured male series to recaptured following original paper \cite[]{tsuda2001movement}. 
	\item 334: Data starts on day 3, thus have added this number of days to our time variable in this case.
\end{itemize}

The R file \verb|s_clean_data.R| attached cleans the data, makes the above amendments and outputs the data in a form amenable for estimation using Stan \citep{carpenter2016stan}.

\begin{figure}[h]
	\centerline{\includegraphics[width=1.25\textwidth]{./Figure_files/mrr_mapGenusCropped.pdf}}
	\caption{\textbf{The location and number of separate time-series for the MRR studies included in this analysis.} The area of each circle indicates the number of time-series at each study site. The colour shows the genus of the mosquitoes in each study.}
	\label{fig:mrr_lifetimes_map}
\end{figure}

\begin{table}[htbp]
	\centering
	\begin{tabular}{lc}
		\toprule
		\textbf{Data characteristic} & \textbf{Count} \\
		\midrule
		Species & 33 \\
		Genera & 3 \\
		\textit{Anopheles} & 94 \\
		\textit{Aedes} & 91 \\
		\textit{Culex} & 47 \\
		Female-only MRR & 179 \\
		Male-only MRR & 35 \\
		Mixed sex MRR & 18 \\
		Pre-release blood-feeding only MRR & 71 \\
		Pre-release sugar-feeding only MRR & 41 \\
		Pre-release both blood- and sugar-feeding MRR & 4 \\
		Pre-release neither blood- and sugar-feeding MRR & 116 \\
		MRR time-series & 232 \\
		\bottomrule
	\end{tabular}%
	\caption{\textbf{Summary of variables across all MRR time-series.} For the `Species' and `Genera' rows, the counts represent the number of distinct entries in the database; otherwise the counts represent the number of time series.}
	\label{tab:mrr_aggregateData}%
\end{table}%

\begin{table}[htbp]
	\centering
	\footnotesize
	\begin{adjustwidth}{-0.5in}{-0.5in}%
		\begin{tabularx}{1.25\textwidth}{l|ccccc}
			\toprule
			\textbf{Data characteristic} & \textbf{Min} & \textbf{Mean} & \textbf{Median} & \textbf{Max} & \textbf{Standard deviation} \\
			\midrule
			Study duration, days & 6.0   & 11.8  & 10.0  & 71.0    & 4.7 \\
			Number of days on which collections took place & 6.0   & 10.1  & 9.0   & 47    & 9.1 \\
			Number of separate release days & 1.0     & 1.9   & 1.0     & 23    & 3.0 \\
			Number released & 66    & 4,929  & 1,297  & 86,200 & 12,043 \\
			Number recaptured & 2     & 163   & 63    & 4,090  & 399 \\
			Recapture percentage & 0\% & 8.6\% & 5.2\% & 57.1\% & 10.04\% \\
			Age at release, days & 0   & 1.8   & 0   & 13    & 2.8 \\
			\bottomrule
		\end{tabularx}%
		\caption{\textbf{Summary of data from individual MRR time-series.}}
		\label{tab:mrr_IndividualData}%
	\end{adjustwidth}
\end{table}%

\begin{figure}[h]
	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_sexReleasesOverTime.pdf}}
	\caption{\textbf{The number of MRR time-series by date and sex of released mosquitoes.} The numbers represent by-decade totals of the 232 time-series we used in the main analysis.}
	\label{fig:mrr_sexReleasesOverTime}
\end{figure}

\subsection{Statistical model}\label{sec:mrr_statistical}
Data for a typical MRR experiment consists of a single release of $N_{R}$ marked mosquitoes followed by a series of marked mosquito recaptures on subsequent days (Fig. \ref{fig:mrr_exampleMRRSeries} shows three such example series). We model the number $y(t)$ of marked mosquitoes recaptured on day $t$ using a negative binomial sampling model,
%
\begin{equation}\label{eq:NB}
y(t) \sim \text{NB}\left((N_{R} - Y(t-1)) S(t) \psi, \kappa\right),
\end{equation}
%
where $Y(t-1)$ is the cumulative number of mosquitoes caught on all days before $t$, $S(t)$ is the probability that an individual mosquito survives and remains in the study area until time $t$, $\psi$ is the daily recapture probability for an individual mosquito which is assumed constant through time, and $\kappa$ is the time-independent shape parameter of the negative binomial distribution that controls the extent of variance in recapture rate likely due mostly to environmental heterogeneity. We chose a parameterisation of the negative binomial such that its mean is given by $\mu(t) = (N_{R} - Y(t-1)) S(t) \psi$, and its variance by $\sigma(t)^2 = \mu(t) + \frac{\mu(t)^2}{\kappa}$.

For some MRR experiments, there were a number of releases ($q\geq 2$) of marked mosquitoes throughout the duration of the study. In contrast to single releases, with multiple releases over time, it is not in general possible to determine the particular release to which a recaptured mosquito belongs. To avoid the complication of directly inferring this quantity, we choose to represent previous recaptures probabilistically. This results in a slightly different mean to that of the single release model,
%
\begin{equation}
\mu(t) = N_{Released}(1-\psi)^{t-1} S(t) \psi,
\end{equation}
%
where the factor of $(1-\psi)^{t-1}$ represents the probability that a mosquito is \textit{not} recaptured on a previous day. Where experiments consisted of two or more releases occurring at distinct points in time we assumed that recaptures of individual mosquitoes from either batch were independent of one another, although with the same sampling parameters ($\psi$ and $\kappa$). This results in an overall mean composed of the sum of those from all $q$ releases,
%
\begin{equation}
\mu(t) = \mu_1(t) + ... + \mu_q(t).
\end{equation}
%

\begin{figure}[h]
	\centerline{\includegraphics[width=1.25\textwidth]{./Figure_files/mrr_exampleMRRSeries.pdf}}
	\caption{\textbf{Example MRR time-series for three different studies from the \cite{guerra2014global} database.} In all the above experiments there was a single release of a batch of marked mosquitoes.}
	\label{fig:mrr_exampleMRRSeries}
\end{figure}

The simplest assumption is that death and dispersal rates are independent of mosquito age, giving an exponential `survival' function,
%
\begin{equation}
S(t) = e^{-\lambda t},
\end{equation}
%
where $\lambda$ is the sum of the rates of death and dispersal from the study area. To estimate the lifespan of mosquitoes we use this model, partly because of the limited evidence we determine in favour of senescence (Figs. \ref{fig:mrr_elpd} and \ref{fig:dissection_elpd}).

We assume that the number of mosquitoes recaptured on one day is independent of the number recaptured on any other day, apart from their joint dependence on $\lambda$, $\psi$ and $\kappa$. This results in a likelihood of the data of the form,
%
\begin{equation}
\mathcal{L}(y(t_1),y(t_2),...,y(t_R)|\lambda,\psi,\kappa) = \prod\limits_{i=1}^{R} p(y(t_i)|\psi,\lambda,\kappa),
\end{equation}
%
where $R$ is the number of individual days where collections took place, and $p(y(t_i)|\psi,\lambda,\kappa)$ is the probability of recapturing $y(t_i)$ mosquitoes on day $t_i$ determined from a negative binomial distribution with parameters $\mu = e^{-\lambda t} \psi$, and $\kappa$.

\subsection{Individual time-series estimates}\label{sec:MRR_individual_analysis}
We first treat each time-series separately, and estimate individual $(\lambda,\psi,\kappa)$ parameters of the statistical model (described in the previous section) for each time-series. To use a Bayesian methodology we must specify priors on this set of parameters. Here we choose to specify independent priors on each parameter of the form: $\lambda\sim \mathcal{N}(-2.32,1)$, $\psi\sim \text{exp}(50)$ (with $\psi$ constrained to be less than 0.1, a threshold below which the majority of daily recapture probabilities in the data are observed) and $\kappa\sim \textnormal{log-normal}(2,1)$. This prior on the rate parameter of the exponential distribution ($\lambda$) corresponds to a wide range of possible lifespans (Fig. \ref{fig:mrr_individualTimeSeries_priors}A), with a mean of 10 days. It was necessary to use an informative prior on $\psi$ to bound it to sensible values and to prevent the correlation of estimates with $\lambda$ leading to unreasonably short estimates of $\lambda$; the prior on $\kappa$ is fairly uninformative and allows a wide range of values (Fig. \ref{fig:mrr_individualTimeSeries_priors}B,C).

\begin{figure}[h]
	\centerline{\includegraphics[width=1.25\textwidth]{./Figure_files/mrr_individualTimeSeries_priors.pdf}}
	\caption{\textbf{The prior probability distributions used for lifespan (A), recapture probability (B) and dispersal parameter (C) in the individual time-series analysis for the MRR studies.} Note: the prior for the recapture probability was truncated at $\psi=0.1$ since most MRR day one recapture fractions were significantly below this threshold.}
	\label{fig:mrr_individualTimeSeries_priors}
\end{figure}

\subsection{Estimating lifespan at the species, genus and overall groupings}\label{sec:MRR_hierarchical}
We synthesise information from across all MRR experiments to produce more robust estimates of mosquito lifespan than can be obtained from considering the individual time-series separately. However there exists considerable heterogeneity across the experiments. This heterogeneity has two sources. There is that arising from variability in experimental methodology. There is also variability from actual differences in lifespan across the different mosquito cohorts; for example due to genetic differences between mosquito populations or due to climatic differences. Because of this heterogeneity we use a Bayesian hierarchical model which is akin to a `random effects' model in classical statistics. This type of model assumes that there is random variation in parameters at the individual time-series level, although each of the parameters is drawn from a common `population-level' distribution. In our case we separately estimate three different model groupings where the population-level distributions correspond to the species, genus and overall (across all studies) levels respectively (Fig. \ref{fig:mrr_methodDiagramHierarchy}). This allows us to estimate average mosquito lifespan for a given species or genus, or alternatively across all time-series, by independent sampling from the posterior predictive distribution from which the individual $\lambda_i$ are drawn.

\begin{figure}[h]
	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_methodDiagramHierarchy.pdf}}
	\caption{\textbf{A representation of the hierarchical Bayesian model, for the constant mortality model case when considering the species-level grouping.} Here `pdf' indicates `probability density function', and $\lambda_i$, $\psi_i$, and $\kappa_i$ represent the force of mortality for the exponential distribution, daily recapture probability and over-dispersion parameter of the negative binomial distribution for time-series $i$ in the dataset; $\mu_i(t)= \psi_i e^{-\lambda_i t}$ is the mean number of mosquitoes recaptured on day $t$.}
	\label{fig:mrr_methodDiagramHierarchy}
\end{figure}

Whilst we assume that the variation in mosquito mortality across experiments is random in nature, we recognise that there may exist systematic sources of variation that we do not include in our model. To account for this potential source of bias we examined how the individual estimates of lifespan correlated with experimental covariates (trapping method and number or density of traps). The data on trapping method in the Guerra et al., (2014) database is less reliable than other data entries because published MRR studies may use multiple trapping methods and typically do not disaggregate captures by method. When we examined the correlation of trapping method with lifespan estimates, however, we did not find systematic variation by method (data not shown). We determined a positive correlation between trapping density and lifespan estimates (Fig. \ref{fig:mrr_lifeSpanVsTrapDensity}), which indicates the impact of mosquito dispersal on estimated lifespan.

A Bayesian framework requires that we incorporate our pre-analysis beliefs into our estimates through the use of priors for all parameters in a particular model. In our hierarchical model we are required to specify prior distributions at two levels of the analysis. The first of these links the individual time-series estimates with the overarching group-level distributions. The parameters of these group-level distributions are then set prior distributions (Fig. \ref{fig:mrr_methodDiagramHierarchy}). As an example, we suppose that the rate parameter $\lambda_i$ of our exponential survival model in experiment $i$ is given by:
%
\begin{align}
\lambda_i = \text{exp}(&c_i + \delta^m_{\text{genus}[i]} \text{sex}^{\text{m}}_i + \delta^{\text{mix}}_{\text{genus}[i]} \text{sex}^{\text{mix}}_i +\\  & \delta^{\text{sugar}}_{\text{genus}[i]} \text{sugar}_i +  \delta^{\text{blood}}_{\text{genus}[i]} \text{blood}_i),
\end{align}
%
where $\delta^m_{\text{genus}[i]}$ measures a genus-specific effect of male-only releases ($\text{sex}^{\text{m}}_i=1$) versus female-only ($\text{sex}^{\text{m}}_i=0$); $\delta^{\text{mix}}_{\text{genus}[i]}$ is the analogous effect for mixed releases versus female-only; $\delta^{\text{sugar}}_{\text{genus}[i]}$ is the analogous effect for mosquitoes that were sugar-fed prior to release versus those that were not fed; and $\delta^{\text{blood}}_{\text{genus}[i]}$ is the analogous effect for mosquitoes that were blood-fed prior to release versus those that were not fed. The parameter $c_i$ is modelled hierarchically, being drawn from a group-level distribution assumed to be normal,
%
\begin{equation}
c_i \sim  \mathcal{N}(\mu_\lambda,\sigma)
\end{equation}
%
We then specify priors on these group-level parameters, $\mu_\lambda \sim \mathcal{N}(-2.32,1)$ and $\sigma \sim \textnormal{log-normal}(-3,1)$. On all $\delta$ parameters, except $\delta^{\text{mix}}_{\text{genus}[i]}$, we specify standard normal priors. We specify a uniform prior on $\delta^{\text{mix}}_{\text{genus}[i]}\sim U(0,\delta^m_{\text{genus}[i]})$ so that the effect of mixed releases lies between the female-only release effect (0) and the male-only release effect ($\delta^m_{\text{genus}[i]}$).

The relatively complex nature of priors in hierarchical models make it important to determine their influence on inferences. We chose the above -- somewhat uninformative -- priors to allow a range of mosquito lifespans in order to minimise their effect on the estimates we report (Fig. \ref{fig:mrr_meanLifespanPrior}). We also chose to set hierarchical priors on the remaining parameters in our models -- $\psi$ the probability of daily recapture, and $\kappa$ the `over-dispersion' parameter of a negative binomial distribution. For $\psi$ we chose an informative prior that placed all probability mass $\psi\leq 10\%$, since in most experiments the fraction of mosquitoes captured was significantly below this value. For $\kappa$ we chose a fairly wide prior that had most of its probability mass below $\kappa=20$ (Fig. \ref{fig:mrr_PsiKappaPriors}). In all cases the hierarchical priors were chosen to have similar implications on lifespan, recapture probability and over-dispersion at the individual time-series level as for the non-hierarchical analysis described previously.

\begin{figure}[h]
	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_lifespan_priors.pdf}}
	\caption{\textbf{Samples from the prior predictive distribution of mean lifespan across the six models of mosquito mortality introduced in Table \ref{tab:mrr_survivalDescription}.} In all cases 10,000 samples were generated using the priors indicated in Table \ref{tab:mrr_priors}, resulting in distributions with a mean close to 10 days.}
	\label{fig:mrr_meanLifespanPrior}
\end{figure}

\begin{figure}[h]
	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_priors_PsiKappa.pdf}}
	\caption{\textbf{Samples from the priors for daily recapture probability ($\psi$) and over-dispersion parameter ($\kappa$).} In both panels, 10,000 samples were generated using the following hierarchical priors: for the individual time series parameters, $\psi_i \sim \text{exp}(r)$ (with $\psi$ constrained to be less than 0.1 - encompassing the upper limit of day one recapture fractions observed in most data series) and $\kappa_i \sim \text{exp}(p)$; for the top-level parameters: $r\sim \text{gamma}(100, 2)$ and $p\sim \text{log-normal}(1.5, 1)$.}\label{fig:mrr_PsiKappaPriors}
\end{figure}

\subsection{Testing for age-dependent mortality in wild mosquitoes}\label{sec:mrr_age_dependence}
Previous work has found evidence of age-dependent mortality in lab populations \citep{styer2007mosquitoes,dawes2009anopheles}, and less-commonly in wild mosquitoes \citep{clements1981analysis}. Furthermore, recent modelling work has examined the implications of departures from a constant risk of mortality \citep{styer2007mosquitoes,hancock2009age,novoseltsev2012age}. To determine whether senescence occurs in wild mosquitoes we re-estimate our model using survival functions, $S(t)$, that allow for a rate of mortality that can vary with age. Specifically, we re-estimate our model using five other models that were previously used in the literature (see Table \ref{tab:mrr_survivalDescription} for a description of these). Each of these models makes different assumptions about how the rate of death and dispersal varies with age, but all can be represented by a general form,
%
\begin{equation}\label{eq:survivalInt}
S(t) = e^{-\int\limits_{0}^{t}\lambda(\tau) \mathrm{d}\tau}.
\end{equation}
%
where we constrain the parameters of our models to preclude the possibility of a hazard that decreases with age ($\frac{\mathrm{d}\lambda(t)}{\mathrm{d}t} \geq 0$). Whilst this could occur if older mosquitoes disperse less than younger mosquitoes, it is unlikely this would outweigh any declines in survival associated with old age.

The parameters of the survival function of each model were assigned hierarchical priors that allowed considerable variation in mosquito lifespan (see Fig. \ref{fig:mrr_meanLifespanPrior}). Otherwise the statistical model we used remained the same as for the constant mortality case. In testing for age-dependent mortality, we did not account for sex or pre-release feeding effects due to the complexity of including these elements in more complex survival models.

\begin{table}
	\scriptsize
	\noindent\makebox[\textwidth]{%
		\begin{tabularx}{1.35\textwidth}{c|cXcX}
			\textbf{Survival function} & \textbf{Hazard rate} & \textbf{Interpretation} & \textbf{Mean time spent in study area, }\textbf{$\overline{T}$} & \textbf{Papers assuming this function} \\
			\midrule
			Exponential & $\lambda$ & Constant mortality risk & $\frac{1}{\lambda}$ & \cite{ross1910prevention}; \cite{anderson1992infectious}; \cite{smith2004statics} \\
			\midrule
			Gompertz &  $\alpha e^{\beta t}$     & Mortality increases with age at an ever-increasing rate &  $\frac{e^{\frac{\alpha}{\beta}} \Gamma(0,\frac{\alpha}{\beta})}{\beta}$    & \cite{clements1981analysis}; \cite{styer2007mosquitoes};\cite{novoseltsev2012age} \\
			\midrule
			Weibull &  $\alpha t^\beta$; $\;\beta\geq 1$     &  Mortality increases with age at an ever-increasing rate     &   $\alpha^{-\frac{1}{\beta}} \Gamma(1+\frac{1}{\beta})$    & \cite{hancock2009age}; and \cite{carey2001insect} considering general insect demography \\
			\midrule
			Gompertz-Makeham &  $\alpha e^{\beta t} + c$      &  Two additive mortality risks: one that increases with age, and another that is age-independent  &  No simple analytic form     & \cite{styer2007mosquitoes} \\
			\midrule
			Logistic & $\frac{\alpha e^{\beta t}}{1+\frac{\alpha s}{b}\left(e^{b t} - 1\right)}$   & Mortality risk increases with age, although at a declining rate  & No simple analytic form   & \cite{styer2007mosquitoes}; \cite{novoseltsev2012age} \\
			\midrule
			Logistic-Makeham &   $\frac{\alpha e^{\beta t}}{1+\frac{\alpha s}{b}\left(e^{b t} - 1\right)} + c$    & Two separate additive mortality risks: an age-dependent risk of the same form as the logistic model, and a constant hazard  &  No simple analytic form  & \cite{styer2007mosquitoes}; \cite{bellan2010importance} \\
			\bottomrule
	\end{tabularx}}
	\caption{\textbf{A description of the survival functions used in this study, arranged in rough order of model complexity (simple-complex from top-bottom.)} All parameters are defined to be non-negative. $\Gamma(\theta)$ and $\Gamma(\theta_1,\theta_2)$ refer to the Euler gamma function, and incomplete gamma function respectively. The `mean time spent in study area' is an estimate of the combined effects of mosquito mortality and dispersal from the area of the study where collections take place, since our data does not provide spatially-resolved data.}\label{tab:mrr_survivalDescription}
\end{table}

\begin{table}
	\footnotesize
	\noindent\makebox[\textwidth]{%
		\begin{tabularx}{1.2\textwidth}{l|c|X}
			\textbf{Survival function} & \textbf{time-series-level priors} & \textbf{Group-level priors}\\
			\midrule
			Exponential & $c\sim \mathcal{N}(\mu_\lambda,\sigma)$ &$\mu_\lambda \sim N(-2.32,1)\text{,}\; \sigma\sim \textnormal{log-normal}(-3,1)$\\
			\midrule
			Gompertz & $\alpha,\beta\sim$ log-normal$(\mu_{\alpha|\beta},0.2)$ &$\mu_\alpha\sim N(-3,1)\text{,}\; \mu_\beta\sim N(-3,1.1)$\\
			\midrule
			Weibull & $\alpha,(\beta-1)\sim$  log-normal$(\mu_{\alpha|\beta},0.2)$ &$\mu_\alpha\sim N(-4.8,1.75)\text{,}\; \mu_\beta\sim N(-4,2)$\\
			\midrule
			Gompertz-Makeham & $\alpha,\beta,c\sim$ log-normal$(\mu_{\alpha|\beta|c},0.2)$ &$\mu_\alpha\sim N(-3,1)\text{,}\; \mu_\beta\sim N(-3,0.5)\text{,}\; \mu_c\sim N(-4.5,0.5)$\\
			\midrule
			Logistic & $\alpha,\beta,s\sim$  log-normal$(\mu_{\alpha|\beta|s},0.2)$ &$\mu_\alpha\sim N(-2.8,1)\text{,}\; \mu_\beta\sim N(-3,1)\text{,}\; \mu_s\sim N(-3,1)$\\
			\midrule
			Logistic-Makeham & $\alpha,\beta,s,c\sim$ log-normal$(\mu_{\alpha|\beta|s|c},0.2)$ &$\mu_\alpha\sim N(-3.2,1)\text{,}\; \mu_\beta\sim N(-3.2,1)\text{,}\; \mu_s\sim N(-3,0.5)\text{,}\; \mu_c\sim N(-4,0.5)$\\
			\bottomrule
	\end{tabularx}}\caption{\textbf{The priors used on parameters of each different survival model.} For the exponential model the `group-level' priors were the same for the genus and `overall' models that were also estimated. The exponential model was the only model that was simple enough to allow the scale parameter of the log-normal ($\sigma$) to be estimated by the data. The notation $\alpha,\beta\sim \text{log-normal}(\mu_{\alpha|\beta},0.2)$ means that $\alpha$ and $\beta$ were assigned independent log-normal priors with location parameters $\mu_\alpha$ and $\mu_\beta$ respectively, and a scale parameter of 0.2 in both cases.}\label{tab:mrr_priors}
\end{table}

\subsection{Model estimation by MCMC}\label{sec:mrr_MCMC}
The likelihood and priors we use result in posterior distributions whose analytic form cannot be calculated with existent computational methods. Instead we use Markov Chain Monte Carlo (MCMC) methods to sample from each posterior distribution. We used \textit{Stan} software \citep{carpenter2016stan} that implements an efficient MCMC algorithm known as NUTS \citep{hoffman2014no}.

To judge convergence of the sampling algorithm to the posterior density, we calculated $\hat{R}$ across all Markov chains -- a ratio that compares the between-chain variation to that within each chain that is commonly used to measure convergence in MCMC \citep{gelman1992inference}. For each class of model, we used the following MCMC parameters,
%
\begin{itemize}
	\item Individual lifespan models (Section \ref{sec:MRR_individual_analysis}): 12 independent chains with 1000 iterations per chain,
	\item Hierarchical lifespan estimates (Section \ref{sec:MRR_hierarchical}): 12 independent chains with 15,000 iterations per chain,
	\item Age-dependent analysis (Section \ref{sec:mrr_age_dependence}): 12 independent chains with 3000 iterations per chain.
\end{itemize}

In all cases, we discarded the first half of these iterations as warm-up \citep{gelman2014bayesian}. At the end of all runs for lifespan estimation, $\hat{R}<1.1$ for all model parameters. For the age-dependent cases, there remained a handful of parameters where $\hat{R}>1.1$ after running simulations for 3000 iterations but this did not affect our inferences, since repeated runs demonstrated the same pattern. We also ensured that across each MCMC run the number of divergent iterations (that can bias the MCMC away from the true posterior density) was minimal. In the majority of cases, the number of divergent iterations was far fewer than 1\% of the total number of samples.

The Stan scripts for the lifespan estimation (i.e. those using an exponential survival model) for the MRR analyses are provided below. The script for the individual time series analysis (Section \ref{sec:MRR_individual_analysis}):

\begin{minted}{stan}
data {
// Single release parameters
int<lower=0> N; // number of observations
int<lower=0> K; // number of groups
int Y[N]; // observations for the releases
int S[K]; // time series lengths
vector[N] t; // time observations (days) 
int Pos[K]; // starting position of each dataset

// Multiple release parameters
int RelFreq[K]; // frequency of releases in each study
// starting position of releases in each study
// in RelNumber
int PosRel[K];
int<lower=0> NReleases; // total number of releases
// numbers released in each study
vector[NReleases] RelNumber;
// time of each release in each study
vector[NReleases] RelTime;

real Age[K];
}

parameters {
real<lower=0.3, upper=20> kappa[K];
real<lower=0> cDecay[K];
real<lower=0, upper=1> psi[K];
}

model {
// Likelihood
for (i in 1:K) {
real lambda[S[i]];

if (RelFreq[i]< 2) // Single release
{
   real numberCaught;
   numberCaught = 0;	
   for (j in 1:S[i])
   {
     lambda[j]= (psi[i] *
                (RelNumber[PosRel[i]] - numberCaught) *
                exp(-cDecay[i] * (t[Pos[i] + j - 1] + Age[i])));
     //1e-5 to prevent loc=0
     Y[Pos[i] + j - 1] ~ neg_binomial_2(
        		lambda[j] + 1e-5, kappa[i]); 
     numberCaught = numberCaught + Y[Pos[i] + j - 1];
   }
} else // Multiple release
  {
    for (j in 1:S[i])
    {
      real lambdaTemp;			
      lambdaTemp = 0;
      for (kk in 1:RelFreq[i])
      {
        if (t[Pos[i] + j - 1]> RelTime[PosRel[i]+kk-1])
        {
           lambdaTemp = (
                   lambdaTemp +
                   RelNumber[PosRel[i] + kk - 1] * 
                   ((1 - psi[i])^(t[Pos[i] + j - 1] -
                   RelTime[PosRel[i]+kk-1]-1)) *
                   psi[i] * exp(-cDecay[i] * 
                   (t[Pos[i] + j - 1] -
                   RelTime[PosRel[i]+kk-1] + Age[i]))
           );
          }
   }
   lambda[j] = lambdaTemp;
   Y[Pos[i] + j - 1] ~ neg_binomial_2(lambda[j]+1e-5, kappa[i]);
   }
 }
}

// Priors
for (i in 1:K)
{
   cDecay[i] ~ lognormal(-2.32, 1);
   kappa[i] ~ lognormal(2, 1);  
   psi[i] ~ exponential(50);
}
}

generated quantities {
vector[K] lifespans;
  for(i in 1:K)
    lifespans[i] = 1 / cDecay[i];
}
\end{minted}

The script for the hierarchical lifespan estimation (Section \ref{sec:MRR_hierarchical}) is shown below:

\begin{minted}{stan}
data {
// Single release parameters
int<lower=0> N; // number of observations
int<lower=0> K; // number of groups
int Y[N]; // observations
int S[K]; // group sizes
vector[N] t; // time observations (days) 
int Pos[K]; // starting position of each dataset

// Multiple release parameters
int RelFreq[K]; // release frequency in each study
int PosRel[K]; // starting position of releases in each study
int<lower=0> NReleases; // release frequencies
vector[NReleases] RelNumber; // numbers released
vector[NReleases] RelTime; // release times

int<lower=0> nSpecies; // number of individual species
int SpeciesIndex[K]; // species index of each study

real Age[K];
int SexM[K];
int SexMix[K];
int Genus[K];
int Blood[K];
int Sugar[K];
}

transformed data{
int SpeciesToGenus[nSpecies];
for(i in 1:K){
  SpeciesToGenus[SpeciesIndex[i]] = Genus[i];
}
}

parameters {
real<lower=0.3,upper=100> kappa[K];
real a[nSpecies];
real<lower=0> p[nSpecies];
real<lower=0> r[nSpecies];
real cDecay[K];
real<lower=0,upper=0.1> psi[K];
real<lower=0> sigma_a[nSpecies];
real delta_M[3];
real delta_Blood[3];
real delta_Sugar[3];
real<lower=0, upper=1> u_mixed[3];
}

transformed parameters{
real delta_Mixed[3];
for(i in 1:3)
  delta_Mixed[i] = u_mixed[i] * delta_M[i];
}

model {
for (i in 1:K) {
  real lambda[S[i]];
  real decay_temp = exp(cDecay[i]+delta_M[Genus[i]] * SexM[i] +
    delta_Mixed[Genus[i]] * SexMix[i] + delta_Sugar[Genus[i]] *
    Sugar[i] + delta_Blood[Genus[i]] * Blood[i]);

  if (RelFreq[i]< 2) // Single release
  {
    real R_times_psi;
    real numberCaught;
    numberCaught = 0;	
    R_times_psi = psi[i];
    for (j in 1:S[i])
    {
    lambda[j]= R_times_psi *
            (RelNumber[PosRel[i]]-numberCaught) *
            exp(-decay_temp * (t[Pos[i] + j - 1] + Age[i]));
    Y[Pos[i] + j - 1] ~ neg_binomial_2(lambda[j] + 1e-5,
                                       kappa[i]);
    numberCaught = numberCaught + Y[Pos[i] + j - 1];
    }
  }
  else // Multiple release
  {
    for (j in 1:S[i])
    {
    real lambdaTemp;			
    lambdaTemp = 0;
    for (kk in 1:RelFreq[i])
    {
      if (t[Pos[i] + j - 1]> RelTime[PosRel[i] + kk - 1])
      {
        lambdaTemp = (lambdaTemp +
          RelNumber[PosRel[i] + kk - 1] * 
          ((1-psi[i])^(t[Pos[i] + j - 1] -
          RelTime[PosRel[i] + kk - 1] - 1)) *
          psi[i] * exp(-decay_temp *
          (t[Pos[i] + j - 1] -
          RelTime[PosRel[i] + kk - 1] + Age[i]));
         );
        }
      }
      lambda[j] = lambdaTemp;
      Y[Pos[i] + j - 1] ~ neg_binomial_2(lambda[j] + 1e-5,
                                         kappa[i]);
      }
    }
}

// Priors
for (i in 1:K)
{
  int speciesTemp;
  speciesTemp = SpeciesIndex[i];
  cDecay[i] ~ normal(a[speciesTemp], sigma_a[speciesTemp]);
  kappa[i] ~ exponential(p[speciesTemp]);  
  psi[i] ~ exponential(r[speciesTemp]);
}
a ~ normal(-2.32, 1);
sigma_a ~ lognormal(-3, 1);
p ~ lognormal(1.5, 1);
r ~ gamma(100, 2);
delta_M ~ normal(0, 1);
delta_Blood ~ normal(0, 1);
delta_Sugar ~ normal(0, 1);
}

generated quantities{
real overallcDecay[nSpecies];
real overallFLife[nSpecies];
real overallMLife[nSpecies];
real overallMixedLife[nSpecies];
real overallFBloodLife[nSpecies];
real overallFBothLife[nSpecies];
real overallFSugarLife[nSpecies];
real overallMSugarLife[nSpecies];

for(i in 1:nSpecies){
  overallcDecay[i] = normal_rng(a[i], sigma_a[i]);
  overallFLife[i] = 1 / exp(overallcDecay[i]);
  overallMLife[i] = 1 / exp(overallcDecay[i] +
                            delta_M[SpeciesToGenus[i]]);
  overallMixedLife[i] = 1 / exp(overallcDecay[i] +
                            delta_Mixed[SpeciesToGenus[i]]);
  overallFBloodLife[i] = 1 / exp(overallcDecay[i] +
                             delta_Blood[SpeciesToGenus[i]]);
  overallFSugarLife[i] = 1 / exp(overallcDecay[i] +
                             delta_Sugar[SpeciesToGenus[i]]);
  overallFBothLife[i] = 1 / exp(overallcDecay[i] +
                             delta_Sugar[SpeciesToGenus[i]] +
                             delta_Blood[SpeciesToGenus[i]]);
  overallMSugarLife[i] = 1 / exp(overallcDecay[i] +
                             delta_M[SpeciesToGenus[i]] +
                             delta_Sugar[SpeciesToGenus[i]]);
}
}
\end{minted}

\subsection{K-Fold cross validation}\label{sec:mrr_kFold}
One way to estimate a model's out-of-sample predictive capability is to use the Akaike Information Criterion that explicitly penalises a model in accordance to its complexity (typically indexed by its number of free parameters) to correct for model over-fit. The way in which this correction takes places however, is fairly heuristic, and less appropriate for hierarchical models. An alternative method, common in the machine learning literature, is known as `cross-validation' (see for example, \cite{kohavi1995study}). Here to estimate out of sample predictive capability, the original data set is partitioned into training and test sets. The model is then fitted to the training set, and its predictive performance measured on the independent test set.

We use cross-validation to compare the predictive power of the species, genus and overall models, and later to determine whether age-dependent mortality occurs. In particular, we use K-Fold cross validation where we repeatedly partition our dataset into a test set composed of a number of individual time-series, and a training set of the remainder (see for example, \cite{marsland2015machine}). We then use the fitted model to measure the predictive performance on the test set. The particular measure we use is the expected log point-wise predictive density (elpd)\citep{vehtari2015efficient} which sums the predictive performance for each data point averaged across all posterior samples.

Since the models are hierarchical we must specify how to draw parameters for a given test sample time-series. Here we chose to draw values of the individual time-series parameters as samples from the population distribution that corresponds to that particular sample's respective grouping. So if the particular test sample time-series pertains to \textit{Ae. aegypti} (and we are using the species-level model) then we draw independent samples for the statistical model's parameters from the estimated \textit{Ae. aegypti} population distribution,
%
\begin{align}
\theta_i \sim p(\theta^{\text{Ae. aegypti}}|data),
\end{align}
%
where $\theta_i$ is the value of the parameters used in test time-series $i$, and $p(\theta^{\text{Ae. aegypti}}|data)$ is the overall posterior distribution across all \textit{Ae. aegypti} experiments.

\subsection{Model checking}
To check the fit of the model to the data, we simulated from the posterior predictive distribution for each data point within each time series of recapture observations (see Figures \ref{fig:mrr_ppc1}, \ref{fig:mrr_ppc2} \&  \ref{fig:mrr_ppc3} and the attached file, \verb|mrr_ppcs_all.pdf| for the full graphs). In the majority of cases, the data lay within the 95\% predictive intervals indicating that the model was a good fit to the data.

{%
	\includegraphics[page=1,scale=.6]{./Figure_files/mrr_ppcs_all}
	\captionof{figure}{The numbers of mosquitoes recaptured (black lines) versus the 95\% central posterior interval of the posterior predictive distribution (blue shading) for a selection of the time series. For the rest of the posterior predictive checks for the MRR models, see the file referenced in the text.}
	\label{fig:mrr_ppc1}
	\par

{%
	\includegraphics[page=2,scale=.6]{./Figure_files/mrr_ppcs_all}
	\captionof{figure}{The numbers of mosquitoes recaptured (black lines) versus the 95\% central posterior interval of the posterior predictive distribution (blue shading) for a selection of the time series. For the rest of the posterior predictive checks for the MRR models, see the file referenced in the text.}
	\label{fig:mrr_ppc2}
	\par
}

{%
	\includegraphics[page=3,scale=.6]{./Figure_files/mrr_ppcs_all}
	\captionof{figure}{The numbers of mosquitoes recaptured (black lines) versus the 95\% central posterior interval of the posterior predictive distribution (blue shading) for a selection of the time series. For the rest of the posterior predictive checks for the MRR models, see the file referenced in the text.}
	\label{fig:mrr_ppc3}
	\par
}

\section{Dissection based estimates}

\subsection{Collection of dissection data}\label{sec:dissection_dissectionData}
A comprehensive search of the literature using Google Scholar (\url{scholar.google.co.uk}) was performed using various combinations of the following keywords: dissection, mosquito, parity, parous, age, and physiological age. No constraints were placed on publication date, location or type. This list was supplemented with a number of author-specific searches for those individuals most prevalent in the literature. In particular we searched for all articles authored by: Charlwood, Muller, Schlein, Samarawickrema, Reisen, Detinova, Polodova, Gillies, and Wilkes. An additional list of potential articles was provided by doing a forward article search on some of the most highly-cited articles in the database: \cite{polovodova1949determination,detinova1962age,gillies1965study,clements1981analysis}. The list of results was then filtered manually by examining the titles and abstracts to produce a candidate list of the articles most likely to contain data on the physiological age of dissected mosquitoes caught in the wild, as determined by the \cite{polovodova1949determination} method.

A relational database was used to store the raw data from the actual experiments, along with the meta-data associated with each of the experiments. In many of the published articles wild mosquitoes were caught and dissected over a period of time, and the raw data thus consists of snapshots of the age structure of the population at regular intervals in time. In the cases where the data was more sparse (fewer than ten individuals, on average, per date), we aggregated across dates and recorded this as a single entry; otherwise we recorded the snapshots of the population at each different date. We record separate series for each species that was captured, or for those that were recorded at separate capture locations (potentially with an alternative collection method), and do not aggregate over these datasets.

For each individual series we recorded the following meta-data: genus, species, collection method, whether or not insecticide was mentioned as being used during or shortly before the time of mosquito collections, and the start and end dates of the experiment. At the article level we recorded the following meta-data: author, year, title, country, location (within country), start and end date, whether insecticide was used during any of the experimental replicates, whether a mrr experiment occurred alongside the dissection study, and the collection location (indoor or outdoor). At either the individual series or article levels we record additional meta-data describing the nature of data collection, for example explaining where the data was contained within the article, whether it was obtained by digitising graphs, and the number of separate dated series. For those few cases ($n=16$ series) where the data was obtained by digitising graphs, we used the WebPlotDigitizer online tool \citep{digitise}.

The data collection method resulted in 568 separate dissection series, across 72 published articles (see Section \ref{sec:appendix_dissectionStudyList} for a list of the studies included in our final dataset). The published datasets cover the period from 1960-2015 with comparable numbers of studies across each decade (Fig. \ref{fig:dissection_timeSeries}). The statistical approach applied to the data relies on the assumption that there is a constant rate of recruitment into the adult population (see Section \ref{sec:dissection_dissectionStats}). If populations fluctuate strongly from month-to-month this assumption will likely be violated. To try to mitigate against such a risk, we aggregate the data across all dates to obtain a single series for each identifier, resulting in 201 such series. To obtain a reasonable level of accuracy on estimates we remove all those individual (aggregated) series where there are fewer than 100 mosquitoes in total in the series. Finally we remove data for any species where there was only one series resulting in 131 series across four genera (\textit{Anopheles}, \textit{Culex}, \textit{Aedes}, and \textit{Mansonia}) and 25 species (Table \ref{tab:dissection_speciesNumbers}). These studies are distributed across a wide range of geographies (Fig. \ref{fig:dissection_map}). The raw data for the dissection analysis are shown in Figures \ref{fig:dissection_exampleData1}, \ref{fig:dissection_exampleData2} and \ref{fig:dissection_exampleData3}.

\begin{figure}[ht]
	\centerline{\includegraphics[width=0.8\textwidth]{./Figure_files/dissection_timeSeries.pdf}}
	\caption{\textbf{The numbers of published studies that estimate gonotrophic age by dissection across the study period.}}\label{fig:dissection_timeSeries}
\end{figure}

\begin{figure}[ht]
	\centerline{\includegraphics[width=1.25\textwidth]{./Figure_files/dissection_map.pdf}}
	\caption{\textbf{The geographic location of each of the dissection databases included in the meta-analysis.} The area of the bubbles indicates the number of unique data series available.}\label{fig:dissection_map}
\end{figure}

\begin{table}[htbp]
	\centering
	\begin{tabular}{lll}
		\toprule
		\textbf{Genus} & \textbf{Species} & \textbf{Frequency} \\
		\midrule
		\textit{Anopheles} & \textit{gambiae s.l.} & 19 \\
		\textit{Culex} & \textit{quinquefasciatus} & 12 \\
		\textit{Anopheles} & \textit{maculipennis} & 11 \\
		\textit{Anopheles} & \textit{farauti s.l.} & 10 \\
		\textit{Anopheles} & \textit{sergentii} & 8 \\
		\textit{Aedes} & \textit{polynesiensis} & 8 \\
		\textit{Culex} & \textit{pipiens} & 6 \\
		\textit{Anopheles} & \textit{culicifacies} & 5 \\
		\textit{Anopheles} & \textit{darlingi} & 5 \\
		\textit{Anopheles} & \textit{quadrimaculatus} & 5 \\
		\textit{Anopheles} & \textit{stephensi} & 4 \\
		\textit{Anopheles} & \textit{melas} & 4 \\
		\textit{Culex} & \textit{annulirostris} & 4 \\
		\textit{Aedes} & \textit{aegypti} & 3 \\
		\textit{Aedes} & \textit{samoanus} & 3 \\
		\textit{Anopheles} & \textit{minimus} & 3 \\
		\textit{Anopheles} & \textit{rivulorum} & 3 \\
		\textit{Culex} & \textit{thalassius} & 3 \\
		\textit{Mansonia} & \textit{uniformis} & 3 \\
		\textit{Anopheles} & \textit{subpictus} & 2 \\
		\textit{Aedes} & \textit{sollicitans} & 2 \\
		\textit{Aedes} & \textit{vexans} & 2 \\
		\textit{Anopheles} & \textit{bellator} & 2 \\
		\textit{Anopheles} & \textit{cruzii} & 2 \\
		\textit{Culex} & \textit{tritaeniorhynchus} & 2 \\
		\bottomrule
		\textit{\textbf{Anopheles}} &       & \textbf{83} \\
		\textit{\textbf{Culex}} &       & \textbf{27} \\
		\textit{\textbf{Aedes}} &       & \textbf{18} \\
		\textit{\textbf{Mansonia}} &       & \textbf{3} \\
		\bottomrule
		\textbf{Total} &  & \textbf{131} \\
	\end{tabular}%
	\caption{\textbf{The numbers of dissection series for each species or genus included in the overall dataset.}}\label{tab:dissection_speciesNumbers}%
\end{table}%

\begin{table}[htbp]
	\centering
	\footnotesize
	\begin{adjustwidth}{-0.5in}{-0.5in}%
		\begin{tabularx}{1.25\textwidth}{l|ccccc}
			\toprule
			\textbf{Variable} & \textbf{Min} & \textbf{Median} & \textbf{Mean} & \textbf{Max} & \textbf{Standard deviation} \\
			\midrule
			Min age of captures, gonotrophic cycles & 0  & 0  & 0.02  & 1.00  & 0.12 \\
			Median age of captures, gonotrophic cycles & 0  & 1.00  & 0.63  & 2.00  & 0.62 \\
			Mean age of captures, gonotrophic cycles & 0.17  & 0.90  & 1.00  & 3.03  & 0.58 \\
			Max age of captures, gonotrophic cycles & 1.00  & 5.00  & 5.77  & 13.00 & 2.94 \\
			Number captured & 100 & 565 & 1,317 & 14,012 & 1,885 \\
			\bottomrule
		\end{tabularx}%
	\end{adjustwidth}
	\caption{\textbf{Summaries of the characteristics of the 131 individual dissection data series that are included in this analysis.} Note when series are censored (see Section \ref{sec:dissection_censored}) we assume that all mosquitoes dissected over the threshold are of this age when calculating summary statistics.}\label{tab:dissection_summaryStats}
\end{table}

\begin{figure}[ht]
	\centerline{\includegraphics[width=1.3\textwidth]{./Figure_files/dissection_parity_data1.png}}
	\caption{\textbf{The normalised physiological age series for nine species in the database.} Each different coloured line represents an individual series. In each case the count for all ages has been normalised by the nulliparous count. In all cases we do not include any data for censored observations (see Section \ref{sec:dissection_censored}).}\label{fig:dissection_exampleData1}
\end{figure}

\begin{figure}[ht]
	\centerline{\includegraphics[width=1.3\textwidth]{./Figure_files/dissection_parity_data2.png}}
	\caption{\textbf{The normalised physiological age series for nine species in the database.} Each different coloured line represents an individual series. In each case the count for all ages has been normalised by the nulliparous count. In all cases we do not include any data for censored observations (see Section \ref{sec:dissection_censored}).}\label{fig:dissection_exampleData2}
\end{figure}

\begin{figure}[ht]
	\centerline{\includegraphics[width=1.3\textwidth]{./Figure_files/dissection_parity_data3.png}}
	\caption{\textbf{The normalised physiological age series for seven species in the database.} Each different coloured line represents an individual series. In each case the count for all ages has been normalised by the nulliparous count. In all cases we do not include any data for censored observations (see Section \ref{sec:dissection_censored}).}\label{fig:dissection_exampleData3}
\end{figure}

\subsection{Statistical analysis of dissection data}\label{sec:dissection_dissectionStats}
We suppose that the number of individuals recruited into the adult female population is constant over time, meaning that the age-structure of the population is stable. Whilst environmental heterogeneity will naturally lead to variation in adult recruitment over time, we hope that by aggregating data across studies undertaken over a range of dates we may reduce the impact of this effect on our results. Suppose that the probability an individual female mosquito survives until age $a$ is given by the survival function $S(a)$, then the number of individuals in the population surviving to this age is given by,
%
\begin{align}\label{eq:dissection_structured}
A(a) = A(0) S(a),
\end{align}
%
where $A(0)$ is the number of female mosquitoes recruited into the population per unit time. Consider one individual experiment where we randomly sample from a wild population that is structured as per eq. (\ref{eq:dissection_structured}). In this case the number of individuals sampled at each age is binomially-distributed,
%
\begin{align}
X(a) \sim \mathcal{B}(A(a),p(a)),
\end{align} 
%
where $p(a)$ is the probability of recapturing a single mosquito of age $a$. In what follows we assume that the probability of capturing a given mosquito is independent of their age, so that $p(a)=p=const$. Since in general $A(a)$ is large and $p(a)$ is small, we can approximate the above using a Poisson distribution,
%
\begin{align}
X(a) \sim \text{Poisson}(A(a)p).
\end{align}
%
However we believe that the assumption of \textit{independent} captures of individual mosquitoes, which underlies the binomial and Poisson models, is likely suspect for the same reasons as for the MRR analysis (see Section \ref{sec:mrr_statistical}). As before we choose to specify a negative binomial sampling distribution that allows for non-independent captures,
%
\begin{align}
X(a) \sim \text{NB}(A(a)p,\kappa),
\end{align} 
%
where we use the parameterisation such that the mean is given by $A(a)p$, and the over-dispersion parameter, $\kappa$, where as $\kappa\rightarrow \infty$ the above sampling distribution approaches a Poisson.

In the field, unfortunately, we do not know the number of mosquitoes recruited into the adult population, $A(a)$, nor the probability of capturing an individual at a given point in time, $p$. Instead we model their product $\Psi=A(0)p$ (the population of adult female mosquitoes of age zero that can be captured) probabilistically resulting in a model,
%
\begin{align}\label{eq:dissertation_negativeBinomial}
X(a) \sim \text{NB}(\Psi S(a),\kappa).
\end{align}
%
The resultant likelihood of a data series consisting of counts: $(y(a_1),y(a_2),...,y(a_R))$, is then calculated by assuming (conditional) independence of the observations,
%
\begin{equation}
\mathcal{L}(y(a_1),y(a_2),...,y(a_R)|S(.),\Psi,\kappa) = \prod\limits_{a=a_1}^{a_R} p(y(a)|S(a),\Psi,\kappa),
\end{equation}
%
where $p(y(a)|S(a),\Psi,\kappa)$ corresponds to the negative binomial probability mass function for a count of $y(a)$ mosquitoes aged $a$ as specified in eqn. (\ref{eq:dissertation_negativeBinomial}), and $R$ is the number of separate physiological age classes in which the count was non-zero.
%
Since we do not know $\Psi$ we must learn it from the data. One approach to estimate this parameter could be use the number of captures of nulliparous mosquitoes, $X(0)$ in place of $\Psi$. We prefer to allow for some uncertainty in this parameter and use the data to estimate its value. However, unfortunately, the negative binomial likelihood allows too much variation in this parameter (because the data are over-dispersed), and instead we specify a likelihood of the form,
%
\begin{align}
X(0) \sim \mathcal{N}(\Psi,\sqrt{\Psi}),
\end{align}
%
solely for the first data point $X(0)$. The above allows for some freedom in $\Psi$ whilst ensuring that the parameter's probability mass lies near enough to $X(0)$ to allow useful model estimates. $\Psi$ is set a uniform prior over the range $[0,\frac{3}{2}X(0)]$.

Since $S(a)$ is monotonically-decreasing we know that, if recruitment to the adult population is constant, then the numbers of nulliparous individuals should exceed the count in subsequent parity states. However in a number of data series there is a relative dearth of nulliparous mosquitoes versus uniparous individuals. This deficiency has been noted in a previously-published study where the authors hypothesise that it is due to the issue of sampling the nulliparous population, since they are more likely to rest outside, compared with parous individuals \citep{gillies1965study}. However there is evidence to suggest that, on average, the first gonotrophic cycle is longer than subsequent cycles (see Section \ref{sec:dissection_gonotrophicData}), meaning that a relative surplus of nulliparous mosquitoes may exist in captured samples \citep{clements1981analysis}.

In our analysis, we chose to remove the counts of nulliparous individuals from the series where the count was low relative to uniparous or higher parity individuals since this may indicate issues with sampling the nulliparous population. Specifically we stipulated that the nulliparous count should exceed 90\% of the count for the uniparous individuals. For those cases where this condition was not met we removed the nulliparous observation and analyse the series of counts for all subsequent pars (uniparous and subsequent parous states).

Here we estimate our model with one of six different survival functions (the same as for the MRR case; Table \ref{tab:mrr_survivalDescription}), each of which makes different assumptions regarding how the force of mortality is affected by mosquito age. To estimate lifespan at the species, genus and overall groupings we then use a hierarchical Bayesian model of the same mathematical form as for the analysis of MRR experiments (see Section \ref{sec:MRR_hierarchical}). However the priors for the analysis of dissection data were modified so that they represented a mean prior lifespan of three gonotrophic cycles, although allowed considerable variation in mean lifespan (Fig. \ref{fig:dissection_lifespanPriorsParity}). The priors used for the over-dispersion parameter ($\kappa$) for the negative binomial likelihood were the same as for the MRR analysis.

To determine whether mosquitoes experience age-dependent mortality we compared the predictive power of each of the models that incorporate a hazard function that depends on age with that from the exponential model. As for the MRR analysis we also used K-Fold cross-validation to perform this comparison, where the data are randomly partitioned into training and test sets (see Section \ref{sec:mrr_kFold}). The model is then fitted to each training set and used to predict the data in the independent test set. Since there are fewer series than for the MRR dataset we used two partitions for each species, where each partition was of roughly the same number of individual series.

\begin{figure}[ht]
	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/dissection_lifespanPriorsParity.pdf}}
	\caption{\textbf{The priors for mean lifespan used to analyse the data for each of the hierarchical models.} The distributional form of the priors is given in Table \ref{tab:dissection_priors}. In all cases the plots show data for 10,000 samples from the relevant prior distribution.}\label{fig:dissection_lifespanPriorsParity}
\end{figure}

\begin{table}
	\footnotesize
	\noindent\makebox[\textwidth]{%
		\begin{tabularx}{1.2\textwidth}{l|c|X}
			\textbf{Survival function} & \textbf{Physiological age series-level priors} & \textbf{Group-level priors}\\
			\midrule
			Exponential & $\lambda\sim \textnormal{log-normal}(\mu_\lambda,\sigma)$ &$\mu_\lambda \sim N(-1.2,1)\text{,}\; \sigma\sim \textnormal{log-normal}(-2,1)$\\
			\midrule
			Gompertz & $\alpha,\beta\sim$ log-normal$(\mu_{\alpha|\beta},0.2)$ &$\mu_\alpha\sim N(-1.5,1)\text{,}\; \mu_\beta\sim N(-2.5,0.5)$\\
			\midrule
			Weibull & $\alpha,(\beta-1)\sim$  log-normal$(\mu_{\alpha|\beta},0.2)$ &$\mu_\alpha\sim N(-2.5,1)\text{,}\; \mu_\beta\sim N(-4,0.5)$\\
			\midrule
			Gompertz-Makeham & $\alpha,\beta,c\sim$ log-normal$(\mu_{\alpha|\beta|c},0.2)$ &$\mu_\alpha\sim N(-1.4,1)\text{,}\; \mu_\beta\sim N(-3,0.4)\text{,}\; \mu_c\sim N(-4.5,0.5)$\\
			\midrule
			Logistic & $\alpha,\beta,s\sim$  log-normal$(\mu_{\alpha|\beta|s},0.2)$ &$\mu_\alpha\sim N(-1.4,1)\text{,}\; \mu_\beta\sim N(-2,1)\text{,}\; \mu_s\sim N(-3,1)$\\
			\midrule
			Logistic-Makeham & $\alpha,\beta,s,c\sim$ log-normal$(\mu_{\alpha|\beta|s|c},0.2)$ &$\mu_\alpha\sim N(-3,1)\text{,}\; \mu_\beta\sim N(-1.3,1)\text{,}\; \mu_s\sim N(-2,0.5)\text{,}\; \mu_c\sim N(-4,0.5)$\\
			\bottomrule
	\end{tabularx}}\caption{\textbf{The priors used on parameters of each different survival model for the dissection data analysis.} For the exponential model the `group-level' priors were the same for the genus and `overall' models that were also estimated. The exponential model was the only model that was simple enough to allow the scale parameter of the log-normal ($\sigma$) to be estimated by the data. The notation $\alpha,\beta\sim \text{log-normal}(\mu_{\alpha|\beta},0.2)$ means that $\alpha$ and $\beta$ were assigned independent log-normal priors with location parameters $\mu_\alpha$ and $\mu_\beta$ respectively, and a scale parameter of 0.2 in both cases.}\label{tab:dissection_priors}
\end{table}

The estimates of mean mosquito lifespan that we present here (Fig. \ref{fig:dissection_lifetimes_exponential}) resulted from the use of the exponential survival model (i.e. no age dependence). This choice was made because we found limited evidence in support of age-dependent mortality (see `Results'). However, since the priors for each of the survival functions were specified to allow for a wide variety of lifespans (Fig. \ref{fig:dissection_lifespanPriorsParity}) the specific choice of survival function made little difference to resultant estimates (data not shown).

To demonstrate that the results we obtain are not sensitive to the particular form of hierarchical prior structure chosen we also provide estimates of the lifespan for each series analysed separately (i.e. non-hierarchically). To do so we assume priors on the rate parameter of the exponential distribution that provides support over a wide range of possible lifespans (Fig. \ref{fig:dissection_lifespanPrior_individual}A). We also specified a prior on the over-dispersion parameter $\kappa$ that was comparable to the hierarchical case (Fig. \ref{fig:dissection_lifespanPrior_individual}B).

\begin{figure}[ht]
	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/dissection_lifespanPrior_individual.pdf}}
	\caption{\textbf{The priors for A. mean lifespan and B. the over-dispersion parameter used to analyse the individual dissection series.} The prior on the rate parameter of the exponential distribution was $\lambda\sim \text{log-normal}(-1.8,1)$. The prior on the dispersion parameter was $\kappa\sim \text{log-normal}(1,1)$.}\label{fig:dissection_lifespanPrior_individual}
\end{figure}

\subsubsection{Analysis of censored data series}\label{sec:dissection_censored}
In some of the published papers the data are censored above a threshold age, $a_T$. In these cases we only know the number, $y(a_T)$, of individuals who were captured and dissected with an estimated physiological age that is greater than or equal to the threshold. Since we do not know the estimated physiological age of individual specimens $i$, we represent this as a parameter, $\aleph_i$, in our statistical model. This means that the joint distribution is a function of $y(a_T)$ different $\aleph_i$ parameters. Since these parameters are not directly of interest, and our chosen MCMC engine, Stan \citep{stan-software:2014}, does not directly allow discrete parameters in models, we marginalise these out of the joint distribution. To do this exactly would require an infinite sum over all the possible ages for all $y(a_T)$ mosquitoes that have been caught whose age exceeds the threshold. Rather than carry out this intractable number of summations we instead make the approximation that all mosquitoes in this group are of the same age $\aleph_i = \aleph,\; \forall i \in (1,...,y(a_T))$. We believe this assumption is justifiable, particularly since the numbers of mosquitoes in each subsequent age category is a strongly-decreasing function of age. This means that whilst we do not know with certainty individuals' ages, it is likely that most will be of the threshold age $a_T$.

This approximation means that to marginalise the parameter $\aleph$ out of the joint distribution, we are only required to do a single summation. Specifically if we have $y(a_T)$ captured individuals of age equal to, or exceeding, some threshold age, $a_T$, the probability of these observations is given by,
%
\begin{align}
q(y(a_T)|S(.),\Psi,\kappa) &= \sum_{\aleph=a_T}^{\infty} p(y(a_T),\aleph|S(\aleph),\Psi,\kappa)\\
&= \sum_{\aleph=a_T}^{\infty} p(y(a_T)|S(\aleph),\Psi,\kappa) \times p(\aleph)
\end{align}
%
where $q(y(a_T)|S(\aleph),\Psi,\kappa)$ is the probability of observing $y(a_T)$ counts for mosquitoes of an age $\aleph$. In practice, since we do not believe mosquitoes live for longer than 20 gonotrophic cycles (the maximum observed in the data was 13), we cut-off the summation at this point, and assume that the discrete prior probability distribution $p(\aleph)$ is uniform over this range. The overall likelihood for the cases where the series are censored therefore has the form,
%
\begin{equation}
\mathcal{L}(y(a_1),y(a_2),...,y(a_T-1),y(a_T)|S(.),\Psi,\kappa) = \left(\prod\limits_{a=a_1}^{a_T-1} p(y(a)|S(a),\Psi,\kappa)\right) q(y(a_T)|S(\aleph),\Psi,\kappa).
\end{equation}
%
In practice, the number of mosquitoes captured in those series that are censored represents a small percentage of total captures (the median is approximately 2\%), and so the effect of the $q(.)$ term above is likely minimal on resultant inferences.

\subsection{Model estimation by MCMC}\label{sec:dissection_MCMC}
As for the MRR analysis, we used \textit{Stan} software \citep{carpenter2016stan} to sample from the posterior distributions of model parameters.

To judge convergence of the sampling algorithm to the posterior density, we calculated $\hat{R}$ across all Markov chains -- a ratio that compares the between-chain variation to that within each chain that is commonly used to measure convergence in MCMC \citep{gelman1992inference}. For each model, we ran 16 independent Markov chains with 200 iterations per chain, discarding the first half of these iterations as warm-up \citep{gelman2014bayesian}. After running the algorithm for the given number of observations, $\hat{R}<1.1$ for all model parameters. We also ensured that across each MCMC run the number of divergent iterations (that can bias the MCMC away from the true posterior density) was minimal. In the majority of cases, the number of divergent iterations was far fewer than 1\% of the total number of samples.

The model used to estimate the lifespan in terms of gonotrophic cycles is provided below.

\begin{minted}{stan}
data{
int N; // number of obs
int K; // number of time series
int Y[N]; // number captured
int S[K]; // length of each series
int threshold[K]; // age of censoring; if -1 then no censoring
int Pos[K]; // position of start of each series in Y
int species[K]; // index variable of species
int nSpecies; // number of species
}

parameters{
real<lower=0> PopSize[K];
real<lower=0> alpha[K];
real alpha_mean[nSpecies];
real<lower=0> alpha_sigma[nSpecies];
real<lower=0> kappa[K];
real<lower=0> p[nSpecies];
}

model{
// Likelihood
for(i in 1:K){
  for(t in 2:S[i]){
    // uncensored
    if(threshold[i] < 0)
      Y[Pos[i]+t-1] ~ neg_binomial_2(PopSize[i] *
			exp(-alpha[i] * (t-1)), kappa[i]);
    // censoring
    else{
      if((t-1) < threshold[i])
        Y[Pos[i]+t-1] ~ neg_binomial_2(PopSize[i] *
			exp(-alpha[i] * (t-1)), kappa[i]);
      else{
        real lLogProb[20];
        for(j in 1:20){
          lLogProb[j] = neg_binomial_2_lpmf(Y[Pos[i]+t-1] |
			PopSize[i] * exp(-alpha[i] * (t-1)),
			kappa[i]);
        }
        target += log_sum_exp(lLogProb);
      }
    }
  }  
}

// Priors
for(i in 1:K){
  int aSpecies;
  aSpecies = species[i];
  Y[Pos[i]] ~ normal(PopSize[i], sqrt(PopSize[i]));
  alpha[i] ~ lognormal(alpha_mean[aSpecies],
  			alpha_sigma[aSpecies]);
  kappa[i] ~ exponential(p[aSpecies]);
}

alpha_mean ~ normal(-1.2, 1);
alpha_sigma ~ lognormal(-2, 1);
p ~ lognormal(1, 1);
}

generated quantities{
real alpha_average[nSpecies];
for(i in 1:nSpecies)
  alpha_average[i] = lognormal_rng(alpha_mean[i],alpha_sigma[i]);
}
\end{minted}

\subsection{Model checking}
As for the MRR experiments, we simulated from the posterior predictive distribution for each series of dissection data. In the majority of the cases, the capture data lay within the 95\% posterior predictive intervals, indicating that our model was a good fit to the data (see Figures \ref{fig:dissection_ppc_all1}, \ref{fig:dissection_ppc_all2} \&  \ref{fig:dissection_ppc_all3} and the attached file, \verb|dissection_ppcs_all.pdf| for the full graphs).

\newgeometry{scale=0.8}
\thispagestyle{empty}
{%
	\centering
	\includegraphics[page=1,scale=.7]{./Figure_files/dissection_ppcs_all}
	\captionof{figure}{\textbf{The numbers of mosquitoes recaptured (black lines) versus the 95\% central posterior interval of the posterior predictive distribution (blue shading) for a selection of the dissection series.} The titles correspond to the unique combination of "identifier - physiological-experiment-id" in the database. For the rest of the posterior predictive checks for the dissection data models, see the file referenced in the text.}
	\label{fig:dissection_ppc_all1}
	\par
}
\restoregeometry

\newgeometry{scale=0.8}
\thispagestyle{empty}
{%
	\centering
	\includegraphics[page=2,scale=.7]{./Figure_files/dissection_ppcs_all}
	\captionof{figure}{\textbf{The numbers of mosquitoes recaptured (black lines) versus the 95\% central posterior interval of the posterior predictive distribution (blue shading) for a selection of the dissection series.} The titles correspond to the unique combination of "identifier - physiological-experiment-id" in the database. For the rest of the posterior predictive checks for the dissection data models, see the file referenced in the text.}
	\label{fig:dissection_ppc_all2}
	\par
}
\restoregeometry

\newgeometry{scale=0.8}
\thispagestyle{empty}
{%
	\centering
	\includegraphics[page=3,scale=.7]{./Figure_files/dissection_ppcs_all}
	\captionof{figure}{\textbf{The numbers of mosquitoes recaptured (black lines) versus the 95\% central posterior interval of the posterior predictive distribution (blue shading) for a selection of the dissection series.} The titles correspond to the unique combination of "identifier - physiological-experiment-id" in the database. For the rest of the posterior predictive checks for the dissection data models, see the file referenced in the text.}
	\label{fig:dissection_ppc_all3}
	\par
}
\restoregeometry

\subsection{Data collection of gonotrophic cycle duration}\label{sec:dissection_gonotrophicData}
To convert the estimates of lifespan in physiological age into chronological age we require estimates of the duration of the gonotrophic cycle. To determine this characteristic we conducted a meta-analysis of previously-published studies that estimate the duration of the gonotrophic cycle. A search of the literature using Google Scholar (\url{scholar.google.co.uk}) was performed using the search term: `gonotrophic cycle duration'. The list of articles was then supplemented with a list of references discussed by \cite{silver2007mosquito}. Based on the abstracts of the resultant list of published studies we then decided whether to search each article for estimates of the duration of the gonotrophic cycle. Overall 79 separate estimates of this parameter were found across 42 published articles (see Section \ref{sec:appendix_gonotrophicDurationStudyList} for a list of the studies included in our dataset). Along with information about the estimates we also recorded study and series meta-data, including the location of the study, the method used for estimation, the species and genus, as well as the temperatures and/or seasons in which the experiments were carried out. The methods used to estimate gonotrophic cycle duration in the literature largely fell into two distinct categories: those based on MRR studies ($n=29$); and those based on observations of mosquitoes in a laboratory setting ($n=42$). In the MRR studies estimates are made of the duration of gonotrophic cycles by dissecting recaptured mosquitoes at each time point using the method of \cite{polovodova1949determination} to count ovariolar dilations. For the laboratory studies the duration of gonotrophic cycles is determined more directly by observing the time taken for mosquitoes to mate then blood-feed and finally oviposit.

Along with point estimates of the parameter we also collected information about the uncertainty in the estimates (if available). In many articles the duration of the gonotrophic cycle was estimated separately for the first versus subsequent cycles, and these estimates were recorded separately. Raw estimates of gonotrophic cycle duration were obtained for species across four different genera (\textit{Aedes}, \textit{Anopheles}, \textit{Culex} and \textit{Masonia}), although there was a bias towards \textit{Anopheles} with $n=47$ estimates out of the total of 79 (Fig. \ref{S-fig:dissection_gonotrophicCycleRaw}).

Since we collected data on the method used to estimate the duration of the gonotrophic cycle we show the raw estimates for the two most common approaches: MRR and laboratory (Fig. \ref{S-fig:dissection_gonotrophicCycleRaw_MRRVsLab}). From these results it is evident that the estimates from the laboratory studies are, on average, higher than those using the MRR method.

\subsection{Statistical analysis of gonotrophic cycle data}\label{sec:dissection_gonotrophicMethod}
As discussed in Section \ref{sec:dissection_gonotrophicData}, there was considerable study-level heterogeneity in the information provided for the estimates of the gonotrophic cycle duration. A number of studies ($n=24$) provided no estimates of uncertainty in gonotrophic cycle duration, whereas the rest gave some indication of confidence or alternatively a range of possible estimates. However the types of uncertainty intervals that were specified varied considerably from study to study, from the vague (but common)  `4-6 days' to the more helpful `5$\pm 1$ day (95\% confidence interval)'.

The heterogeneous nature of the gonotrophic cycle estimates requires a method that explicitly accounts for this characteristic of the data. We decided to model the estimates as representing quantiles from an underlying normal distribution that represents uncertainty over possible durations of the gonotrophic cycle. In those circumstances where lower, central and upper bounds were given explicitly the data was fairly symmetric and so we believe assuming the observations come from an unskewed normal distribution is reasonable. Explicitly we treated each type of uncertainty interval as follows,

\begin{itemize}
	\item Simple range, for example, `4-6 days': treat the lower and upper bounds as the 2.5\% and 97.5\% quantiles of a normal distribution.
	\item Confidence intervals, `5$\pm 1$ day (95\% confidence interval)': treat the lower and upper bounds as the relevant quantiles of a normal distribution.
	\item Point estimates, for example `5 days': treat this as the median of a normal distribution.
\end{itemize}

The benefit of this approach is that we can convert the quantiles from our normal distribution parameterised by a mean ($\mu$) and standard deviation ($\sigma$) to equivalent quantiles from a standard normal,
%
\begin{align}
Z_{50} = \frac{G_{50} - \mu}{\sigma},
\end{align}
%
where $Z_{50}$ and $G_{50}$ indicate 50\% quantiles from a standard normal distribution and a normal distribution respectively. By manipulating the above expression we obtain the following,
%
\begin{align}
G_{50} = \sigma Z_{50} + \mu,
\end{align}
%
which forms a straight line in $(Z,G)$ space with slope $\sigma$ and y-intercept $\mu$. Therefore by plotting the raw observations of $(Z,G)$ quantiles then estimating a linear regression line we can characterise the underlying $\mathcal{N}(\mu,\sigma)$ distribution from which we assume they are drawn.

Due to the relative unavailability of estimates for gonotrophic cycle duration across the different genera we generated a single distribution to represent the uncertainty in this parameter by pooling all the data. This was justified because there was not found to be significant variation in the 1st gonotrophic cycle duration (ANOVA: $F_{3,75}=2.6,p>0.05$).

Using the collected data, we estimated that uncertainty in the first gonotrophic cycle duration is estimated as $\mathcal{N}(4.26, 0.38)$ and, for subsequent cycles, $\mathcal{N}(3.85, 0.35)$.

\subsection{Conversion of lifespan from physiological to calendar age}\label{sec:dissection_conversion}
The estimates of lifespan produced from analysing the dissection data are in terms of physiological age (the number of gonotrophic cycles undertaken). To allow comparison with the estimates from the MRR studies (Fig. \ref{fig:mrr_lifespans}) as well as produce more useful estimates to inform disease transmission dynamical models we convert these estimates to calendar days. To do this we use the estimated parameters of the normal densities that we have assumed represent uncertainty in gonotrophic cycle duration (see Section \ref{sec:dissection_gonotrophicMethod}) and use them to convert our posterior samples of mean physiological lifespan $(L^p_1,...,L^p_S)$ to lifespan in calendar ages $(L^c_1,...,L^c_S)$. To do this we iterate the following for all $i\in (1,...,S)$ posterior samples,

\begin{enumerate}
	\item Sample $G_{1i} \sim \mathcal{N}(\mu_{1i},\sigma_{1i})$, to obtain a duration for the 1st gonotrophic cycle.
	\item Sample $G_{2i} \sim \mathcal{N}(\mu_{2i},\sigma_{2i})$, to obtain a duration for subsequent gonotrophic cycles.
	\item If $L^p_i > 1$, the mean lifespan is longer than one gonotrophic cycle:
	\subitem then $L^c_i = G_{1i} + G_{2i} \times (L^p_i - 1)$.
	\item Else:
	\subitem then $L^c_i = G_{1i}\times L^p_i$.
\end{enumerate}

In the using the above methodology to convert from physiological to calendar age we implicitly assume that for each gonotrophic cycle after the first that the cycles are of the same length (we choose one $G_{2i}$ per sample). However since we allow a different $G_{2i}$ for each sample we nonetheless believe that the above approach allows for sufficient uncertainty in the estimates of subsequent gonotrophic cycle durations. If instead we believed that significant variation in gonotrophic cycle occurred within a particular mosquito's life (as well as between mosquitoes) then we could draw a new value of $G_{2i}$ for each subsequent cycle. This produces results with a little more uncertainty than previously (data not shown).

\section{EIPs of vector species of malaria, dengue fever, chikungunya and Zika}
The following anopheline species were identified as vector species of malaria,

\begin{itemize}
	\item \textbf{Africa}: \textit{A. gambiae s.l.}, \textit{A. funestus} and \textit{A. melas} \citep{sinka2012global}, and \textit{A. rivulorum} \citep{wilkes1996anopheles}.
	\item \textbf{Americas}: \textit{A. albimanus}, \textit{A. albitarsis}, \textit{A. darlingi}, \textit{A. pseudopunctipennis} and \textit{A. quadrimaculatus} \citep{sinka2012global}; \textit{A. bellator}  \citep{forattini1999role,lorenz2012morphometrical}; \textit{A. cruzii} \citep{lorenz2012morphometrical}; and \textit{A. vestitipennis} \citep{sinka2010dominant}.
	\item \textbf{Europe and the Middle-East}: \textit{A. sergentii} \citep{sinka2012global}; and \textit{A. maculipennis} \citep{hackett1935varieties}.
	\item \textbf{Asia}: \textit{A. farauti s.l.}, \textit{A. koliensis}, \textit{A. lesteri}, \textit{A. maculatus}, \textit{A. minimus}, \textit{A. punctulatus}, \textit{A. stephensi} and \textit{A. subpictus s.l.} \citep{sinka2012global}; and \textit{A. culicifacies} \citep{green1980chromosomal}.
\end{itemize}

For dengue fever, chikungunya and Zika, the main vector species are \textit{Ae. aegypti} and \textit{Ae. albopictus} \citep{kraemer2015global,grard2014zika,benelli2016declining}.

\section{Studies included in the MRR meta-analysis}\label{sec:appendix_mrrStudyList}
The following is the subset of studies from the original \cite{guerra2014global} database which were used in the MRR meta-analysis: \cite{marini2010study,baber2010population,lacroix2009dispersal,maciel2008calculating,midega2007estimating,maciel2007daily,elizondo2006gonotrophic,ba2005aspects,fabian2005mark,la2004anopheles,watson2000aedes,harrington2001analysis,tsuda2001movement,muir1998aedes,toure1998mark,quinones1997anopheles,costantini1996density,trpis1995estimates,jensen1994comparison,fernandez1994gonotrophic,jaal1992mark,rodriguez1992gonotrophic,chiang1991capture,jensen1991assessment,eldridge1990daily,macdonald1968mark,pumpuni1989population,charlwood1987mark,charlwood1989capture,birley1989effect,arredondo1998gonotrophic,hii1990estimation,renshaw1994host,milby1989estimation,maciel2007body,loong1990survival,nelson1980dispersal,maciel2006movement,mcdonald1977population,curtis1980preliminary,rawlings1982dispersal,conway1974population,reisen1979anopheles,nelson1978estimates,rawlings1981influence,sempala1981ecology,takagi1995movement,buei1980field,eyles1943experiment,ordonez2001use,pant1973field,charlwood1988evidence,reisen1982anopheles,nayar1980quantitative,carnevale1979etude,eyles1946long,reisen1984impact,charlwood1986capture,trpis1975demonstration,lutwama1994mark,wada1969dispersal,takken1998dispersal,abdel1966study,valerio2012dispersal,zetek1915behavior,takagi1995movement,yasuno1975migration,eyles1943measurement,germain1974evaluation}.

\section{Studies included in the dissection study meta-analysis}\label{sec:appendix_dissectionStudyList}
The following is a list of studies included in the dissection study meta-analysis: \cite{catangui1971studies,charlwood1979studies,chang1991comparative,charlwood2000dry,edalat2015vectorial,de2011survivorship,lines1991human,magesa1991trial,lines1991monitoring,forattini1996studies,samarawickrema1967study,samarawickrema1968biting,de2007parity,charlwood1985assessing,chandra1996age,hoc1995ovariole,vythilingam1997abundance,russell1986culex,chadee2010oviposition,ebsary1977biting,charlwood1980observations,shriram2011population,mahmood1981duration,samarawickrema1987seasonal,uttah2013physiological,ramaiah1992non,buei1982age,chadee1995seasonal,chandra2008age,foll1966conditions,kanda1975epidemiological,mala2014gonotrophic,shalaby1962studies,world1960preliminary,reisen1980anopheles,detinova1962age,smith1994age,jayanetti1987study,ch2013comparative,pant1962distribution,ghosh2010seasonal,mahanta1999temporal,schlein2012diurnal,penilla2002pteridine,schlein2015decrease,snow1978host,detinova1968age,nathan1981bancroftian,charlwood1985blood,schlein2008approach,beier2012attractive,muller2005plant,surendran2006anopheles,muller2010successful,mendis1998cordon,wilkes1996anopheles,chatterjee2000role,muller2010successful,tuchinda1969diurnal,chan1971life,aigbodion2011pelagia,qualls2015indoor,reisen1981anopheles,gillies1972range,gillies1965study,kay1979age,hitchcock1968age}.

\section{Studies included in the gonotrophic cycle duration meta-analysis}\label{sec:appendix_gonotrophicDurationStudyList}
The following is a list of studies included in the gonotrophic study meta-analysis: \cite{gillies1965study,sheppard1969dynamics,de1967time,pant1973field,germain1974evaluation,lowe1973reproductive,fernandez1994gonotrophic,buei1980field,rawlings1982tests,mori1977gonotrophic,sempala1981ecology,reisen1983population,birley1989effect,suzuki1978preliminary,charlwood1987mark,charlwood1986capture,charlwood1979studies,charlwood1995rise,chang1991comparative,edalat2015vectorial,de2011survivorship,samarawickrema1968biting,samarawickrema1967study,charlwood1985assessing,chandra1996age,russell1986culex,chadee2010oviposition,mahmood1981duration,ahumada2004modeling,kenawy1991development,rajagopalan1980population,chandra2008age,scholl1979aedes,mala2014gonotrophic,afrane2005effects,gillies1953duration,quinones1997anopheles,rua2005laboratory,delatte2009influence,arredondo1998gonotrophic,wong2014sampling,ijumba2002malaria}.

 \bibliographystyle{authordate1}
 \bibliography{Malaria}
\newpage

\end{document}

### 2018-11-28 MS.tex
Added author summary as per PLOS guidelines

@@
 \section{Author summary}
+Mosquitoes transmit some of the most important diseases afflicting humans, with malaria alone killing between 0.4-1.2 million people annually, chiefly children in low-income countries.
+The transmission strength of these diseases depends critically on the duration of mosquito lifespans and some of the most successful disease control interventions, including insecticide-treated bednets, explicitly target reductions in mosquito longevity.
+In this study, we conduct meta-analyses of two important classes of field experiments which estimate wild mosquito lifespan: mark-release-recapture studies, where mosquitoes are marked with a dye then released with the number of marked mosquitoes caught monitored over time; and experiments involving dissection of wild-caught females, whose reproductive anatomy is used as a biological clock to determine physiological age.
+In both analyses, we estimate that most mosquito species live less than 10 days, on average, which suggests that relatively few mosquitoes live sufficiently long to transmit disease.
+We find evidence of variation in mosquito mortality across species, with the estimates of lifespan obtained from each method largely corresponding for the few species with data from both experiments.
+Finally, by fitting a range of survival models to the data, we conclude that, for most species, mosquitoes do not experience strong age-related increases in mortality.
 

### 2018-11-28 MS.tex
Changed methods as per discussion with Ace. Only remaining thing is to deal with major vector species repetitions.

@@
 
+MRR estimates the length of time a mosquito remains alive and is still
+in the area available for recapture.
+In dissections of females, the majority of ovarioles have fewer dilations than the number of gonotrophic cycles an individual has experienced, also meaning that estimates derived from these data likely understand true physiological age \citep{hugo2008evaluations}.
+It is unclear which of these methods leads to lower estimates but in both cases we term our estimates lower bounds on lifespan, which we shall refer to as LBL.
+
 \subsection{Lifespan estimates from
@@
 
-MRR estimates the length of time a mosquito remains alive and is still
-in the area available for recapture. It thus provides a lower bound to mean
-lifespan which we shall refer to as LBL. In 187 of the 230 MRR time
+In 187 of the 230 MRR time
 series the estimated LBL was less than 10 days (Fig. \ref{fig:mrr_lifespan_individualEstimates}). The smallest
@@
 
-The estimated mean LBL for female mosquitoes not pre-fed with blood or
-sugar before release for \emph{Culex, Anopheles} and \emph{Aedes} were
+Since female mosquitoes are most epidemiologically relevant, we start by discussing their estimated lifespan.
+Also, since the majority of mosquitoes were unfed with blood or sugar prior to release, unless otherwise stated, our estimates represent quantities for unfed mosquitoes.
+The estimated mean LBL for female mosquitoes for \emph{Culex, Anopheles} and \emph{Aedes} were
 2.9, 6.8 and 8.1 days respectively with an overall estimate of 6.0 days
@@
 
+Using the data collected from a literature search, we estimated that the first gonotrophic cycle duration had a mean of 4.3 days (std.
+error: 0.4 days) and, for subsequent cycles, the mean was 3.9 days (std.
+error: 0.4 days; see SOM).
 To compare the two methods, we converted numbers of gonotrophic cycles
 (physiological age) into lifespan (chronological age) as described in
-the SOM using estimates of gonotrophic cycle duration (Fig.
-\ref{S-fig:dissection_gonotrophicCycleRaw}).
+the SOM using these estimates of gonotrophic cycle duration.
 Table \ref{S-tab:dissection_estimated_lifespans_chron} provides posterior summaries of chronological for the species and genera in the dissection dataset (see also Fig.
@@
 gonotrophic cycles in mosquito samples collected over a specific period
-of time. Overall, we found 568 time series in 72 published articles.
-Because seldom were sufficient ancillary data available to analyse
-trends over time, we aggregated the data from each time series. We
-further omitted time series with fewer than 100 mosquitoes and for
+of time.
+Overall, we found 568 physiological age cross-sections at recorded distinct times in 72 published articles.
+Our statistical approach relies on steady recruitment to the adult mosquito population.
+To guard against the effect of fluctuating population sizes on our analysis, we aggregated the data at a given location across cross-sections taken at different times.
+We further omitted time series with fewer than 100 mosquitoes and for
 species with only one data set leaving 131 studies of mosquitoes in the
 genera \emph{Anopheles, Aedes}, \emph{Culex} and \emph{Mansonia}.
+
+The data which we use provides measures of the age distribution of mosquitoes within each investigated population.
+By assuming that the population sizes were approximately fixed throughout the period of investigation, this allows us to estimate mean lifespan using a statistical model of mortality incorporating the probability of mosquito capture.
+We modelled the number of mosquitoes found by dissection to be of age
+\emph{a} using the negative binomial distribution with mean
+$\Psi S(a)$ and shape parameter $\kappa$, where
+$\Psi$ is the product of the recruitment rate of adult mosquitoes,
+which we assume is constant over time, and the probability of being
+captured for dissection, and \emph{S}(\emph{a}) is the probability of
+surviving until age \emph{a}. We used the number of females that have
+yet to lay eggs (nulliparous) to estimate the recruitment rate as
+described further in the SOM. Initial examination revealed that in some
+data sets the number of nulliparous females was anomalously low,
+something that has been noticed before \citep{gillies1965study}. As some
+studies have suggested that the first gonotrophic cycle tends to be
+longer than the subsequent ones, this is probably due to differences in
+capture probability. In data sets where the fraction of nulliparous
+females was less than 90\% the uniparous (completed on gonotrophic
+cycle) we excluded the nulliparous observation. Data was analysed using a Bayesian framework as with the MRR data with
+minor differences in the specification of the priors (see SOM).
 
@@
 
-We modelled the number of mosquitoes found by dissection to be of age
-\emph{a} using the negative binomial distribution with mean
-$\Psi S(a)$ and shape parameter $\kappa$, where
-$\Psi$ is the product of the recruitment rate of adult mosquitoes,
-which we assume is constant over time, and the probability of being
-captured for dissection, and \emph{S}(\emph{a}) is the probability of
-surviving until age \emph{a}. We used the number of females that have
-yet to lay eggs (nulliparous) to estimate the recruitment rate as
-described further in the SOM. Initial examination revealed that in some
-data sets the number of nulliparous females was anomalously low,
-something that has been noticed before \citep{gillies1965study}. As some
-studies have suggested that the first gonotrophic cycle tends to be
-longer than the subsequent ones, this is probably due to differences in
-capture probability. In data sets where the fraction of nulliparous
-females was less than 90\% the uniparous (completed on gonotrophic
-cycle) we excluded the nulliparous observation.
-
-Data was analysed using a Bayesian framework as with the MRR data with
-minor differences in the specification of the priors (see SOM). Some
-published studies do not distinguish the number of gonotrophic studies
-beyond a threshold (the more ovariole dilations there are the harder it
-is to count them) which is akin to censoring the data. Because this
-censoring involves only a relatively small number of mosquitoes in each
-time series (median = 2\%), and because of the technical difficulties of
-allowing for censoring in our Bayesian estimation procedures, we assume
-the females died at the threshold age.
-
 \section{Acknowledgements}\label{acknowledgements}

### 2018-11-29 MS.tex
Changed order of dissection section

@@
 counted and from this the mean number of cycles before death was
-estimated. Overall, the mean number of cycles completed in a lifetime
-was 1.3 (posterior mean; Fig. \ref{fig:dissection_lifetimes_exponential}; Table \ref{S-tab:dissection_estimated_lifespans}) and across the 131 studies, 95\% of
-the individual time series estimates were less than 3 (Fig. \ref{S-fig:dissection_individualEstimates_allSpecies}). The
+estimated.
+Across the 131 studies, 95\% of the individual time series estimates were less than 3 gonotrophic cycles (Fig.
+\ref{S-fig:dissection_individualEstimates_allSpecies}) and, overall, the mean number of cycles completed in a lifetime
+was 1.3 (posterior mean; Fig. \ref{fig:dissection_lifetimes_exponential}; Table \ref{S-tab:dissection_estimated_lifespans}). The
 estimated greatest number of cycles was for \emph{Anopheles sergentii}

### 2018-11-29 MS.tex
Changes to remove major vector references

@@
 where constant mortality models performed at least as well. The species
-where age-dependent mortality best fit the data included the major
+where age-dependent mortality best fit the data included the
 vector of dengue fever, Zika and chikungunya, \emph{Ae. Aegypti}. These
@@
 drought-adapted and long-lived \emph{A sergentii.} The proportions
-surviving long enough to become infectious for the two major African
-malaria vectors were: for \emph{A. gambiae s.l.}: 10\% (MRR) and 27\%
+surviving long enough to become infectious for \emph{A. gambiae s.l.}: 10\% (MRR) and 27\%
 (dissection); and for \emph{A. funestus s.l.}: 9\% (MRR). Using the
@@
 estimate; 25\%-50\% CI: 7.6-14.0 days) from the analysis of dissection
-studies. This species is a major vector of malaria in the Sahara \citep{sinka2010dominant}, where to act as a disease vector it must persevere
+studies. This species is a vector of malaria in the Sahara \citep{sinka2010dominant}, where to act as a disease vector it must persevere
 through these hard conditions. It is reasonable to hypothesise that this
@@
 potential for blood-feeding and oviposition is greater. The species with
-the greatest discrepancy in the estimates was the major African malaria
-vector \emph{A. gambiae s.l.}, where we estimated LBLs of 4.5 days
+the greatest discrepancy in the estimates was \emph{A. gambiae s.l.}, where we estimated LBLs of 4.5 days
 (mean estimate; 25\%-75\% CI: 3.8-5.1 days for unfed female) from the

### 2018-12-05 supplementary_materials.tex
Changed comparison figures to account for the two extra species that overlap
Made changes to the results MS as per Ace's comments

@@
 		\midrule
-		\textit{Ae. aegypti} & 1     & 55    & 2.2   & 0.14 \\
-		\textit{A. darlingi} & 1     & 6     & 114.8   & 0.00 \\
-		\textit{An. farauti s.l.} & 1     & 14    & 0.9  & 0.35 \\
-		\textit{A. gambiae s.l.} & 1     & 37    & 3.9   & 0.06 \\
-		\textit{Cx. pipiens} & 1     & 13    & 0.6   & 0.47 \\
-		\textit{A. quadrimaculatus} & 1     & 8     & 2.4     & 0.16 \\
-		\textit{Cx. quinquefasciatus} & 1     & 13    & 2.2    & 0.17 \\
-		\textit{A. sergentii} & 1     & 8     & 0.4     & 0.55 \\
-		\textit{Cx. tritaeniorhynchus} & 1     & 5     & 3.5   & 0.12 \\
-		\textit{Ae. vexans} & 1     & 2     & 1.2   & 0.39 \\
+		\textit{Ae. aegypti} & 1     & 55    & 1.9   & 0.17 \\
+		\textit{A. culicifacies s.l.} & 1     & 19    & 2.2   & 0.15 \\
+		\textit{A. darlingi} & 1     & 6     & 80.7   & 0.00 \\
+		\textit{An. farauti s.l.} & 1     & 14    & 2.1  & 0.17 \\
+		\textit{A. gambiae s.l.} & 1     & 37    & 4.3   & 0.05 \\
+		\textit{Cx. pipiens} & 1     & 13    & 0.4   & 0.56 \\
+		\textit{A. quadrimaculatus} & 1     & 8     & 2.2     & 0.18 \\
+		\textit{Cx. quinquefasciatus} & 1     & 13    & 2.2    & 0.16 \\
+		\textit{A. sergentii} & 1     & 8     & 1.0    & 0.36 \\
+		\textit{A. subpictus} & 1     & 2    & 11.4   & 0.08 \\
+		\textit{Cx. tritaeniorhynchus} & 1     & 5     & 3.6   & 0.12 \\
+		\textit{Ae. vexans} & 1     & 2     & 1.1   & 0.41 \\
 	\end{tabular}%
@@
 	\label{fig:mrr_mcPowerAnalysis_senescence}
+\end{figure}
+
+\begin{figure}[ht]
+	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/elpd_comparison.pdf}}
+	\caption{\textbf{Comparing the evidence for age-dependent mortality by fitting the Gompertz model to data from the MRR versus dissection studies for the 12 species with data from both.} The points indicate the estimated predictive accuracy compared with the exponential model (see SOM Methods) and the lower and upper whiskers represent 2.5\% and 97.5\% quantiles on this quantity.}\label{fig:elpd_comparison}
 \end{figure}

### 2019-05-24 writeUp/supplementary_materials_methods.tex
Rejigged layout of repo and added to-do

--- 2018-11-28 supplementary_materials_methods.tex
@@
 \subsubsection{Analysis of censored data series}\label{sec:dissection_censored}
-In some of the published papers the data are censored above a threshold age, $a_T$.
+Some published studies do not distinguish the number of gonotrophic studies
+beyond a threshold, $a_T$ (the more ovariole dilations there are the harder it
+is to count them) which is akin to censoring the data. This
+censoring involves only a relatively small number of mosquitoes in each
+time series (median = 2\%) and so we are confident that these few cases do not substantially affect our inferences.
+We, however, now describe the statistical model used to describe these instances.
 In these cases we only know the number, $y(a_T)$, of individuals who were captured and dissected with an estimated physiological age that is greater than or equal to the threshold.
@@
 
-Using the collected data, we estimated that uncertainty in the first gonotrophic cycle duration is estimated as $\mathcal{N}(4.26, 0.38)$ and, for subsequent cycles, $\mathcal{N}(3.85, 0.35)$.
+Using the collected data, we estimated that the first gonotrophic cycle duration was $\mathcal{N}(4.26, 0.38)$ and, for subsequent cycles, $\mathcal{N}(3.85, 0.35)$.
 

### 2019-06-10 writeUp/supplementary_materials_methods.tex
Trained neural network for longer
Started to write up methods

@@
 
+\section{Detinova parity analysis}
+\cite{massey2016global} provide a global database for bionomic quantities for the dominant vector species of malaria. In this database, there are
+
+\cite{detinova1962age} provided an alternative approach to estimate the lifespan of individual specimens, based on sampling the parous rate in a population (the proportion of adult females that have laid eggs) by separating parous from nulliparous females.
+In what follows, we follow the derivation given by \cite{davidson1954estimation}, to determine an expression for the proportion of nulliparous mosquitoes in terms of the gonotrophic cycle length $G$, here assumed the same for each cycle subsequent to an adult emerging.
+If a constant daily survival rate $p$ is assumed, the proportion of emerging adults surviving to become parous is $p^G$.
+If the rate of recruitment into the adult classes is constant over time, the proportion of nulliparous mosquitoes is,
+%
+\begin{equation}\label{eq:detinova}
+\begin{aligned}
+\mathbb{n} &= \frac{p^G}{p^G + p^{2G} + p^{3G} + \cdots}\\
+&= p^G / \frac{p^G}{1-p^G}\\
+&= 1-p^G.
+\end{aligned}
+\end{equation}
+%
+The daily survival rate can hence be calculated by inverting eq.
+(\ref{eq:detinova}): $p = \sqrt[G]{\mathcal{P}}$, where $\mathcal{P}=1-\mathbb{n}$ is the proportion parous sampled from the population.
+
 \section{EIPs of vector species of malaria, dengue fever, chikungunya and Zika}

### 2019-06-17 writeUp/supplementary_materials_methods.tex
Updated SOM

@@
 \section{Detinova parity analysis}
-\cite{massey2016global} provide a global database for bionomic quantities for the dominant vector species of malaria. In this database, there are
-
-\cite{detinova1962age} provided an alternative approach to estimate the lifespan of individual specimens, based on sampling the parous rate in a population (the proportion of adult females that have laid eggs) by separating parous from nulliparous females.
+\cite{detinova1962age} provided an alternative approach to estimate the lifespan of individual specimens, based on sampling the parous rate in a population (the proportion of adult females that have laid eggs) using dissection to determine parity of each specimen.
 In what follows, we follow the derivation given by \cite{davidson1954estimation}, to determine an expression for the proportion of nulliparous mosquitoes in terms of the gonotrophic cycle length $G$, here assumed the same for each cycle subsequent to an adult emerging.
@@
 (\ref{eq:detinova}): $p = \sqrt[G]{\mathcal{P}}$, where $\mathcal{P}=1-\mathbb{n}$ is the proportion parous sampled from the population.
+
+\cite{massey2016global} provide a global database of bionomic quantities for the dominant vector species of malaria.
+This database contains the number of parous and nulliparous samples in studies where Detinova's dissection approach was applied to field caught speciments ($n=1490$ non-missing observations) across Africa ($n=778$), Asia ($n=549$) and the Americas ($n=163$).
+
+We estimated lifespan at the species and complex / subgroup level (hereafter denoted as \textit{s.l.}) using a Bayesian hierarchical framework.
+The number of parous individuals $X_i^{\{s\}}$ in sample $i$ (consisting of species $s$) was modelled by a binomial sampling distribution,
+%
+\begin{equation}
+X_i^{\{s\}} \sim \mathcal{B}(N, \mathcal{P}_i^{\{s\}}),
+\end{equation}
+%
+where $N$ is the sample size and $\mathcal{P}_i^{\{s\}}$ is the proportion parous in sample $i$ for species $s$. A hierarchical prior structure was used for the parous proportion in each sample,
+%
+\begin{equation}
+\mathcal{P}_i^{\{s\}}\sim \text{beta}\left(\phi^{\{s\}} \kappa^{\{s\}}, (1-\phi^{\{s\}})\kappa^{\{s\}}\right),
+\end{equation}
+%
+where $0\leq\phi^{\{s\}}\leq 1$ is the mean parous proportion and $\kappa^{\{s\}}>1$ is a concentration parameter (both of which are species-specific).
 

### 2019-08-07 writeUp/supplementary_materials_methods.tex
Added Detinova PPCs

@@
 
-\section{Detinova parity analysis}
+\section{Detinova parity analysis}\label{sec:detinova}
 \cite{detinova1962age} provided an alternative approach to estimate the lifespan of individual specimens, based on sampling the parous rate in a population (the proportion of adult females that have laid eggs) using dissection to determine parity of each specimen.
@@
 \cite{massey2016global} provide a global database of bionomic quantities for the dominant vector species of malaria.
-This database contains the number of parous and nulliparous samples in studies where Detinova's dissection approach was applied to field caught speciments ($n=1490$ non-missing observations) across Africa ($n=778$), Asia ($n=549$) and the Americas ($n=163$).
+This database contains the number of parous and nulliparous samples in studies where Detinova's dissection approach was applied to field caught speciments ($n=1490$ non-missing observations for total number of parous individuals) across Africa ($n=778$), Asia ($n=549$) and the Americas ($n=163$).
 
 We estimated lifespan at the species and complex / subgroup level (hereafter denoted as \textit{s.l.}) using a Bayesian hierarchical framework.
-The number of parous individuals $X_i^{\{s\}}$ in sample $i$ (consisting of species $s$) was modelled by a binomial sampling distribution,
+We denote this grouping variable by $s$ in what follows.
+The number of parous individuals $X_i^{\{s\}}$ in sample $i$ (consisting of individuals belonging to group $s$) was modelled by a binomial sampling distribution,
 %
@@
 %
-where $N$ is the sample size and $\mathcal{P}_i^{\{s\}}$ is the proportion parous in sample $i$ for species $s$. A hierarchical prior structure was used for the parous proportion in each sample,
+where $N$ is the sample size and $\mathcal{P}_i^{\{s\}}$ is the proportion parous in sample $i$ for group $s$. A hierarchical prior structure was used for the parous proportion in each sample,
 %
@@
 %
-where $0\leq\phi^{\{s\}}\leq 1$ is the mean parous proportion and $\kappa^{\{s\}}>1$ is a concentration parameter (both of which are species-specific).
+where $0\leq\phi^{\{s\}}\leq 1$ is the mean parous proportion and $\kappa^{\{s\}}>1$ is a concentration parameter (both of which are group-specific).
+
+The priors we used in this model were: $\phi\sim \text{beta}(20, 14)$ and $\kappa\sim \text{Pareto}(1, 1.5)$.
+These were chosen because this yielded a prior predictive distribution with a median value close to 10 (see Fig.
+\ref{fig:detinova_priors}).
+
+\begin{figure}[ht]
+	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/detinova_prior_predictive.pdf}}
+	\caption{\textbf{The prior predictive distribution of lifespan for the analysis of Detinova parity data.} Here, we used the estimates of first gonotrophic cycle duration, $G\sim \mathcal{N}(4.26, 0.38)$, resultant from the approach outlined in section \ref{sec:dissection_gonotrophicMethod}, to convert physiological survival to calendar days.
+The figure shows 100,000 samples from the prior predictive distribution.}\label{fig:detinova_priors}
+\end{figure}
+
+The Stan code used to fit the hierarchical model to data is shown below.
+
+\begin{minted}{stan}
+data{
+  int N;
+  int X[N];
+  int num[N];
+  int species[N];
+  int K;
+}
+
+parameters{
+  vector<lower=0, upper=1>[N] theta;
+  vector<lower=0, upper=1>[K] phi;
+  vector<lower=1>[K] kappa;
+}
+
+model{
+  for(i in 1:N){
+    X[i] ~ binomial(num[i], theta[i]);
+    theta[i] ~ beta(phi[species[i]] * kappa[species[i]],
+                    (1 - phi[species[i]]) * kappa[species[i]]);
+  }
+  kappa ~ pareto(1, 1.5);
+  phi ~ beta(20, 14);
+}
+\end{minted}
+
+\subsection{Model checking}
+To check the fit of the models to data, we simulated from the posterior predictive distribution for each grouping used in analysis.
+This means that we have two sets of posterior predictive figures: those where the check was performed at the species level; and those corresponding to the complex or subgroup level.
+In each of these figures, we plot observed parity versus quantiles of posterior predicted parity (black vertical error bars).
+The following describes which figures correspond to which geography and species grouping.
+
+\begin{itemize}
+	\item \textbf{Africa}: the complex-level posterior predictive distribution graph is shown in Figure \ref{fig:detinova_ppcs_africa_grouped}; the species-level graph in Figure \ref{fig:detinova_ppcs_africa_species}.
+	\item \textbf{Americas}: the complex-level posterior predictive distribution graph is shown in Figure \ref{fig:detinova_ppcs_americas_grouped}; the species-level graph in Figure \ref{fig:detinova_ppcs_americas_species}.
+	\item \textbf{Asia}: the complex-level posterior predictive distribution graph is shown in Figure \ref{fig:detinova_ppcs_asia_grouped}; the species-level graph in Figure \ref{fig:detinova_ppcs_asia_species}.
+\end{itemize}
+
+In most cases, the posterior predictive distribution overlaps with the observed data (when the black error bars overlap with the orange ${\text{observed}=\text{predicted}}$ lines), indicating that the model was a reasonable fit to the data.
+For the most extreme observations of parity - when either very few or most samples where parous - however, the model-estimated parity was moved closer to the grand mean.
+This effect is illustrated in the figures by the linear regression line fitted to the data (blue lines).
+In almost all cases, the regression line had a lower gradient than the orange ${\text{observed}=\text{predicted}}$ lines, indicating a degree of ``shrinkage to the mean''.
+This shrinkage to the mean is a common and desirable product of using hierarchical models and can improve the robustness of results by reducing the influence of outliers \citep{lambert2018student}.
+As such, we do not view this as evidence of model misfit.
+
+\begin{figure}[ht]
+	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/detinova_ppc_africa_grouped.pdf}}
+	\caption{\textbf{Posterior predicted parity versus observed quantities for African anopheline samples at the complex level.} The upper and lower extent of the vertical error bars indicate 97.5\% and 2.5\% credible intervals and points indicate posterior medians.
+The blue lines show linear models fitted to pairs of posterior medians and observed parities, and orange ones show the ${\text{observed}=\text{predicted}}$ line.
+The analysis was run using 200 iterations per chain across 4 chains, with 100 of the initial samples discarded as warm-up.}\label{fig:detinova_ppcs_africa_grouped}
+\end{figure}
+
+\begin{figure}[ht]
+	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/detinova_ppc_africa_species.pdf}}
+	\caption{\textbf{Posterior predicted parity versus observed quantities for African anopheline samples at the species level.} The upper and lower extent of the vertical error bars indicate 97.5\% and 2.5\% credible intervals and points indicate posterior medians.
+The blue lines show linear models fitted to pairs of posterior medians and observed parities, and orange ones show the ${\text{observed}=\text{predicted}}$ line.
+The analysis was run using 200 iterations per chain across 4 chains, with 100 of the initial samples discarded as warm-up.}\label{fig:detinova_ppcs_africa_species}
+\end{figure}
+
+\begin{figure}[ht]
+	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/detinova_ppc_americas_grouped.pdf}}
+	\caption{\textbf{Posterior predicted parity versus observed quantities for American anopheline samples at the complex level.} The upper and lower extent of the vertical error bars indicate 97.5\% and 2.5\% credible intervals and points indicate posterior medians.
+The blue lines show linear models fitted to pairs of posterior medians and observed parities, and orange ones show the ${\text{observed}=\text{predicted}}$ line.
+The analysis was run using 200 iterations per chain across 4 chains, with 100 of the initial samples discarded as warm-up.}\label{fig:detinova_ppcs_americas_grouped}
+\end{figure}
+
+\begin{figure}[ht]
+	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/detinova_ppc_americas_species.pdf}}
+	\caption{\textbf{Posterior predicted parity versus observed quantities for American anopheline samples at the species level.} The upper and lower extent of the vertical error bars indicate 97.5\% and 2.5\% credible intervals and points indicate posterior medians.
+The blue lines show linear models fitted to pairs of posterior medians and observed parities, and orange ones show the ${\text{observed}=\text{predicted}}$ line.
+The analysis was run using 200 iterations per chain across 4 chains, with 100 of the initial samples discarded as warm-up.}\label{fig:detinova_ppcs_americas_species}
+\end{figure}
+
+\begin{figure}[ht]
+	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/detinova_ppc_asia_grouped.pdf}}
+	\caption{\textbf{Posterior predicted parity versus observed quantities for Asian anopheline samples at the complex level.} The upper and lower extent of the vertical error bars indicate 97.5\% and 2.5\% credible intervals and points indicate posterior medians.
+The blue lines show linear models fitted to pairs of posterior medians and observed parities, and orange ones show the ${\text{observed}=\text{predicted}}$ line.
+The analysis was run using 200 iterations per chain across 4 chains, with 100 of the initial samples discarded as warm-up.}\label{fig:detinova_ppcs_asia_grouped}
+\end{figure}
+
+\begin{figure}[ht]
+	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/detinova_ppc_asia_species.pdf}}
+	\caption{\textbf{Posterior predicted parity versus observed quantities for Asian anopheline samples at the species level.} The upper and lower extent of the vertical error bars indicate 97.5\% and 2.5\% credible intervals and points indicate posterior medians.
+The blue lines show linear models fitted to pairs of posterior medians and observed parities, and orange ones show the ${\text{observed}=\text{predicted}}$ line.
+The analysis was run using 200 iterations per chain across 4 chains, with 100 of the initial samples discarded as warm-up.}\label{fig:detinova_ppcs_asia_species}
+\end{figure}
 

### 2019-08-15 writeUp/supplementary_materials_methods.tex
Added more SOM

--- 2019-06-17 writeUp/supplementary_materials_methods.tex
@@
 }
-
+\end{minted}
+
+\begin{minted}{stan}
 // Priors
@@
 
+\subsection{Data and model}\label{sec:detinova_datamodel}
 \cite{massey2016global} provide a global database of bionomic quantities for the dominant vector species of malaria.
-This database contains the number of parous and nulliparous samples in studies where Detinova's dissection approach was applied to field caught speciments ($n=1490$ non-missing observations) across Africa ($n=778$), Asia ($n=549$) and the Americas ($n=163$).
+This database contains the number of parous and nulliparous samples in studies where Detinova's dissection approach was applied to field caught speciments ($n=1490$ non-missing observations for total number of parous individuals) across Africa ($n=778$), Asia ($n=549$) and the Americas ($n=163$).
 
 We estimated lifespan at the species and complex / subgroup level (hereafter denoted as \textit{s.l.}) using a Bayesian hierarchical framework.
-The number of parous individuals $X_i^{\{s\}}$ in sample $i$ (consisting of species $s$) was modelled by a binomial sampling distribution,
+We denote this grouping variable by $s$ in what follows.
+The number of parous individuals $X_i^{\{s\}}$ in sample $i$ (consisting of individuals belonging to group $s$) was modelled by a binomial sampling distribution,
 %
@@
 %
-where $0\leq\phi^{\{s\}}\leq 1$ is the mean parous proportion and $\kappa^{\{s\}}>1$ is a concentration parameter (both of which are species-specific).
+where $0\leq\phi^{\{s\}}\leq 1$ is the mean parous proportion and $\kappa^{\{s\}}>1$ is a concentration parameter (both of which are group-specific).
+
+The priors we used in this model were: $\phi\sim \text{beta}(20, 14)$ and $\kappa\sim \text{Pareto}(1, 1.5)$.
+These were chosen because this yielded a prior predictive distribution with a median value close to 10 (see Fig.
+\ref{fig:detinova_priors}).
+
+\begin{figure}[ht]
+	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/detinova_prior_predictive.pdf}}
+	\caption{\textbf{The prior predictive distribution of lifespan for the analysis of Detinova parity data.} Here, we used the estimates of first gonotrophic cycle duration, $G\sim \mathcal{N}(4.26, 0.38)$, resultant from the approach outlined in section \ref{sec:dissection_gonotrophicMethod}, to convert physiological survival to calendar days.
+The figure shows 100,000 samples from the prior predictive distribution.}\label{fig:detinova_priors}
+\end{figure}
+
+The Stan code used to fit the hierarchical model to data is shown below.
+
+\begin{minted}{stan}
+data{
+  int N; // total obs
+  int X[N]; // number parous
+  int num[N]; // number of samples collected
+  int species[N]; // integer index array with group identity
+  int K; // number of different groups
+}
+
+parameters{
+  vector<lower=0, upper=1>[N] theta;
+  vector<lower=0, upper=1>[K] phi;
+  vector<lower=1>[K] kappa;
+}
+
+model{
+  for(i in 1:N){
+    X[i] ~ binomial(num[i], theta[i]);
+    theta[i] ~ beta(phi[species[i]] * kappa[species[i]],
+                    (1 - phi[species[i]]) * kappa[species[i]]);
+  }
+  kappa ~ pareto(1, 1.5);
+  phi ~ beta(20, 14);
+}
+\end{minted}
+
+\subsection{Data processing}
+In this section, we describe the data processing carried out on the \cite{massey2016global} dataset.
+For each of the three datasets (Africa, Americas and Asia) contained with the SOM of \cite{massey2016global}, we used only those observations which had a non-missing and positive total capture size.
+For each observation, we calculated observed parity ($=\text{parous count}/ \text{total capture size}$) and compared this quantity with the `parity percent' column in the datasets.
+If the difference between these two quantities was lower than 1\%, we considered the discrepancy to be due to numerical rounding and did not investigate further.
+In those cases where the difference exceeded 1\%, we did the following:
+
+\begin{itemize}
+	\item Africa: considering the database entry for \cite{ijumba2002malaria}, there was a difference of about 50\%.
+Reading the paper, it looks like there is a mistake in their Table 5, and that the parous number should be 131 and, as such, we changed this number in the database; considering the entry for \cite{bockarie46service}, there was a difference of around 5\%.
+We could not access the paper here, so assumed that the issue was with the `parity percent' entry.
+	\item Americas: no differences exceeding 1\%.
+	\item Asia: considering the entry for \cite{kitthawee1992relationship}, there was a difference of around 3\%.
+Reading the paper, it looks like there was a (small) error in the database, and the parous number should be 35 (rather than 36) - this adjustment was made for the ensuing analysis.
+\end{itemize}
+
+For the species- and complex-level estimates of lifespan, we considered only those groups with five or more parity experiments.
+For the continent-level and overall estimates, we used all data.
+In quite a few cases, a species was indicated as a complex (for example, `Anopheles gambiae complex'), presumably when insufficient information was available in the paper to identify the species.
+These data were dropped for the species-level analysis but were included for all other analyses.
+
+For the species-level analyses, there were, hence, $n=223$ observations for Africa across 4 species; $n=141$ observations for the Americas across 6 species; and $n=214$ observations for Asia across 8 species.
+For the complex-level analyses, there were, hence, $n=757$ observations for Africa across 3 groups; $n=77$ observations for the Americas across 3 groups; and $n=455$ observations for Asia across 11 groups.
+For the continent-level and overall analyses, there were $n=1490$ observations across Africa ($n=778$), the Americas ($n=163$) and Asia ($n=549$).
+
+\subsection{Model estimation by MCMC and cross validation}
+The statistical models were fit to data using Stan's NUTS algorithm \cite{carpenter2016stan}.
+In all cases, models were run for 2000 iterations on each of 4 chains, with the first half of the iterations discarded as warm-up.
+We monitored diagnosed convergence as per \cite{gelman1992inference}, with $\hat{R}<1.1$ on all parameters.
+In all model runs, there were fewer than a handful of divergent transitions out of 2000 total iterations.
+
+To evaluate model fit to data, we used K-fold cross validation with 4 folds for each analysis.
+This means that, in a run, approximately $3/4$ of data was used to train models, and the remaining data used to evaluate out-of-sample predictive capability through the log-likelihood.
+To evaluate out-of-sample predictive log-likelihood for a given data point, we drew data-level parameters $\mathcal{P}_i^{\{s\}}$ from the relevant top-level distribution; for example, for a species-level fit, we drew data-level parameters from the species-level distribution relevant to the species for the data point in question.
+
+The folds were chosen so that, at all levels of analysis, there was always at least a handful of data points belonging to a particular group in the training set.
+This meant that, to generate meaningful comparisons between the species-level, complex-level and continent-level models, we selected only those data points with at least 5 observations for each species and complex.
+This reduced the number of data points used for model comparison versus those used for model estimation.
+For Africa, we had $n=223$ data points; for the Americas, there were $n=141$; and for Asia, there were $n=214$.
+
+For the comparison between continent-level grouping and overall, we were able to use more data points since we were less constrained by the minimal numbers of data belonging to each group.
+This meant we had $n=1490$ data points, with $n=778$ for Africa, $n=163$ for the Americas, and $n=549$, for Asia.
+
+\subsection{Model checking}
+To check the fit of models to the data, we simulated from the posterior predictive distribution for each grouping used in analysis.
+This means that we have two sets of posterior predictive figures: those where the check was performed at the species level; and those corresponding to the complex or subgroup level.
+In each of these figures, we plot observed empirical cumulative distribution function (eCDF) for the proportion parous versus the equivalent from each posterior predictive dataset.
+
+\begin{itemize}
+	\item \textbf{Africa}: the complex-level posterior predictive distribution graph is shown in Figure \ref{fig:detinova_ecdf_africa_grouped}; the species-level graph in Figure \ref{fig:detinova_ecdf_africa_species}.
+	\item \textbf{Americas}: the complex-level posterior predictive distribution graph is shown in Figure \ref{fig:detinova_ecdf_americas_grouped}; the species-level graph in Figure \ref{fig:detinova_ecdf_americas_species}.
+	\item \textbf{Asia}: the complex-level posterior predictive distribution graph is shown in Figure \ref{fig:detinova_ecdf_asia_grouped}; the species-level graph in Figure \ref{fig:detinova_ecdf_asia_species}.
+\end{itemize}
+
+In all cases, the eCDF for the actual data was bounded by the range of the posterior predictive eCDFs. By this metric, we conclude that the model seems to represent the data well.
+
+\begin{figure}[ht]
+	\centerline{\includegraphics[width=0.7\textwidth]{./Figure_files/detinova_ecdf_africa_grouped.pdf}}
+	\caption{\textbf{Africa, complex-level: empirical cumulative distribution function for observations (orange) and posterior predictive simulations (green).} In each case, we ran 4 Markov chains, with 200 iterations per chain; discarding the first 100 iterations as warm-up.}\label{fig:detinova_ecdf_africa_grouped}
+\end{figure}
+
+\begin{figure}[ht]
+	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/detinova_ecdf_africa_species.pdf}}
+	\caption{\textbf{Africa, species-level: empirical cumulative distribution function for observations (orange) and posterior predictive simulations (green).} In each case, we ran 4 Markov chains, with 200 iterations per chain; discarding the first 100 iterations as warm-up.}\label{fig:detinova_ecdf_africa_species}
+\end{figure}
+
+\begin{figure}[ht]
+	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/detinova_ecdf_americas_grouped.pdf}}
+	\caption{\textbf{Americas, complex-level: empirical cumulative distribution function for observations (orange) and posterior predictive simulations (green).} In each case, we ran 4 Markov chains, with 200 iterations per chain; discarding the first 100 iterations as warm-up.}\label{fig:detinova_ecdf_americas_grouped}
+\end{figure}
+
+\begin{figure}[ht]
+	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/detinova_ecdf_americas_species.pdf}}
+	\caption{\textbf{Americas, species-level: empirical cumulative distribution function for observations (orange) and posterior predictive simulations (green).} In each case, we ran 4 Markov chains, with 200 iterations per chain; discarding the first 100 iterations as warm-up.}\label{fig:detinova_ecdf_americas_species}
+\end{figure}
+
+\begin{figure}[ht]
+	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/detinova_ecdf_asia_grouped.pdf}}
+	\caption{\textbf{Asia, complex-level: empirical cumulative distribution function for observations (orange) and posterior predictive simulations (green).} In each case, we ran 4 Markov chains, with 200 iterations per chain; discarding the first 100 iterations as warm-up.}\label{fig:detinova_ecdf_asia_grouped}
+\end{figure}
+
+\begin{figure}[ht]
+	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/detinova_ecdf_asia_species.pdf}}
+	\caption{\textbf{Asia, species-level: empirical cumulative distribution function for observations (orange) and posterior predictive simulations (green).} In each case, we ran 4 Markov chains, with 200 iterations per chain; discarding the first 100 iterations as warm-up.}\label{fig:detinova_ecdf_asia_species}
+\end{figure}
 

### 2019-08-20 writeUp/parity_results.tex
Updated POA

\documentclass[]{article}

%opening
\title{Parity results}
\author{Ben Lambert}

\begin{document}

\maketitle

\section{Grouping level}
See \verb|grouping_holdouts.xlsx| for fits of species-level versus complex-level vs continent-level and for continent-level vs overall-level. For the former, the only interesting case is for the Americas, where the species-level model fits better than the complex level grouping - driven by \textit{A. albitarsis} which has species with strongly varying lifespan estimates. For the continent vs overall fits, Africa was better fit by a model grouping all species together than one allowing continental differences in lifespan; for Asia, the opposite trend was found, with the data being better fit by a model allowing continental differences in lifespan.

\section{Effect of temperature on lifespan}
In all cases, see Table X (from `\verb|temperature_holdout.xlsx|) for a summary.

\subsection{Mean temperature}
The overall hold-out log-likelihood was: 20,400.3 for the model without temperature terms; -20,164.61 for the model with linear temperature terms; and -19,131.74 for the model with up to quadratic temperature terms. The p values comparing to the model without temperature terms were: for the linear model, $p=0.18 (n=724, z=0.91)$; and for the quadratic model, $p<0.01 (n=724, z=3.64)$. The quadratic fit, however, is almost entirely driven by the result for \textit{A. albitarsis s.l.} (see `\verb|detinova_temperature_fits|'), which had significantly higher lifespan at both low or high mean temperatures. When this species is removed from the results, the difference versus the model without temperature terms is no longer significant ($p=0.38 (n=666, z=0.30)$.

\subsection{Daily (nightly) temperature range}
This is the model where we compare min-max for temperature extracted at midnight each day. The overall hold-out log-likelihood was: 20,400.3 for the model without temperature terms; -19,222.3 ($p=0.01, n=724, z=2.33$) for the model with linear temperature terms; and -18,741.22 ($p<0.01, n=724, z=3.60$) for the model with up to quadratic temperature terms. This result was driven by the fit of \textit{A. gambiae s.l.}, where our estimation determined that lifespan increased with daily temperature range.

\section{Anopheles gambiae investigations}
\subsection{Malaria prevalence vs parity}
There were only $n=142$ studies between 2000-2015 (when MAP estimates are available). For these cases, I've plotted 2-10 prevalence versus parity and find no correlation (see \verb|detinova_gambiae_pfpr.pdf|).

\end{document}

### 2019-08-20 writeUp/supplementary_materials_methods.tex
Updated POA

@@
 (\ref{eq:detinova}): $p = \sqrt[G]{\mathcal{P}}$, where $\mathcal{P}=1-\mathbb{n}$ is the proportion parous sampled from the population.
+The distribution of mosquitoes dying each day follows a geometric distribution, with probability of dying on a given day equal to $1-p$.
+The mean lifespan is, hence, given by $1/(1-p)$.
 
@@
 \begin{equation}
-X_i^{\{s\}} \sim \mathcal{B}(N, \mathcal{P}_i^{\{s\}}),
+X_i^{\{s\}} \sim \mathcal{B}(N_i, \mathcal{P}_i^{\{s\}}),
 \end{equation}
 %
-where $N$ is the sample size and $\mathcal{P}_i^{\{s\}}$ is the proportion parous in sample $i$ for group $s$. A hierarchical prior structure was used for the parous proportion in each sample,
-%
-\begin{equation}
+where $N_i$ is the sample size and $\mathcal{P}_i^{\{s\}}$ is the proportion parous in sample $i$ for group $s$. A hierarchical prior structure was used for the parous proportion in each sample,
+%
+\begin{equation}\label{eq:detinova_prior}
 \mathcal{P}_i^{\{s\}}\sim \text{beta}\left(\phi^{\{s\}} \kappa^{\{s\}}, (1-\phi^{\{s\}})\kappa^{\{s\}}\right),
@@
 The priors we used in this model were: $\phi\sim \text{beta}(20, 14)$ and $\kappa\sim \text{Pareto}(1, 1.5)$.
-These were chosen because this yielded a prior predictive distribution with a median value close to 10 (see Fig.
+These were chosen because this yielded a prior predictive distribution with a median lifespan close to 10 (see Fig.
 \ref{fig:detinova_priors}).
@@
 This meant we had $n=1490$ data points, with $n=778$ for Africa, $n=163$ for the Americas, and $n=549$, for Asia.
+
+To investigate the effect of measures of temperature, we extracted daily data from the ERA-interim re-analysis by ECMWF \citep{dee2011era} through their website using the Python API.
+This data included:
+
+\begin{itemize}
+	\item Two-meter surface temperature of air at 2m above the surface of land, sea or in-land waters.
+	\item Daily minimum two-meter surface temperature.
+	\item Daily maximum two-meter surface temperature.
+\end{itemize}
+
+For this analysis, we included only those studies where the start month listed in the database coincided with the end date - meaning the study took place within a given month - resulting in $n=742$ data points.
+The temperature data was extracted for the GPS position of each study ($\pm 0.1 \text{ degrees}$) for each day of the month over which mosquito sampling took place, with a grid size of 0.05 degrees.
+The Python script to extract this information is given below.
+
+\begin{minted}{python}
+from ecmwfapi import ECMWFDataServer
+import time
+def extract_meanminmax_temperature(a_area, a_date_start,
+		a_date_end):
+  server = ECMWFDataServer()
+  date_range = str(a_date_start) + "/to/" + str(a_date_end)
+  date_file = str(a_date_start) + "_to_" + str(a_date_end)
+  area=(str(a_area["north"]) + '/' + str(a_area["west"]) +
+       '/' + str(a_area["south"]) + '/' + str(a_area["east"])) 
+  area_file = (str(a_area["north"]) + '_' + str(a_area["west"])
+  + '_' + str(a_area["south"]) + '_' + str(a_area["east"]))
+  a_filename = ("../data/raw/weather/temperature_" +
+  	str(area_file) + "_" + date_file + ".nc")
+  server.retrieve({
+    "class": "ei",
+    "dataset": "interim",
+    "date": str(date_range),
+    "expver": "1",
+    "grid": "0.05/0.05",
+    "levtype": "sfc",
+    "area" : str(area),
+    "param": "167.128/201.128/202.128",
+    "step": "3",
+    "stream": "oper",
+    "time": "00:00:00",
+    "type": "fc",
+    "target":  str(a_filename),
+    "format" : "netcdf",
+  })
+\end{minted}
+
+To determine the impact of temperature on mosquito lifespan, we estimated a generalised linear model using a logit link, with hierarchical complex-level intercepts and either no temperature term, and linear temperature terms with or without quadratic temperature terms.
+In each case, these temperature terms were allowed to vary at the complex level (non-hierarchically).
+This meant our likelihood had the form,
+%
+\begin{equation}
+X_i^{\{s\}} \sim \mathcal{B}(N_i, \text{inv-logit}(\alpha_i^{\{s\}})),
+\end{equation}
+%
+where,
+%
+\begin{equation}
+\alpha_i^{\{s\}} = \text{logit} (\mathcal{P}_i^{\{s\}}) + \delta_1^{\{s\}} T_i + \delta_2^{\{s\}} T_i^2.
+\end{equation}
+%
+Here $T_i$ is a measure of temperature at study location $i$ (measures described below); $\mathcal{P}_i$ was set a prior according to eq.
+\eqref{eq:detinova_prior} and $\delta_1^{\{s\}}\sim \mathcal{N}(0, 1)$ and $\delta_2^{\{s\}}\sim \mathcal{N}(0, 1)$.
+
+The temperature measures ($T_i$), we investigated were summary statistics calculated across the daily data for all days within the month each study took place and across all grid points falling within the area centred on the GPS location of each study (described above),
+
+\begin{itemize}
+	\item Average two meter temperature (average of mean daily data for each grid point),
+	\item Minimum two meter temperature (average of minimum daily data for each grid point),
+	\item Maximum two meter temperature (average of maximum daily data for each grid point),
+	\item Standard deviation in two meter temperature (average of standard deviation in temperature for each grid point),
+	\item Daily range in two meter temperature (average of maximum minus minimum daily temperature for each grid point).
+\end{itemize}
 

### 2019-08-28 writeUp/parity_results.tex
Started new write up

@@
 
+\subsection{Country-level estimates}
+I have estimated country-level lifespans for the $n=541$ data points where GPS location was available.
+The lifespans are plotted in \verb|detinova_lifespan_gambiae_country.pdf|, which determines that Senegal has, on average, the longest lifespan.
+K-fold CV, however, indicated that these country-level differences were not significant, since a model grouping at the continent level performed better on out-of-sample prediction.
+This better performance was almost uniformly observed across the countries in the dataset.
+
+\subsection{With or without Reunion}
+On inspection of the plots of lifespan against temperature range (see \verb|detinova_gambiae_daynight_range_lifespan_country.pdf|), it looked as if the results may hinge on the data from Reunion.
+I, hence, repeated the analysis with/without this country.
+With this country included (see \verb|fit_complex_daynightmax.rds|), the models with linear and quadratic temperature terms fitted the data better.
+Without Reunion, the linear model fitted the data best but the difference in fit was weaker (compare \verb|detinova_gambiae_daynight_range_lifespan_country.pdf| with \verb|detinova_gambiae_daynight_range_lifespan_noreunion_country.pdf|) and no longer significant at the 5\% level (\verb|fit_complex_daynightmax_noreunion.rds|).
+
 \end{document}

### 2019-09-05 writeUp/supplementary_materials.tex
Updated POA and SOM

--- 2018-12-05 supplementary_materials.tex
@@
 \begin{figure}[ht]
-	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/dissection_gonotrophicCycleRaw.pdf}}
+	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/gonotrophic_data.pdf}}
 	\caption{\textbf{Raw gonotrophic cycle duration estimates for A.
 the 1st cycle and B.
-subsequent cycles.} The confidence intervals represent a range of uncertainties since they were not available in a standardised form (see Section \ref{sec:dissection_gonotrophicMethod} for a discussion of how we handle this issue).
+subsequent cycles.} When present, the confidence intervals represent our estimates of the 95\% confidence interval on gonotrophic cycle duration (see Section \ref{sec:dissection_gonotrophicMethod} for a discussion of how we handle this issue); if our estimate of the lower limit was below zero, we deemed this unbiological and set it to 1 day for this graph.
 If no statement of uncertainty was given in the estimates then we indicate this by a triangle marker.
-The studies are ordered according to the central estimate of gonotrophic duration.}\label{fig:dissection_gonotrophicCycleRaw}
-\end{figure}
-
-\begin{figure}[ht]
-	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/dissection_gonotrophicCycleRaw_MRRVsLab.pdf}}
-	\caption{\textbf{Raw gonotrophic cycle duration estimates from A.
-the MRR and B.
-from laboratory-based studies.} The confidence intervals represent heterogeneous measures since they were not available in a standardised form (see Section \ref{sec:dissection_gonotrophicMethod} for a discussion of we handle this issue).
-If no statement of uncertainty was given in the estimates then we indicate this by a triangle marker.
-The studies are ordered according to the central estimate of gonotrophic duration.}\label{fig:dissection_gonotrophicCycleRaw_MRRVsLab}
+The studies are ordered within each Genus according to the central estimate of gonotrophic duration.}\label{fig:dissection_gonotrophicCycleRaw}
 \end{figure}

### 2020-01-05 writeUp/supplementary_materials_methods.tex
updated SOM and MS to include higher kappa estimates

@@
 In this analysis, both time series are included.
-	\item 483: There are three studies in the paper \cite[]{rawlings1981influence}, (by dye colour there is one for `M\&Y', another for white, and a third for green), whereas the database only contains one release series.
-In this analysis, we analyse all three time series.
-	\item 578: A null observation was removed on day 12, since isn't present in the original paper \cite[]{eyles1943experiment}.
-	\item 585: It is not clear from the original paper \cite[]{smith1941observations} whether the missing recapture observations on certain days are because no mosquitoes were found on those days, or because no efforts were made to recapture them on those days.
+	\item 483: There are three studies in the paper \cite[]{rawlings1981influence}, (by dye colour: there is one for `M\&Y', another for white, and a third for green), whereas the database only contains one release series.
+In this analysis, we analyse all three time series separately.
+	\item 578: A null observation was removed on day 12, since it isn't present in the original paper \cite[]{eyles1943experiment}.
+	\item 585: It is not clear from the original paper \cite[]{smith1941observations} whether missing recapture observations on certain days are because no mosquitoes were found on those days, or because no efforts were made to recapture them on those days.
 We have assumed that the number recaptured were as stated in the database on those days (missing, because no recapture efforts were made on those days).
 	\item 586: It was unclear from database whether the missing observations are due to lack of recapture, or no effort.
-Have assumed that zero mosquitoes were captured on these days as seems most likely (the paper could not be accessed).
+Have assumed that zero mosquitoes were captured on these days (the paper could not be accessed).
 	\item 342: Changed number of released from 6,000 to 12,000 following the original publication \cite[]{bryan1991anopheles}.
@@
 
-The R file \verb|s_clean_data.R| attached cleans the data, makes the above amendments and outputs the data in a form amenable for estimation using Stan \citep{carpenter2016stan}.
+The R file \verb|s_clean_data.R| cleans the data, makes the above amendments and outputs the data in a form amenable for estimation using Stan \citep{carpenter2016stan}.
 
@@
 	\end{tabular}%
-	\caption{\textbf{Summary of variables across all MRR time-series.} For the `Species' and `Genera' rows, the counts represent the number of distinct entries in the database; otherwise the counts represent the number of time series.}
+	\caption{\textbf{Summary of variables across all MRR time-series used in hierarchical analyses.} For the `Species' and `Genera' rows, the counts represent the number of distinct entries in the database; otherwise the counts represent the number of time series.}
 	\label{tab:mrr_aggregateData}%
@@
 		\end{tabularx}%
-		\caption{\textbf{Summary of data from individual MRR time-series.}}
+		\caption{\textbf{Summary of data from individual MRR time-series used in hierarchical analyses.}}
 		\label{tab:mrr_IndividualData}%
@@
 	\centerline{\includegraphics[width=1.25\textwidth]{./Figure_files/mrr_exampleMRRSeries.pdf}}
-	\caption{\textbf{Example MRR time-series for three different studies from the \cite{guerra2014global} database.} In all the above experiments there was a single release of a batch of marked mosquitoes.}
+	\caption{\textbf{Example MRR time-series for three different studies from the \cite{guerra2014global} database.} In all the above experiments, there was a single release of marked mosquitoes on day zero.}
 	\label{fig:mrr_exampleMRRSeries}
@@
 where $\lambda$ is the sum of the rates of death and dispersal from the study area.
-To estimate the lifespan of mosquitoes we use this model, partly because of the limited evidence we determine in favour of senescence (Figs.
-\ref{fig:mrr_elpd} and \ref{fig:dissection_elpd}).
+To estimate the lifespan of mosquitoes we use this model, partly because of the limited evidence we find for senescence.
 
@@
 \subsection{Individual time-series estimates}\label{sec:MRR_individual_analysis}
-We first treat each time-series separately, and estimate individual $(\lambda,\psi,\kappa)$ parameters of the statistical model (described in the previous section) for each time-series.
-To use a Bayesian methodology we must specify priors on this set of parameters.
-Here we choose to specify independent priors on each parameter of the form: $\lambda\sim \mathcal{N}(-2.32,1)$, $\psi\sim \text{exp}(50)$ (with $\psi$ constrained to be less than 0.1, a threshold below which the majority of daily recapture probabilities in the data are observed) and $\kappa\sim \textnormal{log-normal}(2,1)$.
+We first treat each time-series separately, and estimate individual $(\lambda,\psi,\kappa)$ parameters of the statistical model (described in Section \ref{sec:mrr_statistical}) for each time-series.
+To use a Bayesian methodology, we must specify priors on this set of parameters.
+Here, we choose to specify independent priors on each parameter of the form: $\lambda\sim \mathcal{N}(-2.32,1)$, $\psi\sim \text{exp}(50)$ (with $\psi$ constrained to be less than 0.1, a threshold below which the majority of daily recapture probabilities in the data are observed) and $\kappa\sim \textnormal{log-normal}(2,1)$.
 This prior on the rate parameter of the exponential distribution ($\lambda$) corresponds to a wide range of possible lifespans (Fig.
 \ref{fig:mrr_individualTimeSeries_priors}A), with a mean of 10 days.
-It was necessary to use an informative prior on $\psi$ to bound it to sensible values and to prevent the correlation of estimates with $\lambda$ leading to unreasonably short estimates of $\lambda$; the prior on $\kappa$ is fairly uninformative and allows a wide range of values (Fig.
+It was necessary to use an informative prior on $\psi$ to bound it to sensible values and to prevent unreasonably short estimates of $\lambda$; the prior on $\kappa$ is fairly uninformative and allows a wide range of values (Fig.
 \ref{fig:mrr_individualTimeSeries_priors}B,C).
@@
 We synthesise information from across all MRR experiments to produce more robust estimates of mosquito lifespan than can be obtained from considering the individual time-series separately.
-However there exists considerable heterogeneity across the experiments.
-This heterogeneity has two sources.
-There is that arising from variability in experimental methodology.
-There is also variability from actual differences in lifespan across the different mosquito cohorts; for example due to genetic differences between mosquito populations or due to climatic differences.
-Because of this heterogeneity we use a Bayesian hierarchical model which is akin to a `random effects' model in classical statistics.
+However, there exists considerable heterogeneity across the experiments.
+This heterogeneity has two sources: that arising from variability in experimental methodology; but also that from actual differences in lifespan across the different mosquito cohorts -- for example, due to genetic differences between mosquito populations or due to climatic differences.
+Because of this heterogeneity, we use a Bayesian hierarchical model which is akin to a `random effects' model in classical statistics.
 This type of model assumes that there is random variation in parameters at the individual time-series level, although each of the parameters is drawn from a common `population-level' distribution.
-In our case we separately estimate three different model groupings where the population-level distributions correspond to the species, genus and overall (across all studies) levels respectively (Fig.
+In our case, we separately estimate three different model groupings where the population-level distributions correspond to the species, genus and overall (across all studies) levels respectively (Fig.
 \ref{fig:mrr_methodDiagramHierarchy}).
-This allows us to estimate average mosquito lifespan for a given species or genus, or alternatively across all time-series, by independent sampling from the posterior predictive distribution from which the individual $\lambda_i$ are drawn.
+This allows us to estimate mosquito lifespan for a given species or genus by independent sampling from the posterior predictive distribution from which the individual $\lambda_i$ are drawn.
 
@@
 Whilst we assume that the variation in mosquito mortality across experiments is random in nature, we recognise that there may exist systematic sources of variation that we do not include in our model.
-To account for this potential source of bias we examined how the individual estimates of lifespan correlated with experimental covariates (trapping method and number or density of traps).
-The data on trapping method in the Guerra et al., (2014) database is less reliable than other data entries because published MRR studies may use multiple trapping methods and typically do not disaggregate captures by method.
-When we examined the correlation of trapping method with lifespan estimates, however, we did not find systematic variation by method (data not shown).
-We determined a positive correlation between trapping density and lifespan estimates (Fig.
-\ref{fig:mrr_lifeSpanVsTrapDensity}), which indicates the impact of mosquito dispersal on estimated lifespan.
-
-A Bayesian framework requires that we incorporate our pre-analysis beliefs into our estimates through the use of priors for all parameters in a particular model.
-In our hierarchical model we are required to specify prior distributions at two levels of the analysis.
+To account for this potential source of bias, we examined how the individual estimates of lifespan correlated with experimental covariates (number and density of traps).
+From this, we determined a slight but insignificant positive correlation between trapping number and density with individual lifespan estimates -- suggesting that downward bias due to insufficient trapping may be relatively small.
+
+In our hierarchical model, we must specify prior distributions at two levels of the analysis.
 The first of these links the individual time-series estimates with the overarching group-level distributions.
@@
 \ref{fig:mrr_methodDiagramHierarchy}).
-As an example, we suppose that the rate parameter $\lambda_i$ of our exponential survival model in experiment $i$ is given by:
+As an example, suppose that the rate parameter $\lambda_i$ of our exponential survival model in experiment $i$ is given by:
 %
@@
 \begin{equation}
-c_i \sim  \mathcal{N}(\mu_\lambda,\sigma)
+c_i \sim  \mathcal{N}(\mu_\lambda,\sigma).
 \end{equation}
@@
 On all $\delta$ parameters, except $\delta^{\text{mix}}_{\text{genus}[i]}$, we specify standard normal priors.
-We specify a uniform prior on $\delta^{\text{mix}}_{\text{genus}[i]}\sim U(0,\delta^m_{\text{genus}[i]})$ so that the effect of mixed releases lies between the female-only release effect (0) and the male-only release effect ($\delta^m_{\text{genus}[i]}$).
+We specify a uniform prior on $\delta^{\text{mix}}_{\text{genus}[i]}\sim U(0,\delta^m_{\text{genus}[i]})$, so that the effect of mixed releases lies between the female-only release effect (0) and the male-only release effect ($\delta^m_{\text{genus}[i]}$).
 
 The relatively complex nature of priors in hierarchical models make it important to determine their influence on inferences.
-We chose the above -- somewhat uninformative -- priors to allow a range of mosquito lifespans in order to minimise their effect on the estimates we report (Fig.
+We chose the above -- somewhat uninformative -- priors, to allow a range of mosquito lifespans in order to minimise their effect on the estimates we report (Fig.
 \ref{fig:mrr_meanLifespanPrior}).
 We also chose to set hierarchical priors on the remaining parameters in our models -- $\psi$ the probability of daily recapture, and $\kappa$ the `over-dispersion' parameter of a negative binomial distribution.
-For $\psi$ we chose an informative prior that placed all probability mass $\psi\leq 10\%$, since in most experiments the fraction of mosquitoes captured was significantly below this value.
-For $\kappa$ we chose a fairly wide prior that had most of its probability mass below $\kappa=20$ (Fig.
+For $\psi$, we chose an informative prior that placed all probability mass $\psi\leq 10\%$, since in most experiments the fraction of mosquitoes captured was significantly below this value.
+For $\kappa$, we chose a fairly wide prior that had most of its probability mass below $\kappa=20$ (Fig.
 \ref{fig:mrr_PsiKappaPriors}).
-In all cases the hierarchical priors were chosen to have similar implications on lifespan, recapture probability and over-dispersion at the individual time-series level as for the non-hierarchical analysis described previously.
+In all cases, the hierarchical priors were chosen to have similar implications on lifespan, recapture probability and over-dispersion at the individual time-series level as for the non-hierarchical analysis described previously.
 
@@
 	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_lifespan_priors.pdf}}
-	\caption{\textbf{Samples from the prior predictive distribution of mean lifespan across the six models of mosquito mortality introduced in Table \ref{tab:mrr_survivalDescription}.} In all cases 10,000 samples were generated using the priors indicated in Table \ref{tab:mrr_priors}, resulting in distributions with a mean close to 10 days.}
+	\caption{\textbf{Samples from the prior predictive distribution of mean lifespan across the six models of mosquito mortality introduced in Table \ref{tab:mrr_survivalDescription}.} In all cases, 10,000 samples were generated using the priors indicated in Table \ref{tab:mrr_priors}, resulting in distributions with a mean close to 10 days.}
 	\label{fig:mrr_meanLifespanPrior}
@@
 \subsection{Testing for age-dependent mortality in wild mosquitoes}\label{sec:mrr_age_dependence}
-Previous work has found evidence of age-dependent mortality in lab populations \citep{styer2007mosquitoes,dawes2009anopheles}, and less-commonly in wild mosquitoes \citep{clements1981analysis}.
-Furthermore, recent modelling work has examined the implications of departures from a constant risk of mortality \citep{styer2007mosquitoes,hancock2009age,novoseltsev2012age}.
-To determine whether senescence occurs in wild mosquitoes we re-estimate our model using survival functions, $S(t)$, that allow for a rate of mortality that can vary with age.
+Previous work has found evidence of age-dependent mortality in lab populations \citep{styer2007mosquitoes,dawes2009anopheles}, and less-commonly in wild mosquitoes \citep{clements1981analysis,harrington2008age}.
+Furthermore, modelling work has examined the implications of departures from a constant risk of mortality \citep{styer2007mosquitoes,hancock2009age,novoseltsev2012age}.
+To determine whether senescence occurs in wild mosquitoes, we re-estimate the model using a range of different survival functions, $S(t)$, that permit mortality to increases with age.
 Specifically, we re-estimate our model using five other models that were previously used in the literature (see Table \ref{tab:mrr_survivalDescription} for a description of these).
@@
 The parameters of the survival function of each model were assigned hierarchical priors that allowed considerable variation in mosquito lifespan (see Fig.
-\ref{fig:mrr_meanLifespanPrior}).
-Otherwise the statistical model we used remained the same as for the constant mortality case.
+\ref{fig:mrr_meanLifespanPrior}) but had similar means as to the constant mortality case.
+Otherwise, the statistical model we used remained the same as for the constant mortality case.
 In testing for age-dependent mortality, we did not account for sex or pre-release feeding effects due to the complexity of including these elements in more complex survival models.
@@
 $\Gamma(\theta)$ and $\Gamma(\theta_1,\theta_2)$ refer to the Euler gamma function, and incomplete gamma function respectively.
-The `mean time spent in study area' is an estimate of the combined effects of mosquito mortality and dispersal from the area of the study where collections take place, since our data does not provide spatially-resolved data.}\label{tab:mrr_survivalDescription}
+The `mean time spent in study area' is an estimate of the combined effects of mosquito mortality and dispersal from the area of the study where collections take place, since we do not analyse spatially-resolved data.}\label{tab:mrr_survivalDescription}
 \end{table}
@@
 			\bottomrule
-	\end{tabularx}}\caption{\textbf{The priors used on parameters of each different survival model.} For the exponential model the `group-level' priors were the same for the genus and `overall' models that were also estimated.
+	\end{tabularx}}\caption{\textbf{The priors used on parameters of each different survival model.} For the exponential model, the `group-level' priors were the same for the genus and `overall' models that were also estimated.
 The exponential model was the only model that was simple enough to allow the scale parameter of the log-normal ($\sigma$) to be estimated by the data.
@@
 \subsection{Model estimation by MCMC}\label{sec:mrr_MCMC}
-The likelihood and priors we use result in posterior distributions whose analytic form cannot be calculated with existent computational methods.
-Instead we use Markov Chain Monte Carlo (MCMC) methods to sample from each posterior distribution.
-We used \textit{Stan} software \citep{carpenter2016stan} that implements an efficient MCMC algorithm known as NUTS \citep{hoffman2014no}.
-
-To judge convergence of the sampling algorithm to the posterior density, we calculated $\hat{R}$ across all Markov chains -- a ratio that compares the between-chain variation to that within each chain that is commonly used to measure convergence in MCMC \citep{gelman1992inference}.
+The likelihood and priors used result in posterior distributions whose analytic form cannot be calculated with existent computational methods.
+Instead, we use Markov Chain Monte Carlo (MCMC) methods to sample from each posterior distribution.
+Specifically, we used \textit{Stan} software \citep{carpenter2016stan} that implements an efficient MCMC algorithm known as NUTS \citep{hoffman2014no}.
+
+To judge convergence of the sampling algorithm to the posterior density, we calculated $\hat{R}$ across all Markov chains -- a ratio that compares between-chain variation to that within each chain that is commonly used to measure convergence in MCMC \citep{gelman1992inference}.
 For each class of model, we used the following MCMC parameters,
@@
 For the age-dependent cases, there remained a handful of parameters where $\hat{R}>1.1$ after running simulations for 3000 iterations but this did not affect our inferences, since repeated runs demonstrated the same pattern.
-We also ensured that across each MCMC run the number of divergent iterations (that can bias the MCMC away from the true posterior density) was minimal.
+We also ensured that across each MCMC run, the number of divergent iterations (that can bias the MCMC away from the true posterior density) was minimal.
 In the majority of cases, the number of divergent iterations was far fewer than 1\% of the total number of samples.
 
-The Stan scripts for the lifespan estimation (i.e.
+The Stan scripts for lifespan estimation (i.e.
 those using an exponential survival model) for the MRR analyses are provided below.
@@
 We use cross-validation to compare the predictive power of the species, genus and overall models, and later to determine whether age-dependent mortality occurs.
-In particular, we use K-Fold cross validation where we repeatedly partition our dataset into a test set composed of a number of individual time-series, and a training set of the remainder (see for example, \cite{marsland2015machine}).
+In particular, we use K-Fold cross validation (with $K=4$ folds for the MRR analysis), where we repeatedly partition our dataset into a test set composed of a number of individual time-series, and a training set of the remainder (see for example, \cite{marsland2015machine}).
 We then use the fitted model to measure the predictive performance on the test set.
@@
 
-Since the models are hierarchical we must specify how to draw parameters for a given test sample time-series.
+Since the models are hierarchical, we must specify how to draw parameters for a given test sample time-series.
 Here we chose to draw values of the individual time-series parameters as samples from the population distribution that corresponds to that particular sample's respective grouping.
 So if the particular test sample time-series pertains to \textit{Ae.
-aegypti} (and we are using the species-level model) then we draw independent samples for the statistical model's parameters from the estimated \textit{Ae.
+aegypti} (and we are using the species-level model), then we draw independent samples for the statistical model's parameters from the estimated \textit{Ae.
 aegypti} population distribution,
@@
 
-\section{Dissection based estimates}
-
-\subsection{Collection of dissection data}\label{sec:dissection_dissectionData}
+\section{Polovoda dissection experiments}
+
+\subsection{Collection of dissection data on physiological age}\label{sec:dissection_dissectionData}
 A comprehensive search of the literature using Google Scholar (\url{scholar.google.co.uk}) was performed using various combinations of the following keywords: dissection, mosquito, parity, parous, age, and physiological age.
@@
 This list was supplemented with a number of author-specific searches for those individuals most prevalent in the literature.
-In particular we searched for all articles authored by: Charlwood, Muller, Schlein, Samarawickrema, Reisen, Detinova, Polodova, Gillies, and Wilkes.
+In particular, we searched for all articles authored by: Charlwood, Muller, Schlein, Samarawickrema, Reisen, Detinova, Polodova, Gillies, and Wilkes.
 An additional list of potential articles was provided by doing a forward article search on some of the most highly-cited articles in the database: \cite{polovodova1949determination,detinova1962age,gillies1965study,clements1981analysis}.
@@
 A relational database was used to store the raw data from the actual experiments, along with the meta-data associated with each of the experiments.
-In many of the published articles wild mosquitoes were caught and dissected over a period of time, and the raw data thus consists of snapshots of the age structure of the population at regular intervals in time.
+In many of the published articles, wild mosquitoes were caught and dissected over a period of time, and the raw data thus consists of snapshots of the age structure of the population at regular intervals in time.
 In the cases where the data was more sparse (fewer than ten individuals, on average, per date), we aggregated across dates and recorded this as a single entry; otherwise we recorded the snapshots of the population at each different date.
@@
 
-For each individual series we recorded the following meta-data: genus, species, collection method, whether or not insecticide was mentioned as being used during or shortly before the time of mosquito collections, and the start and end dates of the experiment.
-At the article level we recorded the following meta-data: author, year, title, country, location (within country), start and end date, whether insecticide was used during any of the experimental replicates, whether a mrr experiment occurred alongside the dissection study, and the collection location (indoor or outdoor).
-At either the individual series or article levels we record additional meta-data describing the nature of data collection, for example explaining where the data was contained within the article, whether it was obtained by digitising graphs, and the number of separate dated series.
+For each individual series, we recorded the following meta-data: genus, species, collection method, whether or not insecticide was mentioned as being used during or shortly before the time of mosquito collections, and the start and end dates of the experiment.
+At the article level, we recorded the following meta-data: author, year, title, country, location (within country), start and end date, whether insecticide was used during any of the experimental replicates, whether an MRR experiment occurred alongside the dissection study, and the collection location (indoor or outdoor).
+At either the individual series or article levels, we recorded additional meta-data describing the nature of data collection, for example explaining where the data was contained within the article, whether it was obtained by digitising graphs, and the number of separate dated series.
 For those few cases ($n=16$ series) where the data was obtained by digitising graphs, we used the WebPlotDigitizer online tool \citep{digitise}.
@@
 The statistical approach applied to the data relies on the assumption that there is a constant rate of recruitment into the adult population (see Section \ref{sec:dissection_dissectionStats}).
-If populations fluctuate strongly from month-to-month this assumption will likely be violated.
-To try to mitigate against such a risk, we aggregate the data across all dates to obtain a single series for each identifier, resulting in 201 such series.
-To obtain a reasonable level of accuracy on estimates we remove all those individual (aggregated) series where there are fewer than 100 mosquitoes in total in the series.
-Finally we remove data for any species where there was only one series resulting in 131 series across four genera (\textit{Anopheles}, \textit{Culex}, \textit{Aedes}, and \textit{Mansonia}) and 25 species (Table \ref{tab:dissection_speciesNumbers}).
+If populations fluctuate strongly from month-to-month, this assumption will likely be violated.
+To try to mitigate against such a risk, we aggregated the data across all dates to obtain a single series for each identifier, resulting in 201 such series.
+To obtain a reasonable level of accuracy on estimates, we removed all those individual (aggregated) series where fewer than 100 mosquitoes were collected.
+Finally, we remove data for any species where there was only one series resulting in 131 series across four genera (\textit{Anopheles}, \textit{Culex}, \textit{Aedes}, and \textit{Mansonia}) and 25 species (Table \ref{tab:dissection_speciesNumbers}).
 These studies are distributed across a wide range of geographies (Fig.
@@
 	\centerline{\includegraphics[width=0.8\textwidth]{./Figure_files/dissection_timeSeries.pdf}}
-	\caption{\textbf{The numbers of published studies that estimate gonotrophic age by dissection across the study period.}}\label{fig:dissection_timeSeries}
+	\caption{\textbf{The numbers of published studies that estimate gonotrophic age by Polovodova's dissection method across the study period.}}\label{fig:dissection_timeSeries}
 \end{figure}
@@
 	\centerline{\includegraphics[width=1.25\textwidth]{./Figure_files/dissection_map.pdf}}
-	\caption{\textbf{The geographic location of each of the dissection databases included in the meta-analysis.} The area of the bubbles indicates the number of unique data series available.}\label{fig:dissection_map}
+	\caption{\textbf{The geographic location of each of the Polovodova datasets included in the meta-analysis.} The area of the bubbles indicates the number of unique data series available.}\label{fig:dissection_map}
 \end{figure}
@@
 	\end{tabular}%
-	\caption{\textbf{The numbers of dissection series for each species or genus included in the overall dataset.}}\label{tab:dissection_speciesNumbers}%
+	\caption{\textbf{The numbers of Polovodova dissection series for each species or genus included in the overall dataset.} Note, these counts represent numbers of series analysed by the hierarchical models -- that is, after aggregating the individual series as described in Section \ref{sec:dissection_dissectionData}.}\label{tab:dissection_speciesNumbers}%
 \end{table}%
@@
 	\footnotesize
-	\begin{adjustwidth}{-0.5in}{-0.5in}%
-		\begin{tabularx}{1.25\textwidth}{l|ccccc}
+	\begin{adjustwidth}{-0.3in}{-0.3in}%
+		\begin{tabularx}{1.1\textwidth}{l|ccccc}
 			\toprule
@@
 	\end{adjustwidth}
-	\caption{\textbf{Summaries of the characteristics of the 131 individual dissection data series that are included in this analysis.} Note when series are censored (see Section \ref{sec:dissection_censored}) we assume that all mosquitoes dissected over the threshold are of this age when calculating summary statistics.}\label{tab:dissection_summaryStats}
+	\caption{\textbf{Summaries of the characteristics of the 131 individual Polovodova dissection data series that are included in this analysis.} Note when series are censored (see Section \ref{sec:dissection_censored}) we assume that all mosquitoes dissected over the threshold are of this age when calculating summary statistics.}\label{tab:dissection_summaryStats}
 \end{table}
@@
 	\caption{\textbf{The normalised physiological age series for nine species in the database.} Each different coloured line represents an individual series.
-In each case the count for all ages has been normalised by the nulliparous count.
-In all cases we do not include any data for censored observations (see Section \ref{sec:dissection_censored}).}\label{fig:dissection_exampleData1}
+In each cas,e the count for all ages has been normalised by the nulliparous count.
+In all cases, we do not include any data for censored observations (see Section \ref{sec:dissection_censored}).}\label{fig:dissection_exampleData1}
 \end{figure}
@@
 	\caption{\textbf{The normalised physiological age series for nine species in the database.} Each different coloured line represents an individual series.
-In each case the count for all ages has been normalised by the nulliparous count.
-In all cases we do not include any data for censored observations (see Section \ref{sec:dissection_censored}).}\label{fig:dissection_exampleData2}
+In each case, the count for all ages has been normalised by the nulliparous count.
+In all cases, we do not include any data for censored observations (see Section \ref{sec:dissection_censored}).}\label{fig:dissection_exampleData2}
 \end{figure}
@@
 	\caption{\textbf{The normalised physiological age series for seven species in the database.} Each different coloured line represents an individual series.
-In each case the count for all ages has been normalised by the nulliparous count.
-In all cases we do not include any data for censored observations (see Section \ref{sec:dissection_censored}).}\label{fig:dissection_exampleData3}
-\end{figure}
-
-\subsection{Statistical analysis of dissection data}\label{sec:dissection_dissectionStats}
+In each case, the count for all ages has been normalised by the nulliparous count.
+In all cases, we do not include any data for censored observations (see Section \ref{sec:dissection_censored}).}\label{fig:dissection_exampleData3}
+\end{figure}
+
+\subsection{Statistical analysis of Polovodova dissection data}\label{sec:dissection_dissectionStats}
 We suppose that the number of individuals recruited into the adult female population is constant over time, meaning that the age-structure of the population is stable.
-Whilst environmental heterogeneity will naturally lead to variation in adult recruitment over time, we hope that by aggregating data across studies undertaken over a range of dates we may reduce the impact of this effect on our results.
+Whilst environmental heterogeneity will naturally lead to variation in adult recruitment over time, we hope that by aggregating data across studies undertaken over a range of dates this will reduce the impact of this effect.
 Suppose that the probability an individual female mosquito survives until age $a$ is given by the survival function $S(a)$, then the number of individuals in the population surviving to this age is given by,
@@
 %
-where $A(0)$ is the number of female mosquitoes recruited into the population per unit time.
+where $A(0)$ is the number of adult female mosquitoes recruited into the population per unit time.
 Consider one individual experiment where we randomly sample from a wild population that is structured as per eq.
 (\ref{eq:dissection_structured}).
-In this case the number of individuals sampled at each age is binomially-distributed,
+In this case, the number of individuals sampled at each age is binomially-distributed,
 %
 \begin{align}
-X(a) \sim \mathcal{B}(A(a),p(a)),
+y(a) \sim \mathcal{B}(A(a),p(a)),
 \end{align} 
@@
 where $p(a)$ is the probability of recapturing a single mosquito of age $a$.
-In what follows we assume that the probability of capturing a given mosquito is independent of their age, so that $p(a)=p=const$.
-Since in general $A(a)$ is large and $p(a)$ is small, we can approximate the above using a Poisson distribution,
+In what follows, we assume that the probability of capturing a given mosquito is independent of their age, so that $p(a)=p=const$.
+Since, in general, $A(a)$ is large and $p(a)$ is small, we can approximate the above using a Poisson distribution,
 %
 \begin{align}
-X(a) \sim \text{Poisson}(A(a)p).
+y(a) \sim \text{Poisson}(A(a)p).
 \end{align}
 %
-However we believe that the assumption of \textit{independent} captures of individual mosquitoes, which underlies the binomial and Poisson models, is likely suspect for the same reasons as for the MRR analysis (see Section \ref{sec:mrr_statistical}).
-As before we choose to specify a negative binomial sampling distribution that allows for non-independent captures,
+However, the assumption of \textit{independent} captures of individual mosquitoes, which underlies the binomial and Poisson models, is likely suspect for the same reasons as for the MRR analysis (see Section \ref{sec:mrr_statistical}).
+As before, we instead use a negative binomial sampling distribution that allows for non-independent captures,
 %
 \begin{align}
-X(a) \sim \text{NB}(A(a)p,\kappa),
+y(a) \sim \text{NB}(A(a)p,\kappa),
 \end{align} 
 %
-where we use the parameterisation such that the mean is given by $A(a)p$, and the over-dispersion parameter, $\kappa$, where as $\kappa\rightarrow \infty$ the above sampling distribution approaches a Poisson.
+where we use the parameterisation such that the mean is $A(a)p$, and over-dispersion parameter is $\kappa$, where, as $\kappa\rightarrow \infty$, the above sampling distribution approaches a Poisson.
 
 In the field, unfortunately, we do not know the number of mosquitoes recruited into the adult population, $A(a)$, nor the probability of capturing an individual at a given point in time, $p$.
-Instead we model their product $\Psi=A(0)p$ (the population of adult female mosquitoes of age zero that can be captured) probabilistically resulting in a model,
+Instead, we model their product $\Psi=A(0)p$ (the population of adult female mosquitoes of age zero that are expected to be captured) probabilistically resulting in a model,
 %
 \begin{align}\label{eq:dissertation_negativeBinomial}
-X(a) \sim \text{NB}(\Psi S(a),\kappa).
+y(a) \sim \text{NB}(\Psi S(a),\kappa).
 \end{align}
@@
 where $p(y(a)|S(a),\Psi,\kappa)$ corresponds to the negative binomial probability mass function for a count of $y(a)$ mosquitoes aged $a$ as specified in eqn.
-(\ref{eq:dissertation_negativeBinomial}), and $R$ is the number of separate physiological age classes in which the count was non-zero.
-%
-Since we do not know $\Psi$ we must learn it from the data.
+(\ref{eq:dissertation_negativeBinomial}), and $R$ is the number of separate physiological age classes recorded.
+%
+Since we do not know $\Psi$, we must learn it from the data.
 One approach to estimate this parameter could be use the number of captures of nulliparous mosquitoes, $X(0)$ in place of $\Psi$.
@@
 
-Since $S(a)$ is monotonically-decreasing we know that, if recruitment to the adult population is constant, then the numbers of nulliparous individuals should exceed the count in subsequent parity states.
-However in a number of data series there is a relative dearth of nulliparous mosquitoes versus uniparous individuals.
+Since $S(a)$ is monotonically-decreasing, we know that, if recruitment to the adult population is constant, then the numbers of nulliparous individuals should exceed the count in subsequent parity states.
+However, in a number of data series there is a relative dearth of nulliparous mosquitoes versus uniparous individuals.
 This deficiency has been noted in a previously-published study where the authors hypothesise that it is due to the issue of sampling the nulliparous population, since they are more likely to rest outside, compared with parous individuals \citep{gillies1965study}.
-However there is evidence to suggest that, on average, the first gonotrophic cycle is longer than subsequent cycles (see Section \ref{sec:dissection_gonotrophicData}), meaning that a relative surplus of nulliparous mosquitoes may exist in captured samples \citep{clements1981analysis}.
+However, there is conflicting evidence that suggests that, on average, the first gonotrophic cycle is longer than subsequent cycles (see Section \ref{sec:dissection_gonotrophicData}), meaning that a relative surplus of nulliparous mosquitoes may exist in captured samples \citep{clements1981analysis}.
 
 In our analysis, we chose to remove the counts of nulliparous individuals from the series where the count was low relative to uniparous or higher parity individuals since this may indicate issues with sampling the nulliparous population.
-Specifically we stipulated that the nulliparous count should exceed 90\% of the count for the uniparous individuals.
-For those cases where this condition was not met we removed the nulliparous observation and analyse the series of counts for all subsequent pars (uniparous and subsequent parous states).
-
-Here we estimate our model with one of six different survival functions (the same as for the MRR case; Table \ref{tab:mrr_survivalDescription}), each of which makes different assumptions regarding how the force of mortality is affected by mosquito age.
-To estimate lifespan at the species, genus and overall groupings we then use a hierarchical Bayesian model of the same mathematical form as for the analysis of MRR experiments (see Section \ref{sec:MRR_hierarchical}).
-However the priors for the analysis of dissection data were modified so that they represented a mean prior lifespan of three gonotrophic cycles, although allowed considerable variation in mean lifespan (Fig.
+Specifically, we stipulated that the nulliparous count should exceed 90\% of the count for the uniparous individuals.
+For those cases where this condition was not met, we removed the nulliparous observation and analyse the series of counts for all subsequent pars (uniparous and subsequent parous states).
+
+Here, we estimate our model with one of six different survival functions (the same as for the MRR case; Table \ref{tab:mrr_survivalDescription}), each of which makes different assumptions regarding how the force of mortality is affected by mosquito age.
+To estimate lifespan at the species, genus and overall groupings we then use a hierarchical Bayesian model of the same mathematical form as for the analysis of MRR experiments (i.e.
+assuming exponential survival functions; see Section \ref{sec:MRR_hierarchical}).
+However, the priors for the analysis of dissection data were modified so that they represented a mean prior lifespan of three gonotrophic cycles, although allowed considerable variation in this parameter (Fig.
 \ref{fig:dissection_lifespanPriorsParity}).
@@
 
-To determine whether mosquitoes experience age-dependent mortality we compared the predictive power of each of the models that incorporate a hazard function that depends on age with that from the exponential model.
-As for the MRR analysis we also used K-Fold cross-validation to perform this comparison, where the data are randomly partitioned into training and test sets (see Section \ref{sec:mrr_kFold}).
+To determine whether mosquitoes experience age-dependent mortality, we compared the predictive power of each of the models that incorporate a hazard function that increases with age with that from the exponential model.
+As for the MRR analysis, we also used K-Fold cross-validation to perform this comparison, where the data are randomly partitioned into training and test sets (see Section \ref{sec:mrr_kFold}).
 The model is then fitted to each training set and used to predict the data in the independent test set.
-Since there are fewer series than for the MRR dataset we used two partitions for each species, where each partition was of roughly the same number of individual series.
+Since there are fewer series than for the MRR dataset, we used two partitions for each species, where each partition had roughly the same count of individual series.
 
@@
 	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/dissection_lifespanPriorsParity.pdf}}
-	\caption{\textbf{The priors for mean lifespan used to analyse the data for each of the hierarchical models.} The distributional form of the priors is given in Table \ref{tab:dissection_priors}.
-In all cases the plots show data for 10,000 samples from the relevant prior distribution.}\label{fig:dissection_lifespanPriorsParity}
+	\caption{\textbf{The priors for mean lifespan used to analyse the Polovodova dissection data for each of the hierarchical models.} The distributional form of the priors is given in Table \ref{tab:dissection_priors}.
+In all cases, the plots show data for 10,000 samples from the relevant prior distribution.}\label{fig:dissection_lifespanPriorsParity}
 \end{figure}
@@
 			\bottomrule
-	\end{tabularx}}\caption{\textbf{The priors used on parameters of each different survival model for the dissection data analysis.} For the exponential model the `group-level' priors were the same for the genus and `overall' models that were also estimated.
+	\end{tabularx}}\caption{\textbf{The priors used on parameters of each different survival model for the Polovodova dissection data analysis.} For the exponential model the `group-level' priors were the same for the genus and `overall' models that were also estimated.
 The exponential model was the only model that was simple enough to allow the scale parameter of the log-normal ($\sigma$) to be estimated by the data.
@@
 
-The estimates of mean mosquito lifespan that we present here (Fig.
-\ref{fig:dissection_lifetimes_exponential}) resulted from the use of the exponential survival model (i.e.
+The estimates of mean mosquito lifespan that we present here resulted from the use of the exponential survival model (i.e.
 no age dependence).
@@
 However, since the priors for each of the survival functions were specified to allow for a wide variety of lifespans (Fig.
-\ref{fig:dissection_lifespanPriorsParity}) the specific choice of survival function made little difference to resultant estimates (data not shown).
-
-To demonstrate that the results we obtain are not sensitive to the particular form of hierarchical prior structure chosen we also provide estimates of the lifespan for each series analysed separately (i.e.
+\ref{fig:dissection_lifespanPriorsParity}), the specific choice of survival function made little difference to resultant estimates (data not shown).
+
+To demonstrate that the results we obtained are not sensitive to the particular form of hierarchical prior structure chosen, we also provide estimates of the lifespan for each series analysed separately (i.e.
 non-hierarchically).
-To do so we assume priors on the rate parameter of the exponential distribution that provides support over a wide range of possible lifespans (Fig.
-\ref{fig:dissection_lifespanPrior_individual}A).
+To do so, we assume priors on the rate parameter of the exponential distribution that provides support over a wide range of possible lifespans (Fig.
+\ref{fig:dissection_lifespanPrior_individual}A) but had a mean close to the hierarchical case equivalent.
 We also specified a prior on the over-dispersion parameter $\kappa$ that was comparable to the hierarchical case (Fig.
@@
 \subsubsection{Analysis of censored data series}\label{sec:dissection_censored}
-Some published studies do not distinguish the number of gonotrophic studies
-beyond a threshold, $a_T$ (the more ovariole dilations there are the harder it
-is to count them) which is akin to censoring the data. This
-censoring involves only a relatively small number of mosquitoes in each
-time series (median = 2\%) and so we are confident that these few cases do not substantially affect our inferences.
-We, however, now describe the statistical model used to describe these instances.
-In these cases we only know the number, $y(a_T)$, of individuals who were captured and dissected with an estimated physiological age that is greater than or equal to the threshold.
+Some published studies do not distinguish the number of gonotrophic studies beyond a threshold, $a_T$ (the more ovariole dilations there are the harder it is to count them) which is akin to censoring the data.
+This censoring involves only a relatively small number of mosquitoes in each time series (median = 2\%) and so we are confident that these few cases do not substantially affect our inferences.
+We now describe the statistical model used to describe these instances.
+In these cases, we only know the number, $y(a_T)$, of individuals who were captured and dissected with an estimated physiological age that is greater than or equal to the threshold.
 Since we do not know the estimated physiological age of individual specimens $i$, we represent this as a parameter, $\aleph_i$, in our statistical model.
@@
 To do this exactly would require an infinite sum over all the possible ages for all $y(a_T)$ mosquitoes that have been caught whose age exceeds the threshold.
-Rather than carry out this intractable number of summations we instead make the approximation that all mosquitoes in this group are of the same age $\aleph_i = \aleph,\; \forall i \in (1,...,y(a_T))$.
+Rather than carry out this intractable number of summations, we instead make the approximation that all mosquitoes in this group are of the same age $\aleph_i = \aleph,\; \forall i \in (1,...,y(a_T))$.
 We believe this assumption is justifiable, particularly since the numbers of mosquitoes in each subsequent age category is a strongly-decreasing function of age.
@@
 %
-In practice, the number of mosquitoes captured in those series that are censored represents a small percentage of total captures (the median is approximately 2\%), and so the effect of the $q(.)$ term above is likely minimal on resultant inferences.
+In practice, the number of mosquitoes captured in those series that are censored represents a small percentage of total captures, so the effect of the $q(.)$ term above is likely minimal on resultant inferences.
 
@@
 As for the MRR analysis, we used \textit{Stan} software \citep{carpenter2016stan} to sample from the posterior distributions of model parameters.
-
-To judge convergence of the sampling algorithm to the posterior density, we calculated $\hat{R}$ across all Markov chains -- a ratio that compares the between-chain variation to that within each chain that is commonly used to measure convergence in MCMC \citep{gelman1992inference}.
+To judge convergence of the sampling algorithm to the posterior density, we calculated $\hat{R}$ across all Markov chains \citep{gelman1992inference}.
 For each model, we ran 16 independent Markov chains with 200 iterations per chain, discarding the first half of these iterations as warm-up \citep{gelman2014bayesian}.
 After running the algorithm for the given number of observations, $\hat{R}<1.1$ for all model parameters.
-We also ensured that across each MCMC run the number of divergent iterations (that can bias the MCMC away from the true posterior density) was minimal.
+We also ensured that across each MCMC run, the number of divergent iterations (that can bias the MCMC away from the true posterior density) was minimal.
 In the majority of cases, the number of divergent iterations was far fewer than 1\% of the total number of samples.
@@
 	\includegraphics[page=1,scale=.7]{./Figure_files/dissection_ppcs_all}
-	\captionof{figure}{\textbf{The numbers of mosquitoes recaptured (black lines) versus the 95\% central posterior interval of the posterior predictive distribution (blue shading) for a selection of the dissection series.} The titles correspond to the unique combination of "identifier - physiological-experiment-id" in the database.
+	\captionof{figure}{\textbf{The numbers of mosquitoes recaptured (black lines) versus the 95\% central posterior interval of the posterior predictive distribution (blue shading) for a selection of the Polovodova dissection series.} The titles correspond to the unique combination of "identifier - physiological-experiment-id" in the database.
 For the rest of the posterior predictive checks for the dissection data models, see the file referenced in the text.}
@@
 	\includegraphics[page=2,scale=.7]{./Figure_files/dissection_ppcs_all}
-	\captionof{figure}{\textbf{The numbers of mosquitoes recaptured (black lines) versus the 95\% central posterior interval of the posterior predictive distribution (blue shading) for a selection of the dissection series.} The titles correspond to the unique combination of "identifier - physiological-experiment-id" in the database.
+	\captionof{figure}{\textbf{The numbers of mosquitoes recaptured (black lines) versus the 95\% central posterior interval of the posterior predictive distribution (blue shading) for a selection of the Polovodova dissection series.} The titles correspond to the unique combination of "identifier - physiological-experiment-id" in the database.
 For the rest of the posterior predictive checks for the dissection data models, see the file referenced in the text.}
@@
 	\includegraphics[page=3,scale=.7]{./Figure_files/dissection_ppcs_all}
-	\captionof{figure}{\textbf{The numbers of mosquitoes recaptured (black lines) versus the 95\% central posterior interval of the posterior predictive distribution (blue shading) for a selection of the dissection series.} The titles correspond to the unique combination of "identifier - physiological-experiment-id" in the database.
+	\captionof{figure}{\textbf{The numbers of mosquitoes recaptured (black lines) versus the 95\% central posterior interval of the posterior predictive distribution (blue shading) for a selection of the Polovodova dissection series.} The titles correspond to the unique combination of "identifier - physiological-experiment-id" in the database.
 For the rest of the posterior predictive checks for the dissection data models, see the file referenced in the text.}
@@
 
-\subsection{Data collection of gonotrophic cycle duration}\label{sec:dissection_gonotrophicData}
-To convert the estimates of lifespan in physiological age into chronological age we require estimates of the duration of the gonotrophic cycle.
-To determine this characteristic we conducted a meta-analysis of previously-published studies that estimate the duration of the gonotrophic cycle.
+\subsection{Data collection for gonotrophic cycle duration}\label{sec:dissection_gonotrophicData}
+To convert the estimates of lifespan in physiological age into chronological age requires a duration for the gonotrophic cycle.
+To determine this characteristic, we conducted a meta-analysis of previously-published studies that estimate the duration of the gonotrophic cycle.
 A search of the literature using Google Scholar (\url{scholar.google.co.uk}) was performed using the search term: `gonotrophic cycle duration'.
 The list of articles was then supplemented with a list of references discussed by \cite{silver2007mosquito}.
-Based on the abstracts of the resultant list of published studies we then decided whether to search each article for estimates of the duration of the gonotrophic cycle.
+Based on the abstracts of the resultant list of published studies, we then decided whether to search each article for estimates of the duration of the gonotrophic cycle.
 Overall 79 separate estimates of this parameter were found across 42 published articles (see Section \ref{sec:appendix_gonotrophicDurationStudyList} for a list of the studies included in our dataset).
-Along with information about the estimates we also recorded study and series meta-data, including the location of the study, the method used for estimation, the species and genus, as well as the temperatures and/or seasons in which the experiments were carried out.
-The methods used to estimate gonotrophic cycle duration in the literature largely fell into two distinct categories: those based on MRR studies ($n=29$); and those based on observations of mosquitoes in a laboratory setting ($n=42$).
-In the MRR studies estimates are made of the duration of gonotrophic cycles by dissecting recaptured mosquitoes at each time point using the method of \cite{polovodova1949determination} to count ovariolar dilations.
-For the laboratory studies the duration of gonotrophic cycles is determined more directly by observing the time taken for mosquitoes to mate then blood-feed and finally oviposit.
+Along with information about the estimates, we also recorded study and series meta-data, including the location of the study, the method used for estimation, species and genus.
+
+Whilst compiling our dataset on gonotrophic cycles, \cite{massey2016global} published a database of bionomic quantities for malaria vectors (that is, including only anopheline species).
+Included in this dataset were a number of estimates of gonotrophic cycle duration.
+After removing duplicates with our dataset, we were left with 120 estimates of gonotrophic cycle duration.
+The majority of these estimates were based on laboratory experiments conducted on wild-caught specimens or their $F_1$ or $F_2$ descendents ($n=45$), followed by MRR studies ($n=36$), then experiments on laboratory colonies ($n=11$).
+In MRR studies, estimates are made of the duration of gonotrophic cycles by dissecting recaptured mosquitoes at each time point using the method of \cite{polovodova1949determination} to count ovariolar dilations.
+For laboratory studies, duration of gonotrophic cycles is determined by direct observation.
 
 Along with point estimates of the parameter we also collected information about the uncertainty in the estimates (if available).
-In many articles the duration of the gonotrophic cycle was estimated separately for the first versus subsequent cycles, and these estimates were recorded separately.
-Raw estimates of gonotrophic cycle duration were obtained for species across four different genera (\textit{Aedes}, \textit{Anopheles}, \textit{Culex} and \textit{Masonia}), although there was a bias towards \textit{Anopheles} with $n=47$ estimates out of the total of 79 (Fig.
+In many articles, the duration of the gonotrophic cycle was estimated separately for the first versus subsequent cycles, and these estimates were recorded separately.
+If there was no disaggregation of durations between 1st and subsequent cycles, we assumed that each of these were the same.
+Raw estimates of gonotrophic cycle duration were obtained for species across three different genera (\textit{Aedes}, \textit{Anopheles} and \textit{Culex}), although there was a bias towards \textit{Anopheles} with $n=92$ estimates (Fig.
 \ref{S-fig:dissection_gonotrophicCycleRaw}).
-
-Since we collected data on the method used to estimate the duration of the gonotrophic cycle we show the raw estimates for the two most common approaches: MRR and laboratory (Fig.
-\ref{S-fig:dissection_gonotrophicCycleRaw_MRRVsLab}).
-From these results it is evident that the estimates from the laboratory studies are, on average, higher than those using the MRR method.
 
@@
 As discussed in Section \ref{sec:dissection_gonotrophicData}, there was considerable study-level heterogeneity in the information provided for the estimates of the gonotrophic cycle duration.
-A number of studies ($n=24$) provided no estimates of uncertainty in gonotrophic cycle duration, whereas the rest gave some indication of confidence or alternatively a range of possible estimates.
-However the types of uncertainty intervals that were specified varied considerably from study to study, from the vague (but common)  `4-6 days' to the more helpful `5$\pm 1$ day (95\% confidence interval)'.
+A number of studies ($n=42$) provided no estimates of uncertainty in gonotrophic cycle duration, whereas the rest gave some indication of confidence or alternatively a range of possible estimates.
+However, the types of uncertainty intervals that were specified varied considerably across studies: from the vague (but common), `4-6 days'; to the more helpful, `5$\pm 1$ day (95\% confidence interval)'.
 
@@
 We decided to model the estimates as representing quantiles from an underlying normal distribution that represents uncertainty over possible durations of the gonotrophic cycle.
-In those circumstances where lower, central and upper bounds were given explicitly the data was fairly symmetric and so we believe assuming the observations come from an unskewed normal distribution is reasonable.
+In those circumstances where lower, central and upper bounds were given explicitly, the data was fairly symmetric, so we believe assuming the observations come from an unskewed normal distribution is reasonable.
 Explicitly we treated each type of uncertainty interval as follows,
@@
 \begin{align}
-Z_{50} = \frac{G_{50} - \mu}{\sigma},
+Z_{x} = \frac{G_{x} - \mu}{\sigma},
 \end{align}
 %
-where $Z_{50}$ and $G_{50}$ indicate 50\% quantiles from a standard normal distribution and a normal distribution respectively. By manipulating the above expression we obtain the following,
+where $Z_{x}$ and $G_{x}$ indicate $x$\% quantiles from a standard normal distribution and a normal distribution respectively. By manipulating the above expression we obtain the following,
 %
 \begin{align}
-G_{50} = \sigma Z_{50} + \mu,
+G_{x} = \sigma Z_{x} + \mu,
 \end{align}
@@
 which forms a straight line in $(Z,G)$ space with slope $\sigma$ and y-intercept $\mu$.
-Therefore by plotting the raw observations of $(Z,G)$ quantiles then estimating a linear regression line we can characterise the underlying $\mathcal{N}(\mu,\sigma)$ distribution from which we assume they are drawn.
-
-Due to the relative unavailability of estimates for gonotrophic cycle duration across the different genera we generated a single distribution to represent the uncertainty in this parameter by pooling all the data.
-This was justified because there was not found to be significant variation in the 1st gonotrophic cycle duration (ANOVA: $F_{3,75}=2.6,p>0.05$).
-
-Using the collected data, we estimated that the first gonotrophic cycle duration was $\mathcal{N}(4.26, 0.38)$ and, for subsequent cycles, $\mathcal{N}(3.85, 0.35)$.
+Therefore by estimating a linear regression of $Z$ on $G$, we can characterise the underlying $\mathcal{N}(\mu,\sigma)$ distribution from which they are assumed to be drawn.
+
+An ANOVA using the central estimates of 1st gonotrophic cycle duration indicated that there was significant variation in this quantity between genera (ANOVA: $F_{2,116}=8.7,p<0.01$; Kruskal-Wallis rank sum test: $\chi^2_2=23.9,p<0.01$), so we estimated separate durations for each genus.
+The duration of the 1st gonotrophic cycles were estimated as follows: for \textit{Anopheles}, $\mathcal{N}(3.65, 0.40)$; for \textit{Aedes}, $\mathcal{N}(4.62, 0.51)$; and, for \textit{Culex}, $\mathcal{N}(5.23, 0.47)$.
+For the subsequent durations, the estimated durations were: for \textit{Anopheles}, $\mathcal{N}(3.21, 0.38)$; for \textit{Aedes}, $\mathcal{N}(4.78, 0.59)$; and, for \textit{Culex}, $\mathcal{N}(5.15, 0.47)$.
+Pooling data across all genera, the 1st cycle was estimated as $\mathcal{N}(3.95, 0.43)$ and the subsequent as $\mathcal{N}(3.61, 0.42)$.
+
+For the \textit{Mansonia} observations on physiological lifespan collected, we converted these using the overall (i.e. across all genera) estimates of gonotrophic cycle durations.
 
@@
 The estimates of lifespan produced from analysing the dissection data are in terms of physiological age (the number of gonotrophic cycles undertaken).
-To allow comparison with the estimates from the MRR studies (Fig.
-\ref{fig:mrr_lifespans}) as well as produce more useful estimates to inform disease transmission dynamical models we convert these estimates to calendar days.
-To do this we use the estimated parameters of the normal densities that we have assumed represent uncertainty in gonotrophic cycle duration (see Section \ref{sec:dissection_gonotrophicMethod}) and use them to convert our posterior samples of mean physiological lifespan $(L^p_1,...,L^p_S)$ to lifespan in calendar ages $(L^c_1,...,L^c_S)$.
-To do this we iterate the following for all $i\in (1,...,S)$ posterior samples,
+To allow comparison with the estimates from the MRR studies and to produce more useful estimates to inform disease transmission dynamical models, we convert these estimates to calendar days.
+To do this, we use the estimated parameters of the normal densities that we have assumed represent uncertainty in gonotrophic cycle duration (see Section \ref{sec:dissection_gonotrophicMethod}) and use them to convert our posterior samples of mean physiological lifespan $(L^p_1,...,L^p_S)$ into lifespan in calendar ages $(L^c_1,...,L^c_S)$.
+To do this, we iterate the following for all $i\in (1,...,S)$ posterior samples of physiological lifespan for each genus ($j$),
 
 \begin{enumerate}
-	\item Sample $G_{1i} \sim \mathcal{N}(\mu_{1i},\sigma_{1i})$, to obtain a duration for the 1st gonotrophic cycle.
-	\item Sample $G_{2i} \sim \mathcal{N}(\mu_{2i},\sigma_{2i})$, to obtain a duration for subsequent gonotrophic cycles.
+	\item Sample $G_{1i} \sim \mathcal{N}(\mu_{1j},\sigma_{1i})$, to obtain a duration for the 1st gonotrophic cycle (here, $\mu_{1j}$ and $\sigma_{1j}$ are parameters characterising the mean and std.
+dev.
+of a normal representing 1st cycle duration for genus $j$).
+	\item Sample $G_{2i} \sim \mathcal{N}(\mu_{2j},\sigma_{2j})$, to obtain a duration for subsequent gonotrophic cycles (here, $\mu_2$ and $\sigma_2$ are parameters characterising the mean and std.
+dev.
+of a normal representing subsequent cycle duration for genus $j$).
 	\item If $L^p_i > 1$, the mean lifespan is longer than one gonotrophic cycle:
@@
 
-In the using the above methodology to convert from physiological to calendar age we implicitly assume that for each gonotrophic cycle after the first that the cycles are of the same length (we choose one $G_{2i}$ per sample).
-However since we allow a different $G_{2i}$ for each sample we nonetheless believe that the above approach allows for sufficient uncertainty in the estimates of subsequent gonotrophic cycle durations.
+In the using the above methodology to convert from physiological to calendar age, we implicitly assume that all gonotrophic cycles after the first are of the same length (i.e.
+we choose one $G_{2i}$ per sample).
+However, since we allow a different $G_{2i}$ for each sample, we nonetheless believe that the above approach allows for sufficient uncertainty in the estimates of subsequent gonotrophic cycle durations.
 If instead we believed that significant variation in gonotrophic cycle occurred within a particular mosquito's life (as well as between mosquitoes) then we could draw a new value of $G_{2i}$ for each subsequent cycle.
-This produces results with a little more uncertainty than previously (data not shown).
-
-\section{Detinova parity analysis}\label{sec:detinova}
+This produces results with indistinguishable uncertainty than the results we report (data not shown).
+
+\section{Detinova dissection experiments}\label{sec:detinova}
 \cite{detinova1962age} provided an alternative approach to estimate the lifespan of individual specimens, based on sampling the parous rate in a population (the proportion of adult females that have laid eggs) using dissection to determine parity of each specimen.
@@
 We denote this grouping variable by $s$ in what follows.
-The number of parous individuals $X_i^{\{s\}}$ in sample $i$ (consisting of individuals belonging to group $s$) was modelled by a binomial sampling distribution,
+The number of parous individuals $y_i^{\{s\}}$ in sample $i$ (consisting of individuals belonging to group $s$) was modelled by a binomial sampling distribution,
 %
 \begin{equation}
-X_i^{\{s\}} \sim \mathcal{B}(N_i, \mathcal{P}_i^{\{s\}}),
+y_i^{\{s\}} \sim \mathcal{B}(N_i, \mathcal{P}_i^{\{s\}}),
 \end{equation}
@@
 where $0\leq\phi^{\{s\}}\leq 1$ is the mean parous proportion and $\kappa^{\{s\}}>1$ is a concentration parameter (both of which are group-specific).
-
-The priors we used in this model were: $\phi\sim \text{beta}(20, 14)$ and $\kappa\sim \text{Pareto}(1, 1.5)$.
-These were chosen because this yielded a prior predictive distribution with a median lifespan close to 10 (see Fig.
+The priors we used in this model were: $\phi\sim \text{beta}(20, 15)$ and $\kappa\sim \text{Pareto}(5, 1.5)$.
+These were chosen because this yielded a prior predictive distribution with a mean lifespan close to 10 (see Fig.
 \ref{fig:detinova_priors}).
-
-\begin{figure}[ht]
-	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/detinova_prior_predictive.pdf}}
-	\caption{\textbf{The prior predictive distribution of lifespan for the analysis of Detinova parity data.} Here, we used the estimates of first gonotrophic cycle duration, $G\sim \mathcal{N}(4.26, 0.38)$, resultant from the approach outlined in section \ref{sec:dissection_gonotrophicMethod}, to convert physiological survival to calendar days.
-The figure shows 100,000 samples from the prior predictive distribution.}\label{fig:detinova_priors}
-\end{figure}
 
@@
   vector<lower=0, upper=1>[K] phi;
-  vector<lower=1>[K] kappa;
+  vector<lower=5>[K] kappa;
 }
@@
   }
-  kappa ~ pareto(1, 1.5);
-  phi ~ beta(20, 14);
+  kappa ~ pareto(5, 1.5);
+  phi ~ beta(20, 15);
+}
+
+generated quantities{
+  real lifespan[K];
+  real survival[K];
+  real theta_average[K];
+  for(i in 1:K){
+    real gon = normal_rng(3.65, 0.4);
+    theta_average[i] = beta_rng(phi[i] * kappa[i],
+										       (1 - phi[i]) * kappa[i]);
+    survival[i] = theta_average[i] ^ (1.0 / gon);
+    lifespan[i] = 1.0 / (1.0 - survival[i]); 
+  }
 }
 \end{minted}
+
+\begin{figure}[ht]
+	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/detinova_prior_predictive.pdf}}
+	\caption{\textbf{The prior predictive distribution of lifespan for the analysis of Detinova parity data in terms of (left) physiological age and (right) chronological age.} Here, we used the gonotrophic cycle duration estimated for \textit{Anopheles}, $G\sim \mathcal{N}(3.65, 0.40)$, to convert physiological survival to calendar days.
+The figure shows 100,000 samples from the prior predictive distribution.
+The mean lifespan in terms of gonotrophic cycles is approximately 3; in terms of chronological age, it is approximately 10.}\label{fig:detinova_priors}
+\end{figure}
 
@@
 For the continent-level and overall estimates, we used all data.
-In quite a few cases, a species was indicated as a complex (for example, `Anopheles gambiae complex'), presumably when insufficient information was available in the paper to identify the species.
+In a number of cases, a species was indicated as a complex (for example, `Anopheles gambiae complex'), presumably when insufficient information was available in the paper to identify the species.
 These data were dropped for the species-level analysis but were included for all other analyses.
@@
 The statistical models were fit to data using Stan's NUTS algorithm \cite{carpenter2016stan}.
-In all cases, models were run for 2000 iterations on each of 4 chains, with the first half of the iterations discarded as warm-up.
+In all cases, models were run for 1000 iterations on each of 12 chains, with the first half of the iterations discarded as warm-up.
 We monitored diagnosed convergence as per \cite{gelman1992inference}, with $\hat{R}<1.1$ on all parameters.
-In all model runs, there were fewer than a handful of divergent transitions out of 2000 total iterations.
+In all model runs, there were fewer than a handful of divergent transitions.
 

### 2020-01-12 writeUp/supplementary_materials_methods.tex
added info on max age class

@@
 where $p(y(a)|S(a),\Psi,\kappa)$ corresponds to the negative binomial probability mass function for a count of $y(a)$ mosquitoes aged $a$ as specified in eqn.
-(\ref{eq:dissertation_negativeBinomial}), and $R$ is the number of separate physiological age classes recorded.
+(\ref{eq:dissertation_negativeBinomial}), and $R$ is the number of separate physiological age classes that we include: for series with no censoring (see below for how censored data were handled), we include 14 separate age classes (from nulliparous to 13-parous).
+We recognise that the choice of a "maximum" age class is an approximation, since, in theory, we should include a contribution from each "0" observed until $\infty$.
+In reality, however, the limit was chosen to match the maximum number of cycles observed in the data, and the majority of series had far maxima that were less than 5.
+As such, we believe that our truncation should provide a good approximation to the infinite sum.
 %

### 2020-01-27 writeUp/figures.tex
Rejigged layout of repo and added to-do
updated results in manuscript

--- 2018-11-28 figures.tex
@@
 	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_lifetimes_final_female_without_sugar_not_blood.pdf}}
-	\caption{\textbf{Posterior estimates of female mosquito mean lifespan across species, genus and overall groupings as determined from the MRR data.} The lifespans shown are for mosquitoes that were not fed with sugar or blood before release.
+	\caption{\textbf{Estimates of female mosquito mean lifespan across species, genus and overall groupings as determined from the MRR data.} The lifespans shown are for mosquitoes that were not fed with sugar or blood before release.
 The middle line in each box shows the median estimates and the solid dot indicates the mean.
@@
 	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_sexDifferences_without_sugar_nor_blood.pdf}}
-	\caption{\textbf{Posterior estimates of female and male mosquito mean lifespan across species, genus and overall groupings as determined from the MRR data.} The lifespans shown are for mosquitoes that were not fed with sugar or blood (for females) before release.
+	\caption{\textbf{Estimates of female and male mosquito mean lifespan across species, genus and overall groupings as determined from the MRR data.} The lifespans shown are for mosquitoes that were not fed with sugar or blood (for females) before release.
 The middle line in each box shows the median estimates and the solid dot indicates the mean.
@@
 	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/dissection_lifetimes_exponential.pdf}}
-	\caption{\textbf{Posterior estimates of female mosquito mean physiological lifespan across species, genus and overall groupings as determined from the dissection data.} The lifespans shown are for mosquitoes that were not fed with sugar or blood (for females) before release.
+	\caption{\textbf{Estimates of female mosquito mean physiological lifespan across species, genus and overall groupings as determined from the Polovodova dissection data.} The lifespans shown are for mosquitoes that were not fed with sugar or blood (for females) before release.
 The middle line in each box shows the median estimates and the solid dot indicates the mean.
@@
 	\label{fig:dissection_lifetimes_exponential}
+\end{figure}
+
+\begin{figure}[h]
+	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/detinova_lifespans_combined_parity.pdf}}
+	\caption{\textbf{Estimates of the percentage of female anophelines completing at least one reproductive cycle across species and morphospecies and, by continent, as determined from the Detinova dissection data.} The three coloured backgrounds show data according to continent where experiment was carried out (Americas-blue, Asia-grey, Africa-black).
+The middle line in each box shows the median estimates and the solid dot indicates the mean.
+The left and right box edges show the 25\%, and 75\% posterior quantiles respectively.
+The whiskers show the range of the data, excluding points lying more than 1.5 times the interquartile range away from each edge of the box.
+The numbers before the start of the left whisker indicate the number of individual time-series within each species.}
+	\label{fig:detinova_lifetimes}
 \end{figure}

### 2020-01-28 writeUp/supplementary_materials_methods.tex
corrected detinova parity calculation in SOM

@@
 \cite{detinova1962age} provided an alternative approach to estimate the lifespan of individual specimens, based on sampling the parous rate in a population (the proportion of adult females that have laid eggs) using dissection to determine parity of each specimen.
-In what follows, we follow the derivation given by \cite{davidson1954estimation}, to determine an expression for the proportion of nulliparous mosquitoes in terms of the gonotrophic cycle length $G$, here assumed the same for each cycle subsequent to an adult emerging.
-If a constant daily survival rate $p$ is assumed, the proportion of emerging adults surviving to become parous is $p^G$.
-If the rate of recruitment into the adult classes is constant over time, the proportion of nulliparous mosquitoes is,
-%
-\begin{equation}\label{eq:detinova}
-\begin{aligned}
-\mathbb{n} &= \frac{p^G}{p^G + p^{2G} + p^{3G} + \cdots}\\
-&= p^G / \frac{p^G}{1-p^G}\\
-&= 1-p^G.
-\end{aligned}
+In what follows, we follow the derivation given by \cite{davidson1954estimation}, to determine an expression for the mean lifespan of a mosquito in terms of gonotrophic cycles (all assumed to be of the same duration, $G$).
+If the probability of dying in a given cycle, $p$, is assumed constant (i.e.
+there is no age-dependent mortality), then the probability a mosquito completes $Y$ cycles, is geometrically distributed,
+%
+\begin{equation}
+Pr(Y=k) = (1-p)^k p,
 \end{equation}
 %
-The daily survival rate can hence be calculated by inverting eq.
-(\ref{eq:detinova}): $p = \sqrt[G]{\mathcal{P}}$, where $\mathcal{P}=1-\mathbb{n}$ is the proportion parous sampled from the population.
-The distribution of mosquitoes dying each day follows a geometric distribution, with probability of dying on a given day equal to $1-p$.
-The mean lifespan is, hence, given by $1/(1-p)$.
+where $k= 0,1,2,...$.
+This implies that the proportion of adults surviving to become parous is $Pr(Y\geq 1) = 1 - Pr(Y=0) = 1-p=\mathcal{P}$.
+Therefore, the proportion of parous mosquitoes in any given sample $i$ of wild-caught mosquitoes, $\mathcal{P}^{\{i\}}$ is a direct estimator of the probability of surviving a given cycle, $1-p$.
+The mean lifespan of a mosquito in terms of completed gonotrophic cycles is then given by $(1-p)/p= \mathcal{P}/(1-\mathcal{P})$.
+
+If, instead, we choose chronological time as our measure of lifespan, the above analysis still holds, although with $q$ now representing the probability of dying on a particular day, meaning $(1-q)^G$ is the probability of surviving a cycle.
+The daily survival probability $1-q$ is, hence, equal to $\sqrt[G]{\mathcal{P}}$, and the mean lifespan is given by $q/(1-q)=\sqrt[G]{\mathcal{P}}/(1-\sqrt[G]{\mathcal{P}})$.
 

### 2020-01-29 writeUp/supplementary_materials_methods.tex
updated ms with latest results

@@
     survival[i] = theta_average[i] ^ (1.0 / gon);
-    lifespan[i] = 1.0 / (1.0 - survival[i]); 
+    lifespan[i] = survival[i] / (1.0 - survival[i]); 
   }

### 2020-01-30 writeUp/supplementary_materials_methods.tex
updated POA

@@
 For the species-level analyses, there were, hence, $n=223$ observations for Africa across 4 species; $n=141$ observations for the Americas across 6 species; and $n=214$ observations for Asia across 8 species.
-For the complex-level analyses, there were, hence, $n=757$ observations for Africa across 3 groups; $n=77$ observations for the Americas across 3 groups; and $n=455$ observations for Asia across 11 groups.
+For the complex-level analyses, there were, hence, $n=778$ observations for Africa across 4 groups; $n=77$ observations for the Americas across 3 groups; and $n=455$ observations for Asia across 11 groups.
 For the continent-level and overall analyses, there were $n=1490$ observations across Africa ($n=778$), the Americas ($n=163$) and Asia ($n=549$).

### 2020-05-25 writeUp/figures.tex
version sent to Charles

--- 2018-11-28 figures.tex
@@
 	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_individualEstimates_allSpecies_withoutBalabacensis.pdf}}
-	\caption{\textbf{Individual time-series estimates of adult mosquito mean lifespan from the MRR analysis ordered by species.} The middle point shows the median estimates, and the left and right box whiskers show the 25\%, and 75\% posterior quantiles respectively.
+	\caption{\textbf{MRR: lifespan estimates for each time series.} The middle point shows the median estimates, and the left and right box whiskers show the 25\%, and 75\% posterior quantiles respectively.
 All estimates were obtained using the non-hierarchical exponential survival model as described in text.}
@@
 	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_lifetimes_final_female_without_sugar_not_blood.pdf}}
-	\caption{\textbf{Posterior estimates of female mosquito mean lifespan across species, genus and overall groupings as determined from the MRR data.} The lifespans shown are for mosquitoes that were not fed with sugar or blood before release.
+	\caption{\textbf{MRR: grouped lifespan estimates.} The lifespans shown are for mosquitoes that were not fed with sugar or blood before release.
 The middle line in each box shows the median estimates and the solid dot indicates the mean.
@@
 	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_sexDifferences_without_sugar_nor_blood.pdf}}
-	\caption{\textbf{Posterior estimates of female and male mosquito mean lifespan across species, genus and overall groupings as determined from the MRR data.} The lifespans shown are for mosquitoes that were not fed with sugar or blood (for females) before release.
+	\caption{\textbf{MRR: lifespan estimates by sex.} The lifespans shown are for mosquitoes that were not fed with sugar or blood (for females) before release.
 The middle line in each box shows the median estimates and the solid dot indicates the mean.
@@
 	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/dissection_lifetimes_exponential.pdf}}
-	\caption{\textbf{Posterior estimates of female mosquito mean physiological lifespan across species, genus and overall groupings as determined from the dissection data.} The lifespans shown are for mosquitoes that were not fed with sugar or blood (for females) before release.
+	\caption{\textbf{Polovodova dissection: grouped physiological lifespan estimates.} The lifespans shown are for mosquitoes that were not fed with sugar or blood (for females) before release.
 The middle line in each box shows the median estimates and the solid dot indicates the mean.
@@
 \begin{figure}[h]
-	\centerline{\includegraphics[width=1.3\textwidth]{./Figure_files/comparison.pdf}}
-	\caption{\textbf{Comparison of estimates of female mosquito lifespan between MRR and dissection data.} For the MRR data, we show estimates for females that were neither fed with sugar or blood pre-release.
-The red points indicate the posterior mean estimates and the whiskers show the 25\% and 75\% quantiles.
-In both cases, the estimates were produced using the exponential survival model.}
+	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/detinova_lifespans_combined_parity_no_insecticide.pdf}}
+	\caption{\textbf{Detinova dissection: grouped lifespan estimates for anophelines.} As described in SOM, both species and morphospecies estimates are provided where available.
+The three coloured backgrounds show data according to continent where experiment was carried out (Americas-blue, Asia-grey, Africa-black).
+The middle line in each box shows the median estimates and the solid dot indicates the mean.
+The left and right box edges show the 25\%, and 75\% posterior quantiles respectively.
+The whiskers show the range of the data, excluding points lying more than 1.5 times the interquartile range away from each edge of the box.
+The numbers before the start of the left whisker indicate the number of individual time-series within each species.}
+	\label{fig:detinova_lifetimes}
+\end{figure}
+
+\begin{figure}[h]
+	\centerline{\includegraphics[width=0.7\textwidth]{./Figure_files/gonotrophic_cycle_durations.pdf}}
+	\caption{\textbf{Estimates of gonotrophic cycle duration.} Coloured lines represent different genera, with each line indicating the estimated Gaussian probability density representing uncertainty over the reproductive cycle length.}
+	\label{fig:gonotrophic}
+\end{figure}
+
+\begin{figure}[h]
+	\centerline{\includegraphics[width=1.3\textwidth]{./Figure_files/pairwise_comparison.pdf}}
+	\caption{\textbf{Comparing lifespan estimates.} In A, B and C, we show pairwise estimates of lifespan for those species available in both datasets.
+For the MRR data, we show estimates for females that were neither fed with sugar or blood pre-release.
+Points indicate the posterior mean estimates and the whiskers show 25\% and 75\% quantiles.
+The values of $\rho$ indicate Pearson correlations (in no cases were these significant) and $n$ indicates sample size.}
 	\label{fig:comparison}
@@
 \begin{figure}[h]
+	\centerline{\includegraphics[width=1.3\textwidth]{./Figure_files/detinova_gambiae_daynight_range_lifespan_country.pdf}}
+	\caption{\textbf{Day-night temperature range versus lifespan estimates for \textit{Anopheles gambiae s.l.}.} Points indicate the posterior mean estimates and the whiskers show 25\% and 75\% quantiles.
+Each blue line indicates a posterior estimate of the relationship between lifespan and temperature range; the orange line indicates the posterior median.}
+	\label{fig:detinova_gambiae_daynight_range_lifespan_country}
+\end{figure}
+
+\begin{figure}[h]
+	\centerline{\includegraphics[width=1.3\textwidth]{./Figure_files/eip_all_combined.pdf}}
+	\caption{\textbf{Proportions of mosquitoes surviving EIPs of malaria, dengue, chikungunya and Zika.} Colours indicate the measures for the different diseases; shapes indicate the dataset used to produce the estimate.
+We assumed that the EIPs were: malaria-10 days, chikungunya-2 days, dengue fever-6.5 days and Zika-12.5 days.
+The left and right whiskers indicate the 25\% and 75\% posterior quantiles, and the points indicate the mean.
+All estimates were obtained using the exponential survival model.}
+	\label{fig:eip}
+\end{figure}
+
+\begin{figure}[h]
 	\centerline{\includegraphics[width=1.3\textwidth]{./Figure_files/mrr_elpd_vs_exponential_ordered.pdf}}
-	\caption{\textbf{Predictive accuracy of age-dependent models of mortality versus the exponential model by species for the MRR analysis.} The predictive accuracy was determined using K-Fold cross-validation as described in SOM; the measure of accuracy presented here by the central dots is the difference in estimated expected log pointwise predictive density compared to the exponential model.
+	\caption{\textbf{Detecting senescence in MRR studies.} Here, we consider the predictive accuracy of age-dependent models of mortality versus the exponential model by species.
+The predictive accuracy was determined using K-Fold cross-validation as described in SOM; the measure of accuracy presented here by the central dots is the difference in estimated expected log pointwise predictive density compared to the exponential model.
 The lower and upper whiskers represent the lower and upper bounds of the 95\% confidence interval in predictive accuracy.
@@
 
-\begin{figure}[h]
-	\centerline{\includegraphics[width=1.3\textwidth]{./Figure_files/dissection_elpd_vs_exponential_ordered.pdf}}
-	\caption{\textbf{Predictive accuracy of age-dependent models of mortality versus the exponential model by species for the dissection analysis.} The predictive accuracy was determined using K-Fold cross-validation as described in SOM; the measure of accuracy presented here by the central dots is the difference in estimated expected log pointwise predictive density compared to the exponential model.
-The lower and upper whiskers represent the lower and upper bounds of the 95\% confidence interval in predictive accuracy.
-The species have been ordered so that those with the highest average predictive accuracy across all age-dependent models compared to the exponential model appear at the top.
-If all age-dependent models outperform the exponential, we deem this as evidence for age-dependent mortality (`+'); if a subsample of models perform better, we deem the evidence ambiguous (`?'); and if no age-dependent model exceeds the performance of the exponential, we conclude no evidence in favour of senescence (`-').}
-	\label{fig:dissection_elpd}
-\end{figure}
-
-\begin{figure}[h]
-	\centerline{\includegraphics[width=1.3\textwidth]{./Figure_files/eip_all_combined.pdf}}
-	\caption{\textbf{The MRR and dissection analyses estimates of the proportion of mosquitoes surviving the EIP of malaria, chikungunya, dengue fever and Zika for the disease vectors in our sample.} We assumed that the EIPs were: malaria-10 days, chikungunya-2 days, dengue fever-6.5 days and Zika-12.5 days.
-The left and right whiskers indicate the 25\% and 75\% posterior quantiles, and the points indicate the mean.
-All estimates were obtained using the exponential survival model.}
-	\label{fig:eip}
-\end{figure}
-
 \end{document}

### 2020-11-07 writeUp/supplementary_materials_methods.tex
started reworking

@@
 
-\subsection{Data and model}\label{sec:detinova_datamodel}
-\cite{massey2016global} provide a global database of bionomic quantities for the dominant vector species of malaria.
-This database contains the number of parous and nulliparous samples in studies where Detinova's dissection approach was applied to field caught speciments ($n=1490$ non-missing observations for total number of parous individuals) across Africa ($n=778$), Asia ($n=549$) and the Americas ($n=163$).
-
+\subsection{Data processing}
+In this section, we describe the data processing carried out on the \cite{massey2016global} dataset, which provides a global database of bionomic quantities for the dominant vector species of malaria.
+Included in this database is the number of parous and nulliparous samples in studies where Detinova's dissection approach was applied to field caught specimens.
+
+For each of the three datasets (Africa, Americas and Asia) contained with the SOM of \cite{massey2016global}, we used only those observations which had a non-missing and positive total capture size.
+For each observation, we calculated observed parity ($=\text{parous count}/ \text{total capture size}$) and compared this quantity with the `parity percent' column in the datasets.
+If the difference between these two quantities was lower than 1\%, we considered the discrepancy to be due to numerical rounding and did not investigate further.
+In those cases where the difference exceeded 1\%, we did the following:
+
+\begin{itemize}
+	\item Africa: considering the database entry for \cite{ijumba2002malaria}, there was a difference of about 50\%.
+Reading the paper, it looks like there is a mistake in their Table 5, and that the parous number should be 131 and, as such, we changed this number in the database; considering the entry for \cite{bockarie46service}, there was a difference of around 5\%.
+We could not access the paper here, so assumed that the issue was with the `parity percent' entry.
+	\item Americas: no differences exceeding 1\%.
+	\item Asia: considering the entry for \cite{kitthawee1992relationship}, there was a difference of around 3\%.
+Reading the paper, it looks like there was a (small) error in the database, and the parous number should be 35 (rather than 36) - this adjustment was made for the ensuing analysis.
+\end{itemize}
+
+For the species- and complex-level estimates of lifespan, we considered only those groups with five or more parity experiments.
+For the continent-level and overall estimates, we used all data.
+In a number of cases, a species was indicated as a complex (for example, `Anopheles gambiae complex'), presumably when insufficient information was available in the paper to identify the species.
+These data were dropped for the species-level analysis but were included for all other analyses.
+
+The \cite{massey2016global} database included a variable for each observation determining whether insecticide-based controls were in place at the specified location and time period of the study.
+For $n=883$ observations, this information was recorded, allowing us to estimate the impact of these interventions on lifespan (see \ref{sec:detinova_insecticide}).
+This analysis determined that insecticidal interventions reduced lifespan, and, thus, studies where insecticides were in place ($n=364$) were removed from our main analysis.
+
+For the species-level analyses, there were, hence, $n=198$ observations for Africa across 4 species; $n=107$ observations for the Americas across 6 species; and $n=108$ observations for Asia across 4 species.
+For the complex-level analyses, there were, hence, $n=679$ observations for Africa across 4 groups; $n=121$ observations for the Americas across 5 groups; and $n=318$ observations for Asia across 13 groups.
+For the continent-level and overall analyses, there were $n=1126$ observations across Africa ($n=679$), the Americas ($n=126$) and Asia ($n=321$).
+
+\subsection{Models}\label{sec:detinova_datamodel}
 We estimated lifespan at the species and complex / subgroup level (hereafter denoted as \textit{s.l.}) using a Bayesian hierarchical framework.
@@
 
-\subsection{Data processing}
-In this section, we describe the data processing carried out on the \cite{massey2016global} dataset.
-For each of the three datasets (Africa, Americas and Asia) contained with the SOM of \cite{massey2016global}, we used only those observations which had a non-missing and positive total capture size.
-For each observation, we calculated observed parity ($=\text{parous count}/ \text{total capture size}$) and compared this quantity with the `parity percent' column in the datasets.
-If the difference between these two quantities was lower than 1\%, we considered the discrepancy to be due to numerical rounding and did not investigate further.
-In those cases where the difference exceeded 1\%, we did the following:
-
-\begin{itemize}
-	\item Africa: considering the database entry for \cite{ijumba2002malaria}, there was a difference of about 50\%.
-Reading the paper, it looks like there is a mistake in their Table 5, and that the parous number should be 131 and, as such, we changed this number in the database; considering the entry for \cite{bockarie46service}, there was a difference of around 5\%.
-We could not access the paper here, so assumed that the issue was with the `parity percent' entry.
-	\item Americas: no differences exceeding 1\%.
-	\item Asia: considering the entry for \cite{kitthawee1992relationship}, there was a difference of around 3\%.
-Reading the paper, it looks like there was a (small) error in the database, and the parous number should be 35 (rather than 36) - this adjustment was made for the ensuing analysis.
-\end{itemize}
-
-For the species- and complex-level estimates of lifespan, we considered only those groups with five or more parity experiments.
-For the continent-level and overall estimates, we used all data.
-In a number of cases, a species was indicated as a complex (for example, `Anopheles gambiae complex'), presumably when insufficient information was available in the paper to identify the species.
-These data were dropped for the species-level analysis but were included for all other analyses.
-
-For the species-level analyses, there were, hence, $n=223$ observations for Africa across 4 species; $n=141$ observations for the Americas across 6 species; and $n=214$ observations for Asia across 8 species.
-For the complex-level analyses, there were, hence, $n=778$ observations for Africa across 4 groups; $n=77$ observations for the Americas across 3 groups; and $n=455$ observations for Asia across 11 groups.
-For the continent-level and overall analyses, there were $n=1490$ observations across Africa ($n=778$), the Americas ($n=163$) and Asia ($n=549$).
+\subsubsection{Impact of insecticide interventions on lifespan}\label{sec:detinova_insecticide}
+To determine the impact of insecticidal interventions on lifespan, we modified the model described in \S\ref{sec:detinova_datamodel} to estimate effects at the species level.
+To do so, we allowed a logit-linear impact of insecticide.
+Further, the species-level effects were modelled hierarchically using normal priors, allowing the impact of resistance to vary at the observation level within each study.
+The model to estimate this is shown below (where the variable metering the impact of insecticide at the observational level is `delta').
+
+\begin{minted}{stan}
+data{
+  int N;
+  int X[N];
+  int num[N];
+  int species[N];
+  int K;
+  real z[N];
+}
+
+parameters{
+  vector<lower=0, upper=1>[N] theta;
+  vector<lower=0, upper=1>[K] phi;
+  vector<lower=5>[K] kappa;
+  vector[K] delta_raw;
+  real delta_top;
+  real<lower=0> delta_sigma;
+}
+
+transformed parameters{
+  vector[K] delta;
+  for(i in 1:K)
+    delta[i] = delta_top + delta_sigma * delta_raw[i];
+}
+
+model{
+  real theta_eff;
+  for(i in 1:N){
+    theta_eff = logit(theta[i]) + delta[species[i]] * z[i];
+    X[i] ~ binomial_logit(num[i], theta_eff);
+    theta[i] ~ beta(phi[species[i]] * kappa[species[i]],
+                   (1 - phi[species[i]]) * kappa[species[i]]);
+  }
+  kappa ~ pareto(5, 1.5);
+  phi ~ beta(20, 15);
+  delta_raw ~ normal(0, 1);
+  delta_top ~ normal(0, 1);
+  delta_sigma ~ normal(0, 1);
+}
+\end{minted}
+
+\begin{minted}{stan}
+generated quantities{
+  real lifespan_without_insecticide[K];
+  real lifespan_with_insecticide[K];
+  for(i in 1:K){
+    real theta_eff;
+    real theta_eff_without;
+    real gon = normal_rng(3.65, 0.4);
+    real survival;
+    real survival_without;
+    theta_eff = logit(theta[i]) + delta[species[i]];
+    theta_eff_without = logit(theta[i]);
+    survival = inv_logit(theta_eff)^(1.0/gon);
+    survival_without = inv_logit(theta_eff_without)^(1.0/gon);
+    lifespan_with_insecticide[i] = survival / (1.0 - survival);
+    lifespan_without_insecticide[i] = survival_without /
+                                    (1.0 - survival_without);
+  }
+}
+\end{minted}
 

### 2020-11-14 writeUp/supplementary_materials_methods.tex
updated figures and added analysis navigator

@@
 
-For each individual series, we recorded the following meta-data: genus, species, collection method, whether or not insecticide was mentioned as being used during or shortly before the time of mosquito collections, and the start and end dates of the experiment.
-At the article level, we recorded the following meta-data: author, year, title, country, location (within country), start and end date, whether insecticide was used during any of the experimental replicates, whether an MRR experiment occurred alongside the dissection study, and the collection location (indoor or outdoor).
+For each individual series, we recorded the following meta-data: genus, species, collection method and the start and end dates of the experiment.
+At the article level, we recorded the following meta-data: author, year, title, country, location (within country), start and end date, whether an MRR experiment occurred alongside the dissection study, and the collection location (indoor or outdoor).
 At either the individual series or article levels, we recorded additional meta-data describing the nature of data collection, for example explaining where the data was contained within the article, whether it was obtained by digitising graphs, and the number of separate dated series.
@@
 The statistical models were fit to data using Stan's NUTS algorithm \cite{carpenter2016stan}.
-In all cases, models were run for 1000 iterations on each of 12 chains, with the first half of the iterations discarded as warm-up.
+In all cases, models were run for 4000 iterations on each of 12 chains, with the first half of the iterations discarded as warm-up.
 We monitored diagnosed convergence as per \cite{gelman1992inference}, with $\hat{R}<1.1$ on all parameters.
@@
 This reduced the number of data points used for model comparison versus those used for model estimation.
-For Africa, we had $n=223$ data points; for the Americas, there were $n=141$; and for Asia, there were $n=214$.
+For Africa, we had $n=217$ data points; for the Americas, there were $n=107$; and for Asia, there were $n=108$.
 
 For the comparison between continent-level grouping and overall, we were able to use more data points since we were less constrained by the minimal numbers of data belonging to each group.
-This meant we had $n=1490$ data points, with $n=778$ for Africa, $n=163$ for the Americas, and $n=549$, for Asia.
+This meant we had $n=1126$ data points, with $n=679$ for Africa, $n=126$ for the Americas, and $n=321$, for Asia.
 

### 2020-11-21 writeUp/figures.tex
updated graph titles

@@
 	\centerline{\includegraphics[width=1.3\textwidth]{./Figure_files/mrr_elpd_vs_exponential_ordered.pdf}}
-	\caption{\textbf{Detecting senescence in MRR studies.} Here, we consider the predictive accuracy of age-dependent models of mortality versus the exponential model by species.
+	\caption{\textbf{Evidence for senescence: MRR studies.} Here, we consider the predictive accuracy of age-dependent models of mortality versus the exponential model by species.
 The predictive accuracy was determined using K-Fold cross-validation as described in SOM; the measure of accuracy presented here by the central dots is the difference in estimated expected log pointwise predictive density compared to the exponential model.

### 2020-11-21 writeUp/supplementary_materials.tex
added new ms file
updated dissection estimates with new gonotrophic estimates
updated SOM and MS to include higher kappa estimates
updated graph titles

@@
 \section{MRR experments}
-Figures \ref{fig:mrr_lifeSpanVsRange} and \ref{fig:mrr_lifeSpanVsTrapDensity} show the estimates of lifespan versus the trapping range and spatial density of traps at study location.
-In these cases, the lifespans shown were obtained from the non-hierarchical model with the exponential survival model.
-The spatial density of traps was obtained assuming that the study area was a circular area of radius equal to the maximum trapping distance.
-
-Fig.
-\ref{fig:mrr_female_blood_sugar} shows the lifespans obtained for each pre-release feeding category for each genus and across all studies using the hierarchical Bayesian model with exponential survival model.
-In the genus-level models, the feeding-specific effects (not pre-fed, blood-fed-only and sugar-fed-only) were estimated as independent random effects for each genus with standard normal priors.
-For the overall model, the feeding-specific effects were estimated as independent random effects across all studies, with a standard normal prior.
 
@@
 	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_genus_vs_species_vs_top.pdf}}
-	\caption{\textbf{The estimated predictive accuracy of the species- and overall-level models relative to the genus-level model.} The predictive accuracy was determined using K-Fold cross-validation as described in SOM; the measure of accuracy presented here by the central dots is the difference in estimated expected log pointwise predictive density compared to the exponential model.
+	\caption{\textbf{MRR: predictive accuracy of the species- and overall-level models relative to the genus-level model.} The predictive accuracy was determined using K-Fold cross-validation as described in SOM; the measure of accuracy presented here by the central dots is the difference in estimated expected log pointwise predictive density compared to the exponential model.
 The lower and upper whiskers represent the lower and upper bounds of the 95\% confidence interval in predictive accuracy.}
@@
 	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_lifeSpanVsRange.pdf}}
-	\caption{\textbf{Posterior estimates of mean mosquito lifespan for each time-series versus the trapping range.} The markers show the median posterior estimates, with the lower and upper bounds indicating the 25\% and 75\% quantiles, respectively.
+	\caption{\textbf{MRR: mean mosquito lifespan versus trapping range.} The markers show the median posterior lifespan estimates for each time series with the lower and upper bounds indicating the 25\% and 75\% quantiles, respectively.
 The black line shows a linear regression line estimated using the median posterior lifetimes, with the grey shading indicating 95\% confidence intervals.
@@
 	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_lifespanVsTrapDensity.pdf}}
-	\caption{\textbf{Posterior estimates of mean mosquito lifespan for each time-series versus the trap density.} The markers show the median posterior estimates, with the lower and upper bounds indicating the 25\% and 75\% quantiles, respectively.
+	\caption{\textbf{MRR: mean mosquito lifespan versus trapping density.} The markers show the median posterior lifespan estimates for each time series with the lower and upper bounds indicating the 25\% and 75\% quantiles, respectively.
 The trap density was estimated assuming traps were contained within a circular area with radius equal to the trapping range.
@@
 	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_female_blood_sugar.pdf}}
-	\caption{\textbf{Posterior estimates of female mosquito mean lifespan according to pre-release feeding status across species, genus and overall groupings as determined from the MRR data.} The lifespans shown are for mosquitoes that were not fed with sugar or blood (for females) before release.
+	\caption{\textbf{MRR: female mosquito mean lifespan by pre-release feeding status.} The lifespans shown are for mosquitoes that were not fed with sugar or blood (for females) before release.
 The middle line in each box shows the median estimates and the solid dot indicates the mean.
@@
 	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_lifeSpanVsTemperature.pdf}}
-	\caption{\textbf{Posterior estimates of mean mosquito lifespan for each time-series versus the average monthly temperature for each study location and date.} The markers show the median posterior estimates, with the lower and upper bounds indicating the 25\% and 75\% quantiles, respectively.
+	\caption{\textbf{MRR: mean mosquito lifespan versus average monthly air temperature.} The markers show the median posterior lifespan estimates for each time series with the lower and upper bounds indicating the 25\% and 75\% quantiles, respectively.
 The black line shows a regression line with linear and quadratic terms estimated using the median posterior lifetimes, with the grey shading indicating 95\% confidence intervals.
@@
 	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_ThreeSpeciesVersusTemperature.pdf}}
-	\caption{\textbf{Posterior estimates of mean mosquito lifespan for each time-series versus the average monthly temperature for the four species with the most data.} The markers show the median posterior estimates, with the lower and upper bounds indicating the 25\% and 75\% quantiles, respectively.
+	\caption{\textbf{MRR: mean mosquito lifespan versus average monthly air temperature by species.} Here, we include only those four species with the most data.
+The markers show the median posterior lifespan estimates for each time series with the lower and upper bounds indicating the 25\% and 75\% quantiles, respectively.
 The black line shows a regression line with linear and quadratic terms estimated using the median posterior lifetimes, with the grey shading indicating 95\% confidence intervals.
@@
 
-\section{Dissection analysis}
+\section{Polovodova dissection analysis}
 
@@
 	\centerline{\includegraphics[width=1.3\textwidth]{./Figure_files/dissection_individualEstimates_allSpecies.pdf}}
-	\caption{\textbf{Individual time-series estimates of adult mosquito mean lifespan ordered by species median as determined from the dissection data.} The middle line in each box shows the median estimates.
+	\caption{\textbf{Polovodova dissection: individual time-series estimates of adult mosquito mean lifespan.} The middle line in each box shows the median estimates.
 The left and right box whiskers show the 25\%, and 75\% posterior quantiles respectively.
@@
 	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/dissection_lifetimes_exponential_chron.pdf}}
-	\caption{\textbf{Posterior estimates of female mosquito mean chronological lifespan across species, genus and overall groupings as determined from the dissection data.} The lifespans shown are for mosquitoes that were not fed with sugar or blood (for females) before release.
+	\caption{\textbf{Polovodova dissection: female mosquito mean chronological lifespan estimates.} The lifespans shown are for mosquitoes that were not fed with sugar or blood (for females) before release.
 The middle line in each box shows the median estimates and the solid dot indicates the mean.
@@
 		\hline
-		\textit{Anopheles} & \textit{sergentii} & 4.2 & 7.6 & 10.1 & 13.6 & 24.6 & 11.9 & 9.7 \\
-		\textit{Anopheles} & \textit{gambiae s.l.} & 2.8 & 5.2 & 7.6 & 11.4 & 21.6 & 9.5 & 7.4 \\
-		\textit{Culex} & \textit{thalassius} & 4.6 & 6.3 & 7.4 & 8.6 & 12.0 & 7.9 & 4.1 \\
-		\textit{Anopheles} & \textit{farauti s.l.} & 5.9 & 6.5 & 6.9 & 7.3 & 8.0 & 6.9 & 0.7 \\
-		\textit{Anopheles} & \textit{minimus} & 3.1 & 5.2 & 6.9 & 9.0 & 16.4 & 8.1 & 9.4 \\
-		\textit{Anopheles} & \textit{maculipennis} & 3.3 & 4.8 & 5.9 & 7.5 & 11.4 & 6.5 & 2.7 \\
-		\textit{Culex} & \textit{pipiens} & 4.6 & 5.3 & 5.7 & 6.2 & 6.9 & 5.8 & 0.8 \\
-		\textit{Mansonia} & \textit{uniformis} & 3.8 & 4.4 & 4.8 & 5.2 & 6.0 & 4.8 & 0.8 \\
-		\textit{Anopheles} & \textit{rivulorum} & 3.4 & 4.0 & 4.5 & 4.9 & 5.9 & 4.5 & 0.8 \\
-		\textit{Anopheles} & \textit{melas} & 3.3 & 3.9 & 4.3 & 4.8 & 5.5 & 4.4 & 0.7 \\
-		\textit{Anopheles} & \textit{culicifacies} & 3.4 & 4.0 & 4.3 & 4.7 & 5.4 & 4.4 & 0.6 \\
-		\textit{Aedes} & \textit{aegypti} & 2.1 & 3.5 & 4.3 & 5.0 & 7.9 & 5.0 & 11.5 \\
-		\textit{Anopheles} & \textit{subpictus} & 2.6 & 3.6 & 4.3 & 5.1 & 8.3 & 5.0 & 7.7 \\
-		\textit{Anopheles} & \textit{quadrimaculatus} & 2.9 & 3.6 & 4.1 & 4.7 & 5.6 & 4.2 & 0.9 \\
-		\textit{Anopheles} & \textit{stephensi} & 2.4 & 3.4 & 4.1 & 4.9 & 6.9 & 4.3 & 1.8 \\
-		\textit{Culex} & \textit{quinquefasciatus} & 1.7 & 2.8 & 4.0 & 5.5 & 9.0 & 4.5 & 2.5 \\
-		\textit{Anopheles} & \textit{darlingi} & 2.7 & 3.5 & 3.9 & 4.5 & 5.5 & 4.0 & 1.0 \\
-		\textit{Aedes} & \textit{polynesiensis} & 3.1 & 3.5 & 3.8 & 4.1 & 4.5 & 3.8 & 0.5 \\
-		\textit{Culex} & \textit{tritaeniorhynchus} & 1.3 & 2.6 & 3.8 & 5.7 & 16.8 & 9.8 & 123.0 \\
-		\textit{Culex} & \textit{annulirostris} & 2.3 & 3.1 & 3.7 & 4.4 & 5.9 & 3.9 & 1.8 \\
-		\textit{Aedes} & \textit{samoanus} & 2.6 & 3.0 & 3.3 & 3.7 & 4.4 & 3.4 & 0.9 \\
-		\textit{Aedes} & \textit{sollicitans} & 1.5 & 2.2 & 2.7 & 3.3 & 5.4 & 3.0 & 2.0 \\
-		\textit{Aedes} & \textit{vexans} & 1.7 & 2.1 & 2.5 & 2.9 & 3.9 & 2.7 & 2.7 \\
-		\textit{Anopheles} & \textit{cruzii} & 1.4 & 1.9 & 2.3 & 2.9 & 4.6 & 2.6 & 1.5 \\
-		\textit{Anopheles} & \textit{bellator} & 1.2 & 1.7 & 2.0 & 2.5 & 4.3 & 2.5 & 7.2 \\
+	\textit{Culex} & \textit{thalassius} & 5.7 & 8.2 & 9.4 & 11.0 & 15.6 & 10.1 & 5.2 \\
+	\textit{Anopheles} & \textit{sergentii} & 3.7 & 6.4 & 8.5 & 11.5 & 20.6 & 10.1 & 8.1 \\
+	\textit{Culex} & \textit{pipiens} & 5.7 & 6.6 & 7.2 & 7.8 & 8.9 & 7.2 & 1.0 \\
+	\textit{Anopheles} & \textit{gambiae s.l.} & 2.4 & 4.3 & 6.4 & 9.8 & 18.3 & 8.0 & 6.3 \\
+	\textit{Anopheles} & \textit{farauti s.l.} & 4.8 & 5.5 & 5.9 & 6.3 & 6.9 & 5.9 & 0.6 \\
+	\textit{Anopheles} & \textit{minimus} & 2.7 & 4.4 & 5.8 & 7.7 & 13.7 & 6.9 & 8.1 \\
+	\textit{Anopheles} & \textit{maculipennis} & 2.8 & 4.0 & 5.1 & 6.4 & 9.6 & 5.5 & 2.3 \\
+	\textit{Culex} & \textit{quinquefasciatus} & 2.0 & 3.4 & 4.8 & 6.9 & 11.7 & 5.6 & 3.3 \\
+	\textit{Aedes} & \textit{aegypti} & 2.3 & 3.7 & 4.7 & 5.6 & 9.3 & 5.6 & 12.3 \\
+	\textit{Culex} & \textit{annulirostris} & 2.7 & 3.9 & 4.6 & 5.4 & 7.5 & 4.9 & 2.4 \\
+	\textit{Culex} & \textit{tritaeniorhynchus} & 1.6 & 3.2 & 4.6 & 7.2 & 22.8 & 13.0 & 174.0 \\
+	\textit{Mansonia} & \textit{uniformis} & 3.4 & 4.0 & 4.4 & 4.9 & 5.5 & 4.5 & 0.7 \\
+	\textit{Aedes} & \textit{polynesiensis} & 3.2 & 3.7 & 4.1 & 4.5 & 5.1 & 4.1 & 0.6 \\
+	\textit{Anopheles} & \textit{rivulorum} & 2.8 & 3.4 & 3.8 & 4.2 & 5.0 & 3.9 & 0.7 \\
+	\textit{Anopheles} & \textit{melas} & 2.8 & 3.3 & 3.7 & 4.1 & 4.8 & 3.8 & 0.7 \\
+	\textit{Anopheles} & \textit{culicifacies s.l.} & 2.8 & 3.3 & 3.7 & 4.1 & 4.7 & 3.7 & 0.6 \\
+	\textit{Anopheles} & \textit{subpictus s.l.} & 2.2 & 3.1 & 3.7 & 4.5 & 7.0 & 4.3 & 6.6 \\
+	\textit{Aedes} & \textit{samoanus} & 2.7 & 3.2 & 3.6 & 4.0 & 4.9 & 3.7 & 1.0 \\
+	\textit{Anopheles} & \textit{quadrimaculatus} & 2.4 & 3.1 & 3.5 & 4.0 & 4.8 & 3.6 & 0.8 \\
+	\textit{Anopheles} & \textit{stephensi} & 2.1 & 2.9 & 3.5 & 4.2 & 5.7 & 3.7 & 1.5 \\
+	\textit{Anopheles} & \textit{darlingi} & 2.3 & 2.9 & 3.4 & 3.8 & 4.7 & 3.4 & 0.9 \\
+	\textit{Aedes} & \textit{sollicitans} & 1.6 & 2.3 & 2.9 & 3.6 & 6.1 & 3.3 & 2.5 \\
+	\textit{Aedes} & \textit{vexans} & 1.7 & 2.2 & 2.7 & 3.1 & 4.2 & 3.0 & 3.3 \\
+	\textit{Anopheles} & \textit{cruzii} & 1.2 & 1.6 & 2.0 & 2.5 & 4.0 & 2.2 & 1.3 \\
+	\textit{Anopheles} & \textit{bellator} & 1.0 & 1.4 & 1.7 & 2.1 & 3.6 & 2.1 & 5.4 \\
 		\hline
-		\textit{Anopheles} & \textit{} & 2.5 & 4.1 & 5.7 & 7.9 & 12.7 & 6.4 & 3.4 \\
-		\textit{Mansonia} & \textit{} & 3.7 & 4.4 & 4.8 & 5.1 & 5.9 & 4.8 & 0.8 \\
-		\textit{Culex} & \textit{} & 2.1 & 3.2 & 4.4 & 5.9 & 9.5 & 4.9 & 2.4 \\
-		\textit{Aedes} & \textit{} & 2.4 & 3.1 & 3.5 & 3.9 & 4.7 & 3.5 & 0.7 \\
+		\textit{Culex} & \textit{} & 2.5 & 4.0 & 5.4 & 7.6 & 12.3 & 6.2 & 3.2 \\
+		\textit{Anopheles} & \textit{} & 2.1 & 3.6 & 4.9 & 6.8 & 10.8 & 5.5 & 2.9 \\
+		\textit{Mansonia} & \textit{} & 3.3 & 4.0 & 4.5 & 4.9 & 5.6 & 4.5 & 0.8 \\
+		\textit{Aedes} & \textit{} & 2.5 & 3.3 & 3.8 & 4.3 & 5.2 & 3.8 & 0.9 \\
 		\hline
-		\textbf{Overall} & \textit{} & 2.2 & 3.5 & 5.1 & 6.8 & 10.7 & 5.5 & 2.7 \\
+		\textit{Overall} & \textit{} & 2.0 & 3.3 & 4.7 & 6.4 & 10.2 & 5.2 & 2.6 \\
 	\end{tabular}
@@
 
-\subsection{Comparison of LBLs between the MRR and dissection analyses}
-To conduct a comparison between the LBLs from the MRR analysis and from the dissections, we converted the latter lifespans from gonotrophic cycles to chronological age as described in Section \ref{sec:dissection_conversion}.
-We conducted ANOVA using the individual time series estimates of LBL for each of the 10 species with estimates from both MRR- and dissection-based analysis.
-The results of this analysis are shown in Table \ref{tab:comparison}.
+\section{Detinova parity analysis}
 
-\begin{table}[htbp]
-	\centering
-	\begin{tabular}{l|l|l|l|l}
-		species & \multicolumn{1}{l}{$\text{df}_1$} & \multicolumn{1}{l}{$\text{df}_2$} & \multicolumn{1}{l}{F} & \multicolumn{1}{l}{p} \\
-		\midrule
-		\textit{Ae. aegypti} & 1     & 55    & 1.9   & 0.17 \\
-		\textit{A. culicifacies s.l.} & 1     & 19    & 2.2   & 0.15 \\
-		\textit{A. darlingi} & 1     & 6     & 80.7   & 0.00 \\
-		\textit{An. farauti s.l.} & 1     & 14    & 2.1  & 0.17 \\
-		\textit{A. gambiae s.l.} & 1     & 37    & 4.3   & 0.05 \\
-		\textit{Cx. pipiens} & 1     & 13    & 0.4   & 0.56 \\
-		\textit{A. quadrimaculatus} & 1     & 8     & 2.2     & 0.18 \\
-		\textit{Cx. quinquefasciatus} & 1     & 13    & 2.2    & 0.16 \\
-		\textit{A. sergentii} & 1     & 8     & 1.0    & 0.36 \\
-		\textit{A. subpictus} & 1     & 2    & 11.4   & 0.08 \\
-		\textit{Cx. tritaeniorhynchus} & 1     & 5     & 3.6   & 0.12 \\
-		\textit{Ae. vexans} & 1     & 2     & 1.1   & 0.41 \\
-	\end{tabular}%
-	\caption{\textbf{The results of comparisons between LBLs for those species with data for MRRs and dissection-based studies.} The ANOVA tests were conducted using median posterior LBLs estimated at the time-series level.}
-	\label{tab:comparison}%
-\end{table}%
+\begin{table}[htbp!]
+\begin{tabular}{l|l|l|l|l|l|l|l|l}
+	\textbf{Genus} & \textbf{Species} & \textbf{5$\%$} & \textbf{25$\%$} & \textbf{50$\%$} &
+	\textbf{75$\%$} & \textbf{95$\%$} & \textbf{Mean} & \textbf{Std. dev.} \\
+	\hline
+	\textit{Anopheles} & \textit{funestus} & 5.1 & 8.3 & 11.9 & 18.2 & 41.5 & 16.7 & 20.8 \\
+	\textit{Anopheles} & \textit{funestus s.l.} & 4.4 & 7.8 & 13.2 & 24.0 & 75.5 & 25.5 & 58.3
+	\\
+	\textit{Anopheles} & \textit{nili s.l.} & 5.6 & 8.9 & 12.4 & 17.7 & 36.1 & 16.0 & 17.0 \\
+	\textit{Anopheles} & \textit{gambiae} & 4.0 & 6.1 & 8.3 & 12.0 & 22.1 & 10.3 & 8.5 \\
+	\textit{Anopheles} & \textit{gambiae (Form M)} & 3.5 & 5.6 & 8.2 & 12.4 & 27.4 & 11.2 & 14.0
+	\\
+	\textit{Anopheles} & \textit{arabiensis} & 3.2 & 4.9 & 7.2 & 11.2 & 24.0 & 9.8 & 10.0 \\
+	\textit{Anopheles} & \textit{gambiae s.l.} & 3.4 & 5.7 & 8.8 & 14.4 & 35.7 & 13.3 & 19.7 \\
+	\textit{Anopheles} & \textit{leucosphyrus} & 5.0 & 8.1 & 11.0 & 15.4 & 29.0 & 13.8 & 26.4 \\
+	\textit{Anopheles} & \textit{leucosphyrus s.l.} & 5.0 & 7.9 & 11.0 & 15.8 & 29.7 & 14.1 &
+	25.3 \\
+	\textit{Anopheles} & \textit{aconitus} & 3.9 & 6.6 & 9.5 & 13.7 & 29.4 & 13.4 & 65.1 \\
+	\textit{Anopheles} & \textit{dirus (Sp. A)} & 4.9 & 7.4 & 10.3 & 14.9 & 30.2 & 13.1
+	& 10.7 \\
+	\textit{Anopheles} & \textit{dirus s.l.} & 4.3 & 6.6 & 9.4 & 14.3 & 28.8 & 12.3 & 10.5 \\
+	\textit{Anopheles} & \textit{minimus (Sp. A)} & 4.1 & 6.9 & 10.2 & 16.0 & 37.2 &
+	15.0 & 24.5 \\
+	\textit{Anopheles} & \textit{minimus s.l.} & 3.9 & 5.8 & 7.9 & 11.2 & 20.5 & 9.6 & 6.7 \\
+	\textit{Anopheles} & \textit{barbirostris s.l.} & 3.8 & 5.9 & 7.9 & 11.2 & 23.1 & 10.4 &
+	11.9 \\
+	\textit{Anopheles} & \textit{maculatus} & 3.2 & 5.0 & 7.0 & 10.5 & 21.7 & 9.2 & 8.8 \\
+	\textit{Anopheles} & \textit{maculatus s.l.} & 3.4 & 5.3 & 7.6 & 11.7 & 25.6 & 10.3 & 10.5
+	\\
+	\textit{Anopheles} & \textit{sundaicus s.l.} & 3.2 & 4.9 & 6.9 & 10.6 & 22.4 & 9.3 & 9.3 \\
+	\textit{Anopheles} & \textit{subpictus s.l.} & 2.9 & 4.6 & 6.8 & 11.0 & 25.6 & 10.0 & 13.6
+	\\
+	\textit{Anopheles} & \textit{anthropophagus (lesteri)} & 3.2 & 4.7 & 6.4 & 9.1 & 18.1 & 8.1
+	& 9.3 \\
+	\textit{Anopheles} & \textit{annularis s.l.} & 4.3 & 5.4 & 6.3 & 7.4 & 10.1 & 6.7 & 2.1 \\
+	\textit{Anopheles} & \textit{sinensis} & 3.5 & 4.8 & 6.2 & 8.2 & 13.1 & 7.0 & 3.6 \\
+	\textit{Anopheles} & \textit{fluviatilis s.l.} & 3.6 & 4.8 & 6.0 & 7.5 & 11.0 & 6.5 & 2.6 \\
+	\textit{Anopheles} & \textit{culicifacies s.l.} & 2.6 & 4.1 & 5.9 & 9.4 & 20.1 & 8.3 & 10.2
+	\\
+	\textit{Anopheles} & \textit{farauti (No.1)} & 3.6 & 4.7 & 5.5 & 6.6 & 9.3 & 5.9 &
+	2.1 \\
+	\textit{Anopheles} & \textit{farauti s.l.} & 3.9 & 4.9 & 5.6 & 6.6 & 8.7 & 5.9 & 1.6 \\
+	\textit{Anopheles} & \textit{darlingi} & 3.1 & 4.9 & 7.4 & 12.0 & 27.6 & 10.8 & 13.4 \\
+	\textit{Anopheles} & \textit{pseudopunctipennis s.l.} & 3.9 & 5.4 & 6.5 & 8.0 & 13.0 & 7.4 &
+	4.5 \\
+	\textit{Anopheles} & \textit{aquasalis} & 3.0 & 4.6 & 6.5 & 9.7 & 20.0 & 8.5 & 8.0 \\
+	\textit{Anopheles} & \textit{albitarsis (Sp C)} & 5.2 & 9.5 & 16.4 & 32.6
+	& \text{$>$100} & 66.8 & \text{$>$100} \\
+	\textit{Anopheles} & \textit{albitarsis (Sp A)} & 2.3 & 3.3 & 4.3 & 5.9 & 10.1 &
+	5.1 & 3.3 \\
+	\textit{Anopheles} & \textit{albitarsis (Sp. B)} & 2.2 & 3.1 & 4.1 & 5.4 & 9.0 & 4.7 & 2.7
+	\\
+	\textit{Anopheles} & \textit{albitarsis s.l.} & 2.6 & 4.0 & 5.8 & 9.2 & 19.9 & 8.2 & 9.5 \\
+	\textit{Anopheles} & \textit{albimanus} & 2.7 & 4.1 & 5.5 & 8.0 & 15.0 & 6.8 & 4.8 \\
+	\textit{Anopheles} & \textit{nuneztovari s.l.} & 2.6 & 3.6 & 4.5 & 6.1 & 11.4 & 5.5 & 3.8 \\
+\end{tabular}
+\caption{\textbf{Posterior summaries of mean female mosquito lifespan in terms of chronological time from Detinova-type dissection.} The 5\%, 25\%, 50\%, 75\% and 95\% columns indicate the respective quantiles of the posterior distribution for mean lifespan, and the `Mean' and `Std.
+dev.' columns indicate the posterior mean and standard deviation of mean lifespan.}
+\label{tab:detinova_lifespan_chronological}%
+\end{table}
 
@@
 	\centerline{\includegraphics[width=1.25\textwidth]{./Figure_files/mrr_mcPowerAnalysis_senescence.pdf}}
-	\caption{\textbf{The statistical power to detect senescence for A.
+	\caption{\textbf{MRR: statistical power analysis for senescence detection.} The plot shows the statistical power to detect senescence for A.
 three case study populations as a function of B.
 study length and C.
-number released.} For each parameter set we generated 500 simulated data series using the negative binomial sampling model with a gompertz hazard function ($\lambda(t) = \alpha e^{\beta t}$), and estimated the $\beta$ parameter using maximum likelihood and tested the null hypothesis $\beta=0$ (constant mortality risk) against the alternative $\beta>0$ using an likelihood ratio test with a 5\% test size.
+number released.
+For each parameter set we generated 500 simulated data series using the negative binomial sampling model with a gompertz hazard function ($\lambda(t) = \alpha e^{\beta t}$), and estimated the $\beta$ parameter using maximum likelihood and tested the null hypothesis $\beta=0$ (constant mortality risk) against the alternative $\beta>0$ using an likelihood ratio test with a 5\% test size.
 The error bars show the standard deviation in the power for each parameter set.
@@
 \begin{figure}[ht]
+	\centerline{\includegraphics[width=1.25\textwidth]{./Figure_files/elpd_vs_exponential_ordered_dissection.png}}
+	\caption{\textbf{Polovodova dissection: evidence for senescence.}  Here, we consider the predictive accuracy of age-dependent models of mortality versus the exponential model by species.
+The predictive accuracy was determined using K-Fold cross-validation as described in SOM; the measure of accuracy presented here by the central dots is the difference in estimated expected log pointwise predictive density compared to the exponential model.
+The lower and upper whiskers represent the lower and upper bounds of the 95\% confidence interval in predictive accuracy.
+The species have been ordered so that those with the highest average predictive accuracy across all age-dependent models compared to the exponential model appear at the top.
+If all age-dependent models outperform the exponential, we deem this as evidence for age-dependent mortality (`+'); if a subsample of models perform better, we deem the evidence ambiguous (`?'); and if no age-dependent model exceeds the performance of the exponential, we conclude no evidence in favour of senescence (`-').}
+	\label{fig:dissection_senescence}
+\end{figure}
+
+\begin{figure}[ht]
 	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/elpd_comparison.pdf}}
-	\caption{\textbf{Comparing the evidence for age-dependent mortality by fitting the Gompertz model to data from the MRR versus dissection studies for the 12 species with data from both.} The points indicate the estimated predictive accuracy compared with the exponential model (see SOM Methods) and the lower and upper whiskers represent 2.5\% and 97.5\% quantiles on this quantity.}\label{fig:elpd_comparison}
+	\caption{\textbf{Evidence for senescence: comparing evidence from Gompertz model across MRR and Polovodova dissection analyses.} The points indicate the estimated predictive accuracy compared with the exponential model (see SOM Methods) and the lower and upper whiskers represent 2.5\% and 97.5\% quantiles on this quantity.}\label{fig:elpd_comparison}
 \end{figure}
@@
 	\centerline{\includegraphics[width=1.25\textwidth]{./Figure_files/mrr_mcPowerAnalysis.pdf}}
-	\caption{\textbf{The average error in predicting mean lifespan as a function of study length (left) and number of released mosquitoes (right) in a Monte Carlo analysis.} For each parameter set we generated 400 simulated data series using the negative binomial sampling model with an exponential survival function as described in the text, and estimated the mean lifespan using maximum likelihood.
+	\caption{\textbf{MRR: Monte Carlo analysis of errors in predicting mean lifespan.} This shows the Root Mean Square Error (RMSE) in predicting mean lifespan as a function of study length (left) and number of released mosquitoes (right) in a Monte Carlo analysis.
+For each parameter set we generated 400 simulated data series using the negative binomial sampling model with an exponential survival function as described in the text, and estimated the mean lifespan using maximum likelihood.
 The error bars show the standard deviation in the prediction error.}

### 2020-11-21 writeUp/supplementary_materials_methods.tex
updated graph titles

@@
 
-The R file \verb|s_clean_data.R| cleans the data, makes the above amendments and outputs the data in a form amenable for estimation using Stan \citep{carpenter2016stan}.
-
 \begin{figure}[h]
 	\centerline{\includegraphics[width=1.25\textwidth]{./Figure_files/mrr_mapGenusCropped.pdf}}
-	\caption{\textbf{The location and number of separate time-series for the MRR studies included in this analysis.} The area of each circle indicates the number of time-series at each study site.
+	\caption{\textbf{MRR: location of studies included in the analysis.} The area of each circle indicates the number of time-series at each study site.
 The colour shows the genus of the mosquitoes in each study.}
@@
 	\end{tabular}%
-	\caption{\textbf{Summary of variables across all MRR time-series used in hierarchical analyses.} For the `Species' and `Genera' rows, the counts represent the number of distinct entries in the database; otherwise the counts represent the number of time series.}
+	\caption{\textbf{MRR: summary of time-series used in hierarchical analyses.} For the `Species' and `Genera' rows, the counts represent the number of distinct entries in the database; otherwise the counts represent the number of time series.}
 	\label{tab:mrr_aggregateData}%
@@
 		\end{tabularx}%
-		\caption{\textbf{Summary of data from individual MRR time-series used in hierarchical analyses.}}
+		\caption{\textbf{MRR: summary of data from individual time-series used in hierarchical analyses.}}
 		\label{tab:mrr_IndividualData}%
@@
 	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_sexReleasesOverTime.pdf}}
-	\caption{\textbf{The number of MRR time-series by date and sex of released mosquitoes.} The numbers represent by-decade totals of the 232 time-series we used in the main analysis.}
+	\caption{\textbf{MRR: time-series by date and sex of released mosquitoes.} The numbers represent by-decade totals of the 232 time-series we used in the main analysis.}
 	\label{fig:mrr_sexReleasesOverTime}
@@
 	\centerline{\includegraphics[width=1.25\textwidth]{./Figure_files/mrr_exampleMRRSeries.pdf}}
-	\caption{\textbf{Example MRR time-series for three different studies from the \cite{guerra2014global} database.} In all the above experiments, there was a single release of marked mosquitoes on day zero.}
+	\caption{\textbf{MRR: time-series for three different studies from the \cite{guerra2014global} database.} In all the above experiments, there was a single release of marked mosquitoes on day zero.}
 	\label{fig:mrr_exampleMRRSeries}
@@
 	\centerline{\includegraphics[width=1.25\textwidth]{./Figure_files/mrr_individualTimeSeries_priors.pdf}}
-	\caption{\textbf{The prior probability distributions used for lifespan (A), recapture probability (B) and dispersal parameter (C) in the individual time-series analysis for the MRR studies.} Note: the prior for the recapture probability was truncated at $\psi=0.1$ since most MRR day one recapture fractions were significantly below this threshold.}
+	\caption{\textbf{MRR: prior probability distributions used for lifespan (A), recapture probability (B) and dispersal parameter (C) in the individual time-series analysis.} Note: the prior for the recapture probability was truncated at $\psi=0.1$ since most MRR day one recapture fractions were significantly below this threshold.}
 	\label{fig:mrr_individualTimeSeries_priors}
@@
 	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_methodDiagramHierarchy.pdf}}
-	\caption{\textbf{A representation of the hierarchical Bayesian model, for the constant mortality model case when considering the species-level grouping.} Here `pdf' indicates `probability density function', and $\lambda_i$, $\psi_i$, and $\kappa_i$ represent the force of mortality for the exponential distribution, daily recapture probability and over-dispersion parameter of the negative binomial distribution for time-series $i$ in the dataset; $\mu_i(t)= \psi_i e^{-\lambda_i t}$ is the mean number of mosquitoes recaptured on day $t$.}
+	\caption{\textbf{MRR: structure of hierarchical Bayesian model.} This diagram represents the structure of the model for the constant mortality model case with the species-level grouping.
+Here `pdf' indicates `probability density function', and $\lambda_i$, $\psi_i$, and $\kappa_i$ represent the force of mortality for the exponential distribution, daily recapture probability and over-dispersion parameter of the negative binomial distribution for time-series $i$ in the dataset; $\mu_i(t)= \psi_i e^{-\lambda_i t}$ is the mean number of mosquitoes recaptured on day $t$.}
 	\label{fig:mrr_methodDiagramHierarchy}
@@
 	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_lifespan_priors.pdf}}
-	\caption{\textbf{Samples from the prior predictive distribution of mean lifespan across the six models of mosquito mortality introduced in Table \ref{tab:mrr_survivalDescription}.} In all cases, 10,000 samples were generated using the priors indicated in Table \ref{tab:mrr_priors}, resulting in distributions with a mean close to 10 days.}
+	\caption{\textbf{MRR: prior predictive distribution of mean lifespan across the six models of mosquito mortality.} In all cases, 10,000 samples were generated using the priors indicated in Table \ref{tab:mrr_priors}, resulting in distributions with a mean close to 10 days.}
 	\label{fig:mrr_meanLifespanPrior}
@@
 	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_priors_PsiKappa.pdf}}
-	\caption{\textbf{Samples from the priors for daily recapture probability ($\psi$) and over-dispersion parameter ($\kappa$).} In both panels, 10,000 samples were generated using the following hierarchical priors: for the individual time series parameters, $\psi_i \sim \text{exp}(r)$ (with $\psi$ constrained to be less than 0.1 - encompassing the upper limit of day one recapture fractions observed in most data series) and $\kappa_i \sim \text{exp}(p)$; for the top-level parameters: $r\sim \text{gamma}(100, 2)$ and $p\sim \text{log-normal}(1.5, 1)$.}\label{fig:mrr_PsiKappaPriors}
+	\caption{\textbf{MRR: samples from the priors for daily recapture probability ($\psi$) and over-dispersion parameter ($\kappa$).} In both panels, 10,000 samples were generated using the following hierarchical priors: for the individual time series parameters, $\psi_i \sim \text{exp}(r)$ (with $\psi$ constrained to be less than 0.1 - encompassing the upper limit of day one recapture fractions observed in most data series) and $\kappa_i \sim \text{exp}(p)$; for the top-level parameters: $r\sim \text{gamma}(100, 2)$ and $p\sim \text{log-normal}(1.5, 1)$.}\label{fig:mrr_PsiKappaPriors}
 \end{figure}
@@
 			\bottomrule
-	\end{tabularx}}\caption{\textbf{The priors used on parameters of each different survival model.} For the exponential model, the `group-level' priors were the same for the genus and `overall' models that were also estimated.
+	\end{tabularx}}\caption{\textbf{MRR: priors used on parameters of each survival model.} For the exponential model, the `group-level' priors were the same for the genus and `overall' models that were also estimated.
 The exponential model was the only model that was simple enough to allow the scale parameter of the log-normal ($\sigma$) to be estimated by the data.
@@
 	\includegraphics[page=1,scale=.6]{./Figure_files/mrr_ppcs_all}
-	\captionof{figure}{The numbers of mosquitoes recaptured (black lines) versus the 95\% central posterior interval of the posterior predictive distribution (blue shading) for a selection of the time series.
+	\captionof{figure}{\textbf{MRR: posterior predictive distribution versus actual data.} The figure shows the numbers of mosquitoes recaptured (black lines) versus the 95\% central posterior interval of the posterior predictive distribution (blue shading) for a selection of the time series.
 For the rest of the posterior predictive checks for the MRR models, see the file referenced in the text.}
@@
 	\includegraphics[page=2,scale=.6]{./Figure_files/mrr_ppcs_all}
-	\captionof{figure}{The numbers of mosquitoes recaptured (black lines) versus the 95\% central posterior interval of the posterior predictive distribution (blue shading) for a selection of the time series.
+	\captionof{figure}{\textbf{MRR: posterior predictive distribution versus actual data.} The figure shows the numbers of mosquitoes recaptured (black lines) versus the 95\% central posterior interval of the posterior predictive distribution (blue shading) for a selection of the time series.
 For the rest of the posterior predictive checks for the MRR models, see the file referenced in the text.}
@@
 	\includegraphics[page=3,scale=.6]{./Figure_files/mrr_ppcs_all}
-	\captionof{figure}{The numbers of mosquitoes recaptured (black lines) versus the 95\% central posterior interval of the posterior predictive distribution (blue shading) for a selection of the time series.
+	\captionof{figure}{\textbf{MRR: posterior predictive distribution versus actual data.} The figure shows the numbers of mosquitoes recaptured (black lines) versus the 95\% central posterior interval of the posterior predictive distribution (blue shading) for a selection of the time series.
 For the rest of the posterior predictive checks for the MRR models, see the file referenced in the text.}
@@
 	\centerline{\includegraphics[width=0.8\textwidth]{./Figure_files/dissection_timeSeries.pdf}}
-	\caption{\textbf{The numbers of published studies that estimate gonotrophic age by Polovodova's dissection method across the study period.}}\label{fig:dissection_timeSeries}
+	\caption{\textbf{Polovodova dissection: numbers of published studies that estimate gonotrophic age by this method.}}\label{fig:dissection_timeSeries}
 \end{figure}
@@
 	\centerline{\includegraphics[width=1.25\textwidth]{./Figure_files/dissection_map.pdf}}
-	\caption{\textbf{The geographic location of each of the Polovodova datasets included in the meta-analysis.} The area of the bubbles indicates the number of unique data series available.}\label{fig:dissection_map}
+	\caption{\textbf{Polovodova dissection: the geographic location of each of the datasets included in the meta-analysis.} The area of the bubbles indicates the number of unique data series available.}\label{fig:dissection_map}
 \end{figure}
@@
 	\end{tabular}%
-	\caption{\textbf{The numbers of Polovodova dissection series for each species or genus included in the overall dataset.} Note, these counts represent numbers of series analysed by the hierarchical models -- that is, after aggregating the individual series as described in Section \ref{sec:dissection_dissectionData}.}\label{tab:dissection_speciesNumbers}%
+	\caption{\textbf{Polovodova dissection: numbers of dissection series included in the overall dataset.} Note, these counts represent numbers of series analysed by the hierarchical models -- that is, after aggregating the individual series as described in Section \ref{sec:dissection_dissectionData}.}\label{tab:dissection_speciesNumbers}%
 \end{table}%
@@
 	\end{adjustwidth}
-	\caption{\textbf{Summaries of the characteristics of the 131 individual Polovodova dissection data series that are included in this analysis.} Note when series are censored (see Section \ref{sec:dissection_censored}) we assume that all mosquitoes dissected over the threshold are of this age when calculating summary statistics.}\label{tab:dissection_summaryStats}
+	\caption{\textbf{Polovodova dissection: summaries of the data series included in this analysis.} Note when series are censored (see Section \ref{sec:dissection_censored}) we assume that all mosquitoes dissected over the threshold are of this age when calculating summary statistics.}\label{tab:dissection_summaryStats}
 \end{table}
@@
 	\centerline{\includegraphics[width=1.3\textwidth]{./Figure_files/dissection_parity_data1.png}}
-	\caption{\textbf{The normalised physiological age series for nine species in the database.} Each different coloured line represents an individual series.
+	\caption{\textbf{Polovodova dissection: normalised physiological age series for nine species in the database.} Each different coloured line represents an individual series.
 In each cas,e the count for all ages has been normalised by the nulliparous count.
@@
 	\centerline{\includegraphics[width=1.3\textwidth]{./Figure_files/dissection_parity_data2.png}}
-	\caption{\textbf{The normalised physiological age series for nine species in the database.} Each different coloured line represents an individual series.
+	\caption{\textbf{Polovodova dissection: normalised physiological age series for nine species in the database.} Each different coloured line represents an individual series.
 In each case, the count for all ages has been normalised by the nulliparous count.
@@
 	\centerline{\includegraphics[width=1.3\textwidth]{./Figure_files/dissection_parity_data3.png}}
-	\caption{\textbf{The normalised physiological age series for seven species in the database.} Each different coloured line represents an individual series.
+	\caption{\textbf{Polovodova dissection: normalised physiological age series for nine species in the database.} Each different coloured line represents an individual series.
 In each case, the count for all ages has been normalised by the nulliparous count.
@@
 	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/dissection_lifespanPriorsParity.pdf}}
-	\caption{\textbf{The priors for mean lifespan used to analyse the Polovodova dissection data for each of the hierarchical models.} The distributional form of the priors is given in Table \ref{tab:dissection_priors}.
+	\caption{\textbf{Polovodova dissection: prior predictive mean lifespan for each mortality model.} The distributional form of the priors is given in Table \ref{tab:dissection_priors}.
 In all cases, the plots show data for 10,000 samples from the relevant prior distribution.}\label{fig:dissection_lifespanPriorsParity}
@@
 			\bottomrule
-	\end{tabularx}}\caption{\textbf{The priors used on parameters of each different survival model for the Polovodova dissection data analysis.} For the exponential model the `group-level' priors were the same for the genus and `overall' models that were also estimated.
+	\end{tabularx}}\caption{\textbf{Polovodova dissection: priors used on parameters of each different survival model.} For the exponential model the `group-level' priors were the same for the genus and `overall' models that were also estimated.
 The exponential model was the only model that was simple enough to allow the scale parameter of the log-normal ($\sigma$) to be estimated by the data.
@@
 	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/dissection_lifespanPrior_individual.pdf}}
-	\caption{\textbf{The priors for A.
+	\caption{\textbf{Polovodova dissection: priors for A.
 mean lifespan and B.
@@
 	\includegraphics[page=1,scale=.7]{./Figure_files/dissection_ppcs_all}
-	\captionof{figure}{\textbf{The numbers of mosquitoes recaptured (black lines) versus the 95\% central posterior interval of the posterior predictive distribution (blue shading) for a selection of the Polovodova dissection series.} The titles correspond to the unique combination of "identifier - physiological-experiment-id" in the database.
+	\captionof{figure}{\textbf{Polovodova dissection: posterior predictive checks.} The numbers of mosquitoes recaptured (black lines) versus the 95\% central posterior interval of the posterior predictive distribution (blue shading) for a selection of the Polovodova dissection series.
+The titles correspond to the unique combination of "identifier - physiological-experiment-id" in the database.
 For the rest of the posterior predictive checks for the dissection data models, see the file referenced in the text.}
@@
 	\includegraphics[page=2,scale=.7]{./Figure_files/dissection_ppcs_all}
-	\captionof{figure}{\textbf{The numbers of mosquitoes recaptured (black lines) versus the 95\% central posterior interval of the posterior predictive distribution (blue shading) for a selection of the Polovodova dissection series.} The titles correspond to the unique combination of "identifier - physiological-experiment-id" in the database.
+	\captionof{figure}{\textbf{Polovodova dissection: posterior predictive checks.} The numbers of mosquitoes recaptured (black lines) versus the 95\% central posterior interval of the posterior predictive distribution (blue shading) for a selection of the Polovodova dissection series.
+The titles correspond to the unique combination of "identifier - physiological-experiment-id" in the database.
 For the rest of the posterior predictive checks for the dissection data models, see the file referenced in the text.}
@@
 	\includegraphics[page=3,scale=.7]{./Figure_files/dissection_ppcs_all}
-	\captionof{figure}{\textbf{The numbers of mosquitoes recaptured (black lines) versus the 95\% central posterior interval of the posterior predictive distribution (blue shading) for a selection of the Polovodova dissection series.} The titles correspond to the unique combination of "identifier - physiological-experiment-id" in the database.
+	\captionof{figure}{\textbf{Polovodova dissection: posterior predictive checks.} The numbers of mosquitoes recaptured (black lines) versus the 95\% central posterior interval of the posterior predictive distribution (blue shading) for a selection of the Polovodova dissection series.
+The titles correspond to the unique combination of "identifier - physiological-experiment-id" in the database.
 For the rest of the posterior predictive checks for the dissection data models, see the file referenced in the text.}
@@
 	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/detinova_prior_predictive.pdf}}
-	\caption{\textbf{The prior predictive distribution of lifespan for the analysis of Detinova parity data in terms of (left) physiological age and (right) chronological age.} Here, we used the gonotrophic cycle duration estimated for \textit{Anopheles}, $G\sim \mathcal{N}(3.65, 0.40)$, to convert physiological survival to calendar days.
+	\caption{\textbf{Detinova dissection: prior predictive distribution of lifespan in terms of (left) physiological age and (right) chronological age.} Here, we used the gonotrophic cycle duration estimated for \textit{Anopheles}, $G\sim \mathcal{N}(3.65, 0.40)$, to convert physiological survival to calendar days.
 The figure shows 100,000 samples from the prior predictive distribution.
@@
 	\centerline{\includegraphics[width=0.7\textwidth]{./Figure_files/detinova_ecdf_africa_grouped.pdf}}
-	\caption{\textbf{Africa, complex-level: empirical cumulative distribution function for observations (orange) and posterior predictive simulations (green).} In each case, we ran 4 Markov chains, with 200 iterations per chain; discarding the first 100 iterations as warm-up.}\label{fig:detinova_ecdf_africa_grouped}
+	\caption{\textbf{Detinova dissection: posterior predictive check for Africa, at the complex-level.} Plots show the empirical cumulative distribution functions for observations (orange) and posterior predictive simulations (green).
+In each case, we ran 4 Markov chains, with 200 iterations per chain; discarding the first 100 iterations as warm-up.}\label{fig:detinova_ecdf_africa_grouped}
 \end{figure}
@@
 	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/detinova_ecdf_africa_species.pdf}}
-	\caption{\textbf{Africa, species-level: empirical cumulative distribution function for observations (orange) and posterior predictive simulations (green).} In each case, we ran 4 Markov chains, with 200 iterations per chain; discarding the first 100 iterations as warm-up.}\label{fig:detinova_ecdf_africa_species}
+	\caption{\textbf{Detinova dissection: posterior predictive check for Africa, at the species-level.} Plots show the empirical cumulative distribution functions for observations (orange) and posterior predictive simulations (green).
+In each case, we ran 4 Markov chains, with 200 iterations per chain; discarding the first 100 iterations as warm-up.}\label{fig:detinova_ecdf_africa_species}
 \end{figure}
@@
 	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/detinova_ecdf_americas_grouped.pdf}}
-	\caption{\textbf{Americas, complex-level: empirical cumulative distribution function for observations (orange) and posterior predictive simulations (green).} In each case, we ran 4 Markov chains, with 200 iterations per chain; discarding the first 100 iterations as warm-up.}\label{fig:detinova_ecdf_americas_grouped}
+	\caption{\textbf{Detinova dissection: posterior predictive check for the Americas, at the complex-level.} Plots show the empirical cumulative distribution functions for observations (orange) and posterior predictive simulations (green).
+In each case, we ran 4 Markov chains, with 200 iterations per chain; discarding the first 100 iterations as warm-up.}\label{fig:detinova_ecdf_americas_grouped}
 \end{figure}
@@
 	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/detinova_ecdf_americas_species.pdf}}
-	\caption{\textbf{Americas, species-level: empirical cumulative distribution function for observations (orange) and posterior predictive simulations (green).} In each case, we ran 4 Markov chains, with 200 iterations per chain; discarding the first 100 iterations as warm-up.}\label{fig:detinova_ecdf_americas_species}
+	\caption{\textbf{Detinova dissection: posterior predictive check for the Americas, at the species-level.} Plots show the empirical cumulative distribution functions for observations (orange) and posterior predictive simulations (green).
+In each case, we ran 4 Markov chains, with 200 iterations per chain; discarding the first 100 iterations as warm-up.}\label{fig:detinova_ecdf_americas_species}
 \end{figure}
@@
 	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/detinova_ecdf_asia_grouped.pdf}}
-	\caption{\textbf{Asia, complex-level: empirical cumulative distribution function for observations (orange) and posterior predictive simulations (green).} In each case, we ran 4 Markov chains, with 200 iterations per chain; discarding the first 100 iterations as warm-up.}\label{fig:detinova_ecdf_asia_grouped}
+	\caption{\textbf{Detinova dissection: posterior predictive check for Asia, at the complex-level.} Plots show the empirical cumulative distribution functions for observations (orange) and posterior predictive simulations (green).
+In each case, we ran 4 Markov chains, with 200 iterations per chain; discarding the first 100 iterations as warm-up.}\label{fig:detinova_ecdf_asia_grouped}
 \end{figure}
@@
 	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/detinova_ecdf_asia_species.pdf}}
-	\caption{\textbf{Asia, species-level: empirical cumulative distribution function for observations (orange) and posterior predictive simulations (green).} In each case, we ran 4 Markov chains, with 200 iterations per chain; discarding the first 100 iterations as warm-up.}\label{fig:detinova_ecdf_asia_species}
+	\caption{\textbf{Detinova dissection: posterior predictive check for Asia, at the species-level.} Plots show the empirical cumulative distribution functions for observations (orange) and posterior predictive simulations (green).
+In each case, we ran 4 Markov chains, with 200 iterations per chain; discarding the first 100 iterations as warm-up.}\label{fig:detinova_ecdf_asia_species}
 \end{figure}

### 2020-11-28 writeUp/figures.tex
updated figures and ms

--- 2020-05-25 writeUp/figures.tex
@@
 	\centerline{\includegraphics[width=1.3\textwidth]{./Figure_files/detinova_gambiae_daynight_range_lifespan_country.pdf}}
-	\caption{\textbf{Day-night temperature range versus lifespan estimates for \textit{Anopheles gambiae s.l.}.} Points indicate the posterior mean estimates and the whiskers show 25\% and 75\% quantiles.
+	\caption{\textbf{Detinova dissection: day-night temperature range versus lifespan estimates for \textit{Anopheles gambiae s.l.}.} Points indicate the posterior mean estimates and the whiskers show 25\% and 75\% quantiles.
 Each blue line indicates a posterior estimate of the relationship between lifespan and temperature range; the orange line indicates the posterior median.}
@@
 	\centerline{\includegraphics[width=1.3\textwidth]{./Figure_files/mrr_elpd_vs_exponential_ordered.pdf}}
-	\caption{\textbf{Detecting senescence in MRR studies.} Here, we consider the predictive accuracy of age-dependent models of mortality versus the exponential model by species.
+	\caption{\textbf{MRR: evidence for senescence.} Here, we consider the predictive accuracy of age-dependent models of mortality versus the exponential model by species.
 The predictive accuracy was determined using K-Fold cross-validation as described in SOM; the measure of accuracy presented here by the central dots is the difference in estimated expected log pointwise predictive density compared to the exponential model.

### 2020-11-28 writeUp/supplementary_materials.tex
updated figures and ms

@@
 
-\section{MRR experments}
+\section{Lifespan estimates}
 
@@
 	\end{tabular}
-	\caption{\textbf{Posterior summaries of mean unfed female mosquito lifespan from the MRR analysis.} The 5\%, 25\%, 50\%, 75\% and 95\% columns indicate the respective quantiles of the posterior distribution for mean lifespan, and the `Mean' and `Std.
-dev.' columns indicate the posterior mean and standard deviation of mean lifespan.}
+	\caption{\textbf{MRR: mean unfed female mosquito lifespan estimates.} The 5\%, 25\%, 50\%, 75\% and 95\% columns indicate the respective quantiles of the posterior distribution for mean lifespan, and the `Mean' and `Std.
+dev.' columns indicate the posterior mean and standard deviation of mean lifespan.
+Units of lifespan are days.}
 	\label{tab:mrr_estimated_lifespans}%
@@
 	\label{fig:mrr_genusTopLevel}
+\end{figure}
+
+\begin{figure}[h]
+	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_female_blood_sugar.pdf}}
+	\caption{\textbf{MRR: female mosquito mean lifespan by pre-release feeding status.} The lifespans shown are for mosquitoes that were not fed with sugar or blood (for females) before release.
+The middle line in each box shows the median estimates and the solid dot indicates the mean.
+The left and right box edges show the 25\%, and 75\% posterior quantiles respectively.
+The whiskers show the range of the data, excluding points lying more than 1.5 times the interquartile range away from each edge of the box.
+The numbers before the start of the left whisker indicate the number of individual time-series within each species.
+All estimates were obtained using the hierarchical exponential survival model.}
+	\label{fig:mrr_female_blood_sugar}
 \end{figure}
@@
 \begin{figure}[h]
-	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_female_blood_sugar.pdf}}
-	\caption{\textbf{MRR: female mosquito mean lifespan by pre-release feeding status.} The lifespans shown are for mosquitoes that were not fed with sugar or blood (for females) before release.
-The middle line in each box shows the median estimates and the solid dot indicates the mean.
-The left and right box edges show the 25\%, and 75\% posterior quantiles respectively.
-The whiskers show the range of the data, excluding points lying more than 1.5 times the interquartile range away from each edge of the box.
-The numbers before the start of the left whisker indicate the number of individual time-series within each species.
-All estimates were obtained using the hierarchical exponential survival model.}
-	\label{fig:mrr_female_blood_sugar}
+	\centerline{\includegraphics[width=1.3\textwidth]{./Figure_files/dissection_individualEstimates_allSpecies.pdf}}
+	\caption{\textbf{Polovodova dissection: individual time-series estimates of adult mosquito mean lifespan.} The middle line in each box shows the median estimates.
+The left and right box whiskers show the 25\%, and 75\% posterior quantiles respectively.
+All estimates were obtained using the non-hierarchical exponential survival model.}
+	\label{fig:dissection_individualEstimates_allSpecies}
+\end{figure}
+
+\begin{figure}[h]
+	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/detinova_insecticide.pdf}}
+	\caption{\textbf{Detinova dissection: estimated impact of insecticide on lifespan.} The middle line in each box shows the median estimates.
+The left and right box whiskers show the 10\%, and 90\% posterior quantiles respectively.}
+	\label{fig:detinova_insecticide}
+\end{figure}
+
+\begin{figure}[h]
+	\centerline{\includegraphics[width=1.3\textwidth]{./Figure_files/detinova_individual_lifespan_no_insecticide.pdf}}
+	\caption{\textbf{Detinova dissection: individual time-series estimates of adult mosquito mean lifespan.} The middle line in each box shows the median estimates.
+The left and right box whiskers show the 25\%, and 75\% posterior quantiles respectively.
+Note, that a handful of observations are not shown in the plot since their estimates exceeded the reasonable range (here deemed to be up to 12 cycles).}
+	\label{fig:detinova_lifespan_individual}
 \end{figure}
@@
 All estimates of lifespan were obtained using the non-hierarchical exponential survival model.}\label{fig:mrr_ThreeSpeciesVersusTemperature}
-\end{figure}
-
-\section{Polovodova dissection analysis}
-
-\begin{figure}[h]
-	\centerline{\includegraphics[width=1.3\textwidth]{./Figure_files/dissection_individualEstimates_allSpecies.pdf}}
-	\caption{\textbf{Polovodova dissection: individual time-series estimates of adult mosquito mean lifespan.} The middle line in each box shows the median estimates.
-The left and right box whiskers show the 25\%, and 75\% posterior quantiles respectively.
-All estimates were obtained using the non-hierarchical exponential survival model.}
-	\label{fig:dissection_individualEstimates_allSpecies}
 \end{figure}
@@
 	\end{tabular}
-	\caption{\textbf{Posterior summaries of mean female mosquito lifespan in terms of gonotrophic cycles from dissection studies.} The 5\%, 25\%, 50\%, 75\% and 95\% columns indicate the respective quantiles of the posterior distribution for mean lifespan, and the `Mean' and `Std.
+	\caption{\textbf{Polovodova dissection: mean female mosquito lifespan in terms of gonotrophic cycles.} The 5\%, 25\%, 50\%, 75\% and 95\% columns indicate the respective quantiles of the posterior distribution for mean lifespan, and the `Mean' and `Std.
 dev.' columns indicate the posterior mean and standard deviation of mean lifespan.}
@@
 
-\begin{figure}[h]
-	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/dissection_lifetimes_exponential_chron.pdf}}
-	\caption{\textbf{Polovodova dissection: female mosquito mean chronological lifespan estimates.} The lifespans shown are for mosquitoes that were not fed with sugar or blood (for females) before release.
-The middle line in each box shows the median estimates and the solid dot indicates the mean.
-The left and right box edges show the 25\%, and 75\% posterior quantiles respectively.
-The whiskers show the range of the data, excluding points lying more than 1.5 times the interquartile range away from each edge of the box.
-The numbers before the start of the left whisker indicate the number of individual time-series within each species.
-All estimates were obtained using the hierarchical exponential survival model.}
-	\label{fig:dissection_lifetimes_exponential_chron}
-\end{figure}
+\begin{table}[htbp!]
+	\begin{tabular}{l|l|l|l|l|l|l|l|l}
+		\textbf{Genus} & \textbf{Species} & \textbf{5$\%$} & \textbf{25$\%$} & \textbf{50$\%$} &
+		\textbf{75$\%$} & \textbf{95$\%$} & \textbf{Mean} & \textbf{Std. dev.} \\
+		\hline
+	\textit{Anopheles} & \textit{gambiae (Form M)} & 0.9 & 1.8 & 2.6 & 3.6 & 6.7 & 3.2 & 4.6 \\
+	\textit{Anopheles} & \textit{gambiae} & 0.6 & 1.1 & 1.7 & 2.7 & 5.6 & 2.2 & 2.2 \\
+	\textit{Anopheles} & \textit{arabiensis} & 0.4 & 0.9 & 1.5 & 2.6 & 6.3 & 2.3 & 3.9 \\
+	\textit{Anopheles} & \textit{gambiae s.l.} & 0.7 & 1.7 & 3.1 & 6.3 & 21.2 & 7.3 & 56.1 \\
+	\textit{Anopheles} & \textit{moucheti} & 1.0 & 1.9 & 2.8 & 4.2 & 8.9 & 3.8 & 6.3 \\
+	\textit{Anopheles} & \textit{funestus} & 0.8 & 1.6 & 2.7 & 4.4 & 10.6 & 4.0 & 10.6 \\
+	\textit{Anopheles} & \textit{funestus s.l.} & 0.6 & 1.2 & 2.1 & 3.7 & 9.8 & 3.3 & 4.8 \\
+	\textit{Anopheles} & \textit{nili s.l.} & 1.0 & 1.4 & 1.7 & 2.1 & 2.8 & 1.8 & 0.6 \\
+	\textit{Anopheles} & \textit{subpictus s.l.} & 0.9 & 1.5 & 2.1 & 2.8 & 4.8 & 2.4 & 1.8 \\
+	\textit{Anopheles} & \textit{dirus (formerly Sp. A)} & 0.5 & 1.0 & 1.5 & 2.2 & 4.2 & 1.8 &
+	1.5 \\
+	\textit{Anopheles} & \textit{dirus s.l.} & 0.6 & 1.2 & 2.0 & 3.1 & 7.4 & 2.9 & 6.8 \\
+	\textit{Anopheles} & \textit{aconitus} & 0.5 & 1.1 & 1.8 & 2.8 & 6.3 & 2.5 & 4.7 \\
+	\textit{Anopheles} & \textit{sundaicus s.l.} & 0.4 & 0.9 & 1.6 & 2.9 & 7.5 & 2.7 & 7.4 \\
+	\textit{Anopheles} & \textit{minimus s.l.} & 0.5 & 1.1 & 1.6 & 2.5 & 5.8 & 2.3 & 5.6 \\
+	\textit{Anopheles} & \textit{fluviatilis s.l.} & 0.4 & 0.9 & 1.4 & 2.3 & 5.1 & 2.0 & 3.0 \\
+	\textit{Anopheles} & \textit{maculatus s.l.} & 0.6 & 1.0 & 1.2 & 1.6 & 2.5 & 1.4 & 0.8 \\
+	\textit{Anopheles} & \textit{farauti s.l.} & 0.3 & 0.7 & 1.2 & 2.0 & 4.9 & 1.8 & 2.3 \\
+	\textit{Anopheles} & \textit{annularis s.l.} & 0.3 & 0.7 & 1.2 & 2.0 & 4.7 & 1.7 & 1.9 \\
+	\textit{Anopheles} & \textit{sinensis} & 0.4 & 0.8 & 1.1 & 1.7 & 3.0 & 1.4 & 1.0 \\
+	\textit{Anopheles} & \textit{culicifacies s.l.} & 0.5 & 0.8 & 1.1 & 1.5 & 2.3 & 1.2 & 0.6 \\
+	\textit{Anopheles} & \textit{anthropophagus (lesteri)} & 0.4 & 0.9 & 1.4 & 2.3 & 5.1 & 2.0 &
+	3.0 \\
+	\textit{Anopheles} & \textit{lesteri s.l.} & 0.3 & 0.6 & 1.1 & 1.9 & 4.7 & 1.7 & 2.4 \\
+	\textit{Anopheles} & \textit{barbirostris s.l.} & 0.4 & 0.7 & 1.0 & 1.4 & 2.8 & 1.2 & 1.3 \\
+	\textit{Anopheles} & \textit{albitarsis (formerly Sp C)} & 0.8 & 2.0 & 3.9 & 8.4 &
+	32.6 & 13.3 & \text{$>$100} \\
+	\textit{Anopheles} & \textit{albitarsis (formerly Sp A)} & 0.2 & 0.4 & 0.6 & 1.0 & 2.3 & 0.8
+	& 0.9 \\
+	\textit{Anopheles} & \textit{albitarsis (Sp. B)} & 0.1 & 0.3 & 0.6 & 0.9 & 1.9 & 0.7 & 0.8
+	\\
+	\textit{Anopheles} & \textit{albitarsis s.l.} & 0.4 & 0.9 & 1.5 & 3.0 & 8.3 & 2.8 & 5.8 \\
+	\textit{Anopheles} & \textit{nuneztovari s.l.} & 0.3 & 0.6 & 1.1 & 1.9 & 4.4 & 1.6 & 2.3 \\
+	\textit{Anopheles} & \textit{darlingi} & 0.2 & 0.7 & 1.2 & 2.4 & 6.7 & 2.2 & 4.2 \\
+	\textit{Anopheles} & \textit{albimanus} & 0.2 & 0.5 & 0.9 & 1.6 & 3.6 & 1.3 & 1.6 \\
+	\textit{Anopheles} & \textit{aquasalis} & 0.3 & 0.6 & 1.1 & 1.8 & 4.4 & 1.6 & 4.0 \\
+	\textit{Anopheles} & \textit{aquasalis s.l.} & 0.2 & 0.4 & 0.7 & 1.0 & 2.4 & 0.9 & 1.3 \\
+\end{tabular}
+\caption{\textbf{Detinova dissection: mean female mosquito lifespan estimates in terms of gonotrophic cycles.} The 5\%, 25\%, 50\%, 75\% and 95\% columns indicate the respective quantiles of the posterior distribution for mean lifespan, and the `Mean' and `Std.
+dev.' columns indicate the posterior mean and standard deviation of mean lifespan.}
+\label{tab:detinova_lifespan_gonotrophic}%
+\end{table}
 
@@
 	\end{tabular}
-	\caption{\textbf{Posterior summaries of mean female mosquito lifespan in terms of chronological time from dissection studies.} The 5\%, 25\%, 50\%, 75\% and 95\% columns indicate the respective quantiles of the posterior distribution for mean lifespan, and the `Mean' and `Std.
-dev.' columns indicate the posterior mean and standard deviation of mean lifespan.}
+	\caption{\textbf{Polovodova dissection: mean female mosquito lifespan estimates in terms of chronological time.} The 5\%, 25\%, 50\%, 75\% and 95\% columns indicate the respective quantiles of the posterior distribution for mean lifespan, and the `Mean' and `Std.
+dev.' columns indicate the posterior mean and standard deviation of mean lifespan.
+Units of lifespan are days.}
 	\label{tab:dissection_estimated_lifespans_chron}%
 \end{table}
-
-\section{Detinova parity analysis}
 
@@
 		\hline
-	\textit{Anopheles} & \textit{funestus} & 5.1 & 8.3 & 11.9 & 18.2 & 41.5 & 16.7 & 20.8 \\
-	\textit{Anopheles} & \textit{funestus s.l.} & 4.4 & 7.8 & 13.2 & 24.0 & 75.5 & 25.5 & 58.3
+	\textit{Anopheles} & \textit{gambiae (Form M)} & 3.3 & 6.2 & 8.7 & 11.9 & 22.3 & 10.7 & 14.9
 	\\
-	\textit{Anopheles} & \textit{nili s.l.} & 5.6 & 8.9 & 12.4 & 17.7 & 36.1 & 16.0 & 17.0 \\
-	\textit{Anopheles} & \textit{gambiae} & 4.0 & 6.1 & 8.3 & 12.0 & 22.1 & 10.3 & 8.5 \\
-	\textit{Anopheles} & \textit{gambiae (Form M)} & 3.5 & 5.6 & 8.2 & 12.4 & 27.4 & 11.2 & 14.0
+	\textit{Anopheles} & \textit{gambiae} & 2.0 & 3.8 & 5.8 & 9.0 & 18.4 & 7.6 & 7.1 \\
+	\textit{Anopheles} & \textit{arabiensis} & 1.4 & 3.2 & 5.3 & 8.8 & 20.8 & 7.7 & 11.9 \\
+	\textit{Anopheles} & \textit{gambiae s.l.} & 2.5 & 5.7 & 10.3 & 20.6 & 68.4 & 23.9 &
+	\text{$>$100} \\
+	\textit{Anopheles} & \textit{moucheti} & 3.5 & 6.4 & 9.2 & 13.9 & 29.0 & 12.5 & 19.7 \\
+	\textit{Anopheles} & \textit{funestus} & 2.9 & 5.7 & 9.0 & 14.5 & 34.4 & 13.4 & 31.0 \\
+	\textit{Anopheles} & \textit{funestus s.l.} & 2.0 & 4.3 & 7.1 & 12.5 & 32.2 & 11.1 & 15.6 \\
+	\textit{Anopheles} & \textit{nili s.l.} & 3.6 & 4.8 & 5.8 & 7.1 & 9.6 & 6.1 & 2.0 \\
+	\textit{Anopheles} & \textit{subpictus s.l.} & 3.4 & 5.3 & 7.1 & 9.5 & 16.2 & 8.2 & 6.2 \\
+	\textit{Anopheles} & \textit{dirus (formerly Sp. A)} & 1.9 & 3.6 & 5.1 & 7.4 & 13.8 & 6.2 &
+	5.0 \\
+	\textit{Anopheles} & \textit{dirus s.l.} & 2.0 & 4.4 & 6.8 & 10.5 & 24.1 & 9.8 & 23.0 \\
+	\textit{Anopheles} & \textit{aconitus} & 1.9 & 4.0 & 6.2 & 9.6 & 20.7 & 8.5 & 15.4 \\
+	\textit{Anopheles} & \textit{sundaicus s.l.} & 1.5 & 3.4 & 5.6 & 9.7 & 24.7 & 9.1 & 23.6 \\
+	\textit{Anopheles} & \textit{minimus s.l.} & 1.9 & 3.8 & 5.5 & 8.3 & 19.2 & 7.9 & 17.8 \\
+	\textit{Anopheles} & \textit{fluviatilis s.l.} & 1.6 & 3.2 & 5.0 & 7.8 & 16.9 & 6.8 & 9.6 \\
+	\textit{Anopheles} & \textit{maculatus s.l.} & 2.2 & 3.5 & 4.5 & 5.6 & 8.6 & 4.9 & 2.6 \\
+	\textit{Anopheles} & \textit{farauti s.l.} & 1.2 & 2.6 & 4.3 & 6.9 & 16.0 & 6.1 & 7.5 \\
+	\textit{Anopheles} & \textit{annularis s.l.} & 1.1 & 2.5 & 4.3 & 6.9 & 15.6 & 5.8 & 6.2 \\
+	\textit{Anopheles} & \textit{sinensis} & 1.6 & 2.8 & 4.1 & 5.9 & 10.3 & 4.8 & 3.2 \\
+	\textit{Anopheles} & \textit{culicifacies s.l.} & 1.9 & 3.0 & 4.0 & 5.2 & 8.0 & 4.3 & 2.0 \\
+	\textit{Anopheles} & \textit{anthropophagus (lesteri)} & 1.6 & 3.2 & 5.0 & 7.7 & 16.9 & 6.8
+	& 9.9 \\
+	\textit{Anopheles} & \textit{lesteri s.l.} & 0.9 & 2.2 & 3.9 & 6.5 & 15.6 & 5.7 & 7.8 \\
+	\textit{Anopheles} & \textit{barbirostris s.l.} & 1.4 & 2.5 & 3.5 & 4.9 & 9.6 & 4.3 & 4.1 \\
+	\textit{Anopheles} & \textit{albitarsis (formerly Sp C)} & 3.0 & 6.8 & 13.0 & 27.3
+	& \text{$>$100} & 44.2 & \text{$>$100} \\
+	\textit{Anopheles} & \textit{albitarsis (formerly Sp A)} & 0.6 & 1.3 & 2.2 & 3.7 & 7.7 & 3.0
+	& 3.1 \\
+	\textit{Anopheles} & \textit{albitarsis (Sp. B)} & 0.5 & 1.2 & 2.0 & 3.3 & 6.6 & 2.6 & 2.6
 	\\
-	\textit{Anopheles} & \textit{arabiensis} & 3.2 & 4.9 & 7.2 & 11.2 & 24.0 & 9.8 & 10.0 \\
-	\textit{Anopheles} & \textit{gambiae s.l.} & 3.4 & 5.7 & 8.8 & 14.4 & 35.7 & 13.3 & 19.7 \\
-	\textit{Anopheles} & \textit{leucosphyrus} & 5.0 & 8.1 & 11.0 & 15.4 & 29.0 & 13.8 & 26.4 \\
-	\textit{Anopheles} & \textit{leucosphyrus s.l.} & 5.0 & 7.9 & 11.0 & 15.8 & 29.7 & 14.1 &
-	25.3 \\
-	\textit{Anopheles} & \textit{aconitus} & 3.9 & 6.6 & 9.5 & 13.7 & 29.4 & 13.4 & 65.1 \\
-	\textit{Anopheles} & \textit{dirus (Sp. A)} & 4.9 & 7.4 & 10.3 & 14.9 & 30.2 & 13.1
-	& 10.7 \\
-	\textit{Anopheles} & \textit{dirus s.l.} & 4.3 & 6.6 & 9.4 & 14.3 & 28.8 & 12.3 & 10.5 \\
-	\textit{Anopheles} & \textit{minimus (Sp. A)} & 4.1 & 6.9 & 10.2 & 16.0 & 37.2 &
-	15.0 & 24.5 \\
-	\textit{Anopheles} & \textit{minimus s.l.} & 3.9 & 5.8 & 7.9 & 11.2 & 20.5 & 9.6 & 6.7 \\
-	\textit{Anopheles} & \textit{barbirostris s.l.} & 3.8 & 5.9 & 7.9 & 11.2 & 23.1 & 10.4 &
-	11.9 \\
-	\textit{Anopheles} & \textit{maculatus} & 3.2 & 5.0 & 7.0 & 10.5 & 21.7 & 9.2 & 8.8 \\
-	\textit{Anopheles} & \textit{maculatus s.l.} & 3.4 & 5.3 & 7.6 & 11.7 & 25.6 & 10.3 & 10.5
-	\\
-	\textit{Anopheles} & \textit{sundaicus s.l.} & 3.2 & 4.9 & 6.9 & 10.6 & 22.4 & 9.3 & 9.3 \\
-	\textit{Anopheles} & \textit{subpictus s.l.} & 2.9 & 4.6 & 6.8 & 11.0 & 25.6 & 10.0 & 13.6
-	\\
-	\textit{Anopheles} & \textit{anthropophagus (lesteri)} & 3.2 & 4.7 & 6.4 & 9.1 & 18.1 & 8.1
-	& 9.3 \\
-	\textit{Anopheles} & \textit{annularis s.l.} & 4.3 & 5.4 & 6.3 & 7.4 & 10.1 & 6.7 & 2.1 \\
-	\textit{Anopheles} & \textit{sinensis} & 3.5 & 4.8 & 6.2 & 8.2 & 13.1 & 7.0 & 3.6 \\
-	\textit{Anopheles} & \textit{fluviatilis s.l.} & 3.6 & 4.8 & 6.0 & 7.5 & 11.0 & 6.5 & 2.6 \\
-	\textit{Anopheles} & \textit{culicifacies s.l.} & 2.6 & 4.1 & 5.9 & 9.4 & 20.1 & 8.3 & 10.2
-	\\
-	\textit{Anopheles} & \textit{farauti (No.1)} & 3.6 & 4.7 & 5.5 & 6.6 & 9.3 & 5.9 &
-	2.1 \\
-	\textit{Anopheles} & \textit{farauti s.l.} & 3.9 & 4.9 & 5.6 & 6.6 & 8.7 & 5.9 & 1.6 \\
-	\textit{Anopheles} & \textit{darlingi} & 3.1 & 4.9 & 7.4 & 12.0 & 27.6 & 10.8 & 13.4 \\
-	\textit{Anopheles} & \textit{pseudopunctipennis s.l.} & 3.9 & 5.4 & 6.5 & 8.0 & 13.0 & 7.4 &
-	4.5 \\
-	\textit{Anopheles} & \textit{aquasalis} & 3.0 & 4.6 & 6.5 & 9.7 & 20.0 & 8.5 & 8.0 \\
-	\textit{Anopheles} & \textit{albitarsis (Sp C)} & 5.2 & 9.5 & 16.4 & 32.6
-	& \text{$>$100} & 66.8 & \text{$>$100} \\
-	\textit{Anopheles} & \textit{albitarsis (Sp A)} & 2.3 & 3.3 & 4.3 & 5.9 & 10.1 &
-	5.1 & 3.3 \\
-	\textit{Anopheles} & \textit{albitarsis (Sp. B)} & 2.2 & 3.1 & 4.1 & 5.4 & 9.0 & 4.7 & 2.7
-	\\
-	\textit{Anopheles} & \textit{albitarsis s.l.} & 2.6 & 4.0 & 5.8 & 9.2 & 19.9 & 8.2 & 9.5 \\
-	\textit{Anopheles} & \textit{albimanus} & 2.7 & 4.1 & 5.5 & 8.0 & 15.0 & 6.8 & 4.8 \\
-	\textit{Anopheles} & \textit{nuneztovari s.l.} & 2.6 & 3.6 & 4.5 & 6.1 & 11.4 & 5.5 & 3.8 \\
+	\textit{Anopheles} & \textit{albitarsis s.l.} & 1.2 & 3.1 & 5.4 & 9.9 & 26.9 & 9.3 & 18.6 \\
+	\textit{Anopheles} & \textit{nuneztovari s.l.} & 1.0 & 2.3 & 3.9 & 6.4 & 14.7 & 5.6 & 7.6 \\
+	\textit{Anopheles} & \textit{darlingi} & 0.9 & 2.4 & 4.5 & 8.1 & 22.0 & 7.4 & 13.6 \\
+	\textit{Anopheles} & \textit{albimanus} & 0.9 & 2.0 & 3.3 & 5.4 & 12.0 & 4.5 & 5.3 \\
+	\textit{Anopheles} & \textit{aquasalis} & 1.0 & 2.3 & 3.9 & 6.3 & 14.4 & 5.6 & 12.9 \\
+	\textit{Anopheles} & \textit{aquasalis s.l.} & 0.8 & 1.6 & 2.4 & 3.8 & 8.2 & 3.3 & 4.1 \\
 \end{tabular}
-\caption{\textbf{Posterior summaries of mean female mosquito lifespan in terms of chronological time from Detinova-type dissection.} The 5\%, 25\%, 50\%, 75\% and 95\% columns indicate the respective quantiles of the posterior distribution for mean lifespan, and the `Mean' and `Std.
-dev.' columns indicate the posterior mean and standard deviation of mean lifespan.}
+\caption{\textbf{Detinova dissection: mean female mosquito lifespan estimates in terms of chronological time.} The 5\%, 25\%, 50\%, 75\% and 95\% columns indicate the respective quantiles of the posterior distribution for mean lifespan, and the `Mean' and `Std.
+dev.' columns indicate the posterior mean and standard deviation of mean lifespan.
+Units of lifespan are days.}
 \label{tab:detinova_lifespan_chronological}%

### 2020-12-06 writeUp/supplementary_materials_methods.tex
very nearly at a draft

@@
 
-\subsubsection{Impact of insecticide interventions on lifespan}\label{sec:detinova_insecticide}
-To determine the impact of insecticidal interventions on lifespan, we modified the model described in \S\ref{sec:detinova_datamodel} to estimate effects at the species level.
+\subsection{Model estimation by MCMC and cross validation}
+The statistical models were fit to data using Stan's NUTS algorithm \cite{carpenter2016stan}.
+In all cases, models were run for 4000 iterations on each of 12 chains, with the first half of the iterations discarded as warm-up.
+We monitored diagnosed convergence as per \cite{gelman1992inference}, with $\hat{R}<1.1$ on all parameters.
+In all model runs, there were fewer than a handful of divergent transitions.
+
+To evaluate model fit to data, we used K-fold cross validation with 4 folds for each analysis.
+This means that, in a run, approximately $3/4$ of data was used to train models, and the remaining data used to evaluate out-of-sample predictive capability through the log-likelihood.
+To evaluate out-of-sample predictive log-likelihood for a given data point, we drew data-level parameters $\mathcal{P}_i^{\{s\}}$ from the relevant top-level distribution; for example, for a species-level fit, we drew data-level parameters from the species-level distribution relevant to the species for the data point in question.
+
+The folds were chosen so that, at all levels of analysis, there was always at least a handful of data points belonging to a particular group in the training set.
+This meant that, to generate meaningful comparisons between the species-level, complex-level and continent-level models, we selected only those data points with at least 5 observations for each species and complex.
+This reduced the number of data points used for model comparison versus those used for model estimation.
+For Africa, we had $n=217$ data points; for the Americas, there were $n=107$; and for Asia, there were $n=108$.
+
+For the comparison between continent-level grouping and overall, we were able to use more data points since we were less constrained by the minimal numbers of data belonging to each group.
+This meant we had $n=1126$ data points, with $n=679$ for Africa, $n=126$ for the Americas, and $n=321$, for Asia.
+
+\subsection{Model checking}
+To check the fit of models to the data, we simulated from the posterior predictive distribution for each grouping used in analysis.
+This means that we have two sets of posterior predictive figures: those where the check was performed at the species level; and those corresponding to the complex or subgroup level.
+In each of these figures, we plot observed empirical cumulative distribution function (eCDF) for the proportion parous versus the equivalent from each posterior predictive dataset.
+
+\begin{itemize}
+	\item \textbf{Africa}: the complex-level posterior predictive distribution graph is shown in Figure \ref{fig:detinova_ecdf_africa_grouped}; the species-level graph in Figure \ref{fig:detinova_ecdf_africa_species}.
+	\item \textbf{Americas}: the complex-level posterior predictive distribution graph is shown in Figure \ref{fig:detinova_ecdf_americas_grouped}; the species-level graph in Figure \ref{fig:detinova_ecdf_americas_species}.
+	\item \textbf{Asia}: the complex-level posterior predictive distribution graph is shown in Figure \ref{fig:detinova_ecdf_asia_grouped}; the species-level graph in Figure \ref{fig:detinova_ecdf_asia_species}.
+\end{itemize}
+
+In all cases, the eCDF for the actual data was bounded by the range of the posterior predictive eCDFs. By this metric, we conclude that the model seems to represent the data well.
+
+\begin{figure}[ht]
+	\centerline{\includegraphics[width=0.7\textwidth]{./Figure_files/detinova_ecdf_africa_grouped.pdf}}
+	\caption{\textbf{Detinova dissection: posterior predictive check for Africa, at the complex-level.} Plots show the empirical cumulative distribution functions for observations (orange) and posterior predictive simulations (green).
+In each case, we ran 4 Markov chains, with 200 iterations per chain; discarding the first 100 iterations as warm-up.}\label{fig:detinova_ecdf_africa_grouped}
+\end{figure}
+
+\begin{figure}[ht]
+	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/detinova_ecdf_africa_species.pdf}}
+	\caption{\textbf{Detinova dissection: posterior predictive check for Africa, at the species-level.} Plots show the empirical cumulative distribution functions for observations (orange) and posterior predictive simulations (green).
+In each case, we ran 4 Markov chains, with 200 iterations per chain; discarding the first 100 iterations as warm-up.}\label{fig:detinova_ecdf_africa_species}
+\end{figure}
+
+\begin{figure}[ht]
+	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/detinova_ecdf_americas_grouped.pdf}}
+	\caption{\textbf{Detinova dissection: posterior predictive check for the Americas, at the complex-level.} Plots show the empirical cumulative distribution functions for observations (orange) and posterior predictive simulations (green).
+In each case, we ran 4 Markov chains, with 200 iterations per chain; discarding the first 100 iterations as warm-up.}\label{fig:detinova_ecdf_americas_grouped}
+\end{figure}
+
+\begin{figure}[ht]
+	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/detinova_ecdf_americas_species.pdf}}
+	\caption{\textbf{Detinova dissection: posterior predictive check for the Americas, at the species-level.} Plots show the empirical cumulative distribution functions for observations (orange) and posterior predictive simulations (green).
+In each case, we ran 4 Markov chains, with 200 iterations per chain; discarding the first 100 iterations as warm-up.}\label{fig:detinova_ecdf_americas_species}
+\end{figure}
+
+\begin{figure}[ht]
+	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/detinova_ecdf_asia_grouped.pdf}}
+	\caption{\textbf{Detinova dissection: posterior predictive check for Asia, at the complex-level.} Plots show the empirical cumulative distribution functions for observations (orange) and posterior predictive simulations (green).
+In each case, we ran 4 Markov chains, with 200 iterations per chain; discarding the first 100 iterations as warm-up.}\label{fig:detinova_ecdf_asia_grouped}
+\end{figure}
+
+\begin{figure}[ht]
+	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/detinova_ecdf_asia_species.pdf}}
+	\caption{\textbf{Detinova dissection: posterior predictive check for Asia, at the species-level.} Plots show the empirical cumulative distribution functions for observations (orange) and posterior predictive simulations (green).
+In each case, we ran 4 Markov chains, with 200 iterations per chain; discarding the first 100 iterations as warm-up.}\label{fig:detinova_ecdf_asia_species}
+\end{figure}
+
+\subsection{Association of insecticide interventions on lifespan}\label{sec:detinova_insecticide}
+To determine the association of insecticidal interventions with lifespan, we modified the model described in \S\ref{sec:detinova_datamodel} to estimate effects at the species level.
 To do so, we allowed a logit-linear impact of insecticide.
@@
 The model to estimate this is shown below (where the variable metering the impact of insecticide at the observational level is `delta').
+This model was estimated using 1000 iterations per chain (500 of which were warm-up) across 12 chains.
 
@@
 
-\subsection{Model estimation by MCMC and cross validation}
-The statistical models were fit to data using Stan's NUTS algorithm \cite{carpenter2016stan}.
-In all cases, models were run for 4000 iterations on each of 12 chains, with the first half of the iterations discarded as warm-up.
-We monitored diagnosed convergence as per \cite{gelman1992inference}, with $\hat{R}<1.1$ on all parameters.
-In all model runs, there were fewer than a handful of divergent transitions.
-
-To evaluate model fit to data, we used K-fold cross validation with 4 folds for each analysis.
-This means that, in a run, approximately $3/4$ of data was used to train models, and the remaining data used to evaluate out-of-sample predictive capability through the log-likelihood.
-To evaluate out-of-sample predictive log-likelihood for a given data point, we drew data-level parameters $\mathcal{P}_i^{\{s\}}$ from the relevant top-level distribution; for example, for a species-level fit, we drew data-level parameters from the species-level distribution relevant to the species for the data point in question.
-
-The folds were chosen so that, at all levels of analysis, there was always at least a handful of data points belonging to a particular group in the training set.
-This meant that, to generate meaningful comparisons between the species-level, complex-level and continent-level models, we selected only those data points with at least 5 observations for each species and complex.
-This reduced the number of data points used for model comparison versus those used for model estimation.
-For Africa, we had $n=217$ data points; for the Americas, there were $n=107$; and for Asia, there were $n=108$.
-
-For the comparison between continent-level grouping and overall, we were able to use more data points since we were less constrained by the minimal numbers of data belonging to each group.
-This meant we had $n=1126$ data points, with $n=679$ for Africa, $n=126$ for the Americas, and $n=321$, for Asia.
-
+\subsection{Association of weather covariates with temperature}
 To investigate the effect of measures of temperature, we extracted daily data from the ERA-interim re-analysis by ECMWF \citep{dee2011era} through their website using the Python API.
@@
 
-For this analysis, we included only those studies where the start month listed in the database coincided with the end date - meaning the study took place within a given month - resulting in $n=742$ data points.
+For this analysis, we included only those studies where the start month listed in the database coincided with the end date, meaning the study took place within a given month.
+We also omitted those data points where insecticide use was recorded.
+To ensure that we had sufficient data points for each fold used for model comparison, we considered only those species with over 100 observations.
+This resulted in $n=366$ observations across 2 species complexes: \textit{A.
+gambiae s.l.} ($n=257$) and \textit{A.
+funestus s.l.} ($n=109$).
+
 The temperature data was extracted for the GPS position of each study ($\pm 0.1 \text{ degrees}$) for each day of the month over which mosquito sampling took place, with a grid size of 0.05 degrees.
@@
 	\item Average two meter temperature (average of mean daily data for each grid point),
-	\item Minimum two meter temperature (average of minimum daily data for each grid point),
-	\item Maximum two meter temperature (average of maximum daily data for each grid point),
-	\item Standard deviation in two meter temperature (average of standard deviation in temperature for each grid point),
-	\item Daily range in two meter temperature (average of maximum minus minimum daily temperature for each grid point).
+	\item Daily range in two meter temperature (average of maximum minus minimum daily temperature for each grid point). This likely reflects the average day-night temperature range.
 \end{itemize}
 
-\subsection{Model checking}
-To check the fit of models to the data, we simulated from the posterior predictive distribution for each grouping used in analysis.
-This means that we have two sets of posterior predictive figures: those where the check was performed at the species level; and those corresponding to the complex or subgroup level.
-In each of these figures, we plot observed empirical cumulative distribution function (eCDF) for the proportion parous versus the equivalent from each posterior predictive dataset.
-
-\begin{itemize}
-	\item \textbf{Africa}: the complex-level posterior predictive distribution graph is shown in Figure \ref{fig:detinova_ecdf_africa_grouped}; the species-level graph in Figure \ref{fig:detinova_ecdf_africa_species}.
-	\item \textbf{Americas}: the complex-level posterior predictive distribution graph is shown in Figure \ref{fig:detinova_ecdf_americas_grouped}; the species-level graph in Figure \ref{fig:detinova_ecdf_americas_species}.
-	\item \textbf{Asia}: the complex-level posterior predictive distribution graph is shown in Figure \ref{fig:detinova_ecdf_asia_grouped}; the species-level graph in Figure \ref{fig:detinova_ecdf_asia_species}.
-\end{itemize}
-
-In all cases, the eCDF for the actual data was bounded by the range of the posterior predictive eCDFs. By this metric, we conclude that the model seems to represent the data well.
-
-\begin{figure}[ht]
-	\centerline{\includegraphics[width=0.7\textwidth]{./Figure_files/detinova_ecdf_africa_grouped.pdf}}
-	\caption{\textbf{Detinova dissection: posterior predictive check for Africa, at the complex-level.} Plots show the empirical cumulative distribution functions for observations (orange) and posterior predictive simulations (green).
-In each case, we ran 4 Markov chains, with 200 iterations per chain; discarding the first 100 iterations as warm-up.}\label{fig:detinova_ecdf_africa_grouped}
-\end{figure}
-
-\begin{figure}[ht]
-	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/detinova_ecdf_africa_species.pdf}}
-	\caption{\textbf{Detinova dissection: posterior predictive check for Africa, at the species-level.} Plots show the empirical cumulative distribution functions for observations (orange) and posterior predictive simulations (green).
-In each case, we ran 4 Markov chains, with 200 iterations per chain; discarding the first 100 iterations as warm-up.}\label{fig:detinova_ecdf_africa_species}
-\end{figure}
-
-\begin{figure}[ht]
-	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/detinova_ecdf_americas_grouped.pdf}}
-	\caption{\textbf{Detinova dissection: posterior predictive check for the Americas, at the complex-level.} Plots show the empirical cumulative distribution functions for observations (orange) and posterior predictive simulations (green).
-In each case, we ran 4 Markov chains, with 200 iterations per chain; discarding the first 100 iterations as warm-up.}\label{fig:detinova_ecdf_americas_grouped}
-\end{figure}
-
-\begin{figure}[ht]
-	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/detinova_ecdf_americas_species.pdf}}
-	\caption{\textbf{Detinova dissection: posterior predictive check for the Americas, at the species-level.} Plots show the empirical cumulative distribution functions for observations (orange) and posterior predictive simulations (green).
-In each case, we ran 4 Markov chains, with 200 iterations per chain; discarding the first 100 iterations as warm-up.}\label{fig:detinova_ecdf_americas_species}
-\end{figure}
-
-\begin{figure}[ht]
-	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/detinova_ecdf_asia_grouped.pdf}}
-	\caption{\textbf{Detinova dissection: posterior predictive check for Asia, at the complex-level.} Plots show the empirical cumulative distribution functions for observations (orange) and posterior predictive simulations (green).
-In each case, we ran 4 Markov chains, with 200 iterations per chain; discarding the first 100 iterations as warm-up.}\label{fig:detinova_ecdf_asia_grouped}
-\end{figure}
-
-\begin{figure}[ht]
-	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/detinova_ecdf_asia_species.pdf}}
-	\caption{\textbf{Detinova dissection: posterior predictive check for Asia, at the species-level.} Plots show the empirical cumulative distribution functions for observations (orange) and posterior predictive simulations (green).
-In each case, we ran 4 Markov chains, with 200 iterations per chain; discarding the first 100 iterations as warm-up.}\label{fig:detinova_ecdf_asia_species}
-\end{figure}
+Each of these two models were estimated in Stan using 1000 iterations per chain (500 of each were warm-up) across 4 chains and thinned samples by a factor of 20.
+To determine whether each covariate was predictive for lifespan, we undertook K-fold cross-validation (using 4 folds with a given fold holding roughly equal numbers of each species).
 

### 2020-12-10 writeUp/supplementary_materials.tex
draft sent to ace and charles

@@
 
-\begin{figure}[h]
-	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_lifeSpanVsTemperature.pdf}}
-	\caption{\textbf{MRR: mean mosquito lifespan versus average monthly air temperature.} The markers show the median posterior lifespan estimates for each time series with the lower and upper bounds indicating the 25\% and 75\% quantiles, respectively.
-The black line shows a regression line with linear and quadratic terms estimated using the median posterior lifetimes, with the grey shading indicating 95\% confidence intervals.
-All estimates of lifespan were obtained using the non-hierarchical exponential survival model.}\label{fig:mrr_temperature}
-\end{figure}
-
-\begin{figure}[h]
-	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_ThreeSpeciesVersusTemperature.pdf}}
-	\caption{\textbf{MRR: mean mosquito lifespan versus average monthly air temperature by species.} Here, we include only those four species with the most data.
-The markers show the median posterior lifespan estimates for each time series with the lower and upper bounds indicating the 25\% and 75\% quantiles, respectively.
-The black line shows a regression line with linear and quadratic terms estimated using the median posterior lifetimes, with the grey shading indicating 95\% confidence intervals.
-All estimates of lifespan were obtained using the non-hierarchical exponential survival model.}\label{fig:mrr_ThreeSpeciesVersusTemperature}
-\end{figure}
-
 \begin{table}[htbp]
@@
 \end{table}
-
-\begin{figure}[ht]
-	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/gonotrophic_data.pdf}}
-	\caption{\textbf{Raw gonotrophic cycle duration estimates for A.
-the 1st cycle and B.
-subsequent cycles.} When present, the confidence intervals represent our estimates of the 95\% confidence interval on gonotrophic cycle duration (see Section \ref{sec:dissection_gonotrophicMethod} for a discussion of how we handle this issue); if our estimate of the lower limit was below zero, we deemed this unbiological and set it to 1 day for this graph.
-If no statement of uncertainty was given in the estimates then we indicate this by a triangle marker.
-The studies are ordered within each Genus according to the central estimate of gonotrophic duration.}\label{fig:dissection_gonotrophicCycleRaw}
-\end{figure}
 

### 2020-12-10 writeUp/supplementary_materials_methods.tex
draft sent to ace and charles

@@
 The MRR database in \cite{guerra2014global} contains 393 individual time-series, along with meta-data for a range of factors for each experiment (for example, species, geography, and date of study).
-We  cleaned the data, making a number of amendments to it (see Section \ref{sec:amendments}), and transformed the data into a form amenable for our statistical model.
+We  cleaned the data, making a number of amendments to it (see Section \ref{sec:amendments}) and transformed the data into a form amenable for our statistical model.
 We removed time-series with fewer than six separate recapture observations, resulting in 238 time series.
@@
 We have assumed that the number recaptured were as stated in the database on those days (missing, because no recapture efforts were made on those days).
-	\item 586: It was unclear from database whether the missing observations are due to lack of recapture, or no effort.
+	\item 586: It was unclear from the database whether the missing observations were due to lack of recapture, or no recapture effort being made on those days.
 Have assumed that zero mosquitoes were captured on these days (the paper could not be accessed).
-	\item 342: Changed number of released from 6,000 to 12,000 following the original publication \cite[]{bryan1991anopheles}.
+	\item 342: Changed number of released from 6000 to 12,000 following the original publication \cite[]{bryan1991anopheles}.
 	\item 246: Moved the observations from captures to recaptures following the original publication \cite[]{arredondo1998gonotrophic}.
@@
 \end{itemize}
+
+\begin{figure}[h]
+	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_sexReleasesOverTime.pdf}}
+	\caption{\textbf{MRR: time-series by date and sex of released mosquitoes.} The numbers represent by-decade totals of the 232 time-series we used in the main analysis.}
+	\label{fig:mrr_sexReleasesOverTime}
+\end{figure}
 
@@
 
-\begin{figure}[h]
-	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_sexReleasesOverTime.pdf}}
-	\caption{\textbf{MRR: time-series by date and sex of released mosquitoes.} The numbers represent by-decade totals of the 232 time-series we used in the main analysis.}
-	\label{fig:mrr_sexReleasesOverTime}
-\end{figure}
-
 \subsection{Statistical model}\label{sec:mrr_statistical}
@@
 where $Y(t-1)$ is the cumulative number of mosquitoes caught on all days before $t$, $S(t)$ is the probability that an individual mosquito survives and remains in the study area until time $t$, $\psi$ is the daily recapture probability for an individual mosquito which is assumed constant through time, and $\kappa$ is the time-independent shape parameter of the negative binomial distribution that controls the extent of variance in recapture rate likely due mostly to environmental heterogeneity.
-We chose a parameterisation of the negative binomial such that its mean is given by $\mu(t) = (N_{R} - Y(t-1)) S(t) \psi$, and its variance by $\sigma(t)^2 = \mu(t) + \frac{\mu(t)^2}{\kappa}$.
+We chose a parameterisation of the negative binomial such that its mean is given by $\mu(t) = (N_{R} - Y(t-1)) S(t) \psi$ and its variance by $\sigma(t)^2 = \mu(t) + \frac{\mu(t)^2}{\kappa}$.
 
@@
 where the factor of $(1-\psi)^{t-1}$ represents the probability that a mosquito is \textit{not} recaptured on a previous day.
-Where experiments consisted of two or more releases occurring at distinct points in time we assumed that recaptures of individual mosquitoes from either batch were independent of one another, although with the same sampling parameters ($\psi$ and $\kappa$).
+Where experiments consisted of two or more releases occurring at distinct points in time, we assumed that recaptures of individual mosquitoes from either batch were independent of one another, although with the same sampling parameters ($\psi$ and $\kappa$).
 This results in an overall mean composed of the sum of those from all $q$ releases,
@@
 where $\lambda$ is the sum of the rates of death and dispersal from the study area.
-To estimate the lifespan of mosquitoes we use this model, partly because of the limited evidence we find for senescence.
+To estimate the lifespan of mosquitoes, we use this model, partly because of the limited evidence we find for senescence.
 
@@
 	\centerline{\includegraphics[width=1.25\textwidth]{./Figure_files/mrr_individualTimeSeries_priors.pdf}}
-	\caption{\textbf{MRR: prior probability distributions used for lifespan (A), recapture probability (B) and dispersal parameter (C) in the individual time-series analysis.} Note: the prior for the recapture probability was truncated at $\psi=0.1$ since most MRR day one recapture fractions were significantly below this threshold.}
+	\caption{\textbf{MRR: prior probability distributions used for lifespan (A), recapture probability (B) and dispersal parameter (C) in the individual time-series analysis.} Note: the prior for the recapture probability was truncated at $\psi=0.1$ since most recapture fractions on the first day were significantly below this threshold.}
 	\label{fig:mrr_individualTimeSeries_priors}
@@
 \subsection{Testing for age-dependent mortality in wild mosquitoes}\label{sec:mrr_age_dependence}
-Previous work has found evidence of age-dependent mortality in lab populations \citep{styer2007mosquitoes,dawes2009anopheles}, and less-commonly in wild mosquitoes \citep{clements1981analysis,harrington2008age}.
+Previous work has found evidence of age-dependent mortality in lab populations \citep{styer2007mosquitoes,dawes2009anopheles} and less commonly in wild mosquitoes \citep{clements1981analysis,harrington2008age}.
 Furthermore, modelling work has examined the implications of departures from a constant risk of mortality \citep{styer2007mosquitoes,hancock2009age,novoseltsev2012age}.
@@
 At the end of all runs for lifespan estimation, $\hat{R}<1.1$ for all model parameters.
-For the age-dependent cases, there remained a handful of parameters where $\hat{R}>1.1$ after running simulations for 3000 iterations but this did not affect our inferences, since repeated runs demonstrated the same pattern.
+For the age-dependent cases, there remained a handful of parameters where $1.2>\hat{R}>1.1$ after running simulations for 3000 iterations but this did not affect our inferences, since repeated runs demonstrated the same pattern.
 We also ensured that across each MCMC run, the number of divergent iterations (that can bias the MCMC away from the true posterior density) was minimal.
@@
 Raw estimates of gonotrophic cycle duration were obtained for species across three different genera (\textit{Aedes}, \textit{Anopheles} and \textit{Culex}), although there was a bias towards \textit{Anopheles} with $n=92$ estimates (Fig.
-\ref{S-fig:dissection_gonotrophicCycleRaw}).
+\ref{fig:dissection_gonotrophicCycleRaw}).
+
+\begin{figure}[ht]
+	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/gonotrophic_data.pdf}}
+	\caption{\textbf{Raw gonotrophic cycle duration estimates for A.
+the 1st cycle and B.
+subsequent cycles.} When present, the confidence intervals represent our estimates of the 95\% confidence interval on gonotrophic cycle duration (see Section \ref{sec:dissection_gonotrophicMethod} for a discussion of how we handle this issue); if our estimate of the lower limit was below zero, we deemed this unbiological and set it to 1 day for this graph.
+If no statement of uncertainty was given in the estimates then we indicate this by a triangle marker.
+The studies are ordered within each Genus according to the central estimate of gonotrophic duration.}\label{fig:dissection_gonotrophicCycleRaw}
+\end{figure}
 
@@
 
-In the using the above methodology to convert from physiological to calendar age, we implicitly assume that all gonotrophic cycles after the first are of the same length (i.e.
+In using the above methodology to convert from physiological to calendar age, we implicitly assume that all gonotrophic cycles after the first are of the same length (i.e.
 we choose one $G_{2i}$ per sample).
@@
 Reading the paper, it looks like there is a mistake in their Table 5, and that the parous number should be 131 and, as such, we changed this number in the database; considering the entry for \cite{bockarie46service}, there was a difference of around 5\%.
-We could not access the paper here, so assumed that the issue was with the `parity percent' entry.
+We could not access the paper here so assumed that the issue was with the `parity percent' entry.
 	\item Americas: no differences exceeding 1\%.
 	\item Asia: considering the entry for \cite{kitthawee1992relationship}, there was a difference of around 3\%.
-Reading the paper, it looks like there was a (small) error in the database, and the parous number should be 35 (rather than 36) - this adjustment was made for the ensuing analysis.
+Reading the paper, it looks like there was a (small) error in the database, and the parous number should be 35 (rather than 36) -- this adjustment was made for the ensuing analysis.
 \end{itemize}

### 2021-02-15 writeUp/supplementary_materials.tex
updated order as per Charles comments

@@
 \begin{figure}[h]
-	\centerline{\includegraphics[width=1.3\textwidth]{./Figure_files/dissection_individualEstimates_allSpecies.pdf}}
-	\caption{\textbf{Polovodova dissection: individual time-series estimates of adult mosquito mean lifespan.} The middle line in each box shows the median estimates.
-The left and right box whiskers show the 25\%, and 75\% posterior quantiles respectively.
-All estimates were obtained using the non-hierarchical exponential survival model.}
-	\label{fig:dissection_individualEstimates_allSpecies}
-\end{figure}
-
-\begin{figure}[h]
 	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/detinova_insecticide.pdf}}
@@
 
-\begin{table}[htbp]
-	\begin{tabular}{l|l|l|l|l|l|l|l|l}
-		\textbf{Genus} & \textbf{Species} & \textbf{5$\%$} & \textbf{25$\%$} & \textbf{50$\%$} & \textbf{75$\%$} & \textbf{95\%} &  \textbf{Mean} & \textbf{Std. dev.}\\
-		\hline
-		\textit{Anopheles} & \textit{sergentii} & 1.0 & 1.9 & 2.5 & 3.4 & 6.2 & 3.0 & 2.5 \\
-		\textit{Anopheles} & \textit{gambiae s.l.} & 0.7 & 1.2 & 1.9 & 2.9 & 5.4 & 2.4 & 1.9 \\
-		\textit{Culex} & \textit{thalassius} & 1.1 & 1.6 & 1.8 & 2.1 & 2.9 & 1.9 & 1.0 \\
-		\textit{Anopheles} & \textit{farauti s.l.} & 1.5 & 1.6 & 1.7 & 1.7 & 1.9 & 1.7 & 0.1 \\
-		\textit{Anopheles} & \textit{minimus} & 0.7 & 1.2 & 1.7 & 2.2 & 4.1 & 2.0 & 2.2 \\
-		\textit{Anopheles} & \textit{maculipennis} & 0.8 & 1.1 & 1.4 & 1.8 & 2.9 & 1.6 & 0.7 \\
-		\textit{Culex} & \textit{pipiens} & 1.1 & 1.3 & 1.4 & 1.5 & 1.7 & 1.4 & 0.2 \\
-		\textit{Mansonia} & \textit{uniformis} & 0.9 & 1.1 & 1.1 & 1.2 & 1.4 & 1.1 & 0.2 \\
-		\textit{Anopheles} & \textit{rivulorum} & 0.8 & 1.0 & 1.1 & 1.1 & 1.4 & 1.1 & 0.2 \\
-		\textit{Anopheles} & \textit{melas} & 0.8 & 1.0 & 1.0 & 1.1 & 1.3 & 1.0 & 0.2 \\
-		\textit{Anopheles} & \textit{culicifacies} & 0.9 & 1.0 & 1.0 & 1.1 & 1.2 & 1.0 & 0.1 \\
-		\textit{Anopheles} & \textit{subpictus} & 0.6 & 0.9 & 1.0 & 1.2 & 2.1 & 1.2 & 2.0 \\
-		\textit{Aedes} & \textit{aegypti} & 0.5 & 0.8 & 1.0 & 1.2 & 2.0 & 1.2 & 3.0 \\
-		\textit{Anopheles} & \textit{quadrimaculatus} & 0.7 & 0.9 & 1.0 & 1.1 & 1.3 & 1.0 & 0.2 \\
-		\textit{Anopheles} & \textit{stephensi} & 0.6 & 0.8 & 1.0 & 1.1 & 1.6 & 1.0 & 0.4 \\
-		\textit{Anopheles} & \textit{darlingi} & 0.7 & 0.8 & 0.9 & 1.0 & 1.3 & 1.0 & 0.2 \\
-		\textit{Culex} & \textit{quinquefasciatus} & 0.4 & 0.7 & 0.9 & 1.3 & 2.2 & 1.1 & 0.6 \\
-		\textit{Aedes} & \textit{polynesiensis} & 0.8 & 0.9 & 0.9 & 0.9 & 1.0 & 0.9 & 0.1 \\
-		\textit{Culex} & \textit{tritaeniorhynchus} & 0.3 & 0.6 & 0.9 & 1.3 & 4.2 & 2.5 & 30.8 \\
-		\textit{Culex} & \textit{annulirostris} & 0.5 & 0.8 & 0.9 & 1.0 & 1.4 & 0.9 & 0.5 \\
-		\textit{Aedes} & \textit{samoanus} & 0.6 & 0.7 & 0.8 & 0.9 & 1.0 & 0.8 & 0.2 \\
-		\textit{Aedes} & \textit{sollicitans} & 0.4 & 0.5 & 0.6 & 0.8 & 1.3 & 0.7 & 0.5 \\
-		\textit{Aedes} & \textit{vexans} & 0.4 & 0.5 & 0.6 & 0.7 & 0.9 & 0.6 & 0.7 \\
-		\textit{Anopheles} & \textit{cruzii} & 0.3 & 0.5 & 0.5 & 0.7 & 1.1 & 0.6 & 0.4 \\
-		\textit{Anopheles} & \textit{bellator} & 0.3 & 0.4 & 0.5 & 0.6 & 1.0 & 0.6 & 1.8 \\
-		\hline
-		\textit{Anopheles} & \textit{} & 0.6 & 1.0 & 1.4 & 1.9 & 3.2 & 1.6 & 0.9 \\
-		\textit{Mansonia} & \textit{} & 0.9 & 1.1 & 1.1 & 1.2 & 1.4 & 1.1 & 0.2 \\
-		\textit{Culex} & \textit{} & 0.5 & 0.8 & 1.0 & 1.4 & 2.3 & 1.2 & 0.6 \\
-		\textit{Aedes} & \textit{} & 0.6 & 0.7 & 0.8 & 0.9 & 1.1 & 0.8 & 0.2 \\
-		\hline
-		\textbf{Overall} & \textit{} & 0.5 & 0.8 & 1.2 & 1.7 & 2.7 & 1.3 & 0.7 \\
-	\end{tabular}
-	\caption{\textbf{Polovodova dissection: mean female mosquito lifespan in terms of gonotrophic cycles.} The 5\%, 25\%, 50\%, 75\% and 95\% columns indicate the respective quantiles of the posterior distribution for mean lifespan, and the `Mean' and `Std.
-dev.' columns indicate the posterior mean and standard deviation of mean lifespan.}
-	\label{tab:dissection_estimated_lifespans}%
-\end{table}
+\begin{figure}[h]
+	\centerline{\includegraphics[width=1.3\textwidth]{./Figure_files/dissection_individualEstimates_allSpecies.pdf}}
+	\caption{\textbf{Polovodova dissection: individual time-series estimates of adult mosquito mean lifespan.} The middle line in each box shows the median estimates.
+The left and right box whiskers show the 25\%, and 75\% posterior quantiles respectively.
+All estimates were obtained using the non-hierarchical exponential survival model.}
+	\label{fig:dissection_individualEstimates_allSpecies}
+\end{figure}
 
@@
 
-\begin{table}[htbp!]
+\begin{table}[htbp]
 	\begin{tabular}{l|l|l|l|l|l|l|l|l}
@@
 		\hline
-		\textit{Culex} & \textit{thalassius} & 5.7 & 8.2 & 9.4 & 11.0 & 15.6 & 10.1 & 5.2 \\
-		\textit{Anopheles} & \textit{sergentii} & 3.7 & 6.4 & 8.5 & 11.5 & 20.6 & 10.1 & 8.1 \\
-		\textit{Culex} & \textit{pipiens} & 5.7 & 6.6 & 7.2 & 7.8 & 8.9 & 7.2 & 1.0 \\
-		\textit{Anopheles} & \textit{gambiae s.l.} & 2.4 & 4.3 & 6.4 & 9.8 & 18.3 & 8.0 & 6.3 \\
-		\textit{Anopheles} & \textit{farauti s.l.} & 4.8 & 5.5 & 5.9 & 6.3 & 6.9 & 5.9 & 0.6 \\
-		\textit{Anopheles} & \textit{minimus} & 2.7 & 4.4 & 5.8 & 7.7 & 13.7 & 6.9 & 8.1 \\
-		\textit{Anopheles} & \textit{maculipennis} & 2.8 & 4.0 & 5.1 & 6.4 & 9.6 & 5.5 & 2.3 \\
-		\textit{Culex} & \textit{quinquefasciatus} & 2.0 & 3.4 & 4.8 & 6.9 & 11.7 & 5.6 & 3.3 \\
-		\textit{Aedes} & \textit{aegypti} & 2.3 & 3.7 & 4.7 & 5.6 & 9.3 & 5.6 & 12.3 \\
-		\textit{Culex} & \textit{annulirostris} & 2.7 & 3.9 & 4.6 & 5.4 & 7.5 & 4.9 & 2.4 \\
-		\textit{Culex} & \textit{tritaeniorhynchus} & 1.6 & 3.2 & 4.6 & 7.2 & 22.8 & 13.0 & 174.0 \\
-		\textit{Mansonia} & \textit{uniformis} & 3.4 & 4.0 & 4.4 & 4.9 & 5.5 & 4.5 & 0.7 \\
-		\textit{Aedes} & \textit{polynesiensis} & 3.2 & 3.7 & 4.1 & 4.5 & 5.1 & 4.1 & 0.6 \\
-		\textit{Anopheles} & \textit{rivulorum} & 2.8 & 3.4 & 3.8 & 4.2 & 5.0 & 3.9 & 0.7 \\
-		\textit{Anopheles} & \textit{melas} & 2.8 & 3.3 & 3.7 & 4.1 & 4.8 & 3.8 & 0.7 \\
-		\textit{Anopheles} & \textit{culicifacies s.l.} & 2.8 & 3.3 & 3.7 & 4.1 & 4.7 & 3.7 & 0.6 \\
-		\textit{Anopheles} & \textit{subpictus s.l.} & 2.2 & 3.1 & 3.7 & 4.5 & 7.0 & 4.3 & 6.6 \\
-		\textit{Aedes} & \textit{samoanus} & 2.7 & 3.2 & 3.6 & 4.0 & 4.9 & 3.7 & 1.0 \\
-		\textit{Anopheles} & \textit{quadrimaculatus} & 2.4 & 3.1 & 3.5 & 4.0 & 4.8 & 3.6 & 0.8 \\
-		\textit{Anopheles} & \textit{stephensi} & 2.1 & 2.9 & 3.5 & 4.2 & 5.7 & 3.7 & 1.5 \\
-		\textit{Anopheles} & \textit{darlingi} & 2.3 & 2.9 & 3.4 & 3.8 & 4.7 & 3.4 & 0.9 \\
-		\textit{Aedes} & \textit{sollicitans} & 1.6 & 2.3 & 2.9 & 3.6 & 6.1 & 3.3 & 2.5 \\
-		\textit{Aedes} & \textit{vexans} & 1.7 & 2.2 & 2.7 & 3.1 & 4.2 & 3.0 & 3.3 \\
-		\textit{Anopheles} & \textit{cruzii} & 1.2 & 1.6 & 2.0 & 2.5 & 4.0 & 2.2 & 1.3 \\
-		\textit{Anopheles} & \textit{bellator} & 1.0 & 1.4 & 1.7 & 2.1 & 3.6 & 2.1 & 5.4 \\
+		\textit{Anopheles} & \textit{sergentii} & 1.0 & 1.9 & 2.5 & 3.4 & 6.2 & 3.0 & 2.5 \\
+		\textit{Anopheles} & \textit{gambiae s.l.} & 0.7 & 1.2 & 1.9 & 2.9 & 5.4 & 2.4 & 1.9 \\
+		\textit{Culex} & \textit{thalassius} & 1.1 & 1.6 & 1.8 & 2.1 & 2.9 & 1.9 & 1.0 \\
+		\textit{Anopheles} & \textit{farauti s.l.} & 1.5 & 1.6 & 1.7 & 1.7 & 1.9 & 1.7 & 0.1 \\
+		\textit{Anopheles} & \textit{minimus} & 0.7 & 1.2 & 1.7 & 2.2 & 4.1 & 2.0 & 2.2 \\
+		\textit{Anopheles} & \textit{maculipennis} & 0.8 & 1.1 & 1.4 & 1.8 & 2.9 & 1.6 & 0.7 \\
+		\textit{Culex} & \textit{pipiens} & 1.1 & 1.3 & 1.4 & 1.5 & 1.7 & 1.4 & 0.2 \\
+		\textit{Mansonia} & \textit{uniformis} & 0.9 & 1.1 & 1.1 & 1.2 & 1.4 & 1.1 & 0.2 \\
+		\textit{Anopheles} & \textit{rivulorum} & 0.8 & 1.0 & 1.1 & 1.1 & 1.4 & 1.1 & 0.2 \\
+		\textit{Anopheles} & \textit{melas} & 0.8 & 1.0 & 1.0 & 1.1 & 1.3 & 1.0 & 0.2 \\
+		\textit{Anopheles} & \textit{culicifacies} & 0.9 & 1.0 & 1.0 & 1.1 & 1.2 & 1.0 & 0.1 \\
+		\textit{Anopheles} & \textit{subpictus} & 0.6 & 0.9 & 1.0 & 1.2 & 2.1 & 1.2 & 2.0 \\
+		\textit{Aedes} & \textit{aegypti} & 0.5 & 0.8 & 1.0 & 1.2 & 2.0 & 1.2 & 3.0 \\
+		\textit{Anopheles} & \textit{quadrimaculatus} & 0.7 & 0.9 & 1.0 & 1.1 & 1.3 & 1.0 & 0.2 \\
+		\textit{Anopheles} & \textit{stephensi} & 0.6 & 0.8 & 1.0 & 1.1 & 1.6 & 1.0 & 0.4 \\
+		\textit{Anopheles} & \textit{darlingi} & 0.7 & 0.8 & 0.9 & 1.0 & 1.3 & 1.0 & 0.2 \\
+		\textit{Culex} & \textit{quinquefasciatus} & 0.4 & 0.7 & 0.9 & 1.3 & 2.2 & 1.1 & 0.6 \\
+		\textit{Aedes} & \textit{polynesiensis} & 0.8 & 0.9 & 0.9 & 0.9 & 1.0 & 0.9 & 0.1 \\
+		\textit{Culex} & \textit{tritaeniorhynchus} & 0.3 & 0.6 & 0.9 & 1.3 & 4.2 & 2.5 & 30.8 \\
+		\textit{Culex} & \textit{annulirostris} & 0.5 & 0.8 & 0.9 & 1.0 & 1.4 & 0.9 & 0.5 \\
+		\textit{Aedes} & \textit{samoanus} & 0.6 & 0.7 & 0.8 & 0.9 & 1.0 & 0.8 & 0.2 \\
+		\textit{Aedes} & \textit{sollicitans} & 0.4 & 0.5 & 0.6 & 0.8 & 1.3 & 0.7 & 0.5 \\
+		\textit{Aedes} & \textit{vexans} & 0.4 & 0.5 & 0.6 & 0.7 & 0.9 & 0.6 & 0.7 \\
+		\textit{Anopheles} & \textit{cruzii} & 0.3 & 0.5 & 0.5 & 0.7 & 1.1 & 0.6 & 0.4 \\
+		\textit{Anopheles} & \textit{bellator} & 0.3 & 0.4 & 0.5 & 0.6 & 1.0 & 0.6 & 1.8 \\
 		\hline
-		\textit{Culex} & \textit{} & 2.5 & 4.0 & 5.4 & 7.6 & 12.3 & 6.2 & 3.2 \\
-		\textit{Anopheles} & \textit{} & 2.1 & 3.6 & 4.9 & 6.8 & 10.8 & 5.5 & 2.9 \\
-		\textit{Mansonia} & \textit{} & 3.3 & 4.0 & 4.5 & 4.9 & 5.6 & 4.5 & 0.8 \\
-		\textit{Aedes} & \textit{} & 2.5 & 3.3 & 3.8 & 4.3 & 5.2 & 3.8 & 0.9 \\
+		\textit{Anopheles} & \textit{} & 0.6 & 1.0 & 1.4 & 1.9 & 3.2 & 1.6 & 0.9 \\
+		\textit{Mansonia} & \textit{} & 0.9 & 1.1 & 1.1 & 1.2 & 1.4 & 1.1 & 0.2 \\
+		\textit{Culex} & \textit{} & 0.5 & 0.8 & 1.0 & 1.4 & 2.3 & 1.2 & 0.6 \\
+		\textit{Aedes} & \textit{} & 0.6 & 0.7 & 0.8 & 0.9 & 1.1 & 0.8 & 0.2 \\
 		\hline
-		\textit{Overall} & \textit{} & 2.0 & 3.3 & 4.7 & 6.4 & 10.2 & 5.2 & 2.6 \\
+		\textbf{Overall} & \textit{} & 0.5 & 0.8 & 1.2 & 1.7 & 2.7 & 1.3 & 0.7 \\
 	\end{tabular}
-	\caption{\textbf{Polovodova dissection: mean female mosquito lifespan estimates in terms of chronological time.} The 5\%, 25\%, 50\%, 75\% and 95\% columns indicate the respective quantiles of the posterior distribution for mean lifespan, and the `Mean' and `Std.
-dev.' columns indicate the posterior mean and standard deviation of mean lifespan.
-Units of lifespan are days.}
-	\label{tab:dissection_estimated_lifespans_chron}%
+	\caption{\textbf{Polovodova dissection: mean female mosquito lifespan in terms of gonotrophic cycles.} The 5\%, 25\%, 50\%, 75\% and 95\% columns indicate the respective quantiles of the posterior distribution for mean lifespan, and the `Mean' and `Std.
+dev.' columns indicate the posterior mean and standard deviation of mean lifespan.}
+	\label{tab:dissection_estimated_lifespans}%
 \end{table}
@@
 
+\begin{table}[htbp!]
+	\begin{tabular}{l|l|l|l|l|l|l|l|l}
+		\textbf{Genus} & \textbf{Species} & \textbf{5$\%$} & \textbf{25$\%$} & \textbf{50$\%$} & \textbf{75$\%$} & \textbf{95\%} &  \textbf{Mean} & \textbf{Std. dev.}\\
+		\hline
+		\textit{Culex} & \textit{thalassius} & 5.7 & 8.2 & 9.4 & 11.0 & 15.6 & 10.1 & 5.2 \\
+		\textit{Anopheles} & \textit{sergentii} & 3.7 & 6.4 & 8.5 & 11.5 & 20.6 & 10.1 & 8.1 \\
+		\textit{Culex} & \textit{pipiens} & 5.7 & 6.6 & 7.2 & 7.8 & 8.9 & 7.2 & 1.0 \\
+		\textit{Anopheles} & \textit{gambiae s.l.} & 2.4 & 4.3 & 6.4 & 9.8 & 18.3 & 8.0 & 6.3 \\
+		\textit{Anopheles} & \textit{farauti s.l.} & 4.8 & 5.5 & 5.9 & 6.3 & 6.9 & 5.9 & 0.6 \\
+		\textit{Anopheles} & \textit{minimus} & 2.7 & 4.4 & 5.8 & 7.7 & 13.7 & 6.9 & 8.1 \\
+		\textit{Anopheles} & \textit{maculipennis} & 2.8 & 4.0 & 5.1 & 6.4 & 9.6 & 5.5 & 2.3 \\
+		\textit{Culex} & \textit{quinquefasciatus} & 2.0 & 3.4 & 4.8 & 6.9 & 11.7 & 5.6 & 3.3 \\
+		\textit{Aedes} & \textit{aegypti} & 2.3 & 3.7 & 4.7 & 5.6 & 9.3 & 5.6 & 12.3 \\
+		\textit{Culex} & \textit{annulirostris} & 2.7 & 3.9 & 4.6 & 5.4 & 7.5 & 4.9 & 2.4 \\
+		\textit{Culex} & \textit{tritaeniorhynchus} & 1.6 & 3.2 & 4.6 & 7.2 & 22.8 & 13.0 & 174.0 \\
+		\textit{Mansonia} & \textit{uniformis} & 3.4 & 4.0 & 4.4 & 4.9 & 5.5 & 4.5 & 0.7 \\
+		\textit{Aedes} & \textit{polynesiensis} & 3.2 & 3.7 & 4.1 & 4.5 & 5.1 & 4.1 & 0.6 \\
+		\textit{Anopheles} & \textit{rivulorum} & 2.8 & 3.4 & 3.8 & 4.2 & 5.0 & 3.9 & 0.7 \\
+		\textit{Anopheles} & \textit{melas} & 2.8 & 3.3 & 3.7 & 4.1 & 4.8 & 3.8 & 0.7 \\
+		\textit{Anopheles} & \textit{culicifacies s.l.} & 2.8 & 3.3 & 3.7 & 4.1 & 4.7 & 3.7 & 0.6 \\
+		\textit{Anopheles} & \textit{subpictus s.l.} & 2.2 & 3.1 & 3.7 & 4.5 & 7.0 & 4.3 & 6.6 \\
+		\textit{Aedes} & \textit{samoanus} & 2.7 & 3.2 & 3.6 & 4.0 & 4.9 & 3.7 & 1.0 \\
+		\textit{Anopheles} & \textit{quadrimaculatus} & 2.4 & 3.1 & 3.5 & 4.0 & 4.8 & 3.6 & 0.8 \\
+		\textit{Anopheles} & \textit{stephensi} & 2.1 & 2.9 & 3.5 & 4.2 & 5.7 & 3.7 & 1.5 \\
+		\textit{Anopheles} & \textit{darlingi} & 2.3 & 2.9 & 3.4 & 3.8 & 4.7 & 3.4 & 0.9 \\
+		\textit{Aedes} & \textit{sollicitans} & 1.6 & 2.3 & 2.9 & 3.6 & 6.1 & 3.3 & 2.5 \\
+		\textit{Aedes} & \textit{vexans} & 1.7 & 2.2 & 2.7 & 3.1 & 4.2 & 3.0 & 3.3 \\
+		\textit{Anopheles} & \textit{cruzii} & 1.2 & 1.6 & 2.0 & 2.5 & 4.0 & 2.2 & 1.3 \\
+		\textit{Anopheles} & \textit{bellator} & 1.0 & 1.4 & 1.7 & 2.1 & 3.6 & 2.1 & 5.4 \\
+		\hline
+		\textit{Culex} & \textit{} & 2.5 & 4.0 & 5.4 & 7.6 & 12.3 & 6.2 & 3.2 \\
+		\textit{Anopheles} & \textit{} & 2.1 & 3.6 & 4.9 & 6.8 & 10.8 & 5.5 & 2.9 \\
+		\textit{Mansonia} & \textit{} & 3.3 & 4.0 & 4.5 & 4.9 & 5.6 & 4.5 & 0.8 \\
+		\textit{Aedes} & \textit{} & 2.5 & 3.3 & 3.8 & 4.3 & 5.2 & 3.8 & 0.9 \\
+		\hline
+		\textit{Overall} & \textit{} & 2.0 & 3.3 & 4.7 & 6.4 & 10.2 & 5.2 & 2.6 \\
+	\end{tabular}
+	\caption{\textbf{Polovodova dissection: mean female mosquito lifespan estimates in terms of chronological time.} The 5\%, 25\%, 50\%, 75\% and 95\% columns indicate the respective quantiles of the posterior distribution for mean lifespan, and the `Mean' and `Std.
+dev.' columns indicate the posterior mean and standard deviation of mean lifespan.
+Units of lifespan are days.}
+	\label{tab:dissection_estimated_lifespans_chron}%
+\end{table}
+
 \section{Evidence for senescence}
@@
 	\label{fig:dissection_senescence}
-\end{figure}
-
-\begin{figure}[ht]
-	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/elpd_comparison.pdf}}
-	\caption{\textbf{Evidence for senescence: comparing evidence from Gompertz model across MRR and Polovodova dissection analyses.} The points indicate the estimated predictive accuracy compared with the exponential model (see SOM Methods) and the lower and upper whiskers represent 2.5\% and 97.5\% quantiles on this quantity.}\label{fig:elpd_comparison}
 \end{figure}

### 2021-05-17 writeUp/figures.tex
very nearly at a draft
updated order as per Charles comments
Updated text and a smartened a few figures

@@
 \begin{figure}[h]
-	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/dissection_lifetimes_exponential.pdf}}
-	\caption{\textbf{Polovodova dissection: grouped physiological lifespan estimates.} The lifespans shown are for mosquitoes that were not fed with sugar or blood (for females) before release.
-The middle line in each box shows the median estimates and the solid dot indicates the mean.
-The left and right box edges show the 25\%, and 75\% posterior quantiles respectively.
-The whiskers show the range of the data, excluding points lying more than 1.5 times the interquartile range away from each edge of the box.
-The numbers before the start of the left whisker indicate the number of individual time-series within each species.
-All estimates were obtained using the hierarchical exponential survival model.}
-	\label{fig:dissection_lifetimes_exponential}
-\end{figure}
-
-\begin{figure}[h]
 	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/detinova_lifespans_combined_parity_no_insecticide.pdf}}
@@
 	\label{fig:detinova_lifetimes}
+\end{figure}
+
+\begin{figure}[h]
+	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/dissection_lifetimes_exponential.pdf}}
+	\caption{\textbf{Polovodova dissection: grouped physiological lifespan estimates.} The middle line in each box shows the median estimates and the solid dot indicates the mean.
+The left and right box edges show the 25\%, and 75\% posterior quantiles respectively.
+The whiskers show the range of the data, excluding points lying more than 1.5 times the interquartile range away from each edge of the box.
+The numbers before the start of the left whisker indicate the number of individual time-series within each species.
+All estimates were obtained using the hierarchical exponential survival model.}
+	\label{fig:dissection_lifetimes_exponential}
 \end{figure}
@@
 \begin{figure}[h]
-	\centerline{\includegraphics[width=1.3\textwidth]{./Figure_files/detinova_gambiae_daynight_range_lifespan_country.pdf}}
+	\centerline{\includegraphics[width=1.3\textwidth]{./Figure_files/detinova_gambiae_temp_range_lifespan.pdf}}
 	\caption{\textbf{Detinova dissection: day-night temperature range versus lifespan estimates for \textit{Anopheles gambiae s.l.}.} Points indicate the posterior mean estimates and the whiskers show 25\% and 75\% quantiles.

### 2021-05-17 writeUp/supplementary_materials.tex
Updated text and a smartened a few figures

@@
 \begin{table}[htbp]
+	\begin{tabular}{l|l|l|l|l}
+		\textbf{Analysis} & \textbf{Grouping} & \textbf{Log-likelihood} & \textbf{z} & \textbf{p}\\
+		\hline
+		MRR & genus & -7315.2 & - & -\\
+		& overall & -7511.0 & 11.3 & 0.0\\
+		& species & -7748.9 & 5.5 & 0.0\\
+		\hline
+		Detinova & species & -14,672.4 & - & -\\
+		& species-complex & -15,504.9 & 3.6 & 0.0\\
+		& continent & -16,731.1 & 0.4 & 0.3\\
+		\hline
+		Polovodova & species & -3818.3 & - & -\\
+		& genus & -3848.3 & 0.4 & 0.3\\
+		& overall & -3895.5 & 0.4 & 0.4\\
+	\end{tabular}
+		\caption{\textbf{Model comparison by K-fold cross-validation.} Here, we present the estimates of model fit on hold-out datasets (see SOM) as measured by the sum of the pointwise log-likelihood values averaged across all posterior draws (in the ``Log-likelihood'' column) across each of the three analyses and each grouping of studies considered.
+Within each row-block (``MRR'', ``Detinova'', ``Polovodova''), the models are arranged according to the point estimate of the fit: best at top; worst at bottom.
+The $z$ scores show the scaled difference in the averaged pointwise log-likelihood \citep{vehtari2015efficient} versus the model immediately above it in the ranking, and the ``p'' column shows the corresponding $p$ values representing the cumulative density function from a standard normal.}
+\end{table}
+
+\begin{table}[htbp]
 	\begin{tabular}{l|l|l|l|l|l|l|l|l}
@@
 \begin{figure}[h]
-	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_genus_vs_species_vs_top.pdf}}
-	\caption{\textbf{MRR: predictive accuracy of the species- and overall-level models relative to the genus-level model.} The predictive accuracy was determined using K-Fold cross-validation as described in SOM; the measure of accuracy presented here by the central dots is the difference in estimated expected log pointwise predictive density compared to the exponential model.
-The lower and upper whiskers represent the lower and upper bounds of the 95\% confidence interval in predictive accuracy.}
-	\label{fig:mrr_genusTopLevel}
-\end{figure}
-
-\begin{figure}[h]
-	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_female_blood_sugar.pdf}}
-	\caption{\textbf{MRR: female mosquito mean lifespan by pre-release feeding status.} The lifespans shown are for mosquitoes that were not fed with sugar or blood (for females) before release.
-The middle line in each box shows the median estimates and the solid dot indicates the mean.
+	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_female_blood_sugar_inkscaped.pdf}}
+	\caption{\textbf{MRR: female mosquito mean lifespan by pre-release feeding status.} Labels to the left of the vertical axis indicate genus or overall grouping; the labels to the right of the axis indicate pre-release feeding status.
+The middle line in each box shows the median estimates and the solid point indicates the mean.
 The left and right box edges show the 25\%, and 75\% posterior quantiles respectively.
 The whiskers show the range of the data, excluding points lying more than 1.5 times the interquartile range away from each edge of the box.
-The numbers before the start of the left whisker indicate the number of individual time-series within each species.
 All estimates were obtained using the hierarchical exponential survival model.}
@@
 	\label{fig:mrr_lifeSpanVsRange}
-\end{figure}
-
-\begin{figure}[h]
-	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_lifespanVsTrapDensity.pdf}}
-	\caption{\textbf{MRR: mean mosquito lifespan versus trapping density.} The markers show the median posterior lifespan estimates for each time series with the lower and upper bounds indicating the 25\% and 75\% quantiles, respectively.
-The trap density was estimated assuming traps were contained within a circular area with radius equal to the trapping range.
-The black line shows a linear regression line estimated using the median posterior lifetimes, with the grey shading indicating 95\% confidence intervals.
-All estimates were obtained using the non-hierarchical exponential survival model.}
-	\label{fig:mrr_lifeSpanVsTrapDensity}
 \end{figure}
@@
 	\label{fig:detinova_lifespan_individual}
-\end{figure}
-
-\begin{figure}[h]
-	\centerline{\includegraphics[width=1.3\textwidth]{./Figure_files/dissection_individualEstimates_allSpecies.pdf}}
-	\caption{\textbf{Polovodova dissection: individual time-series estimates of adult mosquito mean lifespan.} The middle line in each box shows the median estimates.
-The left and right box whiskers show the 25\%, and 75\% posterior quantiles respectively.
-All estimates were obtained using the non-hierarchical exponential survival model.}
-	\label{fig:dissection_individualEstimates_allSpecies}
 \end{figure}
@@
 \begin{figure}[ht]
-	\centerline{\includegraphics[width=1.25\textwidth]{./Figure_files/mrr_mcPowerAnalysis_senescence.pdf}}
-	\caption{\textbf{MRR: statistical power analysis for senescence detection.} The plot shows the statistical power to detect senescence for A.
-three case study populations as a function of B.
-study length and C.
-number released.
-For each parameter set we generated 500 simulated data series using the negative binomial sampling model with a gompertz hazard function ($\lambda(t) = \alpha e^{\beta t}$), and estimated the $\beta$ parameter using maximum likelihood and tested the null hypothesis $\beta=0$ (constant mortality risk) against the alternative $\beta>0$ using an likelihood ratio test with a 5\% test size.
-The error bars show the standard deviation in the power for each parameter set.
-The recapture probability parameters ($\psi$ and $\kappa$) used were the averages of those estimated for the actual data.}
-	\label{fig:mrr_mcPowerAnalysis_senescence}
-\end{figure}
-
-\begin{figure}[ht]
 	\centerline{\includegraphics[width=1.25\textwidth]{./Figure_files/elpd_vs_exponential_ordered_dissection.png}}
@@
 
+\begin{figure}[ht]
+	\centerline{\includegraphics[width=1.25\textwidth]{./Figure_files/mrr_mcPowerAnalysis_senescence.pdf}}
+	\caption{\textbf{MRR: statistical power analysis for senescence detection.} The plot shows the statistical power to detect senescence for A.
+three case study populations as a function of B.
+study length and C.
+number released.
+For each parameter set we generated 500 simulated data series using the negative binomial sampling model with a gompertz hazard function ($\lambda(t) = \alpha e^{\beta t}$), and estimated the $\beta$ parameter using maximum likelihood and tested the null hypothesis $\beta=0$ (constant mortality risk) against the alternative $\beta>0$ using an likelihood ratio test with a 5\% test size.
+The error bars show the standard deviation in the power for each parameter set.
+The recapture probability parameters ($\psi$ and $\kappa$) used were the averages of those estimated for the actual data.}
+	\label{fig:mrr_mcPowerAnalysis_senescence}
+\end{figure}
+
+\bibliographystyle{authordate1}
+\bibliography{Malaria}
+
 \end{document}

### 2021-07-31 writeUp/figures.tex
updated figure caption about ordering

@@
 	\caption{\textbf{MRR: lifespan estimates for each time series.} The middle point shows the median estimates, and the left and right box whiskers show the 25\%, and 75\% posterior quantiles respectively.
-All estimates were obtained using the non-hierarchical exponential survival model as described in text.}
+All estimates were obtained using the non-hierarchical exponential survival model as described in text.
+The individual estimates within each species are ordered according to the median lifespan estimates; the species estimates are ordered corresponding to the species-level medians.}
 	\label{fig:mrr_lifespan_individualEstimates}

### 2021-08-01 writeUp/supplementary_materials.tex
updated som

@@
 \section{Power analysis of MRR studies}
-Consider an ideal MRR experiment, where the assumptions of no emigration and harmless marking are fully satisfied.
+Consider an ideal MRR experiment, where mosquitoes do not migrate out of the study area and marking is harmless.
 With these data, how accurately can we estimate mosquito lifespan?
 To address this question we generated artificial MRR data and attempted to estimate the (known) parameters by maximum likelihood (a `Monte Carlo' analysis).
-Specifically, we simulated releases of $N$ mosquitoes that are monitored for $m$ days in each case, and determined how statistical power to estimate lifespan depends on these two parameters (Fig.
-\ref{fig:mrr_mcPowerAnalysis}).
-In order to focus on estimating lifespan, we assumed the mosquitoes experience constant mortality, and recaptures follow the negative binomial sampling model (eqn.
-\ref{eq:NB}) where the recapture probability parameters ($\psi$ and $\kappa$) were the averages of those estimated for the actual data.
+Specifically, we simulated releases of $N$ mosquitoes that are monitored for $m$ days in each case (with collections occurring on each day), and determined how the errors in estimating lifespan depended on these two parameters.
+In order to focus on estimating lifespan, we assumed the mosquitoes experience constant mortality, and recaptures follow the negative binomial sampling model where the recapture probability parameters ($\psi=0.03$ and $\kappa=3.18$) were the averages of the posterior draws across all species in the species-level model.
+We assume that all parameters: the true lifespan, $\psi$ and $\kappa$ were unknown and were estimated via maximum likelihood.
+Here, we investigated two lifespans: 5 days and 10 days indicative of estimates obtained in our analyses.
 
@@
 \ref{fig:mrr_mcPowerAnalysis}B).
-For the parameters we use, there are significant gains in accuracy from releasing $1,000$ rather than $100$ mosquitoes, but modest gains from releasing $10,000$ rather than $1000$ mosquitoes.
+For the parameters we use, there are significant gains in accuracy from releasing $1000$ rather than $100$ mosquitoes, but modest gains from releasing $10,000$ rather than $1000$ mosquitoes.
+
+Need to include information about power analysis lifespan!
 
@@
 	\caption{\textbf{MRR: Monte Carlo analysis of errors in predicting mean lifespan.} This shows the Root Mean Square Error (RMSE) in predicting mean lifespan as a function of study length (left) and number of released mosquitoes (right) in a Monte Carlo analysis.
-For each parameter set we generated 400 simulated data series using the negative binomial sampling model with an exponential survival function as described in the text, and estimated the mean lifespan using maximum likelihood.
+For each parameter set we generated 200 simulated data series using the negative binomial sampling model with an exponential survival function as described in the text, and estimated the mean lifespan using maximum likelihood.
 The error bars show the standard deviation in the prediction error.}

### 2022-05-25 src/s_graph_power_analysis.Rmd
added files necessary for submission

---
title: "Power analysis: 2021"
output: html_notebook
---

```{r}
rm(list=ls())
library(tidyverse)
library(reshape2)
```
Duration power analysis
```{r}
durations <- read.csv("power_duration_t.csv", header = F)$V1
durations_rep <- rep(durations, 2)
lower <- read.csv("power_duration_1.csv", header = F) %>% 
  t() %>% 
  as.data.frame() %>% 
  melt() %>% 
  mutate(duration=durations_rep) %>% 
  rename(lower=value)
middle <- read.csv("power_duration_2.csv", header = F) %>% 
  t() %>% 
  as.data.frame() %>% 
  melt() %>% 
  mutate(duration=durations_rep) %>% 
  rename(middle=value)
upper <- read.csv("power_duration_3.csv", header = F) %>% 
  t() %>% 
  as.data.frame() %>% 
  melt() %>% 
  mutate(duration=durations_rep) %>% 
  rename(upper=value)

lookup <- tribble(~variable, ~lifespan,
                  "V1", "5 days",
                  "V2", "10 days")

combined <- lower %>%
  left_join(middle) %>% 
  left_join(upper) %>% 
  left_join(lookup)

combined %>% 
  ggplot(aes(x=duration, y=middle, colour=lifespan)) +
  geom_pointrange(aes(ymin=lower, ymax=upper)) +
  geom_line() +
  scale_y_log10() +
  xlab("Study duration, days") +
  ylab("Absolute error in lifespan estimate") +
  scale_color_brewer("True\nlifepan", palette = "Dark2")
```

Release size power analysis
```{r}
durations <- read.csv("power_size_n.csv", header = F)$V1
durations_rep <- rep(durations, 2)
lower <- read.csv("power_size_1.csv", header = F) %>% 
  t() %>% 
  as.data.frame() %>% 
  melt() %>% 
  mutate(duration=durations_rep) %>% 
  rename(lower=value)
middle <- read.csv("power_size_2.csv", header = F) %>% 
  t() %>% 
  as.data.frame() %>% 
  melt() %>% 
  mutate(duration=durations_rep) %>% 
  rename(middle=value)
upper <- read.csv("power_size_3.csv", header = F) %>% 
  t() %>% 
  as.data.frame() %>% 
  melt() %>% 
  mutate(duration=durations_rep) %>% 
  rename(upper=value)

lookup <- tribble(~variable, ~lifespan,
                  "V1", "5 days",
                  "V2", "10 days")

combined <- lower %>%
  left_join(middle) %>% 
  left_join(upper) %>% 
  left_join(lookup)

combined %>% 
  ggplot(aes(x=duration, y=middle, colour=lifespan)) +
  geom_pointrange(aes(ymin=lower, ymax=upper)) +
  geom_line() +
  scale_y_log10() +
  scale_x_log10() +
  xlab("Release size") +
  ylab("Absolute error in lifespan estimate") +
  scale_color_brewer(palette = "Dark2")
```

### 2022-05-25 src/s_mcmc_for_parity_for_reference_only.R

@@ -1,0 +1,642 @@
+setwd("C:/Users/Ben/Dropbox/Malaria/Data_analysis/Parity")
+rm(list=ls())
+library(rstan)
+library(shinystan)
+options(mc.cores=parallel::detectCores())
+aDF <- read.csv('aggregatedReordered_v1.csv')
+
+
+## Removing series where there is insecticide, or removing the nulliparous observations if
+## Np < X * P1, or removing series with insufficient observations
+fPrepareSpecies <- function(aDF,aSpeciesNumber,aThreshold,aThresholdSize){
+  aDF_gambiae <- aDF[aDF$SpeciesID==aSpeciesNumber,]
+  aDF_gambiae <- aDF_gambiae[aDF_gambiae$Control==1,]
+  aDF_gambiae <- aDF_gambiae[aDF_gambiae$Total>=aThresholdSize,]
+  lRemoveNP <- aDF_gambiae$Nulliparous<aThreshold*aDF_gambiae$X1P
+  obsDF <- aDF_gambiae[,c(11:24,27)]
+  lThreshold <- ifelse(obsDF$Threshold>0,obsDF$Threshold,-1)
+  lStartingT <- ifelse(lRemoveNP,1,0)
+  
+  ## If there are no studies left then return -1
+  if(nrow(aDF_gambiae)==0){
+    return(list(K=-1))
+  }
+  
+  ## Check whether thresholding has been recorded corrected
+  K <- nrow(obsDF)
+  for(i in 1:K){
+    aTempDF <- obsDF[i,]
+    if(lThreshold[i]>0){
+      if(lThreshold[i] < 13){
+        if(sum(aTempDF[(lThreshold[i]+1):13])>0){
+          print(paste0('an error occurred in ...',i))
+        }
+      }
+      
+    }
+  }
+  
+  mObs <- matrix(nrow = K,ncol = 14)
+  for(i in 1:K){
+    mObs[i,1:13] <- unlist(obsDF[i,1:13])
+    mObs[i,14] <- 0
+    if(lThreshold[i]>0){
+      mObs[i,(lThreshold[i]+1)] <- obsDF[i,14]
+    }
+  }
+  
+  if(sum(mObs) != sum(obsDF[,c(1:14)]))
+    print('an error with sums...')
+  
+  ## stack capture observations
+  lY <- vector()
+  lS <- vector(length = K)
[…]

### 2022-05-29 src/s_graph_power_analysis.Rmd
first draft

@@
   left_join(lookup)
+combined_duration <- combined %>%
+  mutate(type="A.")
 
-combined %>% 
+g1 <- combined %>% 
   ggplot(aes(x=duration, y=middle, colour=lifespan)) +
@@
   xlab("Study duration, days") +
-  ylab("Absolute error in lifespan estimate") +
+  ylab("RMSE in lifespan estimate") +
+  theme(legend.position = c(0.7, 0.83)) +
   scale_color_brewer("True\nlifepan", palette = "Dark2")
@@
   left_join(lookup)
+combined_release <- combined %>%
+  mutate(type="B.")
 
-combined %>% 
+g2 <- combined %>% 
   ggplot(aes(x=duration, y=middle, colour=lifespan)) +
@@
   xlab("Release size") +
-  ylab("Absolute error in lifespan estimate") +
-  scale_color_brewer(palette = "Dark2")
+  ylab("RMSE in lifespan estimate") +
+  theme(legend.position = c(0.7, 0.83)) +
+  scale_color_brewer("True\nlifepan", palette = "Dark2")
 ```
+
+```{r}
+library(cowplot)
+g <- plot_grid(g1, g2, nrow = 1, labels = c("A.", "B."))
+```
+
+```{r}
+save_plot("../writeUp/Figure_files/mrr_power_analysis.pdf", g)
+```

### 2022-05-29 writeUp/supplementary_materials.tex
first draft

@@
 \usepackage{listings}
-\usepackage{minted}
 \usepackage{framed}
@@
 
-\renewcommand\listingscaption{Stan code}
 \definecolor{apricot}{rgb}{0.984,0.81,0.69}
@@
 With these data, how accurately can we estimate mosquito lifespan?
-To address this question we generated artificial MRR data and attempted to estimate the (known) parameters by maximum likelihood (a `Monte Carlo' analysis).
-Specifically, we simulated releases of $N$ mosquitoes that are monitored for $m$ days in each case (with collections occurring on each day), and determined how the errors in estimating lifespan depended on these two parameters.
-In order to focus on estimating lifespan, we assumed the mosquitoes experience constant mortality, and recaptures follow the negative binomial sampling model where the recapture probability parameters ($\psi=0.03$ and $\kappa=3.18$) were the averages of the posterior draws across all species in the species-level model.
+To address this question, we generated artificial MRR data and attempted to estimate the (known) parameters by maximum likelihood (a `Monte Carlo' analysis).
+We simulated releases of $N$ mosquitoes that are monitored for $m$ days in each case (with collections occurring on each day) and determined how the errors in estimating lifespan depended on these two parameters.
+In order to focus on estimating lifespan, we assumed the mosquitoes experience constant mortality, and recaptures follow the negative binomial sampling model where the (known) recapture probability parameters $\psi=0.03$ and $\kappa=3.18$ were the averages of the posterior draws across all species in the species-level model.
 We assume that all parameters: the true lifespan, $\psi$ and $\kappa$ were unknown and were estimated via maximum likelihood.
 Here, we investigated two lifespans: 5 days and 10 days indicative of estimates obtained in our analyses.
+For each simulation, we determined the Root Mean Square Error (RMSE) in predicting mean lifespan.
 
-Unsurprisingly, the error in predicting lifespan declines as the duration of study is increased (Fig.
+Unsurprisingly, the error in predicting lifespan declines as the study duration increases (Fig.
 \ref{fig:mrr_mcPowerAnalysis}A).
 However once a study length is much longer than the lifespan of the majority of mosquitoes, the predictive power cannot be improved by extending the study duration.
-The effect of increasing release size (while holding study length constant) is similarly a case of diminishing returns (Fig.
+The effect of increasing release size (while holding study length constant) similarly exhibits diminishing returns (Fig.
 \ref{fig:mrr_mcPowerAnalysis}B).
@@
 
-Need to include information about power analysis lifespan!
-
 We also conducted a power analysis for the detection of senescence in MRR experiments (Fig.
 \ref{fig:mrr_mcPowerAnalysis_senescence}).
-Here we calculated the power of a maximum likelihood estimator of the `senescence parameter' $\beta$ of the Gompertz survival function (see Table \ref{tab:mrr_survivalDescription}) for case study populations with three different levels of senescence (Fig.
+Here we calculated the power to detect senescence assuming that lifespan follows a Gompertz survival function (see Table \ref{tab:mrr_survivalDescription}).
+We simulated MRR time series for single releases, for case study populations with three different levels of senescence assuming a Gompertz survival function (Fig.
 \ref{fig:mrr_mcPowerAnalysis_senescence}A.).
+We assumed that recaptures followed a negative binomial sampling model with $\psi=0.03$ and $\kappa=3.18$.
+Using each simulated time series of recaptures, we then performed two different maximum likelihood estimations: one, where we (falsely) assumed that lifespan followed an exponential distribution (i.e.
+there was no senescence); and another where we assumed it followed a Gompertz survival function.
+In both cases, we estimated the parameters of the corresponding survival function along with $\psi$ and $\kappa$.
+We then performed a likelihood ratio test (using a 5\% test size) comparing the fit of the Gompertz model to that of the simpler exponential model: in doing so, we determined whether there was evidence of senescence.
+We performed 500 experimental replicates at each parameter set.
+
 These indicate that the power to detect senescence is strongly dependent on study length (Fig.
-\ref{fig:mrr_mcPowerAnalysis_senescence}B.), but is insensitive to release size (Fig.
+\ref{fig:mrr_mcPowerAnalysis_senescence}B.) but is insensitive to release size (Fig.
 \ref{fig:mrr_mcPowerAnalysis_senescence}C.).
 \cite{clements1981analysis} conducted a meta-analysis of MRR and dissection-based experiments and found evidence of senescence that is, at least, qualitatively similar in magnitude to that of the `mild' case considered above.
-For this case detecting senescence with a power of 80\% requires a study length of at least 18 days.
-Since the median study length for experiments included in our analysis was 10 days (Table \ref{tab:mrr_aggregateData}) this could partly explain our failure to detect senescence at the species level.
+For this case, detecting senescence with a power of 80\% requires a study length of at least 18 days.
+Since the median study length for experiments included in our analysis was 10 days, (Table \ref{tab:mrr_aggregateData}) this could partly explain our failure to detect senescence at the species level.
 
 \begin{figure}[ht]
-	\centerline{\includegraphics[width=1.25\textwidth]{./Figure_files/mrr_mcPowerAnalysis.pdf}}
-	\caption{\textbf{MRR: Monte Carlo analysis of errors in predicting mean lifespan.} This shows the Root Mean Square Error (RMSE) in predicting mean lifespan as a function of study length (left) and number of released mosquitoes (right) in a Monte Carlo analysis.
-For each parameter set we generated 200 simulated data series using the negative binomial sampling model with an exponential survival function as described in the text, and estimated the mean lifespan using maximum likelihood.
-The error bars show the standard deviation in the prediction error.}
+	\centerline{\includegraphics[width=1.25\textwidth]{./Figure_files/mrr_power_analysis.pdf}}
+	\caption{\textbf{MRR: Monte Carlo analysis of errors in predicting mean lifespan.} This shows the Root Mean Square Error (RMSE) in predicting mean lifespan as a function of study length (A) and number of released mosquitoes (B) Monte Carlo analysis.
+For each parameter set, we generated 200 simulated data series using the negative binomial sampling model with an exponential survival function as described in the text and estimated the mean lifespan using maximum likelihood.
+The error bars show the 25\%-75\% quantiles, and the points show the medians.}
 	\label{fig:mrr_mcPowerAnalysis}
@@
 \begin{figure}[ht]
-	\centerline{\includegraphics[width=1.25\textwidth]{./Figure_files/mrr_mcPowerAnalysis_senescence.pdf}}
+	\centerline{\includegraphics[width=1.25\textwidth]{./Figure_files/power-analysis-senescence.pdf}}
 	\caption{\textbf{MRR: statistical power analysis for senescence detection.} The plot shows the statistical power to detect senescence for A.
@@
 For each parameter set we generated 500 simulated data series using the negative binomial sampling model with a gompertz hazard function ($\lambda(t) = \alpha e^{\beta t}$), and estimated the $\beta$ parameter using maximum likelihood and tested the null hypothesis $\beta=0$ (constant mortality risk) against the alternative $\beta>0$ using an likelihood ratio test with a 5\% test size.
-The error bars show the standard deviation in the power for each parameter set.
-The recapture probability parameters ($\psi$ and $\kappa$) used were the averages of those estimated for the actual data.}
+The error bars show the standard deviation in the power for each parameter set using groups of 40 simulated time series as bootstrap samples.}
 	\label{fig:mrr_mcPowerAnalysis_senescence}

### 2022-06-22 writeUp/power_analysis_only_hawaii.tex
added final submission

--- 2022-05-29 writeUp/supplementary_materials.tex
@@
 % \raggedright
-\title{A Meta-analysis of Longevity Estimates of Mosquito Vectors of Disease: Supplementary Figures}
-\author{Ben Lambert, Ace North, Charles Godfray}
+\title{Power analysis of MRR studies}
+\author{Ben Lambert}
 
 \begin{document}
-\maketitle
-
-\section{Lifespan estimates}
-
-\begin{table}[htbp]
-	\begin{tabular}{l|l|l|l|l}
-		\textbf{Analysis} & \textbf{Grouping} & \textbf{Log-likelihood} & \textbf{z} & \textbf{p}\\
-		\hline
-		MRR & genus & -7315.2 & - & -\\
-		& overall & -7511.0 & 11.3 & 0.0\\
-		& species & -7748.9 & 5.5 & 0.0\\
-		\hline
-		Detinova & species & -14,672.4 & - & -\\
-		& species-complex & -15,504.9 & 3.6 & 0.0\\
-		& continent & -16,731.1 & 0.4 & 0.3\\
-		\hline
-		Polovodova & species & -3818.3 & - & -\\
-		& genus & -3848.3 & 0.4 & 0.3\\
-		& overall & -3895.5 & 0.4 & 0.4\\
-	\end{tabular}
-		\caption{\textbf{Model comparison by K-fold cross-validation.} Here, we present the estimates of model fit on hold-out datasets (see SOM) as measured by the sum of the pointwise log-likelihood values averaged across all posterior draws (in the ``Log-likelihood'' column) across each of the three analyses and each grouping of studies considered.
-Within each row-block (``MRR'', ``Detinova'', ``Polovodova''), the models are arranged according to the point estimate of the fit: best at top; worst at bottom.
-The $z$ scores show the scaled difference in the averaged pointwise log-likelihood \citep{vehtari2015efficient} versus the model immediately above it in the ranking, and the ``p'' column shows the corresponding $p$ values representing the cumulative density function from a standard normal.}
-\end{table}
-
-\begin{table}[htbp]
-	\begin{tabular}{l|l|l|l|l|l|l|l|l}
-		\textbf{Genus} & \textbf{Species} & \textbf{5$\%$} & \textbf{25$\%$} & \textbf{50$\%$} & \textbf{75$\%$} & \textbf{95\%} &  \textbf{Mean} & \textbf{Std. dev.}\\
-		\hline
-		\textit{Aedes} & \textit{simpsoni s.l.} & 6.3 & 11.3 & 18.3 & 31.8 & 75.0 & 26.9 & 29.5 \\
-		\textit{Aedes} & \textit{communis} & 10.0 & 13.2 & 15.4 & 18.2 & 24.3 & 16.2 & 5.2 \\
-		\textit{Anopheles} & \textit{koliensis} & 6.7 & 10.5 & 15.2 & 24.0 & 56.5 & 21.9 & 25.2 \\
-		\textit{Anopheles} & \textit{punctulatus} & 7.2 & 10.7 & 15.1 & 22.8 & 48.9 & 20.3 & 28.6 \\
-		\textit{Aedes} & \textit{cantans} & 9.7 & 12.0 & 13.9 & 16.1 & 20.6 & 14.4 & 4.7 \\
-		\textit{Aedes} & \textit{albopictus} & 8.0 & 10.0 & 11.6 & 13.7 & 17.7 & 12.1 & 3.4 \\
-		\textit{Anopheles} & \textit{albimanus} & 5.9 & 8.6 & 11.6 & 16.9 & 35.7 & 15.2 & 18.7 \\
-		\textit{Anopheles} & \textit{darlingi} & 4.9 & 7.3 & 10.2 & 16.0 & 37.5 & 14.5 & 14.8 \\
-		\textit{Anopheles} & \textit{maculatus} & 5.7 & 7.9 & 10.2 & 13.7 & 24.3 & 12.2 & 8.7 \\
-		\textit{Anopheles} & \textit{culicifacies s.l.} & 4.6 & 7.3 & 9.1 & 11.4 & 18.2 & 10.1 & 6.9 \\
-		\textit{Anopheles} & \textit{sergenti} & 3.6 & 5.9 & 8.7 & 13.8 & 32.6 & 12.4 & 13.4 \\
-		\textit{Aedes} & \textit{triseriatus} & 5.6 & 6.6 & 7.5 & 8.5 & 10.7 & 7.7 & 1.8 \\
-		\textit{Anopheles} & \textit{farauti s.l.} & 4.2 & 5.4 & 6.4 & 7.7 & 10.5 & 6.8 & 2.1 \\
-		\textit{Aedes} & \textit{africanus} & 3.4 & 5.1 & 6.2 & 7.8 & 11.9 & 7.0 & 9.3 \\
-		\textit{Aedes} & \textit{aegypti} & 2.8 & 4.5 & 6.2 & 8.5 & 13.9 & 7.0 & 3.8 \\
-		\textit{Culex} & \textit{quinquefasciatus} & 4.3 & 5.2 & 5.9 & 7.0 & 9.4 & 6.3 & 1.8 \\
-		\textit{Culex} & \textit{nigripalpus} & 3.5 & 4.6 & 5.8 & 7.5 & 11.3 & 6.4 & 3.1 \\
-		\textit{Aedes} & \textit{notoscriptus} & 2.9 & 4.0 & 5.4 & 8.2 & 20.2 & 7.9 & 9.5 \\
-		\textit{Culex} & \textit{pipiens} & 3.4 & 4.4 & 5.2 & 6.4 & 9.0 & 5.6 & 2.0 \\
-		\textit{Anopheles} & \textit{saperoi} & 3.5 & 4.2 & 4.9 & 5.8 & 8.2 & 5.2 & 2.0 \\
-		\textit{Anopheles} & \textit{gambiae s.l.} & 3.0 & 3.8 & 4.4 & 5.1 & 6.4 & 4.5 & 1.1 \\
-		\textit{Aedes} & \textit{melanimon} & 3.3 & 3.9 & 4.3 & 4.8 & 5.7 & 4.4 & 1.0 \\
-		\textit{Anopheles} & \textit{funestus s.l.} & 2.9 & 3.6 & 4.2 & 4.8 & 6.0 & 4.3 & 1.0 \\
-		\textit{Anopheles} & \textit{lesteri} & 2.4 & 3.2 & 3.9 & 5.0 & 7.5 & 4.4 & 2.1 \\
-		\textit{Anopheles} & \textit{quadrimaculatus} & 3.3 & 3.6 & 3.9 & 4.2 & 4.8 & 4.0 & 0.5 \\
-		\textit{Anopheles} & \textit{pseudopunctipennis s.l.} & 2.3 & 2.9 & 3.5 & 4.3 & 6.7 & 3.8 & 1.9 \\
-		\textit{Aedes} & \textit{vexans} & 1.9 & 2.3 & 2.6 & 3.0 & 4.1 & 2.8 & 1.0 \\
-		\textit{Culex} & \textit{tarsalis} & 1.9 & 2.1 & 2.3 & 2.4 & 2.7 & 2.3 & 0.3 \\
-		\textit{Anopheles} & \textit{vestitipennis} & 1.3 & 1.8 & 2.2 & 2.9 & 4.4 & 2.5 & 1.2 \\
-		\textit{Anopheles} & \textit{albitarsis s.l.} & 1.3 & 1.6 & 1.9 & 2.2 & 2.9 & 2.0 & 1.0 \\
-		\textit{Culex} & \textit{tritaeniorhynchus} & 0.7 & 0.9 & 1.1 & 1.2 & 1.6 & 1.1 & 1.0 \\
-		\textit{Anopheles} & \textit{subpictus s.l.} & 0.4 & 0.6 & 0.8 & 1.1 & 2.4 & 1.1 & 2.6 \\
-		\hline
-		\textit{Aedes} & \text{} & 2.8 & 4.8 & 6.9 & 10.0 & 17.2 & 8.1 & 4.9 \\
-		\textit{Anopheles} & \text{} & 1.5 & 3.0 & 5.0 & 8.4 & 17.8 & 6.8 & 6.4 \\
-		\textit{Culex} & \text{} & 1.1 & 1.8 & 2.5 & 3.5 & 5.6 & 2.9 & 1.5 \\
-		\hline
-		\textbf{Overall} & \text{} & 1.4 & 2.8 & 4.6 & 7.5 & 15.4 & 6.0 & 5.2 \\
-	\end{tabular}
-	\caption{\textbf{MRR: mean unfed female mosquito lifespan estimates.} The 5\%, 25\%, 50\%, 75\% and 95\% columns indicate the respective quantiles of the posterior distribution for mean lifespan, and the `Mean' and `Std.
-dev.' columns indicate the posterior mean and standard deviation of mean lifespan.
-Units of lifespan are days.}
-	\label{tab:mrr_estimated_lifespans}%
-\end{table}%
-
-\begin{figure}[h]
-	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_female_blood_sugar_inkscaped.pdf}}
-	\caption{\textbf{MRR: female mosquito mean lifespan by pre-release feeding status.} Labels to the left of the vertical axis indicate genus or overall grouping; the labels to the right of the axis indicate pre-release feeding status.
-The middle line in each box shows the median estimates and the solid point indicates the mean.
-The left and right box edges show the 25\%, and 75\% posterior quantiles respectively.
-The whiskers show the range of the data, excluding points lying more than 1.5 times the interquartile range away from each edge of the box.
-All estimates were obtained using the hierarchical exponential survival model.}
-	\label{fig:mrr_female_blood_sugar}
-\end{figure}
-
-\begin{figure}[h]
-	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_lifeSpanVsRange.pdf}}
-	\caption{\textbf{MRR: mean mosquito lifespan versus trapping range.} The markers show the median posterior lifespan estimates for each time series with the lower and upper bounds indicating the 25\% and 75\% quantiles, respectively.
-The black line shows a linear regression line estimated using the median posterior lifetimes, with the grey shading indicating 95\% confidence intervals.
-All estimates were obtained using the non-hierarchical exponential survival model.}
-	\label{fig:mrr_lifeSpanVsRange}
-\end{figure}
-
-\begin{figure}[h]
-	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/detinova_insecticide.pdf}}
-	\caption{\textbf{Detinova dissection: estimated impact of insecticide on lifespan.} The middle line in each box shows the median estimates.
-The left and right box whiskers show the 10\%, and 90\% posterior quantiles respectively.}
-	\label{fig:detinova_insecticide}
-\end{figure}
-
-\begin{figure}[h]
-	\centerline{\includegraphics[width=1.3\textwidth]{./Figure_files/detinova_individual_lifespan_no_insecticide.pdf}}
-	\caption{\textbf{Detinova dissection: individual time-series estimates of adult mosquito mean lifespan.} The middle line in each box shows the median estimates.
-The left and right box whiskers show the 25\%, and 75\% posterior quantiles respectively.
-Note, that a handful of observations are not shown in the plot since their estimates exceeded the reasonable range (here deemed to be up to 12 cycles).}
-	\label{fig:detinova_lifespan_individual}
-\end{figure}
-
-\begin{table}[htbp!]
-	\begin{tabular}{l|l|l|l|l|l|l|l|l}
-		\textbf{Genus} & \textbf{Species} & \textbf{5$\%$} & \textbf{25$\%$} & \textbf{50$\%$} &
-		\textbf{75$\%$} & \textbf{95$\%$} & \textbf{Mean} & \textbf{Std. dev.} \\
-		\hline
-		\textit{Anopheles} & \textit{gambiae (Form M)} & 0.9 & 1.8 & 2.6 & 3.6 & 6.7 & 3.2 & 4.6 \\
-		\textit{Anopheles} & \textit{gambiae} & 0.6 & 1.1 & 1.7 & 2.7 & 5.6 & 2.2 & 2.2 \\
-		\textit{Anopheles} & \textit{arabiensis} & 0.4 & 0.9 & 1.5 & 2.6 & 6.3 & 2.3 & 3.9 \\
-		\textit{Anopheles} & \textit{gambiae s.l.} & 0.7 & 1.7 & 3.1 & 6.3 & 21.2 & 7.3 & 56.1 \\
-		\textit{Anopheles} & \textit{moucheti} & 1.0 & 1.9 & 2.8 & 4.2 & 8.9 & 3.8 & 6.3 \\
-		\textit{Anopheles} & \textit{funestus} & 0.8 & 1.6 & 2.7 & 4.4 & 10.6 & 4.0 & 10.6 \\
-		\textit{Anopheles} & \textit{funestus s.l.} & 0.6 & 1.2 & 2.1 & 3.7 & 9.8 & 3.3 & 4.8 \\
-		\textit{Anopheles} & \textit{nili s.l.} & 1.0 & 1.4 & 1.7 & 2.1 & 2.8 & 1.8 & 0.6 \\
-		\textit{Anopheles} & \textit{subpictus s.l.} & 0.9 & 1.5 & 2.1 & 2.8 & 4.8 & 2.4 & 1.8 \\
-		\textit{Anopheles} & \textit{dirus (formerly Sp. A)} & 0.5 & 1.0 & 1.5 & 2.2 & 4.2 & 1.8 &
-		1.5 \\
-		\textit{Anopheles} & \textit{dirus s.l.} & 0.6 & 1.2 & 2.0 & 3.1 & 7.4 & 2.9 & 6.8 \\
-		\textit{Anopheles} & \textit{aconitus} & 0.5 & 1.1 & 1.8 & 2.8 & 6.3 & 2.5 & 4.7 \\
-		\textit{Anopheles} & \textit{sundaicus s.l.} & 0.4 & 0.9 & 1.6 & 2.9 & 7.5 & 2.7 & 7.4 \\
-		\textit{Anopheles} & \textit{minimus s.l.} & 0.5 & 1.1 & 1.6 & 2.5 & 5.8 & 2.3 & 5.6 \\
-		\textit{Anopheles} & \textit{fluviatilis s.l.} & 0.4 & 0.9 & 1.4 & 2.3 & 5.1 & 2.0 & 3.0 \\
-		\textit{Anopheles} & \textit{maculatus s.l.} & 0.6 & 1.0 & 1.2 & 1.6 & 2.5 & 1.4 & 0.8 \\
-		\textit{Anopheles} & \textit{farauti s.l.} & 0.3 & 0.7 & 1.2 & 2.0 & 4.9 & 1.8 & 2.3 \\
-		\textit{Anopheles} & \textit{annularis s.l.} & 0.3 & 0.7 & 1.2 & 2.0 & 4.7 & 1.7 & 1.9 \\
-		\textit{Anopheles} & \textit{sinensis} & 0.4 & 0.8 & 1.1 & 1.7 & 3.0 & 1.4 & 1.0 \\
-		\textit{Anopheles} & \textit{culicifacies s.l.} & 0.5 & 0.8 & 1.1 & 1.5 & 2.3 & 1.2 & 0.6 \\
-		\textit{Anopheles} & \textit{anthropophagus (lesteri)} & 0.4 & 0.9 & 1.4 & 2.3 & 5.1 & 2.0 &
-		3.0 \\
-		\textit{Anopheles} & \textit{lesteri s.l.} & 0.3 & 0.6 & 1.1 & 1.9 & 4.7 & 1.7 & 2.4 \\
-		\textit{Anopheles} & \textit{barbirostris s.l.} & 0.4 & 0.7 & 1.0 & 1.4 & 2.8 & 1.2 & 1.3 \\
-		\textit{Anopheles} & \textit{albitarsis (formerly Sp C)} & 0.8 & 2.0 & 3.9 & 8.4 &
-		32.6 & 13.3 & \text{$>$100} \\
-		\textit{Anopheles} & \textit{albitarsis (formerly Sp A)} & 0.2 & 0.4 & 0.6 & 1.0 & 2.3 & 0.8
-		& 0.9 \\
-		\textit{Anopheles} & \textit{albitarsis (Sp. B)} & 0.1 & 0.3 & 0.6 & 0.9 & 1.9 & 0.7 & 0.8
-		\\
-		\textit{Anopheles} & \textit{albitarsis s.l.} & 0.4 & 0.9 & 1.5 & 3.0 & 8.3 & 2.8 & 5.8 \\
-		\textit{Anopheles} & \textit{nuneztovari s.l.} & 0.3 & 0.6 & 1.1 & 1.9 & 4.4 & 1.6 & 2.3 \\
-		\textit{Anopheles} & \textit{darlingi} & 0.2 & 0.7 & 1.2 & 2.4 & 6.7 & 2.2 & 4.2 \\
-		\textit{Anopheles} & \textit{albimanus} & 0.2 & 0.5 & 0.9 & 1.6 & 3.6 & 1.3 & 1.6 \\
-		\textit{Anopheles} & \textit{aquasalis} & 0.3 & 0.6 & 1.1 & 1.8 & 4.4 & 1.6 & 4.0 \\
-		\textit{Anopheles} & \textit{aquasalis s.l.} & 0.2 & 0.4 & 0.7 & 1.0 & 2.4 & 0.9 & 1.3 \\
-	\end{tabular}
-	\caption{\textbf{Detinova dissection: mean female mosquito lifespan estimates in terms of gonotrophic cycles.} The 5\%, 25\%, 50\%, 75\% and 95\% columns indicate the respective quantiles of the posterior distribution for mean lifespan, and the `Mean' and `Std.
-dev.' columns indicate the posterior mean and standard deviation of mean lifespan.}
-	\label{tab:detinova_lifespan_gonotrophic}%
-\end{table}
-
-\begin{table}[htbp]
-	\begin{tabular}{l|l|l|l|l|l|l|l|l}
-		\textbf{Genus} & \textbf{Species} & \textbf{5$\%$} & \textbf{25$\%$} & \textbf{50$\%$} & \textbf{75$\%$} & \textbf{95\%} &  \textbf{Mean} & \textbf{Std. dev.}\\
-		\hline
-		\textit{Anopheles} & \textit{sergentii} & 1.0 & 1.9 & 2.5 & 3.4 & 6.2 & 3.0 & 2.5 \\
-		\textit{Anopheles} & \textit{gambiae s.l.} & 0.7 & 1.2 & 1.9 & 2.9 & 5.4 & 2.4 & 1.9 \\
-		\textit{Culex} & \textit{thalassius} & 1.1 & 1.6 & 1.8 & 2.1 & 2.9 & 1.9 & 1.0 \\
-		\textit{Anopheles} & \textit{farauti s.l.} & 1.5 & 1.6 & 1.7 & 1.7 & 1.9 & 1.7 & 0.1 \\
-		\textit{Anopheles} & \textit{minimus} & 0.7 & 1.2 & 1.7 & 2.2 & 4.1 & 2.0 & 2.2 \\
-		\textit{Anopheles} & \textit{maculipennis} & 0.8 & 1.1 & 1.4 & 1.8 & 2.9 & 1.6 & 0.7 \\
-		\textit{Culex} & \textit{pipiens} & 1.1 & 1.3 & 1.4 & 1.5 & 1.7 & 1.4 & 0.2 \\
-		\textit{Mansonia} & \textit{uniformis} & 0.9 & 1.1 & 1.1 & 1.2 & 1.4 & 1.1 & 0.2 \\
-		\textit{Anopheles} & \textit{rivulorum} & 0.8 & 1.0 & 1.1 & 1.1 & 1.4 & 1.1 & 0.2 \\
-		\textit{Anopheles} & \textit{melas} & 0.8 & 1.0 & 1.0 & 1.1 & 1.3 & 1.0 & 0.2 \\
-		\textit{Anopheles} & \textit{culicifacies} & 0.9 & 1.0 & 1.0 & 1.1 & 1.2 & 1.0 & 0.1 \\
-		\textit{Anopheles} & \textit{subpictus} & 0.6 & 0.9 & 1.0 & 1.2 & 2.1 & 1.2 & 2.0 \\
-		\textit{Aedes} & \textit{aegypti} & 0.5 & 0.8 & 1.0 & 1.2 & 2.0 & 1.2 & 3.0 \\
-		\textit{Anopheles} & \textit{quadrimaculatus} & 0.7 & 0.9 & 1.0 & 1.1 & 1.3 & 1.0 & 0.2 \\
-		\textit{Anopheles} & \textit{stephensi} & 0.6 & 0.8 & 1.0 & 1.1 & 1.6 & 1.0 & 0.4 \\
-		\textit{Anopheles} & \textit{darlingi} & 0.7 & 0.8 & 0.9 & 1.0 & 1.3 & 1.0 & 0.2 \\
-		\textit{Culex} & \textit{quinquefasciatus} & 0.4 & 0.7 & 0.9 & 1.3 & 2.2 & 1.1 & 0.6 \\
-		\textit{Aedes} & \textit{polynesiensis} & 0.8 & 0.9 & 0.9 & 0.9 & 1.0 & 0.9 & 0.1 \\
-		\textit{Culex} & \textit{tritaeniorhynchus} & 0.3 & 0.6 & 0.9 & 1.3 & 4.2 & 2.5 & 30.8 \\
-		\textit{Culex} & \textit{annulirostris} & 0.5 & 0.8 & 0.9 & 1.0 & 1.4 & 0.9 & 0.5 \\
-		\textit{Aedes} & \textit{samoanus} & 0.6 & 0.7 & 0.8 & 0.9 & 1.0 & 0.8 & 0.2 \\
-		\textit{Aedes} & \textit{sollicitans} & 0.4 & 0.5 & 0.6 & 0.8 & 1.3 & 0.7 & 0.5 \\
-		\textit{Aedes} & \textit{vexans} & 0.4 & 0.5 & 0.6 & 0.7 & 0.9 & 0.6 & 0.7 \\
-		\textit{Anopheles} & \textit{cruzii} & 0.3 & 0.5 & 0.5 & 0.7 & 1.1 & 0.6 & 0.4 \\
-		\textit{Anopheles} & \textit{bellator} & 0.3 & 0.4 & 0.5 & 0.6 & 1.0 & 0.6 & 1.8 \\
-		\hline
-		\textit{Anopheles} & \textit{} & 0.6 & 1.0 & 1.4 & 1.9 & 3.2 & 1.6 & 0.9 \\
-		\textit{Mansonia} & \textit{} & 0.9 & 1.1 & 1.1 & 1.2 & 1.4 & 1.1 & 0.2 \\
-		\textit{Culex} & \textit{} & 0.5 & 0.8 & 1.0 & 1.4 & 2.3 & 1.2 & 0.6 \\
-		\textit{Aedes} & \textit{} & 0.6 & 0.7 & 0.8 & 0.9 & 1.1 & 0.8 & 0.2 \\
-		\hline
-		\textbf{Overall} & \textit{} & 0.5 & 0.8 & 1.2 & 1.7 & 2.7 & 1.3 & 0.7 \\
-	\end{tabular}
-	\caption{\textbf{Polovodova dissection: mean female mosquito lifespan in terms of gonotrophic cycles.} The 5\%, 25\%, 50\%, 75\% and 95\% columns indicate the respective quantiles of the posterior distribution for mean lifespan, and the `Mean' and `Std.
-dev.' columns indicate the posterior mean and standard deviation of mean lifespan.}
-	\label{tab:dissection_estimated_lifespans}%
-\end{table}
-
-\begin{table}[htbp!]
-	\begin{tabular}{l|l|l|l|l|l|l|l|l}
-		\textbf{Genus} & \textbf{Species} & \textbf{5$\%$} & \textbf{25$\%$} & \textbf{50$\%$} &
-		\textbf{75$\%$} & \textbf{95$\%$} & \textbf{Mean} & \textbf{Std. dev.} \\
-		\hline
-		\textit{Anopheles} & \textit{gambiae (Form M)} & 3.3 & 6.2 & 8.7 & 11.9 & 22.3 & 10.7 & 14.9
-		\\
-		\textit{Anopheles} & \textit{gambiae} & 2.0 & 3.8 & 5.8 & 9.0 & 18.4 & 7.6 & 7.1 \\
-		\textit{Anopheles} & \textit{arabiensis} & 1.4 & 3.2 & 5.3 & 8.8 & 20.8 & 7.7 & 11.9 \\
-		\textit{Anopheles} & \textit{gambiae s.l.} & 2.5 & 5.7 & 10.3 & 20.6 & 68.4 & 23.9 &
-		\text{$>$100} \\
-		\textit{Anopheles} & \textit{moucheti} & 3.5 & 6.4 & 9.2 & 13.9 & 29.0 & 12.5 & 19.7 \\
-		\textit{Anopheles} & \textit{funestus} & 2.9 & 5.7 & 9.0 & 14.5 & 34.4 & 13.4 & 31.0 \\
-		\textit{Anopheles} & \textit{funestus s.l.} & 2.0 & 4.3 & 7.1 & 12.5 & 32.2 & 11.1 & 15.6 \\
-		\textit{Anopheles} & \textit{nili s.l.} & 3.6 & 4.8 & 5.8 & 7.1 & 9.6 & 6.1 & 2.0 \\
-		\textit{Anopheles} & \textit{subpictus s.l.} & 3.4 & 5.3 & 7.1 & 9.5 & 16.2 & 8.2 & 6.2 \\
-		\textit{Anopheles} & \textit{dirus (formerly Sp. A)} & 1.9 & 3.6 & 5.1 & 7.4 & 13.8 & 6.2 &
-		5.0 \\
-		\textit{Anopheles} & \textit{dirus s.l.} & 2.0 & 4.4 & 6.8 & 10.5 & 24.1 & 9.8 & 23.0 \\
-		\textit{Anopheles} & \textit{aconitus} & 1.9 & 4.0 & 6.2 & 9.6 & 20.7 & 8.5 & 15.4 \\
-		\textit{Anopheles} & \textit{sundaicus s.l.} & 1.5 & 3.4 & 5.6 & 9.7 & 24.7 & 9.1 & 23.6 \\
-		\textit{Anopheles} & \textit{minimus s.l.} & 1.9 & 3.8 & 5.5 & 8.3 & 19.2 & 7.9 & 17.8 \\
-		\textit{Anopheles} & \textit{fluviatilis s.l.} & 1.6 & 3.2 & 5.0 & 7.8 & 16.9 & 6.8 & 9.6 \\
-		\textit{Anopheles} & \textit{maculatus s.l.} & 2.2 & 3.5 & 4.5 & 5.6 & 8.6 & 4.9 & 2.6 \\
-		\textit{Anopheles} & \textit{farauti s.l.} & 1.2 & 2.6 & 4.3 & 6.9 & 16.0 & 6.1 & 7.5 \\
-		\textit{Anopheles} & \textit{annularis s.l.} & 1.1 & 2.5 & 4.3 & 6.9 & 15.6 & 5.8 & 6.2 \\
-		\textit{Anopheles} & \textit{sinensis} & 1.6 & 2.8 & 4.1 & 5.9 & 10.3 & 4.8 & 3.2 \\
-		\textit{Anopheles} & \textit{culicifacies s.l.} & 1.9 & 3.0 & 4.0 & 5.2 & 8.0 & 4.3 & 2.0 \\
-		\textit{Anopheles} & \textit{anthropophagus (lesteri)} & 1.6 & 3.2 & 5.0 & 7.7 & 16.9 & 6.8
-		& 9.9 \\
-		\textit{Anopheles} & \textit{lesteri s.l.} & 0.9 & 2.2 & 3.9 & 6.5 & 15.6 & 5.7 & 7.8 \\
-		\textit{Anopheles} & \textit{barbirostris s.l.} & 1.4 & 2.5 & 3.5 & 4.9 & 9.6 & 4.3 & 4.1 \\
-		\textit{Anopheles} & \textit{albitarsis (formerly Sp C)} & 3.0 & 6.8 & 13.0 & 27.3
-		& \text{$>$100} & 44.2 & \text{$>$100} \\
-		\textit{Anopheles} & \textit{albitarsis (formerly Sp A)} & 0.6 & 1.3 & 2.2 & 3.7 & 7.7 & 3.0
-		& 3.1 \\
-		\textit{Anopheles} & \textit{albitarsis (Sp. B)} & 0.5 & 1.2 & 2.0 & 3.3 & 6.6 & 2.6 & 2.6
-		\\
-		\textit{Anopheles} & \textit{albitarsis s.l.} & 1.2 & 3.1 & 5.4 & 9.9 & 26.9 & 9.3 & 18.6 \\
-		\textit{Anopheles} & \textit{nuneztovari s.l.} & 1.0 & 2.3 & 3.9 & 6.4 & 14.7 & 5.6 & 7.6 \\
-		\textit{Anopheles} & \textit{darlingi} & 0.9 & 2.4 & 4.5 & 8.1 & 22.0 & 7.4 & 13.6 \\
-		\textit{Anopheles} & \textit{albimanus} & 0.9 & 2.0 & 3.3 & 5.4 & 12.0 & 4.5 & 5.3 \\
-		\textit{Anopheles} & \textit{aquasalis} & 1.0 & 2.3 & 3.9 & 6.3 & 14.4 & 5.6 & 12.9 \\
-		\textit{Anopheles} & \textit{aquasalis s.l.} & 0.8 & 1.6 & 2.4 & 3.8 & 8.2 & 3.3 & 4.1 \\
-	\end{tabular}
-	\caption{\textbf{Detinova dissection: mean female mosquito lifespan estimates in terms of chronological time.} The 5\%, 25\%, 50\%, 75\% and 95\% columns indicate the respective quantiles of the posterior distribution for mean lifespan, and the `Mean' and `Std.
-dev.' columns indicate the posterior mean and standard deviation of mean lifespan.
-Units of lifespan are days.}
-	\label{tab:detinova_lifespan_chronological}%
-\end{table}
-
-\begin{table}[htbp!]
-	\begin{tabular}{l|l|l|l|l|l|l|l|l}
-		\textbf{Genus} & \textbf{Species} & \textbf{5$\%$} & \textbf{25$\%$} & \textbf{50$\%$} & \textbf{75$\%$} & \textbf{95\%} &  \textbf{Mean} & \textbf{Std. dev.}\\
-		\hline
-		\textit{Culex} & \textit{thalassius} & 5.7 & 8.2 & 9.4 & 11.0 & 15.6 & 10.1 & 5.2 \\
-		\textit{Anopheles} & \textit{sergentii} & 3.7 & 6.4 & 8.5 & 11.5 & 20.6 & 10.1 & 8.1 \\
-		\textit{Culex} & \textit{pipiens} & 5.7 & 6.6 & 7.2 & 7.8 & 8.9 & 7.2 & 1.0 \\
-		\textit{Anopheles} & \textit{gambiae s.l.} & 2.4 & 4.3 & 6.4 & 9.8 & 18.3 & 8.0 & 6.3 \\
-		\textit{Anopheles} & \textit{farauti s.l.} & 4.8 & 5.5 & 5.9 & 6.3 & 6.9 & 5.9 & 0.6 \\
-		\textit{Anopheles} & \textit{minimus} & 2.7 & 4.4 & 5.8 & 7.7 & 13.7 & 6.9 & 8.1 \\
-		\textit{Anopheles} & \textit{maculipennis} & 2.8 & 4.0 & 5.1 & 6.4 & 9.6 & 5.5 & 2.3 \\
-		\textit{Culex} & \textit{quinquefasciatus} & 2.0 & 3.4 & 4.8 & 6.9 & 11.7 & 5.6 & 3.3 \\
-		\textit{Aedes} & \textit{aegypti} & 2.3 & 3.7 & 4.7 & 5.6 & 9.3 & 5.6 & 12.3 \\
-		\textit{Culex} & \textit{annulirostris} & 2.7 & 3.9 & 4.6 & 5.4 & 7.5 & 4.9 & 2.4 \\
-		\textit{Culex} & \textit{tritaeniorhynchus} & 1.6 & 3.2 & 4.6 & 7.2 & 22.8 & 13.0 & 174.0 \\
-		\textit{Mansonia} & \textit{uniformis} & 3.4 & 4.0 & 4.4 & 4.9 & 5.5 & 4.5 & 0.7 \\
-		\textit{Aedes} & \textit{polynesiensis} & 3.2 & 3.7 & 4.1 & 4.5 & 5.1 & 4.1 & 0.6 \\
-		\textit{Anopheles} & \textit{rivulorum} & 2.8 & 3.4 & 3.8 & 4.2 & 5.0 & 3.9 & 0.7 \\
-		\textit{Anopheles} & \textit{melas} & 2.8 & 3.3 & 3.7 & 4.1 & 4.8 & 3.8 & 0.7 \\
-		\textit{Anopheles} & \textit{culicifacies s.l.} & 2.8 & 3.3 & 3.7 & 4.1 & 4.7 & 3.7 & 0.6 \\
-		\textit{Anopheles} & \textit{subpictus s.l.} & 2.2 & 3.1 & 3.7 & 4.5 & 7.0 & 4.3 & 6.6 \\
-		\textit{Aedes} & \textit{samoanus} & 2.7 & 3.2 & 3.6 & 4.0 & 4.9 & 3.7 & 1.0 \\
-		\textit{Anopheles} & \textit{quadrimaculatus} & 2.4 & 3.1 & 3.5 & 4.0 & 4.8 & 3.6 & 0.8 \\
-		\textit{Anopheles} & \textit{stephensi} & 2.1 & 2.9 & 3.5 & 4.2 & 5.7 & 3.7 & 1.5 \\
-		\textit{Anopheles} & \textit{darlingi} & 2.3 & 2.9 & 3.4 & 3.8 & 4.7 & 3.4 & 0.9 \\
-		\textit{Aedes} & \textit{sollicitans} & 1.6 & 2.3 & 2.9 & 3.6 & 6.1 & 3.3 & 2.5 \\
-		\textit{Aedes} & \textit{vexans} & 1.7 & 2.2 & 2.7 & 3.1 & 4.2 & 3.0 & 3.3 \\
-		\textit{Anopheles} & \textit{cruzii} & 1.2 & 1.6 & 2.0 & 2.5 & 4.0 & 2.2 & 1.3 \\
-		\textit{Anopheles} & \textit{bellator} & 1.0 & 1.4 & 1.7 & 2.1 & 3.6 & 2.1 & 5.4 \\
-		\hline
-		\textit{Culex} & \textit{} & 2.5 & 4.0 & 5.4 & 7.6 & 12.3 & 6.2 & 3.2 \\
-		\textit{Anopheles} & \textit{} & 2.1 & 3.6 & 4.9 & 6.8 & 10.8 & 5.5 & 2.9 \\
-		\textit{Mansonia} & \textit{} & 3.3 & 4.0 & 4.5 & 4.9 & 5.6 & 4.5 & 0.8 \\
-		\textit{Aedes} & \textit{} & 2.5 & 3.3 & 3.8 & 4.3 & 5.2 & 3.8 & 0.9 \\
-		\hline
-		\textit{Overall} & \textit{} & 2.0 & 3.3 & 4.7 & 6.4 & 10.2 & 5.2 & 2.6 \\
-	\end{tabular}
-	\caption{\textbf{Polovodova dissection: mean female mosquito lifespan estimates in terms of chronological time.} The 5\%, 25\%, 50\%, 75\% and 95\% columns indicate the respective quantiles of the posterior distribution for mean lifespan, and the `Mean' and `Std.
-dev.' columns indicate the posterior mean and standard deviation of mean lifespan.
-Units of lifespan are days.}
-	\label{tab:dissection_estimated_lifespans_chron}%
-\end{table}
-
-\section{Evidence for senescence}
-
-\begin{figure}[ht]
-	\centerline{\includegraphics[width=1.25\textwidth]{./Figure_files/elpd_vs_exponential_ordered_dissection.png}}
-	\caption{\textbf{Polovodova dissection: evidence for senescence.}  Here, we consider the predictive accuracy of age-dependent models of mortality versus the exponential model by species.
-The predictive accuracy was determined using K-Fold cross-validation as described in SOM; the measure of accuracy presented here by the central dots is the difference in estimated expected log pointwise predictive density compared to the exponential model.
-The lower and upper whiskers represent the lower and upper bounds of the 95\% confidence interval in predictive accuracy.
-The species have been ordered so that those with the highest average predictive accuracy across all age-dependent models compared to the exponential model appear at the top.
-If all age-dependent models outperform the exponential, we deem this as evidence for age-dependent mortality (`+'); if a subsample of models perform better, we deem the evidence ambiguous (`?'); and if no age-dependent model exceeds the performance of the exponential, we conclude no evidence in favour of senescence (`-').}
-	\label{fig:dissection_senescence}
-\end{figure}
 

### 2022-06-22 writeUp/supplementary_materials_methods.tex
added final submission

@@
 
-The data collection method resulted in 568 separate dissection series, across 72 published articles (see Section \ref{sec:appendix_dissectionStudyList} for a list of the studies included in our final dataset).
+The data collection method resulted in 568 separate dissection series (see \verb|polovodova_full.csv|), across 73 published articles (see \verb|parity_papers_summary.csv|).
 The published datasets cover the period from 1960-2015 with comparable numbers of studies across each decade (Fig.
@@
 If populations fluctuate strongly from month-to-month, this assumption will likely be violated.
-To try to mitigate against such a risk, we aggregated the data across all dates to obtain a single series for each identifier, resulting in 201 such series.
+To try to mitigate against such a risk, we aggregated the data across all dates to obtain a single series for each identifier, resulting in 201 such series (see \verb|polovodova_aggregated.csv|).
 To obtain a reasonable level of accuracy on estimates, we removed all those individual (aggregated) series where fewer than 100 mosquitoes were collected.
@@
 Based on the abstracts of the resultant list of published studies, we then decided whether to search each article for estimates of the duration of the gonotrophic cycle.
-Overall 79 separate estimates of this parameter were found across 42 published articles (see Section \ref{sec:appendix_gonotrophicDurationStudyList} for a list of the studies included in our dataset).
+Overall 79 separate estimates of this parameter were found across 42 published articles (see \verb|gonotrophic.csv|).
 Along with information about the estimates, we also recorded study and series meta-data, including the location of the study, the method used for estimation, species and genus.
@@
 
-Along with point estimates of the parameter we also collected information about the uncertainty in the estimates (if available).
+Along with point estimates of the parameter, we also collected information about the uncertainty in the estimates (if available).
 In many articles, the duration of the gonotrophic cycle was estimated separately for the first versus subsequent cycles, and these estimates were recorded separately.
